### 2019-10-23 ex3/paper/paper.tex
Add paper draft with correct template

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[conference]{IEEEtran}

\usepackage[utf8]{inputenc}
\usepackage{multirow}
\usepackage[inline]{enumitem}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{natbib}

\usepackage[nomain, toc, acronym]{glossaries}
\glsdisablehyper

\newcommand\note[2]{{\color{#1}#2}}
\newcommand\todo[1]{{\note{red}{TODO: #1}}}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Packet-by-Packet Hardening of Intrusion Detection Systems against Smart Adversaries}

\author{\IEEEauthorblockN{Alexander Hartl,
Maximilian Bachl, Joachim Fabini, Tanja Zseby}
\IEEEauthorblockA{Technische Universität Wien\\
firstname.lastname@tuwien.ac.at}}

%\date{\today}

% \IEEEoverridecommandlockouts
% \IEEEpubid{\begin{minipage}[t]{\textwidth}\ \\[10pt]
%         \centering\normalsize{xxx-x-xxxx-xxxx-x/xx/\$31.00 \copyright 2018 IEEE}
% \end{minipage}}

% \renewcommand*{\bibfont}{\footnotesize}

\maketitle%

\thispagestyle{plain}
\pagestyle{plain}

\newacronym{ml}{ML}{Machine Learning}
\newacronym{dl}{DL}{Deep Learning}
\newacronym{aml}{AML}{Adversarial Machine Learning}
\newacronym{ids}{IDS}{Intrusion Detection System}

\begin{abstract}

With the rise of ubiquitous \gls{ml} systems, malicious actors have been catching up quickly to find new ways to exploit them for profit. Recently developed \gls{aml} techniques focus on computer vision and their applicability to network traffic is not straightforward: Network traffic is not only sequential but there are also many constraints that have to be considered when attacking/defending an \gls{ml}-based Intrusion Detection System. For example, attackers can only modify certain features to evade detection and certain features, like packet interarrival time cannot become less than zero. We show that already the first couple of packets of a flow are of crucial importance and are likely going to be targeted by attackers. As a countermeasure we develop an easy-to-use training procedure and can confirm that it significantly and successfully reduces the attack surface.

\end{abstract}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
%\begin{abstract}
%  A clear and well-documented \LaTeX\ document is presented as an
%  article formatted for publication by ACM in a conference proceedings
%  or journal publication. Based on the ``acmart'' document class, this
%  article presents and explains many of the common variations, as well
%  as many of the formatting elements an author may use in the
%  preparation of the documentation of their work.
%\end{abstract}

\maketitle

\todo{Legend in some figures is misplaced}

\section{Introduction}

There is a significant body of scientific work focusing on the detection of unwanted behavior in networks. In the past, a viable way of performing intrusion detection was to inspect the content of packets themselves and detect if a packet delivers potentially harmful content.

More recently, with the rise of omnipresent encryption, available features have changed. Since payload is not available anymore, the focus now lies on features that are always available to observers such as port numbers, flags of packets etc. Network communication is usually grouped into \textit{flows}, which is commonly defined as a sequence of packets between two applications that may be located at different physical locations.

When analyzing flows, not only the aforementioned features are available but also features related to the timing of the individual packets as well as their size. For example, a specific attack might require the attacker to send three small packets and the victim to reply with two large packets followed by a small one the attacker. Various approaches have been proposed to extract features from flows and then perform anomaly detection with the results \cite{meghdouri_analysis_2018}. While this approach works well, it is problematic that first the whole flow has to be received and only afterwards it can be determined if there was an anomaly.

Thus, we design a network anomaly detection system that operates on a per-packet basis and decides if a packet is anomalous based on features that are available for traffic that is encrypted above the transport layer, like for example TLS or QUIC.
We show that our system has similar performance to other flow-based anomaly detection systems but can commonly detect anomalies before the flow is over.

Then we investigate the question whether adversarial samples can be found for this type of data. This is not a trivially answerable question since adversarial samples have mostly been analyzed in the context of computer vision. Our scenario significantly deviates from this in several ways because \begin{enumerate*}
\item the number of features is significantly smaller and
\item only certain features can be manipulated (for example the destination port cannot be manipulated)
\end{enumerate*}.

We can confirm that adversarial samples can be found even when considering the tight constraints mentioned before. Especially the first packets of a flow are critical for determining the decision of a recurrent \gls{ids}. We develop a simple methods for defending against adversaries who can manipulate arbitrary packets by simply leaving out potentially manipulable information from the classification process. The resulting classifier has slightly reduces accuracy but is still useful to be deployed in a real scenario. Finally we develop a custom adversarial training procedure which especially focuses on easy of use and low resource consumption and aims to keep the number of additionally introduced training hyperparameters low. With this novel strategy we can reduce attack surface and harden the resulting \gls{ids} while keeping similar classification performance like for the original \gls{ml} model.

\section{Related work}

\section{RNN-based classifier}

\begin{table}[b]
\caption{Occurence probability of attack types.} \label{tab:occurrence}
\centering
\begin{tabular}{l r} \toprule
Attack type & Proportion \\
\midrule
Normal                                                         & $0.747468$ \\
DoS / DDoS:DoS Hulk                                            & $0.101014$ \\
PortScan:PortScan - Firewall off                               & $0.069020$ \\
DDoS:LOIT                                                      & $0.040784$ \\
Infiltration:Dropbox download & $0.032982$ \\
DoS / DDoS:DoS GoldenEye                                       & $0.003224$ \\
DoS / DDoS:DoS Slowhttptest                                    & $0.001819$ \\
DoS / DDoS:DoS slowloris                                       & $0.001680$ \\
Brute Force:SSH-Patator                                        & $0.001107$ \\
Botnet:ARES                                                    & $0.000327$ \\
Web Attack:XSS                                                 & $0.000293$ \\
PortScan:PortScan - Firewall on                                & $0.000165$ \\
Brute Force:FTP-Patator                                        & $0.000110$ \\
Web Attack:Sql Injection                                       & $0.000006$ \\
DoS / DDoS:Heartbleed                                          & $0.000001$ \\
\bottomrule
\end{tabular}
\label{tab:occurence}
\end{table}

The dataset we use is the \textit{CIC-IDS 2017} dataset \cite{sharafaldin_toward_2018}, which includes the attacks listed in \autoref{tab:occurence} and contains over two million flows in total. We implement an LSTM-based classifier with three layers of 512 neurons each. As the input features we use \begin{itemize*}
\item source port
\item destination port
\item protocol identifier
\item packet sizes
\item interarrival times of packets
\item packet directions (if packet goes there or back)
\item all TCP flags (0 if flow is not TCP)
\end{itemize*}. Of these features, source port, destination port and protocol identifier are constant over the whole flow while the others vary.

We use Z-score normalization and a train/test split of 2:1.

\autoref{tab:performance_results} shows that our classifier achieves an accuracy that is similar to the one of the random forest and the multilayer perceptron that we developed in the previous exercise. However unlike these classifiers, our recurrent classifier has the advantage that it can detect attacks already before they are over.

\begin{table}
\caption{Detection performance results. Random forest and Multilayer Perceptron are from the previous exercise, LSTM is this implementation.} \label{tab:performance_results}
\centering
\begin{tabular}{l r r r} \toprule
& RF & MLP & LSTM \\ \midrule
Accuracy	&	0.9974 & 0.9970	& 0.9947 \\
Precision	&	0.9965 & 0.9979 & 0.9894	\\
Recall	&	0.9931 & 0.9903	& 0.9897 \\
F1	&	0.9948 & 0.9941	& 0.9896 \\
Youden	&	0.9920 & 0.9896 & 0.9861 \\
\bottomrule
\end{tabular}
\end{table}
%
%\section{Explaining Recurrent Neural Networks}
%From a naive perspective, one might be tempted to reuse explainability methods for recurrent neural networks by considering a flow the sum of its packet features.
%We identify several problems which occur when trying to explain decisions made by recurrent neural networks.
%
%\begin{itemize}
%\item
%\textbf{Feature quantity.}
%The number of features that is fed into a recurrent neural network is found as the product of packet features with the total length of the flow. For long flows, the total number of inputs can therefore grow to a substantial size.
%
%\item
%\textbf{Variable sequence lengths.}
%The length of different flows might differ tremendously. Hence, features which are important for the network's outcome for one sequence might not even exist for other flows. Furthermore, the question arises how the distance between flows of different length can be computed.
%
%\item
%\textbf{Lack of a distance measure.}
%However, even if we restricted our analysis to flows of a constant length, a flow is different from the plain concatenation of its packet features.
%If we, for example, consider a classification problem as in this document, we expect that after a certain amount of time steps the classifier is able to perform the classification with a certain confidence. Arguably, features at the beginning of a flow are thus more important for the classifier. Giving differences at the end of a sequence the same weight as differences at its beginning, therefore seems incorrect.
%
%\item
%\textbf{Multiple prediction outputs.}
%Often a recurrent neural network produces an output at each time step. When trying to apply explainability methods, hence, the question arises which prediction to base the methods on.
%
%The natural choice is to base the methods on the prediction output which occurs at the same time step as the feature under investigation. Not only does this approach require a substantially lower amount of computational resources compared to predictions occurring at a later time step, but we also expect the immediately occurring prediction outcome to exhibit a stronger correlation to the feature under investigation than a prediction outcome which occurring many time steps later. However, due to the complex decision processes that can occur for deep neural networks, it is absolutely  uncertain if the network's final prediction is correctly depicted when following this approach.
%\end{itemize}

%\subsection{Feature Importance Metrics}
%\begin{table}
%\caption{Accuracy drop when replacing feature values with random data.}
%\label{tab:feat_importance}
%\centering
%\begin{tabular}{l r}
%\toprule
%Feature & Accuracy drop \\ \midrule
%Protocol	&	0.1323	\\
%Packet Length	&	0.1203	\\
%SYN Flag	&	0.1067	\\
%ACK Flag	&	0.0805	\\
%Destination port	&	0.0796	\\
%PSH Flag	&	0.0767	\\
%Direction	&	0.0732	\\
%Source port	&	0.0254	\\
%Interarrival time	&	0.0200	\\
%RST Flag	&	0.0116	\\
%FIN Flag	&	0.0100	\\
%ECE	Flag &	0.0000	\\
%URG Flag	&	0.0000	\\
%CWR Flag	&	0.0000	\\
%NS Flag	&	0.0000	\\
%\bottomrule
%\end{tabular}
%\end{table}
%As a first step to understanding the neural network's decisions, we tried to estimate how important individual features are for the model's predictions.
%
%To this end, we evaluated by what extent detection accuracy drops when withholding individual features from the classifier. We thus replaced values of individual features by values randomly drawn from the same distribution as the real feature values. We ensured that features which stay constant for a flow, i.e. source port, destination port and protocol, also stay constant for the distorted flows.
%
%\autoref{tab:feat_importance} shows our results. The results match to a large extent with the expectations one has when classifying network flows. In particular,  the most important features turn out to be the used protocol, packet length and the SYN TCP flag. While the first two features are essential for characterizing flows, a high number of packets with set SYN flag is an important characteristic for port scanning or certain DoS attacks.
%
%Also the fact that most of TCP flags do not play an important role in our dataset, was expected.
%It is interesting to note, however, that the interarrival time of packets does not play a crucial role for the detection of attacks.
%
%\subsection{Explainability Plots}
%During flow extraction we used a usual 5-tuple flow key, which distinguishes flows based on the protocol they use and their source and destination port and IP address. Hence, these features are constant for all packets in one flow, which allows us to use explainability methods which have been proposed for non-recurrent classification methods.
%
%In particular,  in literature the use of Partial Dependency Plots (PDP) is proposed. If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space and $f(X) \in [0,1]$ denotes the neural network's prediction, the PDP for feature $i$ can be defined as
%\begin{equation} \label{eq:pdp}
%\text{PDP}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right)
%\end{equation}
%
%
%To inspect attack types in detail, in this research we experiment with a modified definition of the PDP  (with C being the attack classes):
%\begin{equation} \label{eq:pdp_conditional}
%\text{PDP}_{c,i}(y) = \mathbb E_{\boldsymbol X | C = c} \left(f \left( X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n \right) \right)
%\end{equation}
%
%\begin{figure*}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_8_DoS - DDoS-Heartbleed"}.pdf}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_10_Normal"}.pdf}
%
%\caption{Examples of conditional PD plots for source and destination port. We only analyze these features with the PDP since it is not straightforward to define the PDP for features that can vary for each packet in a flow. The y-axis shows the probability for malicious traffic. }
%\label{fig:pdp}
%\end{figure*}
%
%In this variant of the PDP that we call conditional PDP, we don't take the whole dataset for the computation for the PDP but only a subset that belongs to a certain class. Specifically, we take each attack class and compute the PDP only over samples of this attack class. The results in \autoref{fig:pdp} show that for some attack classes the port numbers play an important role. When looking at the class of normal traffic (non-attack traffic) it becomes apparent that traffic destined to a high source port is generally indicative of an attack. We argue that this is because most services that regular users use have low port numbers. A known issue about PDPs in general is that dependencies of the feature under investigation and the remaining variables are not considered, hence potentially querying the neural network for feature combinations which cannot possibly occur.
%
%As apparent from \autoref{eq:pdp}, for a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.
%
%For this reason, Accumulated Local Effects (ALE) plots have been proposed in literature. The basic idea for ALE is that instead of using all samples, only the closest samples are used for assessing the effects of one feature. ALE plots therefore inherently require a distance measure for finding the closest samples, which, as revealed in the above discussion, in our case is not possible in a straight-forward manner.

%\subsection{Plots for Sequences}
%
%As apparent from the above discussion, trying to explain how packet features influence the model's predictions, poses various difficulties. As a first step to explaining our model's predictions, we investigate how important features at different time steps are for the model's decisions. Intuitively, features at the beginning of a flow should be the most important while the model's predictions should not vary significantly anymore, as soon as the model has come to a decision.
%
%\begin{figure*}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_0_Botnet-ARES"}.pdf}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_14_Web Attack-XSS"}.pdf}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
%
%\caption{Median confidence per time step and first and third quartiles for selected attack types. For the majority of attacks, confidence increases in the first few steps and then stays almost constant at 1.}
%\label{fig:confidence}
%\end{figure*}
%
%To determine if this hypothesis is correct, we plot the attack confidence of the classifier for each time step for each attack class. At each time step we average over all flows and also show the first and third quartiles. \autoref{fig:confidence} shows that this hypothesis is correct for most attack classes. At the first couple of packets the confidence for maliciousness is not very high yet but towards the end of the flow it reaches values close to 1 and stays there.
%
%\begin{figure*}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_0_Botnet-ARES"}.pdf}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_1_Brute Force-FTP-Patator"}.pdf}
%
%\caption{The solid line shows the mean value of the feature for all flows of a specific attack type. The shaded region shows the change in confidence that occurs when the feature is varied.}
%\label{fig:pred_plots2}
%\end{figure*}
%
%Next, we want to investigate if our classifier just memorized its inputs. The most interesting packet features that can be used to visualize this in our use case are the inter-arrival time (IAT) between packets in a flows and the packet length. To depict how the model comes to a decision, we determined the total minimum and maximum for these features that ever occur for all flows. For a given flow $X$ we then varied feature values at each time step between minimum and maximum and plot how the confidence changes when the feature varies (\autoref{fig:pred_plots2}) depict the averaged predictions for different attack families. It seems clear that the classifier didn't just memorize the inputs but actually learned them since there is always a smooth decrease of confidence if we change the feature values. Instead, when memorizing, we would expect a sharp decline of confidence to occur. Another interesting observation is that after the first couple of packets the classifier is certain about its decision and changing features doesn't change its confidence any longer. This allows the conclusion that if an attacker is able to change the first few packets, she could possibly change the decision of the classifier. We will further investigate this question in section \autoref{sec:adv}.
%
%\begin{figure*}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_2_Brute Force-SSH-Patator"}.pdf}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_6_DoS - DDoS-DoS Slowhttptest"}.pdf}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
%\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
%
%\caption{The solid line shows the median value of the feature for all flows of a specific attack type. The shaded region are the first and third quartile.}
%\label{fig:plot_features}
%\end{figure*}
%
%Looking at the plots of the features that vary over time (\autoref{fig:plot_features}) we see that certain attack types have a characteristic pattern in which they send packets by which they are easily recognizable. We want to determine how the pattern of packet sizes and interarrival times is significant for the classification. \autoref{fig:plot_features} clearly shows that some attack types have very characteristic patterns of packet sizes and interarrival times.

\section{Adversarial attacks}
\label{sec:adv}

As we saw in the previous section in \autoref{fig:plot_features}, some attack types do have a characteristic pattern of interarrival times and packet lengths. Furthermore, \autoref{fig:pred_plots2} shows after the first couple of packets, the classifier becomes so sure in its decision, that a change in features doesn't seem to have any after the first packets. The hypothesis that follows is that by manipulating the interarrival times and packet lengths of the first packets, the classifier might be fooled. However, our scenario is quite different from how generating adversarial samples usually works: We have significantly less features than in, for example, an image. Furthermore, only two out of those few features can be manipulated. And even these features cannot manipulated freely: Besides the constraints that packet size and interarrival time cannot become negative there are also more problem-specific constraints:
\begin{itemize}
\item Only packets can be manipulated which are transmitted by the attacker. This means the attacker can only control one direction (except for botnets where both sides are attackers).
\item Packets must not be smaller than initially, as otherwise less information could be transmitted or packets might be rendered syntactically invalid.
\item Interarrival times must not decrease, as otherwise the physical speed of data transmission can be violated in some cases. Since an in-depth analysis of when the reduction of interarrival times is possible, would come with substantial effort, we generally disallowed a reduction of this feature.
\end{itemize}

Several methods for generating adversarial samples have been proposed in the recent literature, achieving different speed-quality tradeoffs. Since the reasons discussed above suggest that finding adversarial flows is particularly difficult in our case, for the present research we preferred methods which produces high-quality adversarial samples while not being as fast as other methods.

We implemented the following methods for generating adversarial flows.

\subsection{Carlini-Wagner}
In particular, we implemented the method by \citeauthor{carlini2017towards}~\cite{carlini2017towards}, which performs gradient descent on the optimization objective
\begin{equation} \label{eq:carliniWagner}
d(X,\tilde X) + \epsilon  \max(Z(\tilde X), 0).
\end{equation}
Here, $d(\cdot)$ is a distance metric, $\epsilon \in \mathbb R^+$ is a parameter governing the tradeoff achieved between attack success and distance from the original flow. Furthermore, $Z(\cdot)$ denotes the neural network's logit output, $X$ denotes the original flow and $\tilde X$ the adversarial flow under optimization.

We used $\ell_1$ norm as distance metric and deployed projected gradient descent (PGD) for meeting the real-world constraints discussed above.

The method minimizes the adversarial sample's distance so that the network's decision is just between attack and normal traffic. In the present context, we need to make sure that the classifier's prediction would actually be normal traffic, even though a certain level of noise will be added to interarrival times due to the network between attacker and victim. Instead of the above equation \ref{eq:carliniWagner} we therefore used the optimization objective
\begin{equation}
d(X,\tilde X) + \epsilon  \max(Z(\tilde X), -0.2),
\end{equation}
i.e. we require a slight level of confidence that our adversarial samples are not classified as attack traffic \todo{What is 0.2 in probability and not in logits? Logits are hard to imagine}.

\subsection{$\ell_\infty$-bounded Projected Gradient Descent}
\cite{madry2017towards} use a method for generating adversarial samples which uses PGD  to minimize the network's negative loss function while constraining the achieved $\ell_\infty$ distance from the original samples.

\subsection{Fast Gradient Sign Method}
Finally, we also tested the Fast Gradient Sign Method (FGSM), which was initially proposed in \cite{goodfellow2014explaining} and perhaps is the most frequently used method for generating adversarial sampling. FGSM can be considered a single pass of PGD on the loss function with an equality constraint on the $\ell_\infty$ distance, i.e. the adversarial sample is found as
\begin{equation}
\tilde X = X + \epsilon \text{sgn}( \nabla_X L(X)),
\end{equation}
where $L(X)$ denotes the network's loss function and $\epsilon$ denotes the achieved $\ell_\infty$ distance.

\subsection{Evaluation}
\begin{figure*}
\includegraphics[width=0.45\textwidth]{adv_plots/cwl1.pdf}
\includegraphics[width=0.45\textwidth]{adv_plots/cwl2.pdf}
\caption{Performance of the Carlini-Wagner method with different $\epsilon$ values for $\ell_1$ and $\ell_2$ distances.}
\label{fig:adv_cw}
\end{figure*}
\autoref{fig:adv_cw} depicts the ratio of attack flows which were successfully misclassified as normal traffic with respect to the final packet along with the resulting distance from the original sample. Evidently, the larger the $\epsilon$ parameter, the better the attack works, but also the distances become larger.

Hence, with acceptable distance values we were able to generate adversarial samples for about 80\% of attack samples.

The figure suggests that a tradeoff value between 0.5 and 1 would be ideal to find adversarial samples with a low enough distance from real samples. This conclusion might be misleading, however, as the distance cannot become larger anymore as soon as an adversarial sample has been found, when performing optimization as depicted by the objective function \eqref{eq:carliniWagner}. Hence, the depicted increase of average distances presumably is due to samples for which no adversarial sample was found for lower $\epsilon$ values.

\begin{table*}
\caption{Attack success ratios per attack family. Depicted are flow detection accuracies for adversarial flows (left numbers) and flow detection accuracies for unmodified flows (right numbers).}
\label{tab:adv_per_family}
\centering
\begin{tabular}{lllllllll}
\toprule
Attack type & FGSM & Carlini-Wagner, $\ell_1$ & Carlini-Wagner, $\ell_2$ & $\ell_\infty$-bounded PGD \\
\midrule
Botnet:ARES	&	0.0000 / 0.0000	&	0.0000 / 0.0000	&	0.0000 / 0.0000	&	0.0000 / 0.0000	\\
Brute Force:FTP-Patator	&	0.2800 / 0.5800	&	0.1600 / 0.5800	&	0.1600 / 0.5800	&	0.2800 / 0.5800	\\
Brute Force:SSH-Patator	&	0.8300 / 1.0000	&	0.0000 / 1.0000	&	0.0000 / 1.0000	&	0.8900 / 1.0000	\\
DDoS:LOIT	&	0.8300 / 0.8600	&	0.6400 / 0.8600	&	0.5600 / 0.8600	&	0.8200 / 0.8600	\\
DoS / DDoS:DoS GoldenEye	&	0.8900 / 1.0000	&	0.5000 / 1.0000	&	0.4200 / 1.0000	&	0.9200 / 1.0000	\\
DoS / DDoS:DoS Hulk	&	0.6700 / 1.0000	&	0.5400 / 1.0000	&	0.4700 / 1.0000	&	1.0000 / 1.0000	\\
DoS / DDoS:DoS Slowhttptest	&	0.8700 / 1.0000	&	0.4100 / 1.0000	&	0.1900 / 1.0000	&	0.7400 / 1.0000	\\
DoS / DDoS:DoS slowloris	&	0.4400 / 0.9700	&	0.3100 / 0.9700	&	0.1300 / 0.9700	&	0.9200 / 0.9700	\\
DoS / DDoS:Heartbleed	&	0.0000 / 0.0000	&	0.0000 / 0.0000	&	0.0000 / 0.0000	&	0.0000 / 0.0000	\\
Infiltration:Dropbox download 	&	0.6100 / 0.9600	&	0.4300 / 0.9600	&	0.3800 / 0.9600	&	0.7400 / 0.9600	\\
PortScan:PortScan - Firewall off	&	0.6800 / 0.9100	&	0.3500 / 0.9100	&	0.3500 / 0.9100	&	0.7200 / 0.9100	\\
PortScan:PortScan - Firewall on	&	0.7400 / 0.8700	&	0.6100 / 0.8700	&	0.5300 / 0.8700	&	0.7300 / 0.8700	\\
Web Attack:Sql Injection	&	0.0000 / 0.6923	&	0.0000 / 0.6923	&	0.0000 / 0.6923	&	0.0000 / 0.6923	\\
Web Attack:XSS	&	0.6700 / 0.9500	&	0.7200 / 0.9500	&	0.6200 / 0.9500	&	0.6700 / 0.9500	\\
\bottomrule

\end{tabular}
\end{table*}

It is interesting to note that for very large values of $\epsilon$ the success ratio starts to decrease again.This phenomenon might be due to numerical effects and might be preventable by a smaller learning rate.

\autoref{fig:fgsm} depicts the performance for FGSM and $\ell_\infty$-bounded PGD. As expected, while FGSM is the fastest of all algorithms, it delivers the worst performance.

Of the investigated algorithms, thus the Carlini-Wagner method with $\ell_2$ distance delivered the best performance.

\autoref{tab:adv_per_family} depicts how successful the attack is for individual attack families. For this table we used a $\ell_\infty$ bound of 0.3 for FGSM and $\ell_\infty$-bounded PGD and a tradeoff of 0.5 for Carlini-Wagner. The table shows significant for detecting different attack families in the first place. However, also for generating adversarial flows some attack families are more susceptible than others.

\begin{figure*}
\includegraphics[width=0.45\textwidth]{adv_plots/fgsm.pdf}
\includegraphics[width=0.45\textwidth]{adv_plots/l_inf_pgd.pdf}
\caption{Performance of FSGM and $\ell_\infty$-bounded PGD for different bounds.}
\label{fig:fgsm}
\end{figure*}

Finally, we investigated what adversarial flows look like and which features are manipulated to circumvent detection by the neural network.

To this end, we generated adversarial flows with the Carlini-Wagner method with $\ell_2$ distance and a tradeoff of 0.5. \autoref{adv_plots} depicts how the method manipulates flows for some exemplary attack categories. Depicted is the mean of both unmodified flows and adversarial flows  of samples of individual attack families considering both interarrival time and packet length.

The plots show that primarily features at the beginning of flows are modified. It is remarkable that it seems most promising to modify single packets while leaving most of the flow unchanged. In other cases, however, like depicted for the infiltration attack at the bottom, adversarial flows assume a complex pattern to circumvent detection.

\begin{figure*}[p]
\includegraphics[width=0.48\textwidth]{../plots/plot_adv/1.pdf}
\includegraphics[width=0.48\textwidth]{../plots/plot_adv/2.pdf}
\includegraphics[width=0.48\textwidth]{../plots/plot_adv/3.pdf}
\includegraphics[width=0.48\textwidth]{../plots/plot_adv/4.pdf}
\includegraphics[width=0.48\textwidth]{../plots/plot_adv/5.pdf}
\includegraphics[width=0.48\textwidth]{../plots/plot_adv/6.pdf}
\caption{Successful manipulations performed by the Carlini-Wagner method.}
\label{adv_plots}
\end{figure*}

\subsection{Defences}

The most obvious defence, which is only possible because of the nature of the dataset, is to simply leave out features which the attacker can manipulate. For this, we try two different approaches:
\begin{itemize}
\item Leaving out all features that are manipulable: packet size and interarrival time
\item Leaving out manipulable features when the attacker can manipulate them: only in the direction from the attacker to the victim (This, however does not prevent adversarial samples for botnets, for which both sides are malicious)
\end{itemize}

Both approaches lead to complete resistance to adversarial samples (as we left out all manipulable features), except for botnets, which can still operate when only leaving out manipulable features in one direction. The results in \autoref{tab:performance_results_no_manipulable} show that -- surprisingly -- there is only a small difference in classification performance when only computing the accuracy for the last packet of each flow (flow accuracy). However, when looking at the packet accuracy, which considers all packets of all flows, there is a significant difference. Thus, apparently the interarrival time and packet size are especially important for determining whether a flow is malicious in the first packets of a flow.

In addition to omitting manipulable features we tried to make our network more robust against adversarial samples by augmenting the training set by adversarial flows generated using the Carlini-Wagner method, labeled as additional attacks. However, against our expectations, this approach did not lead to satisfactory results in the sense that the network would learn to classify adversarial samples as attacks. In fact, classification accuracy stayed about constant during the training of the network.

A possible explanation for this observation might be that the adversarial samples assumed too much the shape of normal flows, so that the network could not reliable tell them apart.

\begin{table*}
\caption{Comparing performance with all features, only safe features (interarrival time and packet size removed) and safe features in one direction (interarrival time and packet size removed for packets coming from source). \textit{packets} means that we consider the neural network output of all packets of the flow while \textit{flows} means that we only consider the last packet. At the last packet the classifier is usually already a lot more confident than at the first packets of a flow and thus metrics are always better in this case.}  \label{tab:performance_results_no_manipulable}
\newcommand{\cmidrulespace}{6pt}
\centering
\begin{tabular}{l l l l l l l} \toprule
& \multicolumn{2}{l}{all} & \multicolumn{2}{l}{safe features} & \multicolumn{2}{l}{safe features one direction} \\
\cmidrule(r){2-3} \cmidrule(lr){4-5} \cmidrule(l){6-7}
& packets & flows & packets & flows & packets & flows \\
\midrule
Accuracy & 0.9857054 & 0.99472587 & 0.96678023 & 0.99337596 & 0.98056073 & 0.99353386 \\
Precision & 0.94963342 & 0.9894422 & 0.87175275 & 0.98798596 & 0.91359787 & 0.98994595 \\
Recall & 0.96735895 & 0.98972051 & 0.94374958 & 0.98581328 & 0.97836672 & 0.9844478 \\
F1 & 0.95841423 & 0.98958134 & 0.90632359 & 0.98689842 & 0.94487366 & 0.98718922 \\
Youden & 0.95682951 & 0.98614231 & 0.91525628 & 0.98175163 & 0.95937772 & 0.9810602 \\
\bottomrule
\end{tabular}
\end{table*}

\section{Conclusions}

We have implemented a recurrent classifier based on LSTM to detect network attacks. While attack detection performance is not better than that of non-recurrent approaches, we can already detect attacks before they are over. Furthermore, the recurrent approach allows us to inspect the influence of single packets on the detection performance and shows which packets are \textit{characteristic} for attacks.

We show that adversarial samples are possible even though the data we use are very different from those that are usually used when finding adversarial samples.

\bibliographystyle{plain}
\bibliography{bibliography}

\section*{Acknowledgements}
The Titan Xp used for this research was donated by the NVIDIA Corporation.

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.

### 2019-10-24 ex3/paper/paper.tex
Added distance for adversarial training

@@
 \end{equation}
-i.e. we require a slight level of confidence that our adversarial samples are not classified as attack traffic \todo{What is 0.2 in probability and not in logits? Logits are hard to imagine}.
+i.e.
+we require a slight level of confidence that our adversarial samples are not classified as attack traffic.
+A logit value of $-0.2$ corresponds to a confidence of 45\% for the sample to be an attack.
 
@@
 
-\subsection{Defences}
+\section{Defences}
 
+\subsection{Reducing features}
 The most obvious defence, which is only possible because of the nature of the dataset, is to simply leave out features which the attacker can manipulate. For this, we try two different approaches:
@@
 
+\subsection{Adversarial training}
 In addition to omitting manipulable features we tried to make our network more robust against adversarial samples by augmenting the training set by adversarial flows generated using the Carlini-Wagner method, labeled as additional attacks.
-However, against our expectations, this approach did not lead to satisfactory results in the sense that the network would learn to classify adversarial samples as attacks.
+
+For this purpose, we added one adversarial sample per attack sample to the training set.
+Note that, since our used method for obtaining adversarial samples is deterministic, this is the maximum number of adversarial samples possible.
+With this training set we then alternated retraining of the neural network and optimization of the adversarial samples using gradient descent.
+
+One question which occurs in this process, is how many CPU cycles to spend on network training and adversarial sample optimization.
+As a natural approach, we decided to use as many backpropagation steps for neural network training as we use for adversarial sample optimization.
+For each adversarial sample optimization step, we performed 10 iterations of gradient descent.
+Hence, all adversarial samples were optimized each 10 epochs of neural network training.
+
+\autoref{fig:avdistance} shows the average distance of a block of adversarial samples from to original, unmodified attack samples throughout the adversarial training procedure.
+The figure clearly indicates that adversarial training is effective, as the distance gradually increases.
+Hence, an attacker would have to modify attack samples more and more, eventually rendering the attack unpractical.
+
+However, the figure also shows that an increase of distance can only be achieved after a considerable amount of training.
+It is, hence, questionable if the gained increase in robustness is enough for reliable attack detection.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{../plots/adv_training/avdistance.pdf}
+\caption{Distance of adversarial samples from original attack samples throughout adversarial training.}
+\label{fig:avdistance}
+\end{figure}
+
+%However, against our expectations, this approach did not lead to satisfactory results in the sense that the network would learn to classify adversarial samples as attacks.
 In fact, classification accuracy stayed about constant during the training of the network.

### 2019-10-25 ex3/paper/paper.tex
Paper draft

--- 2019-10-23 ex3/paper/paper.tex
@@
 \newacronym{ids}{IDS}{Intrusion Detection System}
+\newacronym{rnn}{RNN}{Recurrent Neural Network}
+\newacronym{fgsm}{FGSM}{Fast Gradient Sign Method}
+\newacronym{cw}{CW}{Carlini-Wagner method}
+\newacronym{pgd}{PGD}{Projected Gradient Descent}
 
@@
 We develop a simple methods for defending against adversaries who can manipulate arbitrary packets by simply leaving out potentially manipulable information from the classification process.
-The resulting classifier has slightly reduces accuracy but is still useful to be deployed in a real scenario.
+The resulting classifier is slightly less accurate but is still useful to be deployed in a real scenario.
 Finally we develop a custom adversarial training procedure which especially focuses on easy of use and low resource consumption and aims to keep the number of additionally introduced training hyperparameters low.
@@
 \section{Related work}
+
+\cite{szegedy_intriguing_2014} show that it is possible to create adversarial samples for an image, which look similar to the original image but are classified wrongly.
+\cite{goodfellow_explaining_2015} develop the \gls{fgsm}, which makes easy generation of adversarial samples possible.
+\cite{papernot_crafting_2016} explore the use of \gls{aml} for \glspl{rnn}
 
@@
 
-As we saw in the previous section in \autoref{fig:plot_features}, some attack types do have a characteristic pattern of interarrival times and packet lengths.
-Furthermore, \autoref{fig:pred_plots2} shows after the first couple of packets, the classifier becomes so sure in its decision, that a change in features doesn't seem to have any after the first packets.
-The hypothesis that follows is that by manipulating the interarrival times and packet lengths of the first packets, the classifier might be fooled.
+Since most features such as the TCP flags cannot be easily be manipulated because it might violate the TCP and thus cause communication to fails, we experiment whether by manipulating the interarrival times and packet lengths of the first packets, the classifier might be fooled.
 However, our scenario is quite different from how generating adversarial samples usually works: We have significantly less features than in, for example, an image.
@@
 \subsection{Carlini-Wagner}
-In particular, we implemented the method by \citeauthor{carlini2017towards}~\cite{carlini2017towards}, which performs gradient descent on the optimization objective
+In particular, we implemented \gls{cw} \cite{carlini_towards_2017}, which performs gradient descent on the optimization objective
 \begin{equation} \label{eq:carliniWagner}
@@
 
-We used $\ell_1$ norm as distance metric and deployed projected gradient descent (PGD) for meeting the real-world constraints discussed above.
+We used $\ell_1$ norm as distance metric and \gls{pgd} for meeting the real-world constraints discussed above.
 
@@
 \subsection{$\ell_\infty$-bounded Projected Gradient Descent}
-\cite{madry2017towards} use a method for generating adversarial samples which uses PGD  to minimize the network's negative loss function while constraining the achieved $\ell_\infty$ distance from the original samples.
+\cite{madry_towards_2018} use a method for generating adversarial samples which uses \gls{pgd}  to minimize the network's negative loss function while constraining the achieved $\ell_\infty$ distance from the original samples.
 
 \subsection{Fast Gradient Sign Method}
-Finally, we also tested the Fast Gradient Sign Method (FGSM), which was initially proposed in \cite{goodfellow2014explaining} and perhaps is the most frequently used method for generating adversarial sampling.
-FGSM can be considered a single pass of PGD on the loss function with an equality constraint on the $\ell_\infty$ distance, i.e.
+Finally, we also tested the Fast Gradient Sign Method (FGSM), which was initially proposed in \cite{goodfellow_explaining_2015} and perhaps is the most frequently used method for generating adversarial sampling.
+FGSM can be considered a single pass of \gls{pgd} on the loss function with an equality constraint on the $\ell_\infty$ distance, i.e.
 the adversarial sample is found as
@@
 \includegraphics[width=0.45\textwidth]{adv_plots/cwl2.pdf}
-\caption{Performance of the Carlini-Wagner method with different $\epsilon$ values for $\ell_1$ and $\ell_2$ distances.}
+\caption{Performance of the \gls{cw} with different $\epsilon$ values for $\ell_1$ and $\ell_2$ distances.}
 \label{fig:adv_cw}
@@
 \toprule
-Attack type & FGSM & Carlini-Wagner, $\ell_1$ & Carlini-Wagner, $\ell_2$ & $\ell_\infty$-bounded PGD \\
+Attack type & FGSM & Carlini-Wagner, $\ell_1$ & Carlini-Wagner, $\ell_2$ & $\ell_\infty$-bounded \gls{pgd} \\
 \midrule
@@
 
-\autoref{fig:fgsm} depicts the performance for FGSM and $\ell_\infty$-bounded PGD. As expected, while FGSM is the fastest of all algorithms, it delivers the worst performance.
+\autoref{fig:fgsm} depicts the performance for FGSM and $\ell_\infty$-bounded \gls{pgd}. As expected, while FGSM is the fastest of all algorithms, it delivers the worst performance.
 
-Of the investigated algorithms, thus the Carlini-Wagner method with $\ell_2$ distance delivered the best performance.
+Of the investigated algorithms, thus the \gls{cw} with $\ell_2$ distance delivered the best performance.
 
 \autoref{tab:adv_per_family} depicts how successful the attack is for individual attack families.
-For this table we used a $\ell_\infty$ bound of 0.3 for FGSM and $\ell_\infty$-bounded PGD and a tradeoff of 0.5 for Carlini-Wagner.
+For this table we used a $\ell_\infty$ bound of 0.3 for FGSM and $\ell_\infty$-bounded \gls{pgd} and a tradeoff of 0.5 for Carlini-Wagner.
 The table shows significant for detecting different attack families in the first place.
@@
 \includegraphics[width=0.45\textwidth]{adv_plots/l_inf_pgd.pdf}
-\caption{Performance of FSGM and $\ell_\infty$-bounded PGD for different bounds.}
+\caption{Performance of FSGM and $\ell_\infty$-bounded \gls{pgd} for different bounds.}
 \label{fig:fgsm}
@@
 
-To this end, we generated adversarial flows with the Carlini-Wagner method with $\ell_2$ distance and a tradeoff of 0.5.
+To this end, we generated adversarial flows with the \gls{cw} with $\ell_2$ distance and a tradeoff of 0.5.
 \autoref{adv_plots} depicts how the method manipulates flows for some exemplary attack categories.
@@
 \includegraphics[width=0.48\textwidth]{../plots/plot_adv/6.pdf}
-\caption{Successful manipulations performed by the Carlini-Wagner method.}
+\caption{Successful manipulations performed by the \gls{cw}.}
 \label{adv_plots}
@@
 
-In addition to omitting manipulable features we tried to make our network more robust against adversarial samples by augmenting the training set by adversarial flows generated using the Carlini-Wagner method, labeled as additional attacks.
+In addition to omitting manipulable features we tried to make our network more robust against adversarial samples by augmenting the training set by adversarial flows generated using the \gls{cw}, labeled as additional attacks.
 However, against our expectations, this approach did not lead to satisfactory results in the sense that the network would learn to classify adversarial samples as attacks.

### 2019-10-25 ex3/paper/paper.tex

@@
 
-In addition to omitting manipulable features we tried to make our network more robust against adversarial samples by augmenting the training set by adversarial flows generated using the \gls{cw}, labeled as additional attacks.
-However, against our expectations, this approach did not lead to satisfactory results in the sense that the network would learn to classify adversarial samples as attacks.
+\subsection{Adversarial training}
+In addition to omitting manipulable features we tried to make our network more robust against adversarial samples by augmenting the training set by adversarial flows generated using the Carlini-Wagner method, labeled as additional attacks.
+
+For this purpose, we added one adversarial sample per attack sample to the training set.
+Note that, since our used method for obtaining adversarial samples is deterministic, this is the maximum number of adversarial samples possible.
+With this training set we then alternated retraining of the neural network and optimization of the adversarial samples using gradient descent.
+
+One question which occurs in this process, is how many CPU cycles to spend on network training and adversarial sample optimization.
+As a natural approach, we decided to use as many backpropagation steps for neural network training as we use for adversarial sample optimization.
+For each adversarial sample optimization step, we performed 10 iterations of gradient descent.
+Hence, all adversarial samples were optimized each 10 epochs of neural network training.
+
+\autoref{fig:avdistance} shows the average distance of a block of adversarial samples from to original, unmodified attack samples throughout the adversarial training procedure.
+The figure clearly indicates that adversarial training is effective, as the distance gradually increases.
+Hence, an attacker would have to modify attack samples more and more, eventually rendering the attack unpractical.
+
+However, the figure also shows that an increase of distance can only be achieved after a considerable amount of training.
+It is, hence, questionable if the gained increase in robustness is enough for reliable attack detection.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{../plots/adv_training/avdistance.pdf}
+\caption{Distance of adversarial samples from original attack samples throughout adversarial training.}
+\label{fig:avdistance}
+\end{figure}
+
+%However, against our expectations, this approach did not lead to satisfactory results in the sense that the network would learn to classify adversarial samples as attacks.
 In fact, classification accuracy stayed about constant during the training of the network.

### 2019-10-25 ex3/learn.py
Filtered invalid flows

@@ -553,8 +553,8 @@ def adv_internal(in_training = False):
 
 			orig_mask = (index_tensor <= selection_tensor).byte().to(device)
 			# mask_exact = (index_tensor == selection_tensor).byte().to(device)
-			wrong_direction = (seqs[:,:,5]!=forward_direction)
-			matching_cats = [(cats==bidirectional_cat).squeeze() for bidirectional_cat in bidirectional_categories]
+			wrong_direction = (seqs[:,:,5]!=forward_direction).byte()
+			matching_cats = [(cats==bidirectional_cat).squeeze().byte() for bidirectional_cat in bidirectional_categories]
 			# print("matching_cats.shape", [item.shape for item in matching_cats])
 			not_bidirectional = ~reduce(lambda acc, x: acc | x, matching_cats, torch.ByteTensor([False]).to(device))
 			# print(orig_mask.shape, wrong_direction.shape, not_bidirectional.shape)
@@ -1046,7 +1046,8 @@ if __name__=="__main__":
 		categories_mapping_content = json.load(f)
 	categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
 
-	all_data = [item[:opt.maxLength,:] for item in all_data] # if np.random.rand() < 0.005]
+	# a few flows have have invalid IATs due to dataset issues. Sort those out.
+	all_data = [item[:opt.maxLength,:] for item in all_data if np.all(item[:,4]>=0)]
 	if opt.removeChangeable:
 		all_data = [overwrite_manipulable_entries(item) for item in all_data]
 	random.shuffle(all_data)

### 2019-11-22 ex3/plot_adv.py
Updated plot scripts

@@ -10,13 +10,15 @@ from learn import numpy_sigmoid
 
 DIR_NAME = "plots/plot_adv"
 
-with open("categories_mapping.json", "r") as f:
+dataroot_basename = sys.argv[1].split('_')[0]
+
+with open(dataroot_basename + "_categories_mapping.json", "r") as f:
 	categories_mapping_content = json.load(f)
 categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
 reverse_mapping = {v: k for k, v in mapping.items()}
 # print("reverse_mapping", reverse_mapping)
 
-with open("flows_full_no_ttl_normalization_data.pickle", "rb") as f:
+with open(dataroot_basename+"_full_no_ttl_normalization_data.pickle", "rb") as f:
 	means, stds = pickle.load(f)
 
 # TODO: Implement for more than one adv output so that different tradeoffs can be compared.

### 2019-11-22 ex3/plot_features.py

@@ -15,13 +15,15 @@ from learn import numpy_sigmoid
 
 DIR_NAME = "plots/plot_features"
 
-with open("categories_mapping.json", "r") as f:
+dataroot_basename = sys.argv[1].split('_')[0]
+
+with open(dataroot_basename + "_categories_mapping.json", "r") as f:
 	categories_mapping_content = json.load(f)
 categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
 reverse_mapping = {v: k for k, v in mapping.items()}
 # print("reverse_mapping", reverse_mapping)
 
-with open("flows_full_no_ttl_normalization_data.pickle", "rb") as f:
+with open(dataroot_basename+"_full_no_ttl_normalization_data.pickle", "rb") as f:
 	means, stds = pickle.load(f)
 
 file_name = sys.argv[1]

### 2019-11-22 ex3/plot_new.py

@@ -15,13 +15,15 @@ from learn import numpy_sigmoid
 
 DIR_NAME = "plots/plot_new"
 
-with open("categories_mapping.json", "r") as f:
+dataroot_basename = sys.argv[1].split('_')[0]
+
+with open(dataroot_basename + "_categories_mapping.json", "r") as f:
 	categories_mapping_content = json.load(f)
 categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
 reverse_mapping = {v: k for k, v in mapping.items()}
 # print("reverse_mapping", reverse_mapping)
 
-with open("flows_full_no_ttl_normalization_data.pickle", "rb") as f:
+with open(dataroot_basename+"_full_no_ttl_normalization_data.pickle", "rb") as f:
 	means, stds = pickle.load(f)
 
 file_name = sys.argv[1]

### 2019-11-22 ex3/plot_pdp.py

@@ -10,7 +10,9 @@ from collections import Counter
 
 DIR_NAME = "plots/plot_pdp"
 
-with open("categories_mapping.json", "r") as f:
+dataroot_basename = sys.argv[1].split('_')[0]
+
+with open(dataroot_basename + "_categories_mapping.json", "r") as f:
 	categories_mapping_content = json.load(f)
 categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
 reverse_mapping = {v: k for k, v in mapping.items()}

### 2019-11-22 ex3/plot.py
Updated plot scripts

@@ -10,7 +10,9 @@ import pickle
 DIR_NAME = "plots/plot"
 ORDERING = ['', 'Length min', 'IAT min', 'Length max', 'IAT max']
 
-with open("categories_mapping.json", "r") as f:
+dataroot_basename = sys.argv[1].split('_')[0]
+
+with open(dataroot_basename + "_categories_mapping.json", "r") as f:
 	categories_mapping_content = json.load(f)
 categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
 reverse_mapping = {v: k for k, v in mapping.items()}

### 2019-11-22 ex3/plot2.py

@@ -13,15 +13,15 @@ import json
 import pickle
 from learn import numpy_sigmoid
 
-DIR_NAME = "plots/plot2"
+dataroot_basename = sys.argv[1].split('_')[0]
 
-with open("categories_mapping.json", "r") as f:
+with open(dataroot_basename + "_categories_mapping.json", "r") as f:
 	categories_mapping_content = json.load(f)
 categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
 reverse_mapping = {v: k for k, v in mapping.items()}
 # print("reverse_mapping", reverse_mapping)
 
-with open("flows_full_no_ttl_normalization_data.pickle", "rb") as f:
+with open(dataroot_basename+"_full_no_ttl_normalization_data.pickle", "rb") as f:
 	means, stds = pickle.load(f)
 
 file_name = sys.argv[1]

### 2019-11-22 ex3/plot2_adv.py

@@ -14,12 +14,14 @@ import math
 
 DIR_NAME = "plots/plot2_adv"
 
-with open("categories_mapping.json", "r") as f:
+dataroot_basename = sys.argv[1].split('_')[0]
+
+with open(dataroot_basename + "_categories_mapping.json", "r") as f:
 	categories_mapping_content = json.load(f)
 categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
 reverse_mapping = {v: k for k, v in mapping.items()}
 
-with open("flows_full_no_ttl_normalization_data.pickle", "rb") as f:
+with open(dataroot_basename+"_full_no_ttl_normalization_data.pickle", "rb") as f:
 	means, stds = pickle.load(f)
 
 file_name = sys.argv[1]

### 2019-11-22 ex3/weight_feat_imp.py
Add weight feature importance

@@ -1,0 +1,67 @@
+#!/usr/bin/env python3
+
+import matplotlib.pyplot as plt
+import torch
+import numpy as np
+import sys
+import json
+
+hidden_size = 512
+n_layers = 3
+absolute_values = True
+
+with open('features.json', 'r') as f:
+	features = json.load(f)
+
+parameters = torch.load(sys.argv[1])
+
+#	parameters = lstm_module.state_dict()
+weights_hh = [ parameters['lstm.weight_hh_l%d' % layer].cpu().numpy().reshape(4,hidden_size,hidden_size).sum(axis=0) for layer in range(n_layers) ]
+weights_ih = [ parameters['lstm.weight_ih_l%d' % layer].cpu().numpy().reshape(4,hidden_size,-1).sum(axis=0) for layer in range(n_layers) ]
+h2o = parameters['h2o.weight'].cpu().numpy()
+
+if absolute_values:
+	weights_hh = [ np.abs(weights) for weights in weights_hh ]
+	weights_ih = [ np.abs(weights) for weights in weights_ih ]
+	h2o = np.abs(h2o)
+	
+
+input_featimp = [ weights_ih[0] ]
+for i in range(1,n_layers):
+	input_featimp.append(np.matmul(weights_ih[i], input_featimp[-1]))
+input_featimp.append(np.matmul(h2o, input_featimp[-1]))
+
+output_featimp = [ h2o ]
+for i in range(n_layers-1,-1,-1):
+	output_featimp.append(np.matmul(output_featimp[-1], weights_ih[i]))
+output_featimp.reverse()
+#  assert (input_featimp[-1] == output_featimp[0]).all()
+
+fi_over_time = [ np.abs(input_featimp[-1][0,:]) ]
+hh_prod = weights_hh
+for _ in range(10):
+	hh_prod = [ np.matmul(hh_prod[i], weights_hh[i]) for i in range(n_layers) ]
+	fi_over_time.append(np.abs(np.sum([ np.matmul(output_featimp[i+1], np.matmul(hh_prod[i], input_featimp[i])) for i in range(n_layers)], axis=0)[0,:]))
+
+print ('Packet feature importance:', sorted(zip(features, fi_over_time[0]/np.sum(fi_over_time[0])), key=lambda item: item[1]))
+
+# Highly experimental stuff
+decomposition = [ np.linalg.eig(weights_hh[i]) for i in range(n_layers) ]
[…]

### 2019-11-22 ex3/weight_feat_imp.py
Add weight feature importance

@@ -10,10 +10,12 @@ hidden_size = 512
 n_layers = 3
 absolute_values = True
 
-with open('features.json', 'r') as f:
-	features = json.load(f)
+assert len(sys.argv) > 1, 'No model specified'
 
 parameters = torch.load(sys.argv[1])
+
+with open('features.json', 'r') as f:
+	features = json.load(f)
 
 #	parameters = lstm_module.state_dict()
 weights_hh = [ parameters['lstm.weight_hh_l%d' % layer].cpu().numpy().reshape(4,hidden_size,hidden_size).sum(axis=0) for layer in range(n_layers) ]

### 2019-11-25 ex3/learn.py
Added mutinfo stuff

@@ -550,6 +550,21 @@ def dropout_feature_importance():
 		print("accuracy_for_feature", feature_index, feature_array[feature_index], accuracy-accuracy_for_feature)
 	return (accuracy, accuracy_for_features)
 
+def get_feature_importance_distribution(test_data):
+	distribution = np.concatenate([item[0][:,:] for item in test_data], axis=0).transpose(1,0)
+	if opt.adjustFeatImpDistribution:
+		numerical_features = [0,1,3]
+		# retain original distribution of iat, as iat varies over many magnitudes and we're mainly
+		# interested in feature importance of practical (=low) iats
+		retain_features = [2]
+		for feature in distribution.shape[0]:
+			if feature in numerical_features:
+				distribution[feature,:] = np.linspace(np.min(distribution[feature,:]), np.max(distribution[feature,:]), distribution.shape[1])
+			elif feature not in retain_features:
+				unique = np.unique(distribution[feature,:])
+				distribution[feature,:] = np.random.choice(unique, size=distribution.shape[1])
+	return distribution
+	
 # Right now this function replaces all values of one feature by random values sampled from the distribution of all features and looks how the accuracy changes.
 def feature_importance():
 
@@ -565,19 +580,19 @@ def feature_importance():
 
 	_, test_indices = get_nth_split(dataset, n_fold, fold)
 	test_data = torch.utils.data.Subset(dataset, test_indices)
-	test_x = np.concatenate([item[0][:,:] for item in test_data], axis=0).transpose(1,0)
+	distribution = get_feature_importance_distribution(test_data)
 	test_loader = torch.utils.data.DataLoader(test_data, batch_size=opt.batchSize, shuffle=False, collate_fn=custom_collate)
 
 	attack_numbers = mapping.values()
 
 	results_by_attack_number = [list() for _ in range(min(attack_numbers), max(attack_numbers)+1)]
-	randomized_results_by_attack_number = [[list() for _ in range(test_x.shape[0])] for _ in range(min(attack_numbers), max(attack_numbers)+1)]
[…]

### 2019-11-25 ex3/paper/paper.tex
Readd removed parts

@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Packet-by-Packet Hardening of Intrusion Detection Systems against Smart Adversaries}
+\title{An in-depth study of Explainability and Adversarial Robustness in RNNs}
 
@@
 \end{table}
-%
-%\section{Explaining Recurrent Neural Networks}
-%From a naive perspective, one might be tempted to reuse explainability methods for recurrent neural networks by considering a flow the sum of its packet features.
-%We identify several problems which occur when trying to explain decisions made by recurrent neural networks.
-%
-%\begin{itemize}
-%\item
-%\textbf{Feature quantity.}
-%The number of features that is fed into a recurrent neural network is found as the product of packet features with the total length of the flow.
+\section{Explaining Recurrent Neural Networks}
+From a naive perspective, one might be tempted to reuse explainability methods for recurrent neural networks by considering a flow the sum of its packet features.
+We identify several problems which occur when trying to explain decisions made by recurrent neural networks.
+
+\begin{itemize}
+\item
+\textbf{Feature quantity.}
+The number of features that is fed into a recurrent neural network is found as the product of packet features with the total length of the flow.
 For long flows, the total number of inputs can therefore grow to a substantial size.
-%
-%\item
-%\textbf{Variable sequence lengths.}
-%The length of different flows might differ tremendously.
+
+\item
+\textbf{Variable sequence lengths.}
+The length of different flows might differ tremendously.
 Hence, features which are important for the network's outcome for one sequence might not even exist for other flows.
 Furthermore, the question arises how the distance between flows of different length can be computed.
-%
-%\item
-%\textbf{Lack of a distance measure.}
-%However, even if we restricted our analysis to flows of a constant length, a flow is different from the plain concatenation of its packet features.
-%If we, for example, consider a classification problem as in this document, we expect that after a certain amount of time steps the classifier is able to perform the classification with a certain confidence.
+
+\item
+\textbf{Lack of a distance measure.}
+However, even if we restricted our analysis to flows of a constant length, a flow is different from the plain concatenation of its packet features.
+If we, for example, consider a classification problem as in this document, we expect that after a certain amount of time steps the classifier is able to perform the classification with a certain confidence.
 Arguably, features at the beginning of a flow are thus more important for the classifier.
 Giving differences at the end of a sequence the same weight as differences at its beginning, therefore seems incorrect.
-%
-%\item
-%\textbf{Multiple prediction outputs.}
-%Often a recurrent neural network produces an output at each time step. When trying to apply explainability methods, hence, the question arises which prediction to base the methods on.
-%
-%The natural choice is to base the methods on the prediction output which occurs at the same time step as the feature under investigation.
+
+\item
+\textbf{Multiple prediction outputs.}
+Often a recurrent neural network produces an output at each time step. When trying to apply explainability methods, hence, the question arises which prediction to base the methods on.
+
+The natural choice is to base the methods on the prediction output which occurs at the same time step as the feature under investigation.
 Not only does this approach require a substantially lower amount of computational resources compared to predictions occurring at a later time step, but we also expect the immediately occurring prediction outcome to exhibit a stronger correlation to the feature under investigation than a prediction outcome which occurring many time steps later.
 However, due to the complex decision processes that can occur for deep neural networks, it is absolutely  uncertain if the network's final prediction is correctly depicted when following this approach.
-%\end{itemize}
+\end{itemize}
 
-%\subsection{Feature Importance Metrics}
-%\begin{table}
-%\caption{Accuracy drop when replacing feature values with random data.}
-%\label{tab:feat_importance}
-%\centering
-%\begin{tabular}{l r}
-%\toprule
-%Feature & Accuracy drop \\ \midrule
-%Protocol	&	0.1323	\\
-%Packet Length	&	0.1203	\\
-%SYN Flag	&	0.1067	\\
-%ACK Flag	&	0.0805	\\
-%Destination port	&	0.0796	\\
-%PSH Flag	&	0.0767	\\
-%Direction	&	0.0732	\\
-%Source port	&	0.0254	\\
-%Interarrival time	&	0.0200	\\
-%RST Flag	&	0.0116	\\
-%FIN Flag	&	0.0100	\\
-%ECE	Flag &	0.0000	\\
-%URG Flag	&	0.0000	\\
-%CWR Flag	&	0.0000	\\
-%NS Flag	&	0.0000	\\
-%\bottomrule
-%\end{tabular}
-%\end{table}
-%As a first step to understanding the neural network's decisions, we tried to estimate how important individual features are for the model's predictions.
-%
-%To this end, we evaluated by what extent detection accuracy drops when withholding individual features from the classifier.
+\subsection{Feature Importance Metrics}
+\begin{table}
+\caption{Accuracy drop when replacing feature values with random data.}
+\label{tab:feat_importance}
+\begin{tabular}{l r}
+\toprule
+Feature & Accuracy drop \\ \midrule
+Protocol	&	0.1323	\\
+Packet Length	&	0.1203	\\
+SYN Flag	&	0.1067	\\
+ACK Flag	&	0.0805	\\
+Destination port	&	0.0796	\\
+PSH Flag	&	0.0767	\\
+Direction	&	0.0732	\\
+Source port	&	0.0254	\\
+Interarrival time	&	0.0200	\\
+RST Flag	&	0.0116	\\
+FIN Flag	&	0.0100	\\
+ECE	Flag &	0.0000	\\
+URG Flag	&	0.0000	\\
+CWR Flag	&	0.0000	\\
+NS Flag	&	0.0000	\\
+\bottomrule
+\end{tabular}
+\end{table}
+As a first step to understanding the neural network's decisions, we tried to estimate how important individual features are for the model's predictions.
+
+To this end, we evaluated by what extent detection accuracy drops when withholding individual features from the classifier.
 We thus replaced values of individual features by values randomly drawn from the same distribution as the real feature values.
@@
 source port, destination port and protocol, also stay constant for the distorted flows.
-%
-%\autoref{tab:feat_importance} shows our results.
+
+\autoref{tab:feat_importance} shows our results.
 The results match to a large extent with the expectations one has when classifying network flows.
@@
 While the first two features are essential for characterizing flows, a high number of packets with set SYN flag is an important characteristic for port scanning or certain DoS attacks.
-%
-%Also the fact that most of TCP flags do not play an important role in our dataset, was expected.
-%It is interesting to note, however, that the interarrival time of packets does not play a crucial role for the detection of attacks.
-%
-%\subsection{Explainability Plots}
-%During flow extraction we used a usual 5-tuple flow key, which distinguishes flows based on the protocol they use and their source and destination port and IP address.
+
+Also the fact that most of TCP flags do not play an important role in our dataset, was expected.
+It is interesting to note, however, that the interarrival time of packets does not play a crucial role for the detection of attacks.
+
+\subsection{Explainability Plots}
+During flow extraction we used a usual 5-tuple flow key, which distinguishes flows based on the protocol they use and their source and destination port and IP address.
 Hence, these features are constant for all packets in one flow, which allows us to use explainability methods which have been proposed for non-recurrent classification methods.
-%
-%In particular,  in literature the use of Partial Dependency Plots (PDP) is proposed.
+
+In particular,  in literature the use of Partial Dependency Plots (PDP) is proposed.
 If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space and $f(X) \in [0,1]$ denotes the neural network's prediction, the PDP for feature $i$ can be defined as
-%\begin{equation} \label{eq:pdp}
-%\text{PDP}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right)
-%\end{equation}
-%
-%
-%To inspect attack types in detail, in this research we experiment with a modified definition of the PDP  (with C being the attack classes):
-%\begin{equation} \label{eq:pdp_conditional}
-%\text{PDP}_{c,i}(y) = \mathbb E_{\boldsymbol X | C = c} \left(f \left( X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n \right) \right)
-%\end{equation}
-%
-%\begin{figure*}
-%\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_8_DoS - DDoS-Heartbleed"}.pdf}
-%\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_10_Normal"}.pdf}
-%
-%\caption{Examples of conditional PD plots for source and destination port.
+\begin{equation} \label{eq:pdp}
+\text{PDP}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right)
+\end{equation}
+
+To inspect attack types in detail, in this research we experiment with a modified definition of the PDP  (with C being the attack classes):
+\begin{equation} \label{eq:pdp_conditional}
+\text{PDP}_{c,i}(y) = \mathbb E_{\boldsymbol X | C = c} \left(f \left( X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n \right) \right)
+\end{equation}
+
+\begin{figure*}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_8_DoS - DDoS-Heartbleed"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_10_Normal"}.pdf}
+
+\caption{Examples of conditional PD plots for source and destination port.
 We only analyze these features with the PDP since it is not straightforward to define the PDP for features that can vary for each packet in a flow.
@@
 }
-%\label{fig:pdp}
-%\end{figure*}
-%
-%In this variant of the PDP that we call conditional PDP, we don't take the whole dataset for the computation for the PDP but only a subset that belongs to a certain class.
+\label{fig:pdp}
+\end{figure*}
+
+In this variant of the PDP that we call conditional PDP, we don't take the whole dataset for the computation for the PDP but only a subset that belongs to a certain class.
 Specifically, we take each attack class and compute the PDP only over samples of this attack class.
@@
 A known issue about PDPs in general is that dependencies of the feature under investigation and the remaining variables are not considered, hence potentially querying the neural network for feature combinations which cannot possibly occur.
-%
-%As apparent from \autoref{eq:pdp}, for a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.
-%
-%For this reason, Accumulated Local Effects (ALE) plots have been proposed in literature.
+
+As apparent from \autoref{eq:pdp}, for a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.
+
+For this reason, Accumulated Local Effects (ALE) plots have been proposed in literature.
 The basic idea for ALE is that instead of using all samples, only the closest samples are used for assessing the effects of one feature.
@@
 
-%\subsection{Plots for Sequences}
-%
-%As apparent from the above discussion, trying to explain how packet features influence the model's predictions, poses various difficulties.
+\subsection{Plots for Sequences}
+
+As apparent from the above discussion, trying to explain how packet features influence the model's predictions, poses various difficulties.
 As a first step to explaining our model's predictions, we investigate how important features at different time steps are for the model's decisions.
 Intuitively, features at the beginning of a flow should be the most important while the model's predictions should not vary significantly anymore, as soon as the model has come to a decision.
-%
-%\begin{figure*}
-%\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_0_Botnet-ARES"}.pdf}
-%\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
-%\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_14_Web Attack-XSS"}.pdf}
-%\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
-%
-%\caption{Median confidence per time step and first and third quartiles for selected attack types.
+
+\begin{figure*}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_0_Botnet-ARES"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_14_Web Attack-XSS"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
+
+\caption{Median confidence per time step and first and third quartiles for selected attack types.
 For the majority of attacks, confidence increases in the first few steps and then stays almost constant at 1.}
-%\label{fig:confidence}
-%\end{figure*}
-%
-%To determine if this hypothesis is correct, we plot the attack confidence of the classifier for each time step for each attack class.
+\label{fig:confidence}
+\end{figure*}
+
+To determine if this hypothesis is correct, we plot the attack confidence of the classifier for each time step for each attack class.
 At each time step we average over all flows and also show the first and third quartiles.
@@
 At the first couple of packets the confidence for maliciousness is not very high yet but towards the end of the flow it reaches values close to 1 and stays there.
-%
-%\begin{figure*}
-%\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_0_Botnet-ARES"}.pdf}
-%\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
-%\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
-%\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_1_Brute Force-FTP-Patator"}.pdf}
-%
-%\caption{The solid line shows the mean value of the feature for all flows of a specific attack type. The shaded region shows the change in confidence that occurs when the feature is varied.}
-%\label{fig:pred_plots2}
-%\end{figure*}
-%
-%Next, we want to investigate if our classifier just memorized its inputs.
+
+\begin{figure*}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_0_Botnet-ARES"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_1_Brute Force-FTP-Patator"}.pdf}
+
+\caption{The solid line shows the mean value of the feature for all flows of a specific attack type. The shaded region shows the change in confidence that occurs when the feature is varied.}
+\label{fig:pred_plots2}
+\end{figure*}
+
+Next, we want to investigate if our classifier just memorized its inputs.
 The most interesting packet features that can be used to visualize this in our use case are the inter-arrival time (IAT) between packets in a flows and the packet length.
@@
 We will further investigate this question in section \autoref{sec:adv}.
-%
-%\begin{figure*}
-%\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_2_Brute Force-SSH-Patator"}.pdf}
-%\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_6_DoS - DDoS-DoS Slowhttptest"}.pdf}
-%\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
-%\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
-%
-%\caption{The solid line shows the median value of the feature for all flows of a specific attack type. The shaded region are the first and third quartile.}
-%\label{fig:plot_features}
-%\end{figure*}
-%
-%Looking at the plots of the features that vary over time (\autoref{fig:plot_features}) we see that certain attack types have a characteristic pattern in which they send packets by which they are easily recognizable.
+
+\begin{figure*}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_2_Brute Force-SSH-Patator"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_6_DoS - DDoS-DoS Slowhttptest"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
+
+\caption{The solid line shows the median value of the feature for all flows of a specific attack type. The shaded region are the first and third quartile.}
+\label{fig:plot_features}
+\end{figure*}
+
+Looking at the plots of the features that vary over time (\autoref{fig:plot_features}) we see that certain attack types have a characteristic pattern in which they send packets by which they are easily recognizable.
 We want to determine how the pattern of packet sizes and interarrival times is significant for the classification.

### 2019-11-25 ex3/paper/paper.tex
Change abstract

@@
 With the rise of ubiquitous \gls{ml} systems, malicious actors have been catching up quickly to find new ways to exploit them for profit.
-Recently developed \gls{aml} techniques focus on computer vision and their applicability to network traffic is not straightforward: Network traffic is not only sequential but there are also many constraints that have to be considered when attacking/defending an \gls{ml}-based Intrusion Detection System.
-For example, attackers can only modify certain features to evade detection and certain features, like packet interarrival time cannot become less than zero.
-We show that already the first couple of packets of a flow are of crucial importance and are likely going to be targeted by attackers.
-As a countermeasure we develop an easy-to-use training procedure and can confirm that it significantly and successfully reduces the attack surface.
+Recently developed explainability and \gls{aml} techniques focus on computer vision and their applicability to network traffic is not straightforward: Network packets expose fewer features than an image, are sequential and impose several constraints on their features: For example, destination ports cannot be modified since that could impair communication.
+We extend existing explainability techniques and propose new ones, specifically crafted for sequential data.
+Applying them shows that already the first couple of packets of a communication flow are of crucial importance and are likely going to be targeted by attackers.
+As a countermeasure we develop an easy-to-use adversarial training procedure and can confirm that it significantly and successfully reduces the attack surface.
 

### 2019-11-25 ex3/paper/paper.tex
Add scaffold of Adversarial Robustnet Score (ARSch)

@@
 \newacronym{pgd}{PGD}{Projected Gradient Descent}
+\newacronym{pdp}{PDP}{Partial Dependence Plot}
+\newacronym{arm}{ARM}{Adversarial Robustness Metric}
 
@@
 Various approaches have been proposed to extract features from flows and then perform anomaly detection with the results \cite{meghdouri_analysis_2018}.
-While this approach works well, it is problematic that first the whole flow has to be received and only afterwards it can be determined if there was an anomaly.
+While these approaches works well, it is problematic that first the whole flow has to be received and only afterwards it can be determined if there was an anomaly.
+Thus, we design a network anomaly detection system that operates on a per-packet basis and decides if a packet is anomalous based on features that are available for traffic that is encrypted above the transport layer, like for example TLS or QUIC.
+We show that our system has similar performance to other flow-based anomaly detection systems but can commonly detect anomalies \textit{before} the flow is over.
 
-Thus, we design a network anomaly detection system that operates on a per-packet basis and decides if a packet is anomalous based on features that are available for traffic that is encrypted above the transport layer, like for example TLS or QUIC.
-We show that our system has similar performance to other flow-based anomaly detection systems but can commonly detect anomalies before the flow is over.
+Since common explainability methods are designed to work with images or tabular data but not sequences, we extend existing explainability methods such as \glspl{pdp} \cite{friedman_greedy_2001} for sequential data.
+These show the varying characteristics of different attacks and how many packets it takes on average to detect them.
+Furthermore they aid with visualizing potentially vulnerable features for adversaries.
 
-Then we investigate the question whether adversarial samples can be found for this type of data.
+This leads us to investigate the question whether adversarial samples can be found for this type of data.
 This is not a trivially answerable question since adversarial samples have mostly been analyzed in the context of computer vision.
@@
 
-We can confirm that adversarial samples can be found even when considering the tight constraints mentioned before.
-Especially the first packets of a flow are critical for determining the decision of a recurrent \gls{ids}.
+Surprisingly, we can confirm that adversarial samples can be found even when considering the tight constraints mentioned before.
+Also astonishingly, feature importance method reveal that the manipulable features, that are left for adversaries to modify are not even particularly significant for the classification outcome.
+%Especially the first packets of a flow are critical for determining the decision of a recurrent \gls{ids}.
 We develop a simple methods for defending against adversaries who can manipulate arbitrary packets by simply leaving out potentially manipulable information from the classification process.
@@
 
-In particular,  in literature the use of Partial Dependency Plots (PDP) is proposed.
-If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space and $f(X) \in [0,1]$ denotes the neural network's prediction, the PDP for feature $i$ can be defined as
+In particular,  in literature the use of Partial Dependency Plots (\gls{pdp}) is proposed.
+If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space and $f(X) \in [0,1]$ denotes the neural network's prediction, the \gls{pdp} for feature $i$ can be defined as
 \begin{equation} \label{eq:pdp}
-\text{PDP}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right)
+\text{\gls{pdp}}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right)
 \end{equation}
 
-To inspect attack types in detail, in this research we experiment with a modified definition of the PDP  (with C being the attack classes):
+To inspect attack types in detail, in this research we experiment with a modified definition of the \gls{pdp}  (with C being the attack classes):
 \begin{equation} \label{eq:pdp_conditional}
-\text{PDP}_{c,i}(y) = \mathbb E_{\boldsymbol X | C = c} \left(f \left( X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n \right) \right)
+\text{\gls{pdp}}_{c,i}(y) = \mathbb E_{\boldsymbol X | C = c} \left(f \left( X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n \right) \right)
 \end{equation}
@@
 \caption{Examples of conditional PD plots for source and destination port.
-We only analyze these features with the PDP since it is not straightforward to define the PDP for features that can vary for each packet in a flow.
+We only analyze these features with the \gls{pdp} since it is not straightforward to define the \gls{pdp} for features that can vary for each packet in a flow.
 The y-axis shows the probability for malicious traffic.
@@
 
-In this variant of the PDP that we call conditional PDP, we don't take the whole dataset for the computation for the PDP but only a subset that belongs to a certain class.
-Specifically, we take each attack class and compute the PDP only over samples of this attack class.
+In this variant of the \gls{pdp} that we call conditional \gls{pdp}, we don't take the whole dataset for the computation for the \gls{pdp} but only a subset that belongs to a certain class.
+Specifically, we take each attack class and compute the \gls{pdp} only over samples of this attack class.
 The results in \autoref{fig:pdp} show that for some attack classes the port numbers play an important role.
@@
 We argue that this is because most services that regular users use have low port numbers.
-A known issue about PDPs in general is that dependencies of the feature under investigation and the remaining variables are not considered, hence potentially querying the neural network for feature combinations which cannot possibly occur.
+A known issue about \glspl{pdp} in general is that dependencies of the feature under investigation and the remaining variables are not considered, hence potentially querying the neural network for feature combinations which cannot possibly occur.
 
@@
 \end{tabular}
+
+\todo{In the adv table, round the numbers!!! \\
+For Max: Introduce two-level table\\
+What's the tradeoff for CW?}
 \end{table*}
@@
 \end{figure*}
+
+\section{Adversarial Robustness Metric}
+
+As \autoref{tab:adv_per_family} shows, for most attack types, most samples can be modified to look benign by an adversary by using \gls{cw}.
+High accuracy for a particular attack category does not imply that adversarial attacks are more likely to fail for these attacks.
+Thus we argue that beside the classical metrics such as accuracy, false positives etc., a new metric is needed, which quantifies how easily an adversary can fool the classifier.
+
+Such a method should be easy to compute and easy to interpret.
+We thus propose the \gls{arm}, which roughly works as follows: Use \gls{cw} with a small tradeoff to try to generate adversarial samples for an attack type.
+If more than half of the samples are now wrongly classified, stop.
+Otherwise increase the tradeoff of \gls{cw} and repeat.
+If after a predefined number of iterations -- e.g.
+100 -- still not more than half of the samples are adversarial, also break.
+
+The result is a number of adversarial samples (more than half in the ideal case) and some that could not be made adversarial.
+For example, if in total, there are 100 samples, there might be 54 which could be made adversarial.
+Out of these 54, the 50 with the smallest distance to the original samples are taken.
+Of these samples, the distance of each adversarial sample to the original sample is taken and averaged over all 50 samples.
+The resulting average distance is the \gls{arm}.
+The larger it is the better since then an adversary needs to modify the samples more to make them adversarial.
+If not more then half of the samples could be made adversarial, our metric is undefined since then apparently it is not possible for an adversary conduct adversarial attacks on the majority of samples.
+
+Setting the threshold for our metric to 50\% is arbitrary but we think that it is reasonable since if an attacker can make the majority of samples adversarial, we think that its attack is ``successful''.
+
+\todo{Insert results}
 

### 2019-11-25 ex3/learn.py

@@ -564,7 +564,7 @@ def get_feature_importance_distribution(test_data):
 				unique = np.unique(distribution[feature,:])
 				distribution[feature,:] = np.random.choice(unique, size=distribution.shape[1])
 	return distribution
-	
+
 # Right now this function replaces all values of one feature by random values sampled from the distribution of all features and looks how the accuracy changes.
 def feature_importance():
 
@@ -705,7 +705,7 @@ def mutinfo_feat_imp():
 	SAMPLE_COUNT = 100
 	PDF_FEATURE_BINS = 50
 	PDF_CONFIDENCE_BINS = 2
-	
+
 	n_fold = opt.nFold
 	fold = opt.fold
 	lstm_module.eval()
@@ -721,27 +721,27 @@ def mutinfo_feat_imp():
 	flow_lengths = np.zeros(opt.maxLength, dtype=int)
 
 	start_iterating = time.time()
-	
+
 	for real_ind, sample_ind in zip(test_indices, range(len(test_data))):
 		print (sample_ind)
 		flow, _, flow_categories = test_data[sample_ind]
 		cat = int(flow_categories[0,0])
 		flow_lengths[:flow.shape[0]] += 1
-		
+
 		# process constant features
 		lstm_module.init_hidden(len(constant_features)*SAMPLE_COUNT)
 		input_data = torch.FloatTensor(flow[:,None,:]).repeat(1,len(constant_features)*SAMPLE_COUNT,1)
-		
+
 		for k, feat_ind in enumerate(constant_features):
 			input_data[:,k*SAMPLE_COUNT:(k+1)*SAMPLE_COUNT,feat_ind] = torch.FloatTensor(np.random.choice(distribution[feat_ind,:], size=(input_data.shape[0],SAMPLE_COUNT)))
 
 		packed_input = torch.nn.utils.rnn.pack_padded_sequence(input_data, [input_data.shape[0]] *input_data.shape[1]).to(device)
-		
+
 		for i in range(flow.shape[0]):
 
 			output, _ = lstm_module(packed_input)
 			sigmoided = torch.sigmoid(output[0,:,0]).detach().cpu()
-			
+
 			for k,feat_ind in enumerate(features):
 				bin1 = (torch.round((PDF_FEATURE_BINS-1) * (input_data[i,:,feat_ind] - minmax[feat_ind][0]) / (minmax[feat_ind][1]-minmax[feat_ind][0]))).cpu().numpy().astype(int)
[…]

### 2019-11-27 ex3/learn.py
Modified mutinfo

@@ -282,6 +282,7 @@ def train():
 				with open(adv_filename(), 'wb') as f:
 					pickle.dump(train_data.adv_flows, f)
 
+@torch.no_grad()
 def test():
 
 	n_fold = opt.nFold
@@ -551,21 +552,47 @@ def dropout_feature_importance():
 	return (accuracy, accuracy_for_features)
 
 def get_feature_importance_distribution(test_data):
-	distribution = np.concatenate([item[0][:,:] for item in test_data], axis=0).transpose(1,0)
+	distribution = np.concatenate([item[0] for item in test_data], axis=0).transpose(1,0)
+	minmax = list(zip(np.min(distribution, axis=1), np.max(distribution, axis=1)))
 	if opt.adjustFeatImpDistribution:
 		numerical_features = [0,1,3]
 		# retain original distribution of iat, as iat varies over many magnitudes and we're mainly
 		# interested in feature importance of practical (=low) iats
 		retain_features = [2]
-		for feature in distribution.shape[0]:
+		for feature in range(distribution.shape[0]):
 			if feature in numerical_features:
 				distribution[feature,:] = np.linspace(np.min(distribution[feature,:]), np.max(distribution[feature,:]), distribution.shape[1])
 			elif feature not in retain_features:
 				unique = np.unique(distribution[feature,:])
 				distribution[feature,:] = np.random.choice(unique, size=distribution.shape[1])
-	return distribution
+	return distribution, minmax
+	
+def get_bin_boundaries(distribution, n_bins):
+	# For ideal resolution, choose bin boundaries so that each bin would
+	# get the equal number of hits for given distribution
+	spacing = np.linspace(0, distribution.shape[1]-1, n_bins+1, dtype=int)[1:]
+	boundaries = np.stack([ np.sort(distribution[feat_ind,:])[spacing] for feat_ind in range(distribution.shape[0]) ])
+	boundaries[:,-1] = np.inf
+	return boundaries
+	
+def compute_mutinfo(joint_pdf):
+	joint_pdf = joint_pdf / np.sum(joint_pdf)
+	marg_1 = np.sum(joint_pdf, axis=1)
[…]

### 2019-11-28 ex3/paper/paper.tex
Add bibliography add flow accuracy for feature importance methods in learn.py

@@
 \label{tab:feat_importance}
+\centering
 \begin{tabular}{l r}
 \toprule
-Feature & Accuracy drop \\ \midrule
+Feature & Acc.~drop \\ \midrule
 Protocol	&	0.1323	\\
@@
 As a first step to understanding the neural network's decisions, we tried to estimate how important individual features are for the model's predictions.
-
-To this end, we evaluated by what extent detection accuracy drops when withholding individual features from the classifier.
-We thus replaced values of individual features by values randomly drawn from the same distribution as the real feature values.
+The most commonly used techniques for this which are used by practitioners for \glspl{rnn} \cite{stackexchange_cross_validated_neural_2019} and \gls{dl} \cite{stackexchange_cross_validated_feature_2016} in general are based on adding noise to a particular feature and then evaluating the drop in accuracy that occurs.
+As we find it hard to find out the ``correct'' intensity of noise to add, we slightly vary the proposed approach: We do not add noise drawn from a probability distribution but instead sample the value for a feature from the distribution of all values of this feature.
+This makes the method non-parametric since the standard deviation of the noise does not need to be chosen.
 We ensured that features which stay constant for a flow, i.e.
@@
 While the first two features are essential for characterizing flows, a high number of packets with set SYN flag is an important characteristic for port scanning or certain DoS attacks.
+Also the fact that most of TCP flags do not play an important role in our dataset was expected.
+It is interesting to note, however, that the interarrival time of packets does not play a crucial role for the detection of attacks.
 
-Also the fact that most of TCP flags do not play an important role in our dataset, was expected.
-It is interesting to note, however, that the interarrival time of packets does not play a crucial role for the detection of attacks.
+While the method of adding noise to features/randomizing them is convenient since it is easy to implement and understand, we argue that it has some shortcomings: 
+The \gls{rnn} was never trained for dealing with ``garbage'' values that the noise/randomization methods create.
+For example, completely unrealistic feature combinations could occur that were never observed during training.
+Furthermore, sequences of features could occur that cannot occur in reality.
+
+We think that the ``ideal'' feature importance method for our scenario are Shapley values \cite{shapley_value_1953}: To analyze feature importance with them, the feature under investigation in completely omitted from the dataset and a model is trained without it.
+The resulting accuracy is then compared to the baseline accuracy with all features.
+The effort required to do this is high both in terms of storage space as well as training time since for $n$ features $n$ models need to be separately trained and stored.
+While this seems tractable, when also analyzing the feature importance of combinations of features, the number of models required becomes even larger: Out of $n$ features, $2^n$ unique subsets can be chosen, meaning that the computational demand rises exponentially.
+
+Considering the difficulties that arise from computing Shapley values, we propose a light-weight version of them: When training a model, for each sample, we leave out each feature with a certain probability by setting it to zero.
+On average, one feature gets zeroed out but it is also possible that none or more than one are left out.
+This procedure is equivalent to using dropout \cite{srivastava_dropout:_2014} with probability $\frac{1}{n}$ before the first layer.
+The result is the classifier learning to deal with missing features.
+As it can also happen that several features are left out for a sample, our method can also be used to analyze the effect of multiple features missing and can thus show possibly correlated features.
+
+\todo{Add refs to Stackoverflow posts and add our custom feature importance methods.}
 

### 2019-11-28 ex3/learn.py

@@ -481,7 +481,9 @@ def dropout_feature_importance():
 	attack_numbers = mapping.values()
 
 	results_by_attack_number = [list() for _ in range(min(attack_numbers), max(attack_numbers)+1)]
+	results_flow_by_attack_number = [list() for _ in range(min(attack_numbers), max(attack_numbers)+1)]
 	dropout_results_by_attack_number = [[list() for _ in range(test_x.shape[0])] for _ in range(min(attack_numbers), max(attack_numbers)+1)]
+	dropout_results_flow_by_attack_number = [[list() for _ in range(test_x.shape[0])] for _ in range(min(attack_numbers), max(attack_numbers)+1)]
 
 	for input_data, labels, categories in test_loader:
 
@@ -513,6 +515,7 @@ def dropout_feature_importance():
 			flow_category = int(categories_padded[0, batch_index,:].squeeze().item())
 
 			results_by_attack_number[flow_category].append(flow_output)
+			results_flow_by_attack_number[flow_category].append(flow_output[-1])
 
 		for feature_index in range(test_x.shape[0]):
 			# Now draw individual feature values at random and query the nn for modified flows
@@ -531,24 +534,31 @@ def dropout_feature_importance():
 			# Data is (Sequence Index, Batch Index, Feature Index)
 			for batch_index in range(output.shape[1]):
 				flow_length = seq_lens[batch_index]
-
 				flow_output = (torch.round(sigmoided_output[:flow_length,batch_index,:]) == labels_padded[:flow_length,batch_index,:]).detach().cpu().numpy()
-
 				assert (categories_padded[0, batch_index,:] == categories_padded[:flow_length, batch_index,:]).all()
-
 				flow_category = int(categories_padded[0, batch_index,:].squeeze().item())
-
 				dropout_results_by_attack_number[flow_category][feature_index].append(flow_output)
+				dropout_results_flow_by_attack_number[flow_category][feature_index].append(flow_output[-1])
 
 	accuracy = np.mean(np.concatenate([subitem for item in results_by_attack_number for subitem in item], axis=0))
+	flow_accuracy = np.mean(np.concatenate([subitem for item in results_flow_by_attack_number for subitem in item], axis=0))
 	accuracy_for_features = []
-	with open("features.json", "r") as f:
+	accuracy_flow_for_features = []
+	with open("features_meaningful_names.json", "r") as f:
 		feature_array = json.load(f)
[…]

### 2019-11-29 ex3/paper/paper.tex
Add human-readable names for features
Add feature sensitivity

@@
 \end{table}
+When investigating an \gls{ml} classifier, it is natural to ask which inputs have a large influence on the classifier's prediction.
+We feel the need to distinguish metrics for this purpose based on their main aim:
+
+A large amount of research has been spent on finding \emph{feature importance} metrics, which allow the selection  of high-importance features, providing a reasonably good classification performance while resulting in a more light-weight classifier.
+
+Both adversarial machine learning and explainable machine learning bring up the question to what extent individual features are able to change the prediction outcome.
+While being seemingly similar at first sight, traditional feature importance can yield markedly wrong results for this case.
+To distinguish, we propose the term \emph{feature sensitivity} for such metrics.
+
 As a first step to understanding the neural network's decisions, we tried to estimate how important individual features are for the model's predictions.
@@
 
-\todo{Add refs to Stackoverflow posts and add our custom feature importance methods.}
+\todo{Change table with recent values and include table for dropout\_feature\_importance.}
 

### 2019-12-04 ex3/paper/paper.tex
Change table; forgot to commit

--- 2019-11-28 ex3/paper/paper.tex
@@
 \begin{table}
-\caption{Accuracy drop when replacing feature values with random data.}
+\caption{Accuracy drop when \textbf{Left:} when replacing feature values with random data \textbf{Right:} Feature dropout}
 \label{tab:feat_importance}
@@
 Feature & Acc.~drop \\ \midrule
-Protocol	&	0.1323	\\
-Packet Length	&	0.1203	\\
-SYN Flag	&	0.1067	\\
-ACK Flag	&	0.0805	\\
-Destination port	&	0.0796	\\
-PSH Flag	&	0.0767	\\
-Direction	&	0.0732	\\
-Source port	&	0.0254	\\
-Interarrival time	&	0.0200	\\
-RST Flag	&	0.0116	\\
-FIN Flag	&	0.0100	\\
-ECE	Flag &	0.0000	\\
-URG Flag	&	0.0000	\\
-CWR Flag	&	0.0000	\\
-NS Flag	&	0.0000	\\
+Protocol & 0.207 \\
+Packet Length & 0.165 \\
+SYN Flag & 0.099 \\
+ACK Flag & 0.084 \\
+Direction & 0.073 \\
+Destination port & 0.071 \\
+Source port & 0.06 \\
+RST Flag & 0.057 \\
+PSH Flag & 0.056 \\
+Interarrival time & 0.024 \\
+FIN Flag & 0.012 \\
+URG Flag & 0 \\
+ECE Flag & 0 \\
+CWR Flag & 0 \\
+NS Flag & 0 \\
+\bottomrule
+\end{tabular}
+\begin{tabular}{l r}
+\toprule
+Feature & Acc.~drop \\ \midrule
+Destination port & 0.025 \\
+Source port & 0.003 \\
+RST Flag & 0.001 \\
+ACK Flag & 0.001 \\
+Protocol & 0.001 \\
+Packet Length & 0.001 \\
+Direction & 0.001 \\
+SYN Flag & 0 \\
+Interarrival time & 0 \\
+FIN Flag & 0 \\
+ECE Flag & 0 \\
+URG Flag & 0 \\
+CWR Flag & 0 \\
+NS Flag & 0 \\
+PSH Flag & 0 \\
 \bottomrule
@@
 
-Considering the difficulties that arise from computing Shapley values, we propose a light-weight version of them: When training a model, for each sample, we leave out each feature with a certain probability by setting it to zero.
+Considering the difficulties that arise from computing Shapley values, we propose a light-weight version of them, called \textit{feature dropout}: When training a model, for each sample, we leave out each feature with a certain probability by setting it to zero.
 On average, one feature gets zeroed out but it is also possible that none or more than one are left out.
@@
 As it can also happen that several features are left out for a sample, our method can also be used to analyze the effect of multiple features missing and can thus show possibly correlated features.
+The results in \autoref{tab:feat_importance} show that the randomization feature importance method vastly overestimates features' importance.
+With feature dropout, it becomes apparent that there are very few features that actually contain unique information and that affect accuracy when left out: the destination port and the source port.
+\todo{Add correlation and explain why most features have no importance}
 
-\todo{Add refs to Stackoverflow posts and add our custom feature importance methods.}
+A model trained with feature dropout typically yields slightly lower accuracy than a regularly trained model, even if no features are left out (flow accuracy of 0.9943 vs.~0.9965 \todo{Update results in the first table at the very top since our results now are better than they used to be originally.}).
+We thus recommend training two models: One regular one and one with feature dropout to use for the feature importance.
 

### 2019-12-04 ex3/paper/paper.tex
Restructure and fix feature importance section

@@
 \usepackage{natbib}
+\usepackage{subfig}
 
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{An in-depth study of Explainability and Adversarial Robustness in RNNs}
+\title{An in-depth study of Explainability and Adversarial Robustness for RNNs}
 
@@
 \newacronym{pdp}{PDP}{Partial Dependence Plot}
-\newacronym{arm}{ARM}{Adversarial Robustness Metric}
+\newacronym{ars}{ARS}{Adversarial Robustness Score}
 
@@
 With the rise of ubiquitous \gls{ml} systems, malicious actors have been catching up quickly to find new ways to exploit them for profit.
-Recently developed explainability and \gls{aml} techniques focus on computer vision and their applicability to network traffic is not straightforward: Network packets expose fewer features than an image, are sequential and impose several constraints on their features: For example, destination ports cannot be modified since that could impair communication.
-We extend existing explainability techniques and propose new ones, specifically crafted for sequential data.
+Recently developed \gls{aml} techniques focus on computer vision and their applicability to network traffic is not straightforward: Network packets expose fewer features than an image, are sequential and impose several constraints on their features: For example, destination ports cannot be modified since that could impair communication.
+We show that despite these completely different characteristics, adversarial samples can be reliably generated.
+This shows that traditional evaluation metrics such as accuracy are not sufficient for quantifying the adversarial threat and we propose \gls{ars} as a countermeasure.
+Feature importance methods show that even though the modified features are not very relevant for the classifier, they can still be abused to generate adversarial samples.
+We then extend existing explainability techniques and propose new ones, specifically crafted for sequential data.
 Applying them shows that already the first couple of packets of a communication flow are of crucial importance and are likely going to be targeted by attackers.
-As a countermeasure we develop an easy-to-use adversarial training procedure and can confirm that it significantly and successfully reduces the attack surface.
+As a countermeasure we develop an easy-to-use adversarial training procedure and can confirm that it significantly and successfully reduces the attack surface and improves \gls{ars}.
 
@@
 In the past, a viable way of performing intrusion detection was to inspect the content of packets themselves and detect if a packet delivers potentially harmful content.
-
 More recently, with the rise of omnipresent encryption, available features have changed.
@@
 
-Since common explainability methods are designed to work with images or tabular data but not sequences, we extend existing explainability methods such as \glspl{pdp} \cite{friedman_greedy_2001} for sequential data.
-These show the varying characteristics of different attacks and how many packets it takes on average to detect them.
-Furthermore they aid with visualizing potentially vulnerable features for adversaries.
-
-This leads us to investigate the question whether adversarial samples can be found for this type of data.
+Then we investigate the question whether adversarial samples can be found for this type of data.
 This is not a trivially answerable question since adversarial samples have mostly been analyzed in the context of computer vision.
@@
 Surprisingly, we can confirm that adversarial samples can be found even when considering the tight constraints mentioned before.
-Also astonishingly, feature importance method reveal that the manipulable features, that are left for adversaries to modify are not even particularly significant for the classification outcome.
-%Especially the first packets of a flow are critical for determining the decision of a recurrent \gls{ids}.
+Thus we argue that traditional methods like accuracy are not sufficient for evaluating security-sensitive \gls{ml} systems.
+The new performance score \gls{ars} precisely captures the notion of how easily an adversary can generate adversarial samples.
 We develop a simple methods for defending against adversaries who can manipulate arbitrary packets by simply leaving out potentially manipulable information from the classification process.
 The resulting classifier is slightly less accurate but is still useful to be deployed in a real scenario.
-Finally we develop a custom adversarial training procedure which especially focuses on easy of use and low resource consumption and aims to keep the number of additionally introduced training hyperparameters low.
+We develop a custom adversarial training procedure which especially focuses on easy of use and low resource consumption and aims to keep the number of additionally introduced training hyperparameters low.
 With this novel strategy we can reduce attack surface and harden the resulting \gls{ids} while keeping similar classification performance like for the original \gls{ml} model.
+
+Astonishingly, feature importance method reveal that the manipulable features, that are left for adversaries to modify are not even particularly significant for the classification outcome.
+%Especially the first packets of a flow are critical for determining the decision of a recurrent \gls{ids}.
+
+Since common explainability methods are designed to work with images or tabular data but not sequences, we extend existing explainability methods such as \glspl{pdp} \cite{friedman_greedy_2001} for sequential data.
+These show the varying characteristics of different attacks and they aid with visualizing potentially vulnerable features for adversaries.
 
@@
 & RF & MLP & LSTM \\ \midrule
-Accuracy	&	0.9974 & 0.9970	& 0.9947 \\
-Precision	&	0.9965 & 0.9979 & 0.9894	\\
-Recall	&	0.9931 & 0.9903	& 0.9897 \\
-F1	&	0.9948 & 0.9941	& 0.9896 \\
-Youden	&	0.9920 & 0.9896 & 0.9861 \\
+Accuracy	&	0.9974 & 0.9970	& 0.9965 \\
+Precision	&	0.9965 & 0.9979 & 0.9956	\\
+Recall	&	0.9931 & 0.9903	& 0.9905 \\
+F1	&	0.9948 & 0.9941	& 0.9930 \\
+Youden	&	0.9920 & 0.9896 & 0.9890 \\
 \bottomrule
@@
 \end{table}
-\section{Explaining Recurrent Neural Networks}
-From a naive perspective, one might be tempted to reuse explainability methods for recurrent neural networks by considering a flow the sum of its packet features.
-We identify several problems which occur when trying to explain decisions made by recurrent neural networks.
-
-\begin{itemize}
-\item
-\textbf{Feature quantity.}
-The number of features that is fed into a recurrent neural network is found as the product of packet features with the total length of the flow.
-For long flows, the total number of inputs can therefore grow to a substantial size.
-
-\item
-\textbf{Variable sequence lengths.}
-The length of different flows might differ tremendously.
-Hence, features which are important for the network's outcome for one sequence might not even exist for other flows.
-Furthermore, the question arises how the distance between flows of different length can be computed.
-
-\item
-\textbf{Lack of a distance measure.}
-However, even if we restricted our analysis to flows of a constant length, a flow is different from the plain concatenation of its packet features.
-If we, for example, consider a classification problem as in this document, we expect that after a certain amount of time steps the classifier is able to perform the classification with a certain confidence.
-Arguably, features at the beginning of a flow are thus more important for the classifier.
-Giving differences at the end of a sequence the same weight as differences at its beginning, therefore seems incorrect.
-
-\item
-\textbf{Multiple prediction outputs.}
-Often a recurrent neural network produces an output at each time step. When trying to apply explainability methods, hence, the question arises which prediction to base the methods on.
-
-The natural choice is to base the methods on the prediction output which occurs at the same time step as the feature under investigation.
-Not only does this approach require a substantially lower amount of computational resources compared to predictions occurring at a later time step, but we also expect the immediately occurring prediction outcome to exhibit a stronger correlation to the feature under investigation than a prediction outcome which occurring many time steps later.
-However, due to the complex decision processes that can occur for deep neural networks, it is absolutely  uncertain if the network's final prediction is correctly depicted when following this approach.
-\end{itemize}
-
-\subsection{Feature Importance Metrics}
-\begin{table}
-\caption{Accuracy drop when \textbf{Left:} when replacing feature values with random data \textbf{Right:} Feature dropout}
-\label{tab:feat_importance}
-\centering
-\begin{tabular}{l r}
-\toprule
-Feature & Acc.~drop \\ \midrule
-Protocol & 0.207 \\
-Packet Length & 0.165 \\
-SYN Flag & 0.099 \\
-ACK Flag & 0.084 \\
-Direction & 0.073 \\
-Destination port & 0.071 \\
-Source port & 0.06 \\
-RST Flag & 0.057 \\
-PSH Flag & 0.056 \\
-Interarrival time & 0.024 \\
-FIN Flag & 0.012 \\
-URG Flag & 0 \\
-ECE Flag & 0 \\
-CWR Flag & 0 \\
-NS Flag & 0 \\
-\bottomrule
-\end{tabular}
-\begin{tabular}{l r}
-\toprule
-Feature & Acc.~drop \\ \midrule
-Destination port & 0.025 \\
-Source port & 0.003 \\
-RST Flag & 0.001 \\
-ACK Flag & 0.001 \\
-Protocol & 0.001 \\
-Packet Length & 0.001 \\
-Direction & 0.001 \\
-SYN Flag & 0 \\
-Interarrival time & 0 \\
-FIN Flag & 0 \\
-ECE Flag & 0 \\
-URG Flag & 0 \\
-CWR Flag & 0 \\
-NS Flag & 0 \\
-PSH Flag & 0 \\
-\bottomrule
-\end{tabular}
-\end{table}
-When investigating an \gls{ml} classifier, it is natural to ask which inputs have a large influence on the classifier's prediction.
-We feel the need to distinguish metrics for this purpose based on their main aim:
-
-A large amount of research has been spent on finding \emph{feature importance} metrics, which allow the selection  of high-importance features, providing a reasonably good classification performance while resulting in a more light-weight classifier.
-
-Both adversarial machine learning and explainable machine learning bring up the question to what extent individual features are able to change the prediction outcome.
-While being seemingly similar at first sight, traditional feature importance can yield markedly wrong results for this case.
-To distinguish, we propose the term \emph{feature sensitivity} for such metrics.
-
-As a first step to understanding the neural network's decisions, we tried to estimate how important individual features are for the model's predictions.
-The most commonly used techniques for this which are used by practitioners for \glspl{rnn} \cite{stackexchange_cross_validated_neural_2019} and \gls{dl} \cite{stackexchange_cross_validated_feature_2016} in general are based on adding noise to a particular feature and then evaluating the drop in accuracy that occurs.
-As we find it hard to find out the ``correct'' intensity of noise to add, we slightly vary the proposed approach: We do not add noise drawn from a probability distribution but instead sample the value for a feature from the distribution of all values of this feature.
-This makes the method non-parametric since the standard deviation of the noise does not need to be chosen.
-We ensured that features which stay constant for a flow, i.e.
-source port, destination port and protocol, also stay constant for the distorted flows.
-
-\autoref{tab:feat_importance} shows our results.
-The results match to a large extent with the expectations one has when classifying network flows.
-In particular, the most important features turn out to be the used protocol, packet length and the SYN TCP flag.
-While the first two features are essential for characterizing flows, a high number of packets with set SYN flag is an important characteristic for port scanning or certain DoS attacks.
-Also the fact that most of TCP flags do not play an important role in our dataset was expected.
-It is interesting to note, however, that the interarrival time of packets does not play a crucial role for the detection of attacks.
-
-While the method of adding noise to features/randomizing them is convenient since it is easy to implement and understand, we argue that it has some shortcomings: 
-The \gls{rnn} was never trained for dealing with ``garbage'' values that the noise/randomization methods create.
-For example, completely unrealistic feature combinations could occur that were never observed during training.
-Furthermore, sequences of features could occur that cannot occur in reality.
-
-We think that the ``ideal'' feature importance method for our scenario are Shapley values \cite{shapley_value_1953}: To analyze feature importance with them, the feature under investigation in completely omitted from the dataset and a model is trained without it.
-The resulting accuracy is then compared to the baseline accuracy with all features.
-The effort required to do this is high both in terms of storage space as well as training time since for $n$ features $n$ models need to be separately trained and stored.
-While this seems tractable, when also analyzing the feature importance of combinations of features, the number of models required becomes even larger: Out of $n$ features, $2^n$ unique subsets can be chosen, meaning that the computational demand rises exponentially.
-
-Considering the difficulties that arise from computing Shapley values, we propose a light-weight version of them, called \textit{feature dropout}: When training a model, for each sample, we leave out each feature with a certain probability by setting it to zero.
-On average, one feature gets zeroed out but it is also possible that none or more than one are left out.
-This procedure is equivalent to using dropout \cite{srivastava_dropout:_2014} with probability $\frac{1}{n}$ before the first layer.
-The result is the classifier learning to deal with missing features.
-As it can also happen that several features are left out for a sample, our method can also be used to analyze the effect of multiple features missing and can thus show possibly correlated features.
-The results in \autoref{tab:feat_importance} show that the randomization feature importance method vastly overestimates features' importance.
-With feature dropout, it becomes apparent that there are very few features that actually contain unique information and that affect accuracy when left out: the destination port and the source port.
-\todo{Add correlation and explain why most features have no importance}
-
-A model trained with feature dropout typically yields slightly lower accuracy than a regularly trained model, even if no features are left out (flow accuracy of 0.9943 vs.~0.9965 \todo{Update results in the first table at the very top since our results now are better than they used to be originally.}).
-We thus recommend training two models: One regular one and one with feature dropout to use for the feature importance.
-
-\subsection{Explainability Plots}
-During flow extraction we used a usual 5-tuple flow key, which distinguishes flows based on the protocol they use and their source and destination port and IP address.
-Hence, these features are constant for all packets in one flow, which allows us to use explainability methods which have been proposed for non-recurrent classification methods.
-
-In particular,  in literature the use of Partial Dependency Plots (\gls{pdp}) is proposed.
-If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space and $f(X) \in [0,1]$ denotes the neural network's prediction, the \gls{pdp} for feature $i$ can be defined as
-\begin{equation} \label{eq:pdp}
-\text{\gls{pdp}}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right)
-\end{equation}
-
-To inspect attack types in detail, in this research we experiment with a modified definition of the \gls{pdp}  (with C being the attack classes):
-\begin{equation} \label{eq:pdp_conditional}
-\text{\gls{pdp}}_{c,i}(y) = \mathbb E_{\boldsymbol X | C = c} \left(f \left( X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n \right) \right)
-\end{equation}
-
-\begin{figure*}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_8_DoS - DDoS-Heartbleed"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_10_Normal"}.pdf}
-
-\caption{Examples of conditional PD plots for source and destination port.
-We only analyze these features with the \gls{pdp} since it is not straightforward to define the \gls{pdp} for features that can vary for each packet in a flow.
-The y-axis shows the probability for malicious traffic.
-}
-\label{fig:pdp}
-\end{figure*}
-
-In this variant of the \gls{pdp} that we call conditional \gls{pdp}, we don't take the whole dataset for the computation for the \gls{pdp} but only a subset that belongs to a certain class.
-Specifically, we take each attack class and compute the \gls{pdp} only over samples of this attack class.
-The results in \autoref{fig:pdp} show that for some attack classes the port numbers play an important role.
-When looking at the class of normal traffic (non-attack traffic) it becomes apparent that traffic destined to a high source port is generally indicative of an attack.
-We argue that this is because most services that regular users use have low port numbers.
-A known issue about \glspl{pdp} in general is that dependencies of the feature under investigation and the remaining variables are not considered, hence potentially querying the neural network for feature combinations which cannot possibly occur.
-
-As apparent from \autoref{eq:pdp}, for a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.
-
-For this reason, Accumulated Local Effects (ALE) plots have been proposed in literature.
-The basic idea for ALE is that instead of using all samples, only the closest samples are used for assessing the effects of one feature.
-ALE plots therefore inherently require a distance measure for finding the closest samples, which, as revealed in the above discussion, in our case is not possible in a straight-forward manner.
-
-\subsection{Plots for Sequences}
-
-As apparent from the above discussion, trying to explain how packet features influence the model's predictions, poses various difficulties.
-As a first step to explaining our model's predictions, we investigate how important features at different time steps are for the model's decisions.
-Intuitively, features at the beginning of a flow should be the most important while the model's predictions should not vary significantly anymore, as soon as the model has come to a decision.
-
-\begin{figure*}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_0_Botnet-ARES"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_14_Web Attack-XSS"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
-
-\caption{Median confidence per time step and first and third quartiles for selected attack types.
-For the majority of attacks, confidence increases in the first few steps and then stays almost constant at 1.}
-\label{fig:confidence}
-\end{figure*}
-
-To determine if this hypothesis is correct, we plot the attack confidence of the classifier for each time step for each attack class.
-At each time step we average over all flows and also show the first and third quartiles.
-\autoref{fig:confidence} shows that this hypothesis is correct for most attack classes.
-At the first couple of packets the confidence for maliciousness is not very high yet but towards the end of the flow it reaches values close to 1 and stays there.
-
-\begin{figure*}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_0_Botnet-ARES"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_1_Brute Force-FTP-Patator"}.pdf}
-
-\caption{The solid line shows the mean value of the feature for all flows of a specific attack type. The shaded region shows the change in confidence that occurs when the feature is varied.}
-\label{fig:pred_plots2}
-\end{figure*}
-
-Next, we want to investigate if our classifier just memorized its inputs.
-The most interesting packet features that can be used to visualize this in our use case are the inter-arrival time (IAT) between packets in a flows and the packet length.
-To depict how the model comes to a decision, we determined the total minimum and maximum for these features that ever occur for all flows.
-For a given flow $X$ we then varied feature values at each time step between minimum and maximum and plot how the confidence changes when the feature varies (\autoref{fig:pred_plots2}) depict the averaged predictions for different attack families.
-It seems clear that the classifier didn't just memorize the inputs but actually learned them since there is always a smooth decrease of confidence if we change the feature values.
-Instead, when memorizing, we would expect a sharp decline of confidence to occur.
-Another interesting observation is that after the first couple of packets the classifier is certain about its decision and changing features doesn't change its confidence any longer.
-This allows the conclusion that if an attacker is able to change the first few packets, she could possibly change the decision of the classifier.
-We will further investigate this question in section \autoref{sec:adv}.
-
-\begin{figure*}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_2_Brute Force-SSH-Patator"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_6_DoS - DDoS-DoS Slowhttptest"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
-
-\caption{The solid line shows the median value of the feature for all flows of a specific attack type. The shaded region are the first and third quartile.}
-\label{fig:plot_features}
-\end{figure*}
-
-Looking at the plots of the features that vary over time (\autoref{fig:plot_features}) we see that certain attack types have a characteristic pattern in which they send packets by which they are easily recognizable.
-We want to determine how the pattern of packet sizes and interarrival times is significant for the classification.
-\autoref{fig:plot_features} clearly shows that some attack types have very characteristic patterns of packet sizes and interarrival times.
 
@@
 Such a method should be easy to compute and easy to interpret.
-We thus propose the \gls{arm}, which roughly works as follows: Use \gls{cw} with a small tradeoff to try to generate adversarial samples for an attack type.
+We thus propose the \gls{ars}, which roughly works as follows: Use \gls{cw} with a small tradeoff to try to generate adversarial samples for an attack type.
 If more than half of the samples are now wrongly classified, stop.
@@
 Of these samples, the distance of each adversarial sample to the original sample is taken and averaged over all 50 samples.
-The resulting average distance is the \gls{arm}.
+The resulting average distance is the \gls{ars}.
 The larger it is the better since then an adversary needs to modify the samples more to make them adversarial.
@@
 
+\section{Explaining Recurrent Neural Networks}
+From a naive perspective, one might be tempted to reuse explainability methods for recurrent neural networks by considering a flow the sum of its packet features.
+We identify several problems which occur when trying to explain decisions made by recurrent neural networks.
+
+\begin{itemize}
+\item
+\textbf{Feature quantity.}
+The number of features that is fed into a recurrent neural network is found as the product of packet features with the total length of the flow.
+For long flows, the total number of inputs can therefore grow to a substantial size.
+
+\item
+\textbf{Variable sequence lengths.}
+The length of different flows might differ tremendously.
+Hence, features which are important for the network's outcome for one sequence might not even exist for other flows.
+Furthermore, the question arises how the distance between flows of different length can be computed.
+
+\item
+\textbf{Lack of a distance measure.}
+However, even if we restricted our analysis to flows of a constant length, a flow is different from the plain concatenation of its packet features.
+If we, for example, consider a classification problem as in this document, we expect that after a certain amount of time steps the classifier is able to perform the classification with a certain confidence.
+Arguably, features at the beginning of a flow are thus more important for the classifier.
+Giving differences at the end of a sequence the same weight as differences at its beginning, therefore seems incorrect.
+
+\item
+\textbf{Multiple prediction outputs.}
+Often a recurrent neural network produces an output at each time step. When trying to apply explainability methods, hence, the question arises which prediction to base the methods on.
+
+The natural choice is to base the methods on the prediction output which occurs at the same time step as the feature under investigation.
+Not only does this approach require a substantially lower amount of computational resources compared to predictions occurring at a later time step, but we also expect the immediately occurring prediction outcome to exhibit a stronger correlation to the feature under investigation than a prediction outcome which occurring many time steps later.
+However, due to the complex decision processes that can occur for deep neural networks, it is absolutely  uncertain if the network's final prediction is correctly depicted when following this approach.
+\end{itemize}
+
+%\subfloat[Throughput of the fq \gls{qdisc}\label{fig:fqThroughput}
+%]{
+%\includegraphics[width=\columnwidth]{{{plots/throughput_1_fq_cubic_10_20_60_0.5_bw_1571240937211.pcap}}}
+%}{}
+
+\subsection{Feature Importance Metrics}
+\begin{table}
+\caption{Accuracy drop for:}
+\label{tab:feat_importance}
+\subfloat[Random replacement method\label{fig:featRandom}
+]{
+\begin{tabular}{l r}
+\toprule
+Feature & Acc.~drop \\ \midrule
+Protocol & 0.207 \\
+Packet Length & 0.165 \\
+SYN Flag & 0.099 \\
+ACK Flag & 0.084 \\
+Direction & 0.073 \\
+Destination port & 0.071 \\
+Source port & 0.06 \\
+RST Flag & 0.057 \\
+PSH Flag & 0.056 \\
+Interarrival time & 0.024 \\
+FIN Flag & 0.012 \\
+URG Flag & 0 \\
+ECE Flag & 0 \\
+CWR Flag & 0 \\
+NS Flag & 0 \\
+\bottomrule
+\end{tabular}
+}{}
+\subfloat[Feature dropout\label{fig:featDropout}
+]{
+\begin{tabular}{l r}
+\toprule
+Feature & Acc.~drop \\ \midrule
+Destination port & 0.025 \\
+Source port & 0.003 \\
+RST Flag & 0.001 \\
+ACK Flag & 0.001 \\
+Protocol & 0.001 \\
+Packet Length & 0.001 \\
+Direction & 0.001 \\
+SYN Flag & 0 \\
+Interarrival time & 0 \\
+FIN Flag & 0 \\
+ECE Flag & 0 \\
+URG Flag & 0 \\
+CWR Flag & 0 \\
+NS Flag & 0 \\
+PSH Flag & 0 \\
+\bottomrule
+\end{tabular}
+}{}
+\end{table}
+When investigating an \gls{ml} classifier, it is natural to ask which inputs have a large influence on the classifier's prediction.
+We feel the need to distinguish metrics for this purpose based on their main aim:
+
+A large amount of research has been spent on finding \emph{feature importance} metrics, which allow the selection  of high-importance features, providing a reasonably good classification performance while resulting in a more light-weight classifier.
+
+Both adversarial machine learning and explainable machine learning bring up the question to what extent individual features are able to change the prediction outcome.
+While being seemingly similar at first sight, traditional feature importance can yield markedly wrong results for this case.
+To distinguish, we propose the term \emph{feature sensitivity} for such metrics.
+
+As a first step to understanding the neural network's decisions, we tried to estimate how important individual features are for the model's predictions.
+The most commonly used techniques for this which are used by practitioners for \glspl{rnn} \cite{stackexchange_cross_validated_neural_2019} and \gls{dl} \cite{stackexchange_cross_validated_feature_2016} in general are based on adding noise to a particular feature and then evaluating the drop in accuracy that occurs.
+We argue that it is hard to find out the ``correct'' intensity of noise to add, we slightly vary the proposed approach: We do not add noise drawn from a probability distribution but instead sample the value for a feature from the distribution of all values of this feature.
+This makes the method non-parametric since the standard deviation of the noise does not need to be chosen.
+We ensured that features which stay constant for a flow, i.e.
+source port, destination port and protocol, also stay constant throughout the whole flow when randomizing the feature.
+
+Table~\ref{fig:featRandom} shows our results.
+The results match to a large extent with the expectations one has when classifying network flows.
+In particular, the most important features turn out to be the used protocol, packet length and the SYN TCP flag.
+While the first two features are essential for characterizing flows, a high number of packets with set SYN flag is an important characteristic for port scanning or certain DoS attacks.
+Also the fact that most of TCP flags do not play an important role in our dataset was expected.
+It is interesting to note, however, that the interarrival time of packets does not play a crucial role for the detection of attacks.
+
+While the method of adding noise to features/randomizing them is convenient since it is easy to implement and understand, we argue that it has some shortcomings: 
+The \gls{rnn} was never trained for dealing with ``garbage'' values that the noise/randomization methods create.
+For example, completely unrealistic feature combinations could occur that were never observed during training.
+Furthermore, sequences of features could occur that cannot occur in reality.
+
+We think that the ``ideal'' feature importance method for our scenario are Shapley values \cite{shapley_value_1953}: To analyze feature importance with them, the feature under investigation in completely omitted from the dataset and a model is trained without it.
+The resulting accuracy is then compared to the baseline accuracy with all features.
+The effort required to do this is high both in terms of storage space as well as training time since for $n$ features $n$ models need to be separately trained and stored.
+While this seems tractable, when also analyzing the feature importance of combinations of features, the number of models required becomes even larger: Out of $n$ features, $2^n$ unique subsets can be chosen, meaning that the computational demand rises exponentially.
+
+Considering the difficulties that arise from computing Shapley values, we propose a light-weight version of them, called \textit{feature dropout}: When training a model, for each sample, we leave out each feature with a certain probability by setting it to zero.
+On average, one feature gets zeroed out but it is also possible that none or more than one are left out.
+This procedure is equivalent to using dropout \cite{srivastava_dropout:_2014} with probability $\frac{1}{n}$ before the first layer.
+The result is the classifier learning to deal with missing features.
+As it can also happen that several features are left out for a sample, our method can also be used to analyze the effect of multiple features missing and can thus show possibly correlated features.
+The results in Table~\ref{fig:featDropout} show unlike the randomization feature importance method, the feature dropout method does not vastly overestimate features' importance.
+With feature dropout, it becomes apparent that there are very few features that actually contain unique information and that affect accuracy when left out: the destination port and the source port.
+\todo{Add correlation and explain why most features have no importance}
+
+A model trained with feature dropout typically yields slightly lower accuracy than a regularly trained model, even if no features are left out (flow accuracy of 0.9943 vs.~0.9965.
+We thus recommend training two models: One regular one and one with feature dropout to use for the feature importance.
+
+\subsection{Explainability Plots}
+During flow extraction we used a usual 5-tuple flow key, which distinguishes flows based on the protocol they use and their source and destination port and IP address.
+Hence, these features are constant for all packets in one flow, which allows us to use explainability methods which have been proposed for non-recurrent classification methods.
+
+In particular,  in literature the use of Partial Dependency Plots (\gls{pdp}) is proposed.
+If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space and $f(X) \in [0,1]$ denotes the neural network's prediction, the \gls{pdp} for feature $i$ can be defined as
+\begin{equation} \label{eq:pdp}
+\text{\gls{pdp}}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right)
+\end{equation}
+
+To inspect attack types in detail, in this research we experiment with a modified definition of the \gls{pdp}  (with C being the attack classes):
+\begin{equation} \label{eq:pdp_conditional}
+\text{\gls{pdp}}_{c,i}(y) = \mathbb E_{\boldsymbol X | C = c} \left(f \left( X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n \right) \right)
+\end{equation}
+
+\begin{figure*}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_8_DoS - DDoS-Heartbleed"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_10_Normal"}.pdf}
+
+\caption{Examples of conditional PD plots for source and destination port.
+We only analyze these features with the \gls{pdp} since it is not straightforward to define the \gls{pdp} for features that can vary for each packet in a flow.
+The y-axis shows the probability for malicious traffic.
+}
+\label{fig:pdp}
+\end{figure*}
+
+In this variant of the \gls{pdp} that we call conditional \gls{pdp}, we don't take the whole dataset for the computation for the \gls{pdp} but only a subset that belongs to a certain class.
+Specifically, we take each attack class and compute the \gls{pdp} only over samples of this attack class.
+The results in \autoref{fig:pdp} show that for some attack classes the port numbers play an important role.
+When looking at the class of normal traffic (non-attack traffic) it becomes apparent that traffic destined to a high source port is generally indicative of an attack.
+We argue that this is because most services that regular users use have low port numbers.
+A known issue about \glspl{pdp} in general is that dependencies of the feature under investigation and the remaining variables are not considered, hence potentially querying the neural network for feature combinations which cannot possibly occur.
+
+As apparent from \autoref{eq:pdp}, for a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.
+
+For this reason, Accumulated Local Effects (ALE) plots have been proposed in literature.
+The basic idea for ALE is that instead of using all samples, only the closest samples are used for assessing the effects of one feature.
+ALE plots therefore inherently require a distance measure for finding the closest samples, which, as revealed in the above discussion, in our case is not possible in a straight-forward manner.
+
+\subsection{Plots for Sequences}
+
+As apparent from the above discussion, trying to explain how packet features influence the model's predictions, poses various difficulties.
+As a first step to explaining our model's predictions, we investigate how important features at different time steps are for the model's decisions.
+Intuitively, features at the beginning of a flow should be the most important while the model's predictions should not vary significantly anymore, as soon as the model has come to a decision.
+
+\begin{figure*}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_0_Botnet-ARES"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_14_Web Attack-XSS"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
+
+\caption{Median confidence per time step and first and third quartiles for selected attack types.
+For the majority of attacks, confidence increases in the first few steps and then stays almost constant at 1.}
+\label{fig:confidence}
+\end{figure*}
+
+To determine if this hypothesis is correct, we plot the attack confidence of the classifier for each time step for each attack class.
+At each time step we average over all flows and also show the first and third quartiles.
+\autoref{fig:confidence} shows that this hypothesis is correct for most attack classes.
+At the first couple of packets the confidence for maliciousness is not very high yet but towards the end of the flow it reaches values close to 1 and stays there.
+
+\begin{figure*}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_0_Botnet-ARES"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_1_Brute Force-FTP-Patator"}.pdf}
+
+\caption{The solid line shows the mean value of the feature for all flows of a specific attack type. The shaded region shows the change in confidence that occurs when the feature is varied.}
+\label{fig:pred_plots2}
+\end{figure*}
+
+Next, we want to investigate if our classifier just memorized its inputs.
+The most interesting packet features that can be used to visualize this in our use case are the inter-arrival time (IAT) between packets in a flows and the packet length.
+To depict how the model comes to a decision, we determined the total minimum and maximum for these features that ever occur for all flows.
+For a given flow $X$ we then varied feature values at each time step between minimum and maximum and plot how the confidence changes when the feature varies (\autoref{fig:pred_plots2}) depict the averaged predictions for different attack families.
+It seems clear that the classifier didn't just memorize the inputs but actually learned them since there is always a smooth decrease of confidence if we change the feature values.
+Instead, when memorizing, we would expect a sharp decline of confidence to occur.
+Another interesting observation is that after the first couple of packets the classifier is certain about its decision and changing features doesn't change its confidence any longer.
+This allows the conclusion that if an attacker is able to change the first few packets, she could possibly change the decision of the classifier.
+We will further investigate this question in section \autoref{sec:adv}.
+
+\begin{figure*}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_2_Brute Force-SSH-Patator"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_6_DoS - DDoS-DoS Slowhttptest"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
+
+\caption{The solid line shows the median value of the feature for all flows of a specific attack type. The shaded region are the first and third quartile.}
+\label{fig:plot_features}
+\end{figure*}
+
+Looking at the plots of the features that vary over time (\autoref{fig:plot_features}) we see that certain attack types have a characteristic pattern in which they send packets by which they are easily recognizable.
+We want to determine how the pattern of packet sizes and interarrival times is significant for the classification.
+\autoref{fig:plot_features} clearly shows that some attack types have very characteristic patterns of packet sizes and interarrival times.
+
 \section{Conclusions}

### 2019-12-04 ex3/paper/paper.tex
Moved defences, added some stuff

@@
 
-\section{Defences}
-
-\subsection{Reducing features}
-The most obvious defence, which is only possible because of the nature of the dataset, is to simply leave out features which the attacker can manipulate. For this, we try two different approaches:
-\begin{itemize}
-\item Leaving out all features that are manipulable: packet size and interarrival time
-\item Leaving out manipulable features when the attacker can manipulate them: only in the direction from the attacker to the victim (This, however does not prevent adversarial samples for botnets, for which both sides are malicious)
-\end{itemize}
-
-Both approaches lead to complete resistance to adversarial samples (as we left out all manipulable features), except for botnets, which can still operate when only leaving out manipulable features in one direction.
-The results in \autoref{tab:performance_results_no_manipulable} show that -- surprisingly -- there is only a small difference in classification performance when only computing the accuracy for the last packet of each flow (flow accuracy).
-However, when looking at the packet accuracy, which considers all packets of all flows, there is a significant difference.
-Thus, apparently the interarrival time and packet size are especially important for determining whether a flow is malicious in the first packets of a flow.
-
-\subsection{Adversarial training}
-In addition to omitting manipulable features we tried to make our network more robust against adversarial samples by augmenting the training set by adversarial flows generated using the Carlini-Wagner method, labeled as additional attacks.
-
-For this purpose, we added one adversarial sample per attack sample to the training set.
-Note that, since our used method for obtaining adversarial samples is deterministic, this is the maximum number of adversarial samples possible.
-With this training set we then alternated retraining of the neural network and optimization of the adversarial samples using gradient descent.
-
-One question which occurs in this process, is how many CPU cycles to spend on network training and adversarial sample optimization.
-As a natural approach, we decided to use as many backpropagation steps for neural network training as we use for adversarial sample optimization.
-For each adversarial sample optimization step, we performed 10 iterations of gradient descent.
-Hence, all adversarial samples were optimized each 10 epochs of neural network training.
-
-\autoref{fig:avdistance} shows the average distance of a block of adversarial samples from to original, unmodified attack samples throughout the adversarial training procedure.
-The figure clearly indicates that adversarial training is effective, as the distance gradually increases.
-Hence, an attacker would have to modify attack samples more and more, eventually rendering the attack unpractical.
-
-However, the figure also shows that an increase of distance can only be achieved after a considerable amount of training.
-It is, hence, questionable if the gained increase in robustness is enough for reliable attack detection.
-
-\begin{figure}
-\includegraphics[width=\columnwidth]{../plots/adv_training/avdistance.pdf}
-\caption{Distance of adversarial samples from original attack samples throughout adversarial training.}
-\label{fig:avdistance}
-\end{figure}
-
-%However, against our expectations, this approach did not lead to satisfactory results in the sense that the network would learn to classify adversarial samples as attacks.
-In fact, classification accuracy stayed about constant during the training of the network.
-
-A possible explanation for this observation might be that the adversarial samples assumed too much the shape of normal flows, so that the network could not reliable tell them apart.
-
-\begin{table*}
-\caption{Comparing performance with all features, only safe features (interarrival time and packet size removed) and safe features in one direction (interarrival time and packet size removed for packets coming from source).
-\textit{packets} means that we consider the neural network output of all packets of the flow while \textit{flows} means that we only consider the last packet.
-At the last packet the classifier is usually already a lot more confident than at the first packets of a flow and thus metrics are always better in this case.}  \label{tab:performance_results_no_manipulable}
-\newcommand{\cmidrulespace}{6pt}
-\centering
-\begin{tabular}{l l l l l l l} \toprule
-& \multicolumn{2}{l}{all} & \multicolumn{2}{l}{safe features} & \multicolumn{2}{l}{safe features one direction} \\
-\cmidrule(r){2-3} \cmidrule(lr){4-5} \cmidrule(l){6-7}
-& packets & flows & packets & flows & packets & flows \\
-\midrule
-Accuracy & 0.9857054 & 0.99472587 & 0.96678023 & 0.99337596 & 0.98056073 & 0.99353386 \\
-Precision & 0.94963342 & 0.9894422 & 0.87175275 & 0.98798596 & 0.91359787 & 0.98994595 \\
-Recall & 0.96735895 & 0.98972051 & 0.94374958 & 0.98581328 & 0.97836672 & 0.9844478 \\
-F1 & 0.95841423 & 0.98958134 & 0.90632359 & 0.98689842 & 0.94487366 & 0.98718922 \\
-Youden & 0.95682951 & 0.98614231 & 0.91525628 & 0.98175163 & 0.95937772 & 0.9810602 \\
-\bottomrule
-\end{tabular}
-\end{table*}
-
 \section{Explaining Recurrent Neural Networks}
@@
 
+\subsubsection{Neural Network Weights}
+In previous works, a simple method for determining feature importance in neural networks, has been summing up neural network weights leading from an input to the prediction outcome.
+The weights method can be considered for both feature importance and feature sensitivity, however, clearly, especially in the case of complex network architectures, this method is likely to provide wrong results.
+Hence, we provide results for the weights method mainly for comparison.
+Note that an LSTM cell alone has four weights leading from one input to an output.
+
+\subsubsection{Input Perturbation}
+
 As a first step to understanding the neural network's decisions, we tried to estimate how important individual features are for the model's predictions.
@@
 
-Table~\ref{fig:featRandom} shows our results.
-The results match to a large extent with the expectations one has when classifying network flows.
-In particular, the most important features turn out to be the used protocol, packet length and the SYN TCP flag.
-While the first two features are essential for characterizing flows, a high number of packets with set SYN flag is an important characteristic for port scanning or certain DoS attacks.
-Also the fact that most of TCP flags do not play an important role in our dataset was expected.
-It is interesting to note, however, that the interarrival time of packets does not play a crucial role for the detection of attacks.
-
+\subsubsection{Feature Dropout}
 While the method of adding noise to features/randomizing them is convenient since it is easy to implement and understand, we argue that it has some shortcomings: 
@@
 We thus recommend training two models: One regular one and one with feature dropout to use for the feature importance.
+
+\subsubsection{Mutual Information}
+Input Perturbation has some shortcomings when considering feature sensitivity.
+Assuming that the test dataset is representative for production use, for feature importance it is reasonable to equate the distribution of the perturbed feature with the feature distribution.
+On the other hand, when evaluating feature sensitivity, e.g.
+for analyzing potential for adversarial samples, the attacker is not limited by this distribution and frequently is able to choose arbitrary values in the feature space.
+
+Furthermore, let $f_A(x)$  denote the joint probability for classification as attack with a feature value $x$.
+Accuracy then is $\int f_A(x) dx$ for attacks and $1-\int f_A(x) dx$ for benign traffic.
+Consider \autoref{fig:mutinfo}.
+Evidently, both distributions
+yield the same accuracy, but clearly the lower distribution has a larger influence on the prediction. 
+To capture such dependencies, we propose to use mutual information to determine feature sensitivity.
+Mutual Information between two random variables $X,Y$ is defined as 
+\begin{equation}
+I_{X,Y} = \mathbb E \left\{ \log\left(\frac{f_{X,Y}(x,y)}{f_X(x)f_Y(y)}\right) \right\},
+\end{equation}
+with $f_X(x), f_Y(y)$ and $f_{X,Y}(x,y)$ denoting the distribution of $X,Y$ and their joint distribution, respectively.
+To obtain feature sensitivity, we compute mutual information between an input variable and the prediction output for one flow at one particular time step, and average over the result for the test dataset.
+
+\subsubsection{Results}
+Table~\ref{fig:featRandom} shows our results.
+The results match to a large extent with the expectations one has when classifying network flows.
+In particular, the most important features turn out to be the used protocol, packet length and the SYN TCP flag.
+While the first two features are essential for characterizing flows, a high number of packets with set SYN flag is an important characteristic for port scanning or certain DoS attacks.
+Also the fact that most of TCP flags do not play an important role in our dataset was expected.
+It is interesting to note, however, that the interarrival time of packets does not play a crucial role for the detection of attacks.
 
@@
 
+\section{Defences}
+
+\subsection{Reducing features}
+The most obvious defence, which is only possible because of the nature of the dataset, is to simply leave out features which the attacker can manipulate. For this, we try two different approaches:
+\begin{itemize}
+\item Leaving out all features that are manipulable: packet size and interarrival time
+\item Leaving out manipulable features when the attacker can manipulate them: only in the direction from the attacker to the victim (This, however does not prevent adversarial samples for botnets, for which both sides are malicious)
+\end{itemize}
+
+Both approaches lead to complete resistance to adversarial samples (as we left out all manipulable features), except for botnets, which can still operate when only leaving out manipulable features in one direction.
+The results in \autoref{tab:performance_results_no_manipulable} show that -- surprisingly -- there is only a small difference in classification performance when only computing the accuracy for the last packet of each flow (flow accuracy).
+However, when looking at the packet accuracy, which considers all packets of all flows, there is a significant difference.
+Thus, apparently the interarrival time and packet size are especially important for determining whether a flow is malicious in the first packets of a flow.
+
+\subsection{Adversarial training}
+In addition to omitting manipulable features we tried to make our network more robust against adversarial samples by augmenting the training set by adversarial flows generated using the Carlini-Wagner method, labeled as additional attacks.
+
+For this purpose, we added one adversarial sample per attack sample to the training set.
+Note that, since our used method for obtaining adversarial samples is deterministic, this is the maximum number of adversarial samples possible.
+With this training set we then alternated retraining of the neural network and optimization of the adversarial samples using gradient descent.
+
+One question which occurs in this process, is how many CPU cycles to spend on network training and adversarial sample optimization.
+As a natural approach, we decided to use as many backpropagation steps for neural network training as we use for adversarial sample optimization.
+For each adversarial sample optimization step, we performed 10 iterations of gradient descent.
+Hence, all adversarial samples were optimized each 10 epochs of neural network training.
+
+\autoref{fig:avdistance} shows the average distance of a block of adversarial samples from to original, unmodified attack samples throughout the adversarial training procedure.
+The figure clearly indicates that adversarial training is effective, as the distance gradually increases.
+Hence, an attacker would have to modify attack samples more and more, eventually rendering the attack unpractical.
+
+However, the figure also shows that an increase of distance can only be achieved after a considerable amount of training.
+It is, hence, questionable if the gained increase in robustness is enough for reliable attack detection.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{../plots/adv_training/avdistance.pdf}
+\caption{Distance of adversarial samples from original attack samples throughout adversarial training.}
+\label{fig:avdistance}
+\end{figure}
+
+%However, against our expectations, this approach did not lead to satisfactory results in the sense that the network would learn to classify adversarial samples as attacks.
+In fact, classification accuracy stayed about constant during the training of the network.
+
+A possible explanation for this observation might be that the adversarial samples assumed too much the shape of normal flows, so that the network could not reliable tell them apart.
+
+\begin{table*}
+\caption{Comparing performance with all features, only safe features (interarrival time and packet size removed) and safe features in one direction (interarrival time and packet size removed for packets coming from source).
+\textit{packets} means that we consider the neural network output of all packets of the flow while \textit{flows} means that we only consider the last packet.
+At the last packet the classifier is usually already a lot more confident than at the first packets of a flow and thus metrics are always better in this case.}  \label{tab:performance_results_no_manipulable}
+\newcommand{\cmidrulespace}{6pt}
+\centering
+\begin{tabular}{l l l l l l l} \toprule
+& \multicolumn{2}{l}{all} & \multicolumn{2}{l}{safe features} & \multicolumn{2}{l}{safe features one direction} \\
+\cmidrule(r){2-3} \cmidrule(lr){4-5} \cmidrule(l){6-7}
+& packets & flows & packets & flows & packets & flows \\
+\midrule
+Accuracy & 0.9857054 & 0.99472587 & 0.96678023 & 0.99337596 & 0.98056073 & 0.99353386 \\
+Precision & 0.94963342 & 0.9894422 & 0.87175275 & 0.98798596 & 0.91359787 & 0.98994595 \\
+Recall & 0.96735895 & 0.98972051 & 0.94374958 & 0.98581328 & 0.97836672 & 0.9844478 \\
+F1 & 0.95841423 & 0.98958134 & 0.90632359 & 0.98689842 & 0.94487366 & 0.98718922 \\
+Youden & 0.95682951 & 0.98614231 & 0.91525628 & 0.98175163 & 0.95937772 & 0.9810602 \\
+\bottomrule
+\end{tabular}
+\end{table*}
+
 \section{Conclusions}

### 2019-12-04 ex3/learn.py
Changed adv_until_less_than_half

@@ -37,6 +37,16 @@ def output_scores(y_true, y_pred):
 def numpy_sigmoid(x):
 	return 1/(1+np.exp(-x))
 
+def pretty_print(*args, **kwargs):
+	# Print tables in a format which can be parsed easily
+	outputs = []
+	for arg in args:
+		try:
+			outputs.append('%.6f' % arg)
+		except:
+			outputs.append(arg)
+	print('\t'.join(outputs), **kwargs)
+
 class OurDataset(Dataset):
 	def __init__(self, data, labels, categories):
 		self.data = data
@@ -192,7 +202,7 @@ def train():
 	train_data = torch.utils.data.Subset(dataset, train_indices)
 	if opt.advTraining:
 		train_data = AdvDataset(train_data)
-		adv_generator = adv_internal(in_training=True)
+		adv_generator = adv_internal(in_training=True, iterations=10)
 	train_loader = torch.utils.data.DataLoader(train_data, batch_size=opt.batchSize, shuffle=True, collate_fn=custom_collate)
 
 	optimizer = optim.SGD(lstm_module.parameters(), lr=opt.lr)
@@ -201,6 +211,8 @@ def train():
 	writer = SummaryWriter()
 
 	samples = 0
+	save_period = int(round(len(train_indices) * opt.modelSavePeriod))
+	
 	for i in range(1, sys.maxsize):
 		if opt.advTraining:
 			train_data.adv_flows, train_data.categories, av_distance = next(adv_generator)
@@ -274,13 +286,15 @@ def train():
 				writer.add_scalar("adv_confidence", torch.mean(confidences[mask]), samples)
 				writer.add_scalar("adv_end_confidence", torch.mean(confidences[mask_exact]), samples)
 
-
-		# Save after every epoch
-		if i % 1 == 0:
-			torch.save(lstm_module.state_dict(), '%s/lstm_module_%d.pth' % (writer.log_dir, i))
-			if opt.advTraining:
-				with open(adv_filename(), 'wb') as f:
-					pickle.dump(train_data.adv_flows, f)
+			if samples % save_period < output.shape[1]:
+				if len(train_indices) % save_period == 0:
+					filename = 'lstm_module_%d.pth' % i
+				else:
+					filename = 'lstm_module_%.3f.pth' % (samples / len(train_indices))
+				torch.save(lstm_module.state_dict(), '%s/%s' % (writer.log_dir, filename))
+				if opt.advTraining:
+					with open(adv_filename(), 'wb') as f:
[…]

### 2019-12-04 ex3/learn.py
Readded accidentally removed code

@@ -1162,7 +1162,7 @@ def adv_until_less_than_half():
 		results_dict = list(adv_internal(False, tradeoff, lr, iterations, next_filter))[0]
 		modified_flows_by_attack, modified_results_by_attack, original_flows_by_attack, original_results_by_attack = results_dict["modified_flows_by_attack_number"], results_dict["results_by_attack_number"], results_dict["orig_flows_by_attack_number"], results_dict["orig_results_by_attack_number"]
 		ratio_modified_by_attack_number = np.array([np.mean(np.round(numpy_sigmoid(np.array([item[-1] for item in modified_results])))) if len(modified_results)>0 else -float("inf") for modified_results in modified_results_by_attack])
-		next_filter = [index for index, item in enumerate(ratio_modified_by_attack_number) if item <= THRESHOLD ]#or prev_ratios[-1][item] <= THRESHOLD]
+		next_filter = [index for index, item in enumerate(ratio_modified_by_attack_number) if item <= THRESHOLD or prev_ratios[-1][item] <= THRESHOLD]
 		if i==0:
 			orig_results = original_results_by_attack
 			orig_flows = original_flows_by_attack

### 2019-12-04 ex3/paper/paper.tex
Modified paper

@@
 
-The dataset we use is the \textit{CIC-IDS 2017} dataset \cite{sharafaldin_toward_2018}, which includes the attacks listed in \autoref{tab:occurence} and contains over two million flows in total.
-We implement an LSTM-based classifier with three layers of 512 neurons each.
-As the input features we use \begin{itemize*}
+In this research, we use the \textit{CIC-IDS-2017} \cite{sharafaldin_toward_2018} and \textit{UNSW-NB15} \cite{moustafa_unsw-nb15:_2015} datasets which each include more than 2 million flows of network data, containing both benign traffic and a large number of different attacks.
+Attacks contained in CIC-IDS-2017 are shown in \autoref{tab:occurence}.
+
+We implement a three-layer LSTM-based classifier of 512 neurons each. As the input features we use \begin{itemize*}
 \item source port
@@
 
-\autoref{tab:performance_results} shows that our classifier achieves an accuracy that is similar to the one of the random forest and the multilayer perceptron that we developed in the previous exercise.
+\autoref{tab:performance_results} shows that our classifier achieves an accuracy that is similar previous work based on these datasets~\cite{meghdouri_analysis_2018}.
 However unlike these classifiers, our recurrent classifier has the advantage that it can detect attacks already before they are over.
@@
 \begin{table}
-\caption{Detection performance results. Random forest and Multilayer Perceptron are from the previous exercise, LSTM is this implementation.} \label{tab:performance_results}
+% Evaluated using
+% ./python learn.py --dataroot flows.pickle --net runs/Oct26_00-03-50_gpu/lstm_module_1284.pth --function test --batchSize 512
+% ./python learn.py --dataroot flows15.pickle --net runs/Oct28_15-41-46_gpu/lstm_module_997.pth --function test --batchSize 512
+% on Dec. 4th
+\caption{Detection performance on a packet and flow level.} \label{tab:performance_results}
 \centering
-\begin{tabular}{l r r r} \toprule
-& RF & MLP & LSTM \\ \midrule
-Accuracy	&	0.9974 & 0.9970	& 0.9965 \\
-Precision	&	0.9965 & 0.9979 & 0.9956	\\
-Recall	&	0.9931 & 0.9903	& 0.9905 \\
-F1	&	0.9948 & 0.9941	& 0.9930 \\
-Youden	&	0.9920 & 0.9896 & 0.9890 \\
+\begin{tabular}{l r r r r} \toprule
+& \multicolumn{2}{c}{CIC-IDS-17} & \multicolumn{2}{c}{UNSW-NB-15} \\
+	&	Packet	&	Flow	&	Packet	&	Flow	\\ \midrule
+Accuracy	&	0.991	&	0.997	&		&		\\
+Precision	&	0.970	&	0.997	&		&		\\
+Recall	&	0.978	&	0.991	&		&		\\
+F1	&	0.974	&	0.994	&		&		\\
+Youden	&	0.972	&	0.990	&		&		\\
+
 \bottomrule
@@
 \subsubsection{Mutual Information}
+\begin{figure}
+\includegraphics[width=\columnwidth]{../plots/mutinfo_example.pdf}
+\caption{Distributions $f_A(x)$ yielding an identical accuracy drop.}
+\label{fig:mutinfo_example}
+\end{figure}
 Input Perturbation has some shortcomings when considering feature sensitivity.
@@
 Accuracy then is $\int f_A(x) dx$ for attacks and $1-\int f_A(x) dx$ for benign traffic.
-Consider \autoref{fig:mutinfo}.
+Consider \autoref{fig:mutinfo_example}.
 Evidently, both distributions
-yield the same accuracy, but clearly the lower distribution has a larger influence on the prediction. 
-To capture such dependencies, we propose to use mutual information to determine feature sensitivity.
+yield the same accuracy, but clearly the right-side distribution has a larger influence on the prediction, as in the right-side case the prediction deterministically depends on the feature value.
+To capture such dependencies, we propose to use mutual information instead of accuracy drop to determine feature sensitivity.
 Mutual Information between two random variables $X,Y$ is defined as 

### 2019-12-04 ex3/paper/paper.tex
Add 2015 accuracy

@@
 	&	Packet	&	Flow	&	Packet	&	Flow	\\	\midrule
-Accuracy	&	0.991	&	0.997	&		&		\\
-Precision	&	0.970	&	0.997	&		&		\\
-Recall	&	0.978	&	0.991	&		&		\\
-F1	&	0.974	&	0.994	&		&		\\
-Youden	&	0.972	&	0.990	&		&		\\
+Accuracy	&	0.991	&	0.997	&	0.995	&	0.983	\\	
+Precision	&	0.970	&	0.997	&	0.834	&	0.786	\\	
+Recall	&	0.978	&	0.991	&	0.876	&	0.726	\\	
+F1	&	0.974	&	0.994	&	0.855	&	0.755	\\	
+Youden	&	0.972	&	0.990	&	0.873	&	0.719	\\
 

### 2019-12-05 ex3/paper/paper.tex
Add stuff for feature correlation

@@
 The new performance score \gls{ars} precisely captures the notion of how easily an adversary can generate adversarial samples.
-We develop a simple methods for defending against adversaries who can manipulate arbitrary packets by simply leaving out potentially manipulable information from the classification process.
-The resulting classifier is slightly less accurate but is still useful to be deployed in a real scenario.
-We develop a custom adversarial training procedure which especially focuses on easy of use and low resource consumption and aims to keep the number of additionally introduced training hyperparameters low.
-With this novel strategy we can reduce attack surface and harden the resulting \gls{ids} while keeping similar classification performance like for the original \gls{ml} model.
 
@@
 
-\section{Related work}
-
-\cite{szegedy_intriguing_2014} show that it is possible to create adversarial samples for an image, which look similar to the original image but are classified wrongly.
-\cite{goodfellow_explaining_2015} develop the \gls{fgsm}, which makes easy generation of adversarial samples possible.
-\cite{papernot_crafting_2016} explore the use of \gls{aml} for \glspl{rnn}
+Based on the insight gained by the feature importance and explainability methods, we develop a simple methods for defending against adversaries who can manipulate arbitrary packets by simply leaving out potentially manipulable information from the classification process.
+The resulting classifier is slightly less accurate but is still useful to be deployed in a real scenario.
+We develop a custom adversarial training procedure which especially focuses on easy of use and low resource consumption and aims to keep the number of additionally introduced training hyperparameters low.
+With this novel strategy we can reduce attack surface and harden the resulting \gls{ids} while keeping similar classification performance like for the original \gls{ml} model.
 
@@
 
-Several methods for generating adversarial samples have been proposed in the recent literature, achieving different speed-quality tradeoffs.
-Since the reasons discussed above suggest that finding adversarial flows is particularly difficult in our case, for the present research we preferred methods which produces high-quality adversarial samples while not being as fast as other methods.
+Several methods for generating adversarial samples have been proposed in the recent literature, achieving different speed-quality tradeoffs: 
+\cite{szegedy_intriguing_2014} show that it is possible to create adversarial samples for an image, which look similar to the original image but are classified wrongly.
+\cite{goodfellow_explaining_2015} develop the \gls{fgsm}, which makes easy generation of adversarial samples possible.
+\cite{papernot_crafting_2016} explore the use of \gls{aml} for \glspl{rnn}
 
-We implemented the following methods for generating adversarial flows.
+As finding adversarial flows is particularly difficult in our case for the reasons we discussed above, for the present research we preferred methods which produces high-quality adversarial samples while not being as fast as other methods.
+We implemented the several methods for generating adversarial flows.
 
@@
 With feature dropout, it becomes apparent that there are very few features that actually contain unique information and that affect accuracy when left out: the destination port and the source port.
-\todo{Add correlation and explain why most features have no importance}
-
 A model trained with feature dropout typically yields slightly lower accuracy than a regularly trained model, even if no features are left out (flow accuracy of 0.9943 vs.~0.9965.
 We thus recommend training two models: One regular one and one with feature dropout to use for the feature importance.
+
+Feature dropout also allows us to find correlated features or -- more general -- features that contain common information:
+
+% numerator = max(float(baseline_accuracy-accuracy_for_feature_combination), 0)
+% 			denominator = max(float(baseline_accuracy-accuracy_for_feature_1), 0) + max(float(baseline_accuracy-accuracy_for_feature_2), 0)
+\begin{equation}
+\text{score} = \frac{\text{acc}_\text{base}-\text{acc}_\text{-both\_features}}{\left(\text{acc}_\text{base}-\text{acc}_\text{-feature\_1}\right) + \left(\text{acc}_\text{base}-\text{acc}_\text{-feature\_2}\right)}
+\end{equation}
+With $\text{acc}_\text{base}$ we denote the average accuracy of the classifier with all features included, with $\text{acc}_\text{-feature\_i}$ the average accuracy if feature $i$ is omitted and with $\text{acc}_\text{-both\_features}$ the average accuracy if both features are omitted.
+The higher the resulting number, the larger the information that both features share.
 

### 2019-12-05 ex3/learn.py

@@ -212,7 +212,7 @@ def train():
 
 	samples = 0
 	save_period = int(round(len(train_indices) * opt.modelSavePeriod))
-	
+
 	for i in range(1, sys.maxsize):
 		if opt.advTraining:
 			train_data.adv_flows, train_data.categories, av_distance = next(adv_generator)
@@ -469,14 +469,14 @@ def dropout_feature_correlation():
 				else:
 					overall_score = float("nan")
 			correlations[(feat1, feat2)] = overall_score
-			pretty_print("feat1", feat1, feature_array[feat1], feat2, "feat2", feature_array[feat2], "acc_feat1", baseline_accuracy-accuracy_by_feature[feat1], "acc_feat2", baseline_accuracy-accuracy_by_feature[feat2], "joint_acc", joint_accuracy[(feat1, feat2)], "score", overall_score)
+			pretty_print("feat1", feat1, feature_array[feat1], feat2, "feat2", feature_array[feat2], "acc_feat1", baseline_accuracy-accuracy_by_feature[feat1], "acc_feat2", baseline_accuracy-accuracy_by_feature[feat2], "joint_acc", baseline_accuracy-joint_accuracy[(feat1, feat2)], "score", overall_score)
 
 	N_MOST_CORRELATED_TO_SHOW = len(list(correlations.items()))
 	sorted_correlations = list(sorted([item for item in list(correlations.items()) if not math.isnan(item[1])], key=lambda item: item[1], reverse=True))[:N_MOST_CORRELATED_TO_SHOW]
 	print("highest", N_MOST_CORRELATED_TO_SHOW, "scores:")
 	for (feat1, feat2), score in sorted_correlations:
-		pretty_print("feat1", feat1, feature_array[feat1], feat2, "feat2", feature_array[feat2], "acc_feat1", baseline_accuracy-accuracy_by_feature[feat1], "acc_feat2", baseline_accuracy-accuracy_by_feature[feat2], "joint_acc", joint_accuracy[(feat1, feat2)], "score", score)
-	
+		pretty_print("feat1", feat1, feature_array[feat1], feat2, "feat2", feature_array[feat2], "acc_feat1", baseline_accuracy-accuracy_by_feature[feat1], "acc_feat2", baseline_accuracy-accuracy_by_feature[feat2], "joint_acc", baseline_accuracy-joint_accuracy[(feat1, feat2)], "score", score)
+
[…]

### 2019-12-05 ex3/learn.py
Tiny change in mutinfo

@@ -762,7 +762,7 @@ def mutinfo_feat_imp():
 	fold = opt.fold
 	lstm_module.eval()
 
-	with open("features.json", "r") as f:
+	with open("features_meaningful_names.json", "r") as f:
 		feature_array = json.load(f)
 
 	_, test_indices = get_nth_split(dataset, n_fold, fold)

### 2019-12-05 ex3/weight_feat_imp.py
Modified weight feat imp

@@ -5,6 +5,7 @@ import torch
 import numpy as np
 import sys
 import json
+from learn import pretty_print
 
 hidden_size = 512
 n_layers = 3
@@ -14,8 +15,8 @@ assert len(sys.argv) > 1, 'No model specified'
 
 parameters = torch.load(sys.argv[1])
 
-with open('features.json', 'r') as f:
-	features = json.load(f)
+with open('features_meaningful_names.json', 'r') as f:
+	feature_array = json.load(f)
 
 #	parameters = lstm_module.state_dict()
 weights_hh = [ parameters['lstm.weight_hh_l%d' % layer].cpu().numpy().reshape(4,hidden_size,hidden_size).sum(axis=0) for layer in range(n_layers) ]
@@ -45,8 +46,11 @@ for _ in range(10):
 	hh_prod = [ np.matmul(hh_prod[i], weights_hh[i]) for i in range(n_layers) ]
 	fi_over_time.append(np.abs(np.sum([ np.matmul(output_featimp[i+1], np.matmul(hh_prod[i], input_featimp[i])) for i in range(n_layers)], axis=0)[0,:]))
 
-print ('Packet feature importance:', sorted(zip(features, fi_over_time[0]/np.sum(fi_over_time[0])), key=lambda item: item[1]))
+print ('Packet feature importance:')
 
+for feature_index, feature_importance in sorted(enumerate(fi_over_time[0]/np.sum(fi_over_time[0])), key=lambda item: item[1]):
+	pretty_print("weight feat. imp. for", feature_index, feature_array[feature_index], feature_importance)
+	
 # Highly experimental stuff
 decomposition = [ np.linalg.eig(weights_hh[i]) for i in range(n_layers) ]
 valid_eigenvalues = [ [ j for j in range(hidden_size) if np.imag(decomposition[i][0][j]) == 0 ] for i in range(n_layers) ]
@@ -57,13 +61,12 @@ layer_with_max_eigenvalue = max(range(n_layers), key=lambda i: decomposition[i][
 eigenvector = np.real(decomposition[layer_with_max_eigenvalue][1][max_eigenvalues[layer_with_max_eigenvalue]])
 flow_feat_imp = np.abs(np.matmul(output_featimp[layer_with_max_eigenvalue+1], np.matmul(np.matmul(eigenvector[:,None], eigenvector[None,:]), input_featimp[layer_with_max_eigenvalue]))).flatten()
[…]

### 2019-12-05 ex3/paper/paper.tex
Few paper changes

@@
 
-Then we investigate the question whether adversarial samples can be found for this type of data.
+With the recent rise of interest in \gls{aml} techniques, also \gls{aml} for \glspl{ids} have been investigated. E.g. \cite{Bachl2019} investigates remedies for poisoning attacks on \glspl{ids}.
+In this research, we investigate whether adversarial samples can be found for \gls{ml} based on \glspl{rnn}.
 This is not a trivially answerable question since adversarial samples have mostly been analyzed in the context of computer vision.
@@
 \item the number of features is significantly smaller and
-\item only certain features can be manipulated (for example the destination port cannot be manipulated)
+\item only certain features can be manipulated by an adversary (for example the destination port cannot be manipulated)
 \end{enumerate*}.
 
-Surprisingly, we can confirm that adversarial samples can be found even when considering the tight constraints mentioned before.
-Thus we argue that traditional methods like accuracy are not sufficient for evaluating security-sensitive \gls{ml} systems.
-The new performance score \gls{ars} precisely captures the notion of how easily an adversary can generate adversarial samples.
+Surprisingly, we can confirm that adversarial samples can be found even when considering thse tight constraints. % mentioned before.
+Thus, we argue that traditional methods like accuracy are not sufficient for evaluating security-sensitive \gls{ml} systems.
+We propose the \acrfull{ars} as a new performance score, which precisely captures the notion of how easily an adversary can generate adversarial samples.
 
-Astonishingly, feature importance method reveal that the manipulable features, that are left for adversaries to modify are not even particularly significant for the classification outcome.
+Astonishingly, feature importance methods reveal that the manipulable features, that are left for adversaries to modify are not even particularly significant for the classification outcome.
 %Especially the first packets of a flow are critical for determining the decision of a recurrent \gls{ids}.
@@
 
-\autoref{tab:performance_results} shows that our classifier achieves an accuracy that is similar previous work based on these datasets~\cite{meghdouri_analysis_2018}.
+\autoref{tab:performance_results} shows that our classifier achieves an accuracy that is similar previous work based on these datasets~\cite{meghdouri_analysis_2018,Bachl2019}.
 However unlike these classifiers, our recurrent classifier has the advantage that it can detect attacks already before they are over.

### 2019-12-05 ex3/paper/paper.tex
Several small paper changes

@@
 More recently, with the rise of omnipresent encryption, available features have changed.
-Since payload is not available anymore, the focus now lies on features that are always available to observers such as port numbers, flags of packets etc.
-Network communication is usually grouped into \textit{flows}, which is commonly defined as a sequence of packets between two applications that may be located at different physical locations.
+Since payload is not available anymore, the focus now lies on features that are always available to observers such as port numbers, protocol flags etc.
+Network communication is usually grouped into \textit{flows}, which are commonly defined as a sequence of packets sharing certain properties, between two applications that may be located at different physical locations.
 
@@
 Various approaches have been proposed to extract features from flows and then perform anomaly detection with the results \cite{meghdouri_analysis_2018}.
-While these approaches works well, it is problematic that first the whole flow has to be received and only afterwards it can be determined if there was an anomaly.
-Thus, we design a network anomaly detection system that operates on a per-packet basis and decides if a packet is anomalous based on features that are available for traffic that is encrypted above the transport layer, like for example TLS or QUIC.
+While these approaches work well, it is problematic that first the whole flow has to be received and only afterwards it can be determined if there was an anomaly.
+Thus, we design a network anomaly detection system that operates on a per-packet basis and decides if a packet is anomalous based on features that are available even for traffic that is encrypted above the transport layer, like for example TLS or QUIC.
 We show that our system has similar performance to other flow-based anomaly detection systems but can commonly detect anomalies \textit{before} the flow is over.
 
-With the recent rise of interest in \gls{aml} techniques, also \gls{aml} for \glspl{ids} have been investigated. E.g. \cite{Bachl2019} investigates remedies for poisoning attacks on \glspl{ids}.
+With the recent rise of interest in \gls{aml} techniques, also \gls{aml} for \glspl{ids} has been investigated.
+For example, \cite{Bachl2019} investigates remedies for poisoning attacks on \glspl{ids}.
 In this research, we investigate whether adversarial samples can be found for \gls{ml} based on \glspl{rnn}.
@@
 
-Surprisingly, we can confirm that adversarial samples can be found even when considering thse tight constraints. % mentioned before.
+Surprisingly, we can confirm that adversarial samples can be found even when considering these tight constraints. % mentioned before.
 Thus, we argue that traditional methods like accuracy are not sufficient for evaluating security-sensitive \gls{ml} systems.
@@
 
-Based on the insight gained by the feature importance and explainability methods, we develop a simple methods for defending against adversaries who can manipulate arbitrary packets by simply leaving out potentially manipulable information from the classification process.
+Based on the insights gained by the feature importance and explainability methods, we develop a simple method for defending against adversaries who can manipulate arbitrary packets by simply leaving out potentially manipulable information from the classification process.
 The resulting classifier is slightly less accurate but is still useful to be deployed in a real scenario.
-We develop a custom adversarial training procedure which especially focuses on easy of use and low resource consumption and aims to keep the number of additionally introduced training hyperparameters low.
-With this novel strategy we can reduce attack surface and harden the resulting \gls{ids} while keeping similar classification performance like for the original \gls{ml} model.
+Furthermore, we develop a custom adversarial training procedure which especially focuses on ease of use and low resource consumption and aims to keep the number of additionally introduced training hyperparameters low.
+With this novel strategy we can reduce attack surface and harden the resulting \gls{ids} while retaining similar classification performance like for the original \gls{ml} model.
 
@@
 
-We implement a three-layer LSTM-based classifier of 512 neurons each. As the input features we use \begin{itemize*}
+We implement a three-layer LSTM-based classifier with 512 neurons at each layer. As the input features we use \begin{itemize*}
 \item source port
@@
 \end{itemize*}. Of these features, source port, destination port and protocol identifier are constant over the whole flow while the others vary.
-
 We use Z-score normalization and a train/test split of 2:1.
 
-\autoref{tab:performance_results} shows that our classifier achieves an accuracy that is similar previous work based on these datasets~\cite{meghdouri_analysis_2018,Bachl2019}.
-However unlike these classifiers, our recurrent classifier has the advantage that it can detect attacks already before they are over.
+\autoref{tab:performance_results} shows that our classifier achieves an accuracy that is similar to previous work based on these datasets~\cite{meghdouri_analysis_2018,Bachl2019}.
+However, unlike these classifiers, our recurrent classifier has the advantage that it can detect attacks already before they are over.
 
@@
 
-Since most features such as the TCP flags cannot be easily be manipulated because it might violate the TCP and thus cause communication to fails, we experiment whether by manipulating the interarrival times and packet lengths of the first packets, the classifier might be fooled.
-However, our scenario is quite different from how generating adversarial samples usually works: We have significantly less features than in, for example, an image.
+Most features such as, e.g., TCP flags cannot be easily manipulated because it might violate the TCP and thus cause communication to fail.
+Hence, we experiment whether by manipulating the interarrival times and packet lengths of the first packets, the classifier might be fooled.
+Our scenario is quite different from how generating adversarial samples usually works: We have significantly less features than in, for example, an image.
 Furthermore, only two out of those few features can be manipulated.
-And even these features cannot manipulated freely: Besides the constraints that packet size and interarrival time cannot become negative there are also more problem-specific constraints:
+And even these features cannot be manipulated freely: Besides the constraints that packet size and interarrival time cannot become negative there are also more problem-specific constraints:
 \begin{itemize}
@@
 \subsection{Carlini-Wagner}
-In particular, we implemented \gls{cw} \cite{carlini_towards_2017}, which performs gradient descent on the optimization objective
+We implemented the \gls{cw} \cite{carlini_towards_2017}, which performs gradient descent on the optimization objective
 \begin{equation} \label{eq:carliniWagner}
@@
 \end{equation}
-Here, $d(\cdot)$ is a distance metric, $\epsilon \in \mathbb R^+$ is a parameter governing the tradeoff achieved between attack success and distance from the original flow.
+Here, $d(\cdot)$ is a distance metric and $\epsilon \in \mathbb R^+$ is a parameter governing the tradeoff achieved between attack success and distance from the original flow.
 Furthermore, $Z(\cdot)$ denotes the neural network's logit output, $X$ denotes the original flow and $\tilde X$ the adversarial flow under optimization.
@@
 \end{equation}
-i.e.
-we require a slight level of confidence that our adversarial samples are not classified as attack traffic.
-A logit value of $-0.2$ corresponds to a confidence of 45\% for the sample to be an attack.
+i.e. we require a slight level of confidence that our adversarial samples are not classified as attack traffic. A logit value of $-0.2$ corresponds to a confidence of 55\% for the sample to be benign.
 
 \subsection{$\ell_\infty$-bounded Projected Gradient Descent}
-\cite{madry_towards_2018} use a method for generating adversarial samples which uses \gls{pgd}  to minimize the network's negative loss function while constraining the achieved $\ell_\infty$ distance from the original samples.
+\cite{madry_towards_2018} uses a method for generating adversarial samples which deploys \gls{pgd}  to minimize the network's negative loss function while constraining the achieved $\ell_\infty$ distance from the original samples.
 
 \subsection{Fast Gradient Sign Method}
-Finally, we also tested the Fast Gradient Sign Method (FGSM), which was initially proposed in \cite{goodfellow_explaining_2015} and perhaps is the most frequently used method for generating adversarial sampling.
+Finally, we also tested the Fast Gradient Sign Method (FGSM), which was initially proposed in \cite{goodfellow_explaining_2015} and perhaps is the most frequently used method for generating adversarial samples.
 FGSM can be considered a single pass of \gls{pgd} on the loss function with an equality constraint on the $\ell_\infty$ distance, i.e.
@@
 \end{figure*}
-\autoref{fig:adv_cw} depicts the ratio of attack flows which were successfully misclassified as normal traffic with respect to the final packet along with the resulting distance from the original sample.
+\autoref{fig:adv_cw} depicts the ratio of attack flows which were successfully misclassified as benign with respect to the final packet along with the resulting distance from the original sample.
 Evidently, the larger the $\epsilon$ parameter, the better the attack works, but also the distances become larger.
@@
 For this table we used a $\ell_\infty$ bound of 0.3 for FGSM and $\ell_\infty$-bounded \gls{pgd} and a tradeoff of 0.5 for Carlini-Wagner.
-The table shows significant for detecting different attack families in the first place.
+The table shows significant differences for detecting different attack families in the first place.
 However, also for generating adversarial flows some attack families are more susceptible than others.
@@
 \section{Explaining Recurrent Neural Networks}
-From a naive perspective, one might be tempted to reuse explainability methods for recurrent neural networks by considering a flow the sum of its packet features.
-We identify several problems which occur when trying to explain decisions made by recurrent neural networks.
+From a naive perspective, one might be tempted to reuse explainability methods for \glspl{rnn} by considering a flow the sum of its packet features.
+We identify several difficulties which occur when trying to explain decisions made by \glspl{rnn}.
 
@@
 \textbf{Feature quantity.}
-The number of features that is fed into a recurrent neural network is found as the product of packet features with the total length of the flow.
+The number of features that is fed into an \gls{rnn} is found as the product of packet features with the total length of the flow.
 For long flows, the total number of inputs can therefore grow to a substantial size.
@@
 The length of different flows might differ tremendously.
-Hence, features which are important for the network's outcome for one sequence might not even exist for other flows.
+Hence, features which are important for the network's outcome for one flow might not even exist for other flows.
 Furthermore, the question arises how the distance between flows of different length can be computed.
@@
 \textbf{Lack of a distance measure.}
-However, even if we restricted our analysis to flows of a constant length, a flow is different from the plain concatenation of its packet features.
+However, even if we restricted the analysis to flows of a constant length, a flow is different from the plain concatenation of its packet features.
 If we, for example, consider a classification problem as in this document, we expect that after a certain amount of time steps the classifier is able to perform the classification with a certain confidence.
 Arguably, features at the beginning of a flow are thus more important for the classifier.
-Giving differences at the end of a sequence the same weight as differences at its beginning, therefore seems incorrect.
+Giving feature differences at the end of a flow the same weight as differences at its beginning, therefore seems incorrect.
 
@@
 \textbf{Multiple prediction outputs.}
-Often a recurrent neural network produces an output at each time step. When trying to apply explainability methods, hence, the question arises which prediction to base the methods on.
+Often an \gls{rnn} produces an output at each time step. When trying to apply explainability methods, hence, the question arises which output to base the methods on.
 
 The natural choice is to base the methods on the prediction output which occurs at the same time step as the feature under investigation.
-Not only does this approach require a substantially lower amount of computational resources compared to predictions occurring at a later time step, but we also expect the immediately occurring prediction outcome to exhibit a stronger correlation to the feature under investigation than a prediction outcome which occurring many time steps later.
+Not only does this approach require a substantially lower amount of computational resources compared to predictions occurring at a later time step, but we also expect the immediately occurring prediction outcome to exhibit a stronger correlation to the feature under investigation, compared to a prediction outcome occurring many time steps later.
 However, due to the complex decision processes that can occur for deep neural networks, it is absolutely  uncertain if the network's final prediction is correctly depicted when following this approach.
@@
 
+% where does this come from :)
 %\subfloat[Throughput of the fq \gls{qdisc}\label{fig:fqThroughput}
@@
 \end{table}
+As a first step to understanding the neural network's decisions, we tried to estimate how important individual features are for the model's predictions.
 When investigating an \gls{ml} classifier, it is natural to ask which inputs have a large influence on the classifier's prediction.
@@
 
-Both adversarial machine learning and explainable machine learning bring up the question to what extent individual features are able to change the prediction outcome.
+On the other hand, both adversarial machine learning and explainable machine learning bring up the question to what extent individual features are able to change the prediction outcome.
 While being seemingly similar at first sight, traditional feature importance can yield markedly wrong results for this case.
@@
 
+To analyze features for our network, we use the following approaches.
+
 \subsubsection{Neural Network Weights}
-In previous works, a simple method for determining feature importance in neural networks, has been summing up neural network weights leading from an input to the prediction outcome.
+In previous works, a simple method for determining feature importance in neural networks, has been summing up neural network weights leading from a dertain input to the prediction outcome.
 The weights method can be considered for both feature importance and feature sensitivity, however, clearly, especially in the case of complex network architectures, this method is likely to provide wrong results.
@@
 
-As a first step to understanding the neural network's decisions, we tried to estimate how important individual features are for the model's predictions.
-The most commonly used techniques for this which are used by practitioners for \glspl{rnn} \cite{stackexchange_cross_validated_neural_2019} and \gls{dl} \cite{stackexchange_cross_validated_feature_2016} in general are based on adding noise to a particular feature and then evaluating the drop in accuracy that occurs.
+ The most commonly used techniques for feature importance which are used by practitioners for \glspl{rnn} \cite{stackexchange_cross_validated_neural_2019} and \gls{dl} \cite{stackexchange_cross_validated_feature_2016} in general are based on adding noise to a particular feature and then evaluating the drop in accuracy that occurs.
 We argue that it is hard to find out the ``correct'' intensity of noise to add, we slightly vary the proposed approach: We do not add noise drawn from a probability distribution but instead sample the value for a feature from the distribution of all values of this feature.
-This makes the method non-parametric since the standard deviation of the noise does not need to be chosen.
+This makes the method non-parametric since the distribution of the noise does not need to be chosen.
 We ensured that features which stay constant for a flow, i.e.
@@
 
-We think that the ``ideal'' feature importance method for our scenario are Shapley values \cite{shapley_value_1953}: To analyze feature importance with them, the feature under investigation in completely omitted from the dataset and a model is trained without it.
+We think that the ``ideal'' feature importance method for our scenario are Shapley values \cite{shapley_value_1953}: To analyze feature importance with them, the feature under investigation is completely omitted from the dataset and a model is trained without it.
 The resulting accuracy is then compared to the baseline accuracy with all features.
@@
 As it can also happen that several features are left out for a sample, our method can also be used to analyze the effect of multiple features missing and can thus show possibly correlated features.
-The results in Table~\ref{fig:featDropout} show unlike the randomization feature importance method, the feature dropout method does not vastly overestimate features' importance.
+The results in Table~\ref{fig:featDropout} show that, unlike the randomization feature importance method, the feature dropout method does not vastly overestimate features' importance.
 With feature dropout, it becomes apparent that there are very few features that actually contain unique information and that affect accuracy when left out: the destination port and the source port.

### 2019-12-06 ex3/paper/paper.tex
Modified pdp/adv plots

@@
 \subsection{Evaluation}
-\begin{figure*}
-\includegraphics[width=0.45\textwidth]{adv_plots/cwl1.pdf}
-\includegraphics[width=0.45\textwidth]{adv_plots/cwl2.pdf}
-\caption{Performance of the \gls{cw} with different $\epsilon$ values for $\ell_1$ and $\ell_2$ distances.}
-\label{fig:adv_cw}
-\end{figure*}
+% \begin{figure*}
+% \includegraphics[width=0.45\textwidth]{adv_plots/cwl1.pdf}
+% \includegraphics[width=0.45\textwidth]{adv_plots/cwl2.pdf}
+% \caption{Performance of the \gls{cw} with different $\epsilon$ values for $\ell_1$ and $\ell_2$ distances.}
+% \label{fig:adv_cw}
+% \end{figure*}
 \autoref{fig:adv_cw} depicts the ratio of attack flows which were successfully misclassified as benign with respect to the final packet along with the resulting distance from the original sample.
@@
 
-\begin{figure*}
-\includegraphics[width=0.45\textwidth]{adv_plots/fgsm.pdf}
-\includegraphics[width=0.45\textwidth]{adv_plots/l_inf_pgd.pdf}
-\caption{Performance of FSGM and $\ell_\infty$-bounded \gls{pgd} for different bounds.}
-\label{fig:fgsm}
-\end{figure*}
+% \begin{figure*}
+% \includegraphics[width=0.45\textwidth]{adv_plots/fgsm.pdf}
+% \includegraphics[width=0.45\textwidth]{adv_plots/l_inf_pgd.pdf}
+% \caption{Performance of FSGM and $\ell_\infty$-bounded \gls{pgd} for different bounds.}
+% \label{fig:fgsm}
+% \end{figure*}
 
@@
 \subsubsection{Results}
+\begin{figure*}
+\includegraphics[width=.49\textwidth]{figures/feat_imp_flow_2017.pdf}\hspace{.02\textwidth}
+\includegraphics[width=.49\textwidth]{figures/feat_imp_flow_2015.pdf}
+\caption{Feature importance metrics considering the final flow prediction for CIC-IDS-2017 (left side) and UNSW-NB15 (right side).}
+\label{fig:feat_imp}
+\end{figure*}
 Table~\ref{fig:featRandom} shows our results.
@@
 \begin{figure*}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_0_Botnet-ARES"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_1_Brute Force-FTP-Patator"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot2_adv/flows_pred_plots2_outcomes_flows_adv_1.0_notBidirectional_outcomes_0_3.pickle_0_3.pickle_0_Botnet-ARES"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot2_adv/flows_pred_plots2_outcomes_flows_adv_1.0_notBidirectional_outcomes_0_3.pickle_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
+%\includegraphics[width=0.48\textwidth]{{"../plots/plot2_adv/flows_pred_plots2_outcomes_flows_adv_1.0_notBidirectional_outcomes_0_3.pickle_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
+%\includegraphics[width=0.48\textwidth]{{"../plots/plot2_adv/flows_pred_plots2_outcomes_flows_adv_1.0_notBidirectional_outcomes_0_3.pickle_0_3.pickle_1_Brute Force-FTP-Patator"}.pdf}
 

### 2019-12-06 ex3/plot2_adv.py

@@ -3,6 +3,7 @@
 import matplotlib.pyplot as plt
 import matplotlib
 import matplotlib.colors as mc
+from matplotlib.lines import Line2D
 import colorsys
 import numpy as np
 import sys
@@ -14,6 +15,11 @@ import math
 
 DIR_NAME = "plots/plot2_adv"
 
+USE_LOG_SCALE = True
+OMIT_UPPER_AXIS = True
+ONLY_ONE_LEGEND = True
+
+plt.rcParams["font.family"] = "serif"
 dataroot_basename = sys.argv[1].split('_')[0]
 
 with open(dataroot_basename + "_categories_mapping.json", "r") as f:
@@ -21,7 +27,7 @@ with open(dataroot_basename + "_categories_mapping.json", "r") as f:
 categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
 reverse_mapping = {v: k for k, v in mapping.items()}
 
-with open(dataroot_basename+"_full_no_ttl_normalization_data.pickle", "rb") as f:
+with open(dataroot_basename+"_normalization_data.pickle", "rb") as f:
 	means, stds = pickle.load(f)
 
 file_name = sys.argv[1]
@@ -44,7 +50,8 @@ adv_orig_flows_by_attack_number = adv_loaded["orig_flows_by_attack_number"]
 
 colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
 ORDERING = ["original", "adversarial"]
-FEATURE_NAMES = ["packet length", "iat"]
+FEATURE_NAMES = ["Packet length / B", "IAT / s"]
+FEATURE_SCALES = [1, 0.001] 
 
 def brighten(rgb, how_much=0.0):
 	hls = list(colorsys.rgb_to_hls(*rgb))
@@ -57,7 +64,6 @@ colors_rgb = [matplotlib.colors.to_rgb(item) for item in plt.rcParams['axes.prop
 COLOR_MAP_ELEMENTS = 100
 brightness_map = list(np.linspace(1.0, 0.5, num=COLOR_MAP_ELEMENTS))
 colors_rgb_ranges = [matplotlib.colors.ListedColormap([brighten(color, item) for item in brightness_map]) for color in colors_rgb]
-FEATURE_NAMES = ["packet length", "iat"]
 
 
[…]

### 2019-12-06 ex3/learn.py
Final mutinfo changes (probably)

@@ -646,7 +646,7 @@ def feature_importance():
 		# Joint pdf for feature values in individual time steps and confidence in the same time step
 		per_packet_pdf = np.zeros([opt.maxLength,distribution.shape[0],PDF_FEATURE_BINS,PDF_CONFIDENCE_BINS])
 		# Joint pdf for constant feature values and end confidence
-		per_flow_pdf = np.zeros([max(constant_features)+1,PDF_FEATURE_BINS,PDF_CONFIDENCE_BINS])
+		per_flow_pdf = np.zeros([distribution.shape[0],PDF_FEATURE_BINS,PDF_CONFIDENCE_BINS])
 
 	for input_data, labels, categories in test_loader:
 
@@ -691,6 +691,7 @@ def feature_importance():
 				input_data_cloned = torch.nn.utils.rnn.pack_padded_sequence(input_data_padded, input_data_lens, enforce_sorted=False)
 			else:
 				input_data_cloned.data.data[:,feature_index] = torch.FloatTensor(np.random.choice(distribution[feature_index], size=(input_data_cloned.data.data.shape[0]))).to(device)
+			input_data_padded_np = input_data_padded.cpu().numpy()
 			output, seq_lens = lstm_module(input_data_cloned)
 
 			sigmoided_output = torch.sigmoid(output.detach())
@@ -701,12 +702,13 @@ def feature_importance():
 
 				if opt.mutinfo:
 					if feature_index in constant_features:
-						bin1 = np.argmax(input_data_padded[0,batch_index,feature_index] <= bin_boundaries[feat_ind,:], axis=1)
-						#  bin1 = int(torch.round((PDF_FEATURE_BINS-1) * (input_data_padded[0,batch_index,feature_index] - minmax[feature_index][0]) / (minmax[feature_index][1]-minmax[feature_index][0])))
-						bin2 = int(torch.round((PDF_CONFIDENCE_BINS-1) * sigmoided_output[flow_length-1,batch_index,0]))
-						per_flow_pdf[feature_index,bin1,bin2] += 1
+						bin1 = np.argmax(input_data_padded_np[0,batch_index,feature_index] <= bin_boundaries[feature_index,:], axis=0)
+					else:
+						bin1 = np.argmax(input_data_padded_np[:flow_length,batch_index,feature_index,None] <= bin_boundaries[None,feature_index,:], axis=1)
[…]

### 2019-12-06 ex3/paper/paper.tex
Added feature importance results

@@
 \begin{figure*}
-\includegraphics[width=.49\textwidth]{figures/feat_imp_flow_2017.pdf}\hspace{.02\textwidth}
-\includegraphics[width=.49\textwidth]{figures/feat_imp_flow_2015.pdf}
+\includegraphics[width=.49\textwidth]{../plots/importance/feat_imp_flow_2017.pdf}\hspace{.02\textwidth}
+\includegraphics[width=.49\textwidth]{../plots/importance/feat_imp_flow_2015.pdf}
 \caption{Feature importance metrics considering the final flow prediction for CIC-IDS-2017 (left side) and UNSW-NB15 (right side).}

### 2019-12-06 ex3/barplot.py

@@ -1,0 +1,49 @@
+#!/usr/bin/env python3
+
+# An ad-hoc script for plotting tables copy-pasted from LibreOffice Calc
+
+import fileinput
+import numpy as np
+import matplotlib.pyplot as plt
+import sys
+from matplotlib.transforms import Affine2D
+
+group_names = []
+values = []
+
+plt.rcParams['font.family'] = 'serif'
+plt.rcParams['savefig.format'] = 'pdf'
+plt.figure(figsize=(5,4))
+
+for line in fileinput.input():
+	field = line.rstrip().split('\t')
+	group_names.append(field[0])
+	values.append(field[1:])
+
+feature_names = values[0]
+group_names = group_names[1:]
+values = values[1:]
+
+print (group_names)
+print (values)
+values = np.array(values, dtype=float)
+values[values < 0] = 0
+values /= np.sum(values,axis=0)
+
+width = 0.95 / values.shape[1]
+
+order = np.argsort(-np.mean(values, axis=1))
+x = np.arange(len(group_names))
+for i in range(values.shape[1]):
+	plt.bar(x +width*(i-values.shape[1]/2), values[order,i], width, label=feature_names[i])
+plt.legend()
+plt.xticks(x, [group_names[i] for i in order], rotation=45, horizontalalignment='right')
+for tick in plt.gca().xaxis.get_major_ticks():
+	label = tick.label1
+	label.set_transform(label.get_transform() + Affine2D().translate(10,0))
+plt.ylabel('Normalized metric')
+plt.tight_layout()
+
+plt.savefig(sys.stdout.buffer, bbox_inches = 'tight', pad_inches = 0)
+plt.show()
+

### 2019-12-07 ex3/learn.py
Fix yet another bug in correlation

@@ -460,8 +460,8 @@ def dropout_feature_correlation():
 			accuracy_for_feature_2 = accuracy_flow_by_feature[feat2]
 			accuracy_for_feature_combination = np.mean(np.concatenate([feature for attack_type in dropout_flow_results_by_attack_number for feature in attack_type[(feat1,feat2)]]))
 			joint_accuracy[(feat1, feat2)] = accuracy_for_feature_combination
-			numerator = max(float(baseline_flow_accuracy-accuracy_for_feature_combination), 0)
-			denominator = max(float(baseline_flow_accuracy-accuracy_for_feature_1), 0) + max(float(baseline_flow_accuracy-accuracy_for_feature_2), 0)
+			numerator = max(float(baseline_flow_accuracy-accuracy_for_feature_combination), 0.0)
+			denominator = max(float(baseline_flow_accuracy-accuracy_for_feature_1), 0.0) + max(float(baseline_flow_accuracy-accuracy_for_feature_2), 0.0)
 			try:
 				overall_score = numerator/denominator
 			# accuracy doesn't fall if either feat1 or feat2 are omitted
@@ -473,13 +473,13 @@ def dropout_feature_correlation():
 				else:
 					overall_score = float("nan")
 			correlations[(feat1, feat2)] = overall_score
-			pretty_print("feat1", feat1, feature_array[feat1], feat2, "feat2", feature_array[feat2], "acc_feat1", baseline_flow_accuracy-accuracy_by_feature[feat1], "acc_feat2", baseline_flow_accuracy-accuracy_by_feature[feat2], "joint_acc", baseline_flow_accuracy-joint_accuracy[(feat1, feat2)], "score", overall_score, sep=" ")
+			pretty_print("feat1", feat1, feature_array[feat1], feat2, "feat2", feature_array[feat2], "acc_feat1", baseline_flow_accuracy-accuracy_flow_by_feature[feat1], "acc_feat2", baseline_flow_accuracy-accuracy_flow_by_feature[feat2], "joint_acc", baseline_flow_accuracy-joint_accuracy[(feat1, feat2)], "score", overall_score, sep=" ")
 
 	N_MOST_CORRELATED_TO_SHOW = len(list(correlations.items()))
 	sorted_correlations = list(sorted([item for item in list(correlations.items()) if not math.isnan(item[1])], key=lambda item: item[1], reverse=True))[:N_MOST_CORRELATED_TO_SHOW]
[…]

### 2019-12-07 ex3/paper/paper.tex
Make references consistent and add one line to feature correlation text

@@
 With the recent rise of interest in \gls{aml} techniques, also \gls{aml} for \glspl{ids} has been investigated.
-For example, \cite{Bachl2019} investigates remedies for poisoning attacks on \glspl{ids}.
+For example, \cite{bachl_walling_2019} investigates remedies for poisoning attacks on \glspl{ids}.
 In this research, we investigate whether adversarial samples can be found for \gls{ml} based on \glspl{rnn}.
@@
 
-\autoref{tab:performance_results} shows that our classifier achieves an accuracy that is similar to previous work based on these datasets~\cite{meghdouri_analysis_2018,Bachl2019}.
+\autoref{tab:performance_results} shows that our classifier achieves an accuracy that is similar to previous work based on these datasets~\cite{meghdouri_analysis_2018,bachl_walling_2019}.
 However, unlike these classifiers, our recurrent classifier has the advantage that it can detect attacks already before they are over.
@@
 The higher the resulting number, the larger the information that both features share.
+If the score is one, this means that both features contain no joint information.
 

### 2019-12-07 ex3/paper/paper.tex
Add some more feature correlation

@@
 And even these features cannot be manipulated freely: Besides the constraints that packet size and interarrival time cannot become negative there are also more problem-specific constraints:
-\begin{itemize}
+\begin{itemize}[topsep=0pt,wide,labelwidth=!,labelindent=0pt]
 \item Only packets can be manipulated which are transmitted by the attacker. This means the attacker can only control one direction (except for botnets where both sides are attackers).
@@
 % \end{figure*}
+\todo{Max: Fix me. Don't know how the reference became invalid...}
 \autoref{fig:adv_cw} depicts the ratio of attack flows which were successfully misclassified as benign with respect to the final packet along with the resulting distance from the original sample.
@@
 \centering
-\begin{tabular}{lllllllll}
+\begin{tabular}{llllllllll}
 \toprule
-Attack type & FGSM & Carlini-Wagner, $\ell_1$ & Carlini-Wagner, $\ell_2$ & $\ell_\infty$-bounded \gls{pgd} \\
+Attack type & \multicolumn{2}{l}{FGSM} & \multicolumn{2}{l}{Carlini-Wagner, $\ell_1$} & \multicolumn{2}{l}{Carlini-Wagner, $\ell_2$} & \multicolumn{2}{l}{$\ell_\infty$-bounded \gls{pgd}} \\
+\cmidrule(r){2-3} \cmidrule(lr){4-5} \cmidrule(l){6-7} \cmidrule(l){8-9}
+& adv. & orig. & adv. & orig. & adv. & orig. & adv. & orig. \\
 \midrule
-Botnet:ARES	&	0.0000 / 0.0000	&	0.0000 / 0.0000	&	0.0000 / 0.0000	&	0.0000 / 0.0000	\\
-Brute Force:FTP-Patator	&	0.2800 / 0.5800	&	0.1600 / 0.5800	&	0.1600 / 0.5800	&	0.2800 / 0.5800	\\
-Brute Force:SSH-Patator	&	0.8300 / 1.0000	&	0.0000 / 1.0000	&	0.0000 / 1.0000	&	0.8900 / 1.0000	\\
-DDoS:LOIT	&	0.8300 / 0.8600	&	0.6400 / 0.8600	&	0.5600 / 0.8600	&	0.8200 / 0.8600	\\
-DoS / DDoS:DoS GoldenEye	&	0.8900 / 1.0000	&	0.5000 / 1.0000	&	0.4200 / 1.0000	&	0.9200 / 1.0000	\\
-DoS / DDoS:DoS Hulk	&	0.6700 / 1.0000	&	0.5400 / 1.0000	&	0.4700 / 1.0000	&	1.0000 / 1.0000	\\
-DoS / DDoS:DoS Slowhttptest	&	0.8700 / 1.0000	&	0.4100 / 1.0000	&	0.1900 / 1.0000	&	0.7400 / 1.0000	\\
-DoS / DDoS:DoS slowloris	&	0.4400 / 0.9700	&	0.3100 / 0.9700	&	0.1300 / 0.9700	&	0.9200 / 0.9700	\\
-DoS / DDoS:Heartbleed	&	0.0000 / 0.0000	&	0.0000 / 0.0000	&	0.0000 / 0.0000	&	0.0000 / 0.0000	\\
-Infiltration:Dropbox download 	&	0.6100 / 0.9600	&	0.4300 / 0.9600	&	0.3800 / 0.9600	&	0.7400 / 0.9600	\\
-PortScan:PortScan - Firewall off	&	0.6800 / 0.9100	&	0.3500 / 0.9100	&	0.3500 / 0.9100	&	0.7200 / 0.9100	\\
-PortScan:PortScan - Firewall on	&	0.7400 / 0.8700	&	0.6100 / 0.8700	&	0.5300 / 0.8700	&	0.7300 / 0.8700	\\
-Web Attack:Sql Injection	&	0.0000 / 0.6923	&	0.0000 / 0.6923	&	0.0000 / 0.6923	&	0.0000 / 0.6923	\\
-Web Attack:XSS	&	0.6700 / 0.9500	&	0.7200 / 0.9500	&	0.6200 / 0.9500	&	0.6700 / 0.9500	\\
+Botnet:ARES	&	0.0000 & 0.0000	&	0.0000 & 0.0000	&	0.0000 & 0.0000	&	0.0000 & 0.0000	\\
+Brute Force:FTP-Patator	&	0.2800 & 0.5800	&	0.1600 & 0.5800	&	0.1600 & 0.5800	&	0.2800 & 0.5800	\\
+Brute Force:SSH-Patator	&	0.8300 & 1.0000	&	0.0000 & 1.0000	&	0.0000 & 1.0000	&	0.8900 & 1.0000	\\
+DDoS:LOIT	&	0.8300 & 0.8600	&	0.6400 & 0.8600	&	0.5600 & 0.8600	&	0.8200 & 0.8600	\\
+DoS / DDoS:DoS GoldenEye	&	0.8900 & 1.0000	&	0.5000 & 1.0000	&	0.4200 & 1.0000	&	0.9200 & 1.0000	\\
+DoS / DDoS:DoS Hulk	&	0.6700 & 1.0000	&	0.5400 & 1.0000	&	0.4700 & 1.0000	&	1.0000 & 1.0000	\\
+DoS / DDoS:DoS Slowhttptest	&	0.8700 & 1.0000	&	0.4100 & 1.0000	&	0.1900 & 1.0000	&	0.7400 & 1.0000	\\
+DoS / DDoS:DoS slowloris	&	0.4400 & 0.9700	&	0.3100 & 0.9700	&	0.1300 & 0.9700	&	0.9200 & 0.9700	\\
+DoS / DDoS:Heartbleed	&	0.0000 & 0.0000	&	0.0000 & 0.0000	&	0.0000 & 0.0000	&	0.0000 & 0.0000	\\
+Infiltration:Dropbox download 	&	0.6100 & 0.9600	&	0.4300 & 0.9600	&	0.3800 & 0.9600	&	0.7400 & 0.9600	\\
+PortScan:PortScan - Firewall off	&	0.6800 & 0.9100	&	0.3500 & 0.9100	&	0.3500 & 0.9100	&	0.7200 & 0.9100	\\
+PortScan:PortScan - Firewall on	&	0.7400 & 0.8700	&	0.6100 & 0.8700	&	0.5300 & 0.8700	&	0.7300 & 0.8700	\\
+Web Attack:Sql Injection	&	0.0000 & 0.6923	&	0.0000 & 0.6923	&	0.0000 & 0.6923	&	0.0000 & 0.6923	\\
+Web Attack:XSS	&	0.6700 & 0.9500	&	0.7200 & 0.9500	&	0.6200 & 0.9500	&	0.6700 & 0.9500	\\
 \bottomrule
@@
 
-\begin{itemize}
+\begin{itemize}[topsep=0pt,wide,labelwidth=!,labelindent=0pt]
 \item
@@
 \textbf{Multiple prediction outputs.}
-Often an \gls{rnn} produces an output at each time step. When trying to apply explainability methods, hence, the question arises which output to base the methods on.
-
+Often an \gls{rnn} produces an output at each time step.
+When trying to apply explainability methods, hence, the question arises which output to base the methods on.
 The natural choice is to base the methods on the prediction output which occurs at the same time step as the feature under investigation.
@@
 
-% numerator = max(float(baseline_accuracy-accuracy_for_feature_combination), 0)
-% 			denominator = max(float(baseline_accuracy-accuracy_for_feature_1), 0) + max(float(baseline_accuracy-accuracy_for_feature_2), 0)
 \begin{equation}
@@
 If the score is one, this means that both features contain no joint information.
+For example, the score between RST and the protocol identifier is 8.5; the highest of all pairs of features.
+While this may seem unintuitive at the first glance, it likely stems from the fact that if the protocol identifier (TCP or UDP) is missing, then if RST is 1 at some point of the flow, the flow has to be TCP.
 
@@
 \includegraphics[width=\columnwidth]{../plots/mutinfo_example.pdf}
-\caption{Distributions $f_A(x)$ yielding an identical accuracy drop.}
+\caption{Two distributions yielding an identical accuracy drop.}
 \label{fig:mutinfo_example}
@@
 
-Furthermore, let $f_A(x)$  denote the joint probability for classification as attack with a feature value $x$.
+Furthermore, let $f_A(x)$  denote the joint probability \todo{Max: Ist joint probability nicht immer wenn man mehrere Ereignisse hat?
+Ich hätte hier probability density function geschrieben...} for classification as attack with a feature value $x$.
 Accuracy then is $\int f_A(x) dx$ for attacks and $1-\int f_A(x) dx$ for benign traffic.
-Consider \autoref{fig:mutinfo_example}.
-Evidently, both distributions
-yield the same accuracy, but clearly the right-side distribution has a larger influence on the prediction, as in the right-side case the prediction deterministically depends on the feature value.
+\autoref{fig:mutinfo_example} shows two different distributions
+yielding the same accuracy, but clearly the right-side distribution has a larger influence on the prediction, as in the right-side case the prediction deterministically depends on the feature value.
 To capture such dependencies, we propose to use mutual information instead of accuracy drop to determine feature sensitivity.
@@
 The most obvious defence, which is only possible because of the nature of the dataset, is to simply leave out features which the attacker can manipulate. For this, we try two different approaches:
-\begin{itemize}
+\begin{itemize}[topsep=0pt,wide,labelwidth=!,labelindent=0pt]
 \item Leaving out all features that are manipulable: packet size and interarrival time

### 2019-12-07 ex3/paper/paper.tex
Several text changes

@@
 \newacronym{ars}{ARS}{Adversarial Robustness Score}
+\newacronym{ttl}{TTL}{Time-to-Live}
+\newacronym{dos}{DoS}{Denial-of-Service}
+\newacronym{iat}{IAT}{Interarrival time}
+\newacronym{tcp}{TCP}{Transmission control protocol}
 
@@
 
-With the rise of ubiquitous \gls{ml} systems, malicious actors have been catching up quickly to find new ways to exploit them for profit.
+\glspl{rnn} seem like an attractive choice for performing intrusion detection on network data.
+However, with the rise of ubiquitous \gls{ml} systems, malicious actors have been catching up quickly to find new ways to exploit them for profit.
 Recently developed \gls{aml} techniques focus on computer vision and their applicability to network traffic is not straightforward: Network packets expose fewer features than an image, are sequential and impose several constraints on their features: For example, destination ports cannot be modified since that could impair communication.
-We show that despite these completely different characteristics, adversarial samples can be reliably generated.
-This shows that traditional evaluation metrics such as accuracy are not sufficient for quantifying the adversarial threat and we propose \gls{ars} as a countermeasure.
+
+We show that despite these completely different characteristics, adversarial samples can be reliably generated for \glspl{rnn}.
+Since traditional evaluation metrics such as accuracy are not sufficient for quantifying the adversarial threat, we propose the \gls{ars}.
+%as a countermeasure.
+% AH.
+the score alone is not yet a countermeasure
+To understand the classifier's potential for misclassification, we extend existing explainability techniques and propose new ones, particularly suitable % specifically crafted %AH most can also be used for non-sequential models, so lets not restrict applicability too much
+for sequential data.
+Applying them shows that already the first packets of a communication flow are of crucial importance and are likely going to be targeted by attackers.
 Feature importance methods show that even though the modified features are not very relevant for the classifier, they can still be abused to generate adversarial samples.
-We then extend existing explainability techniques and propose new ones, specifically crafted for sequential data.
-Applying them shows that already the first couple of packets of a communication flow are of crucial importance and are likely going to be targeted by attackers.
-As a countermeasure we develop an easy-to-use adversarial training procedure and can confirm that it significantly and successfully reduces the attack surface and improves \gls{ars}.
+As a countermeasure, we develop an easy-to-use adversarial training procedure %and can confirm that it
+which significantly and successfully reduces the attack surface and improves \gls{ars}.
 
@@
 \item packet directions (if packet goes there or back)
-\item all TCP flags (0 if flow is not TCP)
-\end{itemize*}. Of these features, source port, destination port and protocol identifier are constant over the whole flow while the others vary.
+\item all \gls{tcp} flags (0 if flow is not \gls{tcp})
+\end{itemize*}.
+We omitted \gls{ttl} values, as they are likely to lead to unwanted prediction behaviour \cite{bachl_walling_2019}.
+Among the used features, source port, destination port and protocol identifier are constant over the whole flow while the others vary.
 We use Z-score normalization and a train/test split of 2:1.
@@
 \label{sec:adv}
-
-Most features such as, e.g., TCP flags cannot be easily manipulated because it might violate the TCP and thus cause communication to fail.
+We now investigate whether known \gls{aml} methods can be used for generating adversarial flows, i.e. minimally different attack flows which are classified as benign.
+In our scenario, most features such as, e.g., \gls{tcp} flags cannot be easily manipulated because their manipulation might violate the \gls{tcp} and thus cause communication to fail.
 Hence, we experiment whether by manipulating the interarrival times and packet lengths of the first packets, the classifier might be fooled.
@@
 
-Several methods for generating adversarial samples have been proposed in the recent literature, achieving different speed-quality tradeoffs: 
+Several \gls{aml} methods have been proposed in the recent literature, achieving different speed-quality tradeoffs: 
 \cite{szegedy_intriguing_2014} show that it is possible to create adversarial samples for an image, which look similar to the original image but are classified wrongly.
@@
 \subsubsection{Neural Network Weights}
-In previous works, a simple method for determining feature importance in neural networks, has been summing up neural network weights leading from a dertain input to the prediction outcome.
+In previous works, a simple method for determining feature importance in neural networks, has been summing up neural network weights leading from a certain input to the prediction outcome.
 The weights method can be considered for both feature importance and feature sensitivity, however, clearly, especially in the case of complex network architectures, this method is likely to provide wrong results.
@@
 For example, the score between RST and the protocol identifier is 8.5; the highest of all pairs of features.
-While this may seem unintuitive at the first glance, it likely stems from the fact that if the protocol identifier (TCP or UDP) is missing, then if RST is 1 at some point of the flow, the flow has to be TCP.
+While this may seem unintuitive at the first glance, it likely stems from the fact that if the protocol identifier (\gls{tcp} or UDP) is missing, then if RST is 1 at some point of the flow, the flow has to be \gls{tcp}.
 
@@
 \end{figure*}
-Table~\ref{fig:featRandom} shows our results.
-The results match to a large extent with the expectations one has when classifying network flows.
-In particular, the most important features turn out to be the used protocol, packet length and the SYN TCP flag.
+\autoref{fig:featRandom} and \autoref{fig:feat_imp} show the results, which match to a large extent with domain-specific expectations for classifying network flows.
+In particular, rarely used \gls{tcp} flags like NS or URG are unimportant for the classifier.
+On the other hand, destination port and protocol are essential for characterizing flows by hinting at the type of traffic.
+Interarrival time and packet length are important for estimating the amount of transferred information and several flags hint at certain attacks like \gls{dos} attacks.
+
+The weights method is able to reveal features with a very low importance to a certain degree, but disagrees with the other methods to a large extent.
+Less anticipated, however, also input perturbation does not exhibit a considerable correlation with feature dropout.
+Considering its functioning, feature dropout is the most reliable method for evaluating importance for removing features.
+It is remarkable that none of the other methods is able to depict the distinct peak of importance for the destination port.
+
+It is not surprising that mutual information disagrees with feature dropout, since both their aim and their functioning are substantially different.
+For example, mutual information shows that the protocol field can have a substantial impact on the classification even though an identical accuracy can be achieved when omitting it.
+
+% In particular, the most important features turn out to be the used protocol, packet length and the SYN TCP flag.
 While the first two features are essential for characterizing flows, a high number of packets with set SYN flag is an important characteristic for port scanning or certain DoS attacks.
@@
 It is interesting to note, however, that the interarrival time of packets does not play a crucial role for the detection of attacks.
+
+Note that metrics for UNSW-NB15 differ substantially from CIC-IDS2017.
+However, due to the large number of different network attacks and network configurations, it is easily possible that different features are relevant for both datasets.
+We consider the question of model transferability of substantial importance for the presented application, but out of scope for the present research.
 

### 2019-12-07 ex3/paper/paper.tex
Tiny abstract changes

@@
 
-\glspl{rnn} seem like an attractive choice for performing intrusion detection on network data.
-However, with the rise of ubiquitous \gls{ml} systems, malicious actors have been catching up quickly to find new ways to exploit them for profit.
-Recently developed \gls{aml} techniques focus on computer vision and their applicability to network traffic is not straightforward: Network packets expose fewer features than an image, are sequential and impose several constraints on their features: For example, destination ports cannot be modified since that could impair communication.
+\glspl{rnn} yield attractive properties for performing intrusion detection on network data.
+However, with the rise of ubiquitous \gls{ml} systems, malicious actors have been catching up quickly to find new ways to exploit \gls{ml} for profit.
+Recently developed \gls{aml} techniques focus on computer vision and their applicability to network traffic is not straightforward: Network packets expose fewer features than an image, are sequential and impose several constraints on their features.
+For example, destination ports cannot be modified since that could impair communication.
 
@@
 the score alone is not yet a countermeasure
-To understand the classifier's potential for misclassification, we extend existing explainability techniques and propose new ones, particularly suitable % specifically crafted %AH most can also be used for non-sequential models, so lets not restrict applicability too much
+To understand a classifier's potential for misclassification, we extend existing explainability techniques and propose new ones, particularly suitable % specifically crafted %AH most can also be used for non-sequential models, so lets not restrict applicability too much
 for sequential data.
@@
 \end{figure*}
-\autoref{fig:featRandom} and \autoref{fig:feat_imp} show the results, which match to a large extent with domain-specific expectations for classifying network flows.
+Table~\ref{fig:featRandom} and \autoref{fig:feat_imp} show the results, which match to a large extent with domain-specific expectations for classifying network flows.
 In particular, rarely used \gls{tcp} flags like NS or URG are unimportant for the classifier.

### 2019-12-07 ex3/paper/paper.tex
Abbreviated IAT

@@
 \item packet sizes
-\item interarrival times of packets
+\item \glspl{iat} of packets
 \item packet directions (if packet goes there or back)
@@
 In our scenario, most features such as, e.g., \gls{tcp} flags cannot be easily manipulated because their manipulation might violate the \gls{tcp} and thus cause communication to fail.
-Hence, we experiment whether by manipulating the interarrival times and packet lengths of the first packets, the classifier might be fooled.
+Hence, we experiment whether by manipulating the \glspl{iat} and packet lengths of the first packets, the classifier might be fooled.
 Our scenario is quite different from how generating adversarial samples usually works: We have significantly less features than in, for example, an image.
 Furthermore, only two out of those few features can be manipulated.
-And even these features cannot be manipulated freely: Besides the constraints that packet size and interarrival time cannot become negative there are also more problem-specific constraints:
+And even these features cannot be manipulated freely: Besides the constraints that packet size and \gls{iat} cannot become negative there are also more problem-specific constraints:
 \begin{itemize}[topsep=0pt,wide,labelwidth=!,labelindent=0pt]
@@
 \item Packets must not be smaller than initially, as otherwise less information could be transmitted or packets might be rendered syntactically invalid.
-\item Interarrival times must not decrease, as otherwise the physical speed of data transmission can be violated in some cases.
-Since an in-depth analysis of when the reduction of interarrival times is possible, would come with substantial effort, we generally disallowed a reduction of this feature.
+\item \glspl{iat} must not decrease, as otherwise the physical speed of data transmission can be violated in some cases.
+Since an in-depth analysis of when the reduction of \glspl{iat} is possible, would come with substantial effort, we generally disallowed a reduction of this feature.
 \end{itemize}
@@
 The method minimizes the adversarial sample's distance so that the network's decision is just between attack and normal traffic.
-In the present context, we need to make sure that the classifier's prediction would actually be normal traffic, even though a certain level of noise will be added to interarrival times due to the network between attacker and victim.
+In the present context, we need to make sure that the classifier's prediction would actually be normal traffic, even though a certain level of noise will be added to \glspl{iat} due to the network between attacker and victim.
 Instead of the above equation \ref{eq:carliniWagner} we therefore used the optimization objective
@@
 \autoref{adv_plots} depicts how the method manipulates flows for some exemplary attack categories.
-Depicted is the mean of both unmodified flows and adversarial flows  of samples of individual attack families considering both interarrival time and packet length.
+Depicted is the mean of both unmodified flows and adversarial flows  of samples of individual attack families considering both \gls{iat} and packet length.
 
@@
 On the other hand, destination port and protocol are essential for characterizing flows by hinting at the type of traffic.
-Interarrival time and packet length are important for estimating the amount of transferred information and several flags hint at certain attacks like \gls{dos} attacks.
+\gls{iat} and packet length are important for estimating the amount of transferred information and several flags hint at certain attacks like \gls{dos} attacks.
 
@@
 Next, we want to investigate if our classifier just memorized its inputs.
-The most interesting packet features that can be used to visualize this in our use case are the inter-arrival time (IAT) between packets in a flows and the packet length.
+The most interesting packet features that can be used to visualize this in our use case are the \gls{iat} between packets in a flows and the packet length.
 To depict how the model comes to a decision, we determined the total minimum and maximum for these features that ever occur for all flows.
@@
 Looking at the plots of the features that vary over time (\autoref{fig:plot_features}) we see that certain attack types have a characteristic pattern in which they send packets by which they are easily recognizable.
-We want to determine how the pattern of packet sizes and interarrival times is significant for the classification.
-\autoref{fig:plot_features} clearly shows that some attack types have very characteristic patterns of packet sizes and interarrival times.
+We want to determine how the pattern of packet sizes and \glspl{iat} is significant for the classification.
+\autoref{fig:plot_features} clearly shows that some attack types have very characteristic patterns of packet sizes and \glspl{iat}.
 
@@
 \begin{itemize}[topsep=0pt,wide,labelwidth=!,labelindent=0pt]
-\item Leaving out all features that are manipulable: packet size and interarrival time
+\item Leaving out all features that are manipulable: packet size and \gls{iat}
 \item Leaving out manipulable features when the attacker can manipulate them: only in the direction from the attacker to the victim (This, however does not prevent adversarial samples for botnets, for which both sides are malicious)
@@
 However, when looking at the packet accuracy, which considers all packets of all flows, there is a significant difference.
-Thus, apparently the interarrival time and packet size are especially important for determining whether a flow is malicious in the first packets of a flow.
+Thus, apparently the \gls{iat} and packet size are especially important for determining whether a flow is malicious in the first packets of a flow.
 
@@
 \begin{table*}
-\caption{Comparing performance with all features, only safe features (interarrival time and packet size removed) and safe features in one direction (interarrival time and packet size removed for packets coming from source).
+\caption{Comparing performance with all features, only safe features (\gls{iat} and packet size removed) and safe features in one direction (\gls{iat} and packet size removed for packets coming from source).
 \textit{packets} means that we consider the neural network output of all packets of the flow while \textit{flows} means that we only consider the last packet.

### 2019-12-07 ex3/learn.py
Fix pred_plots2

@@ -1378,13 +1378,13 @@ def pred_plots2():
 	_, test_indices = get_nth_split(dataset, n_fold, fold)
 	subset = torch.utils.data.Subset(dataset, test_indices)
 
+	attack_numbers = mapping.values()
+	
 	if opt.pathToAdvOutput == "":
-		features = get_feature_ranges(subset, sampling_density=100)
+		features = [ get_feature_ranges(subset, sampling_density=100) ] * (max(attack_numbers)+1)
 	else:
 		features = get_feature_ranges_from_adv(sampling_density=100)
 	# print("features", features)
-
-	attack_numbers = mapping.values()
 
 	results_by_attack_number = [list() for _ in range(min(attack_numbers), max(attack_numbers)+1)]
 	flows_by_attack_number = [list() for _ in range(min(attack_numbers), max(attack_numbers)+1)]
@@ -1413,7 +1413,7 @@ def pred_plots2():
 			lstm_module.forgetting = True
 
 			input_data = torch.FloatTensor(flow[i,:][None,None,:]).repeat(1,max([len(item) for item in features])*max([(len(item[0][1])-1) if len(item) > 0 else 0 for item in features]),1)
-			for k, (feat_ind, values) in enumerate(features[cat] if opt.pathToAdvOutput!="" else features):
+			for k, (feat_ind, values) in enumerate(features[cat]):
 
 				for j in range(values.size-1):
 					# print("input_data.shape", input_data.shape, "k*values.size+j", k*values.size+j)
@@ -1428,7 +1428,7 @@ def pred_plots2():
 			output, _ = lstm_module(packed_input)
 			sigmoided = torch.sigmoid(output[0,:,0]).detach().cpu().tolist()
 
-			for k, (feat_ind, values) in enumerate(features[cat] if opt.pathToAdvOutput!="" else features):
+			for k, (feat_ind, values) in enumerate(features[cat]):
 				for j in range(values.size-1):
 					prediction_ranges[i,k,:] = sigmoided[k*(values.size-1):(k+1)*(values.size-1)]
 

### 2019-12-07 ex3/paper/paper.tex
Modified paper

@@
 More recently, with the rise of omnipresent encryption, available features have changed.
-Since payload is not available anymore, the focus now lies on features that are always available to observers such as port numbers, protocol flags etc.
+Since payload is no longer avilable, the focus now lies on features that are always available to network monitoring equipment such as port numbers, protocol flags, packet size etc.
+
 Network communication is usually grouped into \textit{flows}, which are commonly defined as a sequence of packets sharing certain properties, between two applications that may be located at different physical locations.
-
-When analyzing flows, not only the aforementioned features are available but also features related to the timing of the individual packets as well as their size.
-For example, a specific attack might require the attacker to send three small packets and the victim to reply with two large packets followed by a small one the attacker.
-Various approaches have been proposed to extract features from flows and then perform anomaly detection with the results \cite{meghdouri_analysis_2018}.
-While these approaches work well, it is problematic that first the whole flow has to be received and only afterwards it can be determined if there was an anomaly.
-Thus, we design a network anomaly detection system that operates on a per-packet basis and decides if a packet is anomalous based on features that are available even for traffic that is encrypted above the transport layer, like for example TLS or QUIC.
+When analyzing flows, not only the aforementioned features are available but also features related to the timing of the individual packets.
+% as well as their size.
+%AH: don't see why pkt size is not available without flows
+For example, a specific attack might require the attacker to send three small packets and the victim to reply with two large packets followed by a small one to the attacker.
+Various approaches have been proposed to extract features from flows and then perform anomaly detection with the extracted flows \cite{meghdouri_analysis_2018}.
+While these approaches frequently work well, it is problematic that the whole flow has to be received first and only afterwards an anomaly can be detected.
+Thus, we design a network \gls{ids} that operates on a per-packet basis and decides if a packet is anomalous based on features that are available even for traffic that is encrypted above the transport layer, like for example TLS or QUIC.
+At the same time, an \gls{rnn}-based \gls{ids} has the benefit of avoiding tedious feature selection procedures while providing any available information to the classifier.
 We show that our system has similar performance to other flow-based anomaly detection systems but can commonly detect anomalies \textit{before} the flow is over.
@@
 For example, \cite{bachl_walling_2019} investigates remedies for poisoning attacks on \glspl{ids}.
-In this research, we investigate whether adversarial samples can be found for \gls{ml} based on \glspl{rnn}.
+In this research, we investigate whether adversarial samples can be found for our \gls{rnn}-based model.
 This is not a trivially answerable question since adversarial samples have mostly been analyzed in the context of computer vision.
@@
 Surprisingly, we can confirm that adversarial samples can be found even when considering these tight constraints. % mentioned before.
-Thus, we argue that traditional methods like accuracy are not sufficient for evaluating security-sensitive \gls{ml} systems.
+Thus, we argue that traditional methods like, e.g., accuracy are not sufficient for evaluating security-sensitive \gls{ml} systems.
 We propose the \acrfull{ars} as a new performance score, which precisely captures the notion of how easily an adversary can generate adversarial samples.
 
-Astonishingly, feature importance methods reveal that the manipulable features, that are left for adversaries to modify are not even particularly significant for the classification outcome.
+In practice, a further crucial requirement for \gls{ml}-based \glspl{ids} 
+is that the classifiers decisions can be interpreted, recently giving rise to increasing interest in explainability methods.
+Also in this case, common methods are designed to work with images or tabular data but not with sequences.
+Hence, we extend existing explainability methods such as \glspl{pdp} \cite{friedman_greedy_2001} for sequential data.
+These show the varying characteristics of different attacks and they aid with visualizing potentially vulnerable features for adversaries.
+Astonishingly, feature importance methods reveal that the manipulable features, that are left for modification by adversaries are not even particularly significant for the classification outcome.
 %Especially the first packets of a flow are critical for determining the decision of a recurrent \gls{ids}.
 
-Since common explainability methods are designed to work with images or tabular data but not sequences, we extend existing explainability methods such as \glspl{pdp} \cite{friedman_greedy_2001} for sequential data.
+%Since common explainability methods are designed to work with images or tabular data but not sequences, we extend existing explainability methods such as \glspl{pdp} \cite{friedman_greedy_2001} for sequential data.
 These show the varying characteristics of different attacks and they aid with visualizing potentially vulnerable features for adversaries.
 
-Based on the insights gained by the feature importance and explainability methods, we develop a simple method for defending against adversaries who can manipulate arbitrary packets by simply leaving out potentially manipulable information from the classification process.
-The resulting classifier is slightly less accurate but is still useful to be deployed in a real scenario.
+Based on the insights gained by the feature importance and explainability methods, we propose two defence methods.
+%develop a simple method for defending against adversaries who can manipulate arbitrary packets by simply leaving out potentially manipulable information from the classification process.
+On the one hand, by simply leaving out manipulable features, we obtain a classifier which is slightly less accurate but is still useful to be deployed in a real scenario.
 Furthermore, we develop a custom adversarial training procedure which especially focuses on ease of use and low resource consumption and aims to keep the number of additionally introduced training hyperparameters low.
-With this novel strategy we can reduce attack surface and harden the resulting \gls{ids} while retaining similar classification performance like for the original \gls{ml} model.
+With this novel strategy we can reduce the attack surface and harden the resulting \gls{ids} while retaining similar classification performance.
+% like for the original \gls{ml} model.
 
-\section{RNN-based classifier}
+\section{An \gls{rnn}-based classifier}
 
@@
 
-In this research, we use the \textit{CIC-IDS-2017} \cite{sharafaldin_toward_2018} and \textit{UNSW-NB15} \cite{moustafa_unsw-nb15:_2015} datasets which each include more than 2 million flows of network data, containing both benign traffic and a large number of different attacks.
-Attacks contained in CIC-IDS-2017 are shown in \autoref{tab:occurence}.
-
-We implement a three-layer LSTM-based classifier with 512 neurons at each layer. As the input features we use \begin{itemize*}
-\item source port
-\item destination port
-\item protocol identifier
-\item packet sizes
-\item \glspl{iat} of packets
-\item packet directions (if packet goes there or back)
-\item all \gls{tcp} flags (0 if flow is not \gls{tcp})
-\end{itemize*}.
+We implemented a three-layer LSTM-based classifier with 512 neurons at each layer. As the input features we use 
+source port, destination port, protocol identifier, packet length, \gls{iat} to the previous packet in the flow, packet direction (if packet goes there or back) and all \gls{tcp} flags (0 if the flow is not \gls{tcp}).
+% \begin{itemize*}
+% \item source port
+% \item destination port
+% \item protocol identifier
+% \item packet sizes
+% \item \glspl{iat} of packets
+% \item packet directions (if packet goes there or back)
+% \item all \gls{tcp} flags (0 if the flow is not \gls{tcp})
+% \end{itemize*}.
 We omitted \gls{ttl} values, as they are likely to lead to unwanted prediction behaviour \cite{bachl_walling_2019}.
@@
 
+For evaluation, we use the \textit{CIC-IDS-2017} \cite{sharafaldin_toward_2018} and \textit{UNSW-NB15} \cite{moustafa_unsw-nb15:_2015} datasets which each include more than 2 million flows of network data, containing both benign traffic and a large number of different attacks.
+Attacks contained in CIC-IDS-2017 are shown in \autoref{tab:occurence}.
 \autoref{tab:performance_results} shows that our classifier achieves an accuracy that is similar to previous work based on these datasets~\cite{meghdouri_analysis_2018,bachl_walling_2019}.
@@
 
-As finding adversarial flows is particularly difficult in our case for the reasons we discussed above, for the present research we preferred methods which produces high-quality adversarial samples while not being as fast as other methods.
-We implemented the several methods for generating adversarial flows.
+As finding adversarial flows is particularly difficult in our case for the reasons we discussed above, for the present research we preferred methods  producing high-quality adversarial samples while not being as fast as other methods.
+We implemented the following \gls{aml} methods.
 
@@
 This allows the conclusion that if an attacker is able to change the first few packets, she could possibly change the decision of the classifier.
-We will further investigate this question in section \autoref{sec:adv}.
+We will further investigate this question in \autoref{sec:adv}.
 

### 2019-12-07 ex3/paper/paper.tex
Addresse Max' comment

@@
 \subsection{Evaluation}
-% \begin{figure*}
-% \includegraphics[width=0.45\textwidth]{adv_plots/cwl1.pdf}
-% \includegraphics[width=0.45\textwidth]{adv_plots/cwl2.pdf}
-% \caption{Performance of the \gls{cw} with different $\epsilon$ values for $\ell_1$ and $\ell_2$ distances.}
-% \label{fig:adv_cw}
-% \end{figure*}
-\todo{Max: Fix me. Don't know how the reference became invalid...}
+\begin{figure*}
+\includegraphics[width=0.45\textwidth]{adv_plots/cwl1.pdf}
+\includegraphics[width=0.45\textwidth]{adv_plots/cwl2.pdf}
+\caption{Performance of the \gls{cw} with different $\epsilon$ values for $\ell_1$ and $\ell_2$ distances.}
+\label{fig:adv_cw}
+\end{figure*}
+\todo{compress figures somehow}
 \autoref{fig:adv_cw} depicts the ratio of attack flows which were successfully misclassified as benign with respect to the final packet along with the resulting distance from the original sample.
@@
 
-% \begin{figure*}
-% \includegraphics[width=0.45\textwidth]{adv_plots/fgsm.pdf}
-% \includegraphics[width=0.45\textwidth]{adv_plots/l_inf_pgd.pdf}
-% \caption{Performance of FSGM and $\ell_\infty$-bounded \gls{pgd} for different bounds.}
-% \label{fig:fgsm}
-% \end{figure*}
+\begin{figure*}
+\includegraphics[width=0.45\textwidth]{adv_plots/fgsm.pdf}
+\includegraphics[width=0.45\textwidth]{adv_plots/l_inf_pgd.pdf}
+\caption{Performance of FSGM and $\ell_\infty$-bounded \gls{pgd} for different bounds.}
+\label{fig:fgsm}
+\end{figure*}
 
@@
 
-Furthermore, let $f_A(x)$  denote the joint probability \todo{Max: Ist joint probability nicht immer wenn man mehrere Ereignisse hat?
-Ich hätte hier probability density function geschrieben...} for classification as attack with a feature value $x$.
-Accuracy then is $\int f_A(x) dx$ for attacks and $1-\int f_A(x) dx$ for benign traffic.
+Furthermore, let $f(A,x)$  denote the joint probability
+%\todo{Max: Ist joint probability nicht immer wenn man mehrere Ereignisse hat? Ich hätte hier probability density function geschrieben...}
+%AH. changed notation and wording
+for classification as attack and a feature value $x$. Accuracy then is $\int f(A,x) dx$ for attacks and $1-\int f(A,x) dx$ for benign traffic.
 \autoref{fig:mutinfo_example} shows two different distributions

### 2019-12-07 ex3/mutinfo_example.py

@@ -1,0 +1,34 @@
+#!/usr/bin/env python3
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+plt.rcParams["font.family"] = "serif"
+
+resolution = 10000
+x = np.arange(0,1,1/resolution)
+
+y = (np.mod(np.arange(resolution), np.ceil(resolution/3.5)) < np.ceil(resolution/3.5)/2).astype(int)
+y2 = y.astype(float)
+y2[np.argmin(y):] = .5
+
+plt.figure(figsize=(5,1.3))
+plt.subplot(1,2,1)
+plt.plot(x,y2, c='k')
+plt.xlim((0,1))
+plt.yticks([0,.5,1])
+plt.ylim((-0.1,1.1))
+plt.xlabel('$x$')
+plt.ylabel('$f(A,x)$')
+
+plt.subplot(1,2,2)
+plt.plot(x,y, c='k')
+plt.xlim((0,1))
+plt.yticks([0,.5,1])
+plt.ylim((-0.1,1.1))
+plt.xlabel('$x$')
+plt.ylabel('$f(A,x)$')
+
+plt.tight_layout()
+plt.savefig('plots/mutinfo_example.pdf', bbox_inches = 'tight', pad_inches = 0)
+plt.show()

### 2019-12-07 ex3/paper/paper.tex
Add a few figures

@@
 \begin{figure*}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_0_Botnet-ARES"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_14_Web Attack-XSS"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_pred_plots2_outcomes_0_3.pickle_5_DoS - DDoS-DoS Hulk"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_pred_plots2_outcomes_0_3.pickle_6_DoS - DDoS-DoS Slowhttptest"}.pdf}
+% \includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_14_Web Attack-XSS"}.pdf}
+% \includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
 
@@
 For the majority of attacks, confidence increases in the first few steps and then stays almost constant at 1.}
+\todo{as we cannot depict all figure, one averaged figure would be more meaningful}
 \label{fig:confidence}

### 2019-12-07 ex3/barplot.py
Tiny figure change

@@ -40,7 +40,7 @@ plt.legend()
 plt.xticks(x, [group_names[i] for i in order], rotation=45, horizontalalignment='right')
 for tick in plt.gca().xaxis.get_major_ticks():
 	label = tick.label1
-	label.set_transform(label.get_transform() + Affine2D().translate(10,0))
+	label.set_transform(label.get_transform() + Affine2D().translate(8,0))
 plt.ylabel('Normalized metric')
 plt.tight_layout()
 

### 2019-12-08 ex3/paper/paper.tex
Left out some figures/tables

@@
 which significantly and successfully reduces the attack surface and improves \gls{ars}.
+
+\todo{Max: Abstract too long}
 
@@
 \includegraphics[width=0.45\textwidth]{adv_plots/l_inf_pgd.pdf}
-\caption{Performance of FSGM and $\ell_\infty$-bounded \gls{pgd} for different bounds.}
+\caption{Performance of FSGM and $\ell_\infty$-bounded \gls{pgd} for different bounds.
+\todo{Max: Omit}}
 \label{fig:fgsm}
@@
 
-Finally, we investigated what adversarial flows look like and which features are manipulated to circumvent detection by the neural network.
-
-To this end, we generated adversarial flows with the \gls{cw} with $\ell_2$ distance and a tradeoff of 0.5.
+%Finally, we investigated what adversarial flows look like and which features are manipulated to circumvent detection by the neural network.
+%
+%To this end, we generated adversarial flows with the \gls{cw} with $\ell_2$ distance and a tradeoff of 0.5.
 \autoref{adv_plots} depicts how the method manipulates flows for some exemplary attack categories.
 Depicted is the mean of both unmodified flows and adversarial flows  of samples of individual attack families considering both \gls{iat} and packet length.
-
-The plots show that primarily features at the beginning of flows are modified.
+%
+%The plots show that primarily features at the beginning of flows are modified.
 It is remarkable that it seems most promising to modify single packets while leaving most of the flow unchanged.
@@
 
-\begin{figure*}[p]
-\includegraphics[width=0.48\textwidth]{../plots/plot_adv/1.pdf}
-\includegraphics[width=0.48\textwidth]{../plots/plot_adv/2.pdf}
-\includegraphics[width=0.48\textwidth]{../plots/plot_adv/3.pdf}
-\includegraphics[width=0.48\textwidth]{../plots/plot_adv/4.pdf}
-\includegraphics[width=0.48\textwidth]{../plots/plot_adv/5.pdf}
-\includegraphics[width=0.48\textwidth]{../plots/plot_adv/6.pdf}
-\caption{Successful manipulations performed by the \gls{cw}.}
-\label{adv_plots}
-\end{figure*}
+%\begin{figure*}[p]
+%\includegraphics[width=0.48\textwidth]{../plots/plot_adv/1.pdf}
+%\includegraphics[width=0.48\textwidth]{../plots/plot_adv/2.pdf}
+%\includegraphics[width=0.48\textwidth]{../plots/plot_adv/3.pdf}
+%\includegraphics[width=0.48\textwidth]{../plots/plot_adv/4.pdf}
+%\includegraphics[width=0.48\textwidth]{../plots/plot_adv/5.pdf}
+%\includegraphics[width=0.48\textwidth]{../plots/plot_adv/6.pdf}
+%\caption{Successful manipulations performed by the \gls{cw}. \todo{Max: Just keep one plot I'd say. Maximum two.}}
+%\label{adv_plots}
+%\end{figure*}
 
@@
 \end{equation}
-
-\begin{figure*}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_8_DoS - DDoS-Heartbleed"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_10_Normal"}.pdf}
-
-\caption{Examples of conditional PD plots for source and destination port.
+%
+%\begin{figure*}
+%\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_8_DoS - DDoS-Heartbleed"}.pdf}
+%\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_10_Normal"}.pdf}
+%
+%\caption{Examples of conditional PD plots for source and destination port.
 We only analyze these features with the \gls{pdp} since it is not straightforward to define the \gls{pdp} for features that can vary for each packet in a flow.
-The y-axis shows the probability for malicious traffic.
-}
-\label{fig:pdp}
-\end{figure*}
-
+The y-axis shows the probability for malicious traffic.}
+%\todo{Max: Remove these? Not really interesting or related to the rest of the paper. Refer to other paper where we talk about this already.}
+%\label{fig:pdp}
+%\end{figure*}
+%
 In this variant of the \gls{pdp} that we call conditional \gls{pdp}, we don't take the whole dataset for the computation for the \gls{pdp} but only a subset that belongs to a certain class.
 Specifically, we take each attack class and compute the \gls{pdp} only over samples of this attack class.
-The results in \autoref{fig:pdp} show that for some attack classes the port numbers play an important role.
+The results (omitted for brevity) show that for some attack classes the port numbers play an important role.
 When looking at the class of normal traffic (non-attack traffic) it becomes apparent that traffic destined to a high source port is generally indicative of an attack.
@@
 A known issue about \glspl{pdp} in general is that dependencies of the feature under investigation and the remaining variables are not considered, hence potentially querying the neural network for feature combinations which cannot possibly occur.
+For a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.
 
-As apparent from \autoref{eq:pdp}, for a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.
-
-For this reason, Accumulated Local Effects (ALE) plots have been proposed in literature.
+%For this reason, Accumulated Local Effects (ALE) plots have been proposed in literature.
 The basic idea for ALE is that instead of using all samples, only the closest samples are used for assessing the effects of one feature.
@@
 
-\begin{figure*}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_pred_plots2_outcomes_0_3.pickle_5_DoS - DDoS-DoS Hulk"}.pdf}
+\begin{figure}
+%\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_pred_plots2_outcomes_0_3.pickle_5_DoS - DDoS-DoS Hulk"}.pdf}
 \includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_pred_plots2_outcomes_0_3.pickle_6_DoS - DDoS-DoS Slowhttptest"}.pdf}
@@
 
-\caption{Median confidence per time step and first and third quartiles for selected attack types.
-For the majority of attacks, confidence increases in the first few steps and then stays almost constant at 1.}
-\todo{as we cannot depict all figure, one averaged figure would be more meaningful}
+\caption{Median confidence per time step and first and third quartiles. For the majority of attacks types, confidence increases in the first few steps and then stays almost constant at 1.}
 \label{fig:confidence}
-\end{figure*}
+\end{figure}
 
@@
 \includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_2_Brute Force-SSH-Patator"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_6_DoS - DDoS-DoS Slowhttptest"}.pdf}
+%\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_6_DoS - DDoS-DoS Slowhttptest"}.pdf}
 \includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
+%\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
 
 \caption{The solid line shows the median value of the feature for all flows of a specific attack type. The shaded region are the first and third quartile.}
+\todo{Max: Maybe here two figures are enough?}
 \label{fig:plot_features}
@@
 Both approaches lead to complete resistance to adversarial samples (as we left out all manipulable features), except for botnets, which can still operate when only leaving out manipulable features in one direction.
-The results in \autoref{tab:performance_results_no_manipulable} show that -- surprisingly -- there is only a small difference in classification performance when only computing the accuracy for the last packet of each flow (flow accuracy).
-However, when looking at the packet accuracy, which considers all packets of all flows, there is a significant difference.
-Thus, apparently the \gls{iat} and packet size are especially important for determining whether a flow is malicious in the first packets of a flow.
+The results show that -- surprisingly -- there is only a small difference in classification performance when only computing the accuracy for the last packet of each flow (flow accuracy): Flow accuracy stays over 99.93\%.
+However, packet accuracy is only 98\% when leaving out the features in one direction and 96.7\% when leaving them out in both directions, Thus, apparently the \gls{iat} and packet size are especially important for determining whether a flow is malicious in the first packets of a flow.
 
@@
 
+\todo{Max: Insert results that show that accuracy on normal data is still ok after adversarial training.}
+
 \begin{figure}
@@
 
-\begin{table*}
-\caption{Comparing performance with all features, only safe features (\gls{iat} and packet size removed) and safe features in one direction (\gls{iat} and packet size removed for packets coming from source).
+%\begin{table*}
+%\caption{Comparing performance with all features, only safe features (\gls{iat} and packet size removed) and safe features in one direction (\gls{iat} and packet size removed for packets coming from source).
 \textit{packets} means that we consider the neural network output of all packets of the flow while \textit{flows} means that we only consider the last packet.
-At the last packet the classifier is usually already a lot more confident than at the first packets of a flow and thus metrics are always better in this case.}  \label{tab:performance_results_no_manipulable}
-\newcommand{\cmidrulespace}{6pt}
-\centering
-\begin{tabular}{l l l l l l l} \toprule
-& \multicolumn{2}{l}{all} & \multicolumn{2}{l}{safe features} & \multicolumn{2}{l}{safe features one direction} \\
-\cmidrule(r){2-3} \cmidrule(lr){4-5} \cmidrule(l){6-7}
-& packets & flows & packets & flows & packets & flows \\
-\midrule
-Accuracy & 0.9857054 & 0.99472587 & 0.96678023 & 0.99337596 & 0.98056073 & 0.99353386 \\
-Precision & 0.94963342 & 0.9894422 & 0.87175275 & 0.98798596 & 0.91359787 & 0.98994595 \\
-Recall & 0.96735895 & 0.98972051 & 0.94374958 & 0.98581328 & 0.97836672 & 0.9844478 \\
-F1 & 0.95841423 & 0.98958134 & 0.90632359 & 0.98689842 & 0.94487366 & 0.98718922 \\
-Youden & 0.95682951 & 0.98614231 & 0.91525628 & 0.98175163 & 0.95937772 & 0.9810602 \\
-\bottomrule
-\end{tabular}
-\end{table*}
+At the last packet the classifier is usually already a lot more confident than at the first packets of a flow and thus metrics are always better in this case.
+\todo{Max: Is this that interesting?
+Remove tables and inline the most interesting results?}}  \label{tab:performance_results_no_manipulable}
+%\newcommand{\cmidrulespace}{6pt}
+%\centering
+%\begin{tabular}{l l l l l l l} \toprule
+%& \multicolumn{2}{l}{all} & \multicolumn{2}{l}{safe features} & \multicolumn{2}{l}{safe features one direction} \\
+%\cmidrule(r){2-3} \cmidrule(lr){4-5} \cmidrule(l){6-7}
+%& packets & flows & packets & flows & packets & flows \\
+%\midrule
+%Accuracy & 0.9857054 & 0.99472587 & 0.96678023 & 0.99337596 & 0.98056073 & 0.99353386 \\
+%Precision & 0.94963342 & 0.9894422 & 0.87175275 & 0.98798596 & 0.91359787 & 0.98994595 \\
+%Recall & 0.96735895 & 0.98972051 & 0.94374958 & 0.98581328 & 0.97836672 & 0.9844478 \\
+%F1 & 0.95841423 & 0.98958134 & 0.90632359 & 0.98689842 & 0.94487366 & 0.98718922 \\
+%Youden & 0.95682951 & 0.98614231 & 0.91525628 & 0.98175163 & 0.95937772 & 0.9810602 \\
+%\bottomrule
+%\end{tabular}
+%\end{table*}
 
-\section{Conclusions}
+\section{Conclusion}
 
@@
 We show that adversarial samples are possible even though the data we use are very different from those that are usually used when finding adversarial samples.
+
+\todo{Max: Need more conclusion. Shouldn't be the same like abstract and intro.}
 

### 2019-12-08 ex3/learn.py
Fixed adv_until_less_than_half bug

@@ -1203,10 +1203,11 @@ def adv_until_less_than_half():
 
 				successfully_changed_flows_mask = (np.round(numpy_sigmoid(np.array([item[-1] for item in prev_results[i][attack_index]]))) == 0).flatten()
 				# TODO: l1 or l2 norm?
-				distances = np.array([np.linalg.norm((orig_flows[attack_index][flow_index]-flow).flatten(), ord=2) for flow_index, flow in enumerate(prev_flows[i][attack_index])])
+				# guess it makes sense to use the same distance metric as in training
+				distances = np.array([np.linalg.norm((orig_flows[attack_index][flow_index]-flow).flatten(), ord=opt.order) for flow_index, flow in enumerate(prev_flows[i][attack_index])])
 				argsorted_distances = np.argsort(distances)
 				correct_indices = argsorted_distances[successfully_changed_flows_mask[argsorted_distances]]
-				lower_part = correct_indices[:int(math.ceil(len(distances)*min(ratio, THRESHOLD)))]
+				lower_part = correct_indices[:int(math.ceil(len(distances)*min(1-ratio, THRESHOLD)))]
 
 				distances_per_packet = [dist/len(flow) for dist, flow in zip(distances[lower_part], prev_flows[i][attack_index])]
 				distances_flows[attack_index] = float(np.mean(distances[lower_part]))

### 2019-12-08 ex3/paper/paper.tex
Updated figures

@@
 
-\begin{table*}
-\caption{Attack success ratios per attack family. Depicted are flow detection accuracies for adversarial flows (left numbers) and flow detection accuracies for unmodified flows (right numbers).}
-\label{tab:adv_per_family}
-\centering
-\begin{tabular}{llllllllll}
-\toprule
-Attack type & \multicolumn{2}{l}{FGSM} & \multicolumn{2}{l}{Carlini-Wagner, $\ell_1$} & \multicolumn{2}{l}{Carlini-Wagner, $\ell_2$} & \multicolumn{2}{l}{$\ell_\infty$-bounded \gls{pgd}} \\
-\cmidrule(r){2-3} \cmidrule(lr){4-5} \cmidrule(l){6-7} \cmidrule(l){8-9}
-& adv. & orig. & adv. & orig. & adv. & orig. & adv. & orig. \\
-\midrule
-Botnet:ARES	&	0.0000 & 0.0000	&	0.0000 & 0.0000	&	0.0000 & 0.0000	&	0.0000 & 0.0000	\\
-Brute Force:FTP-Patator	&	0.2800 & 0.5800	&	0.1600 & 0.5800	&	0.1600 & 0.5800	&	0.2800 & 0.5800	\\
-Brute Force:SSH-Patator	&	0.8300 & 1.0000	&	0.0000 & 1.0000	&	0.0000 & 1.0000	&	0.8900 & 1.0000	\\
-DDoS:LOIT	&	0.8300 & 0.8600	&	0.6400 & 0.8600	&	0.5600 & 0.8600	&	0.8200 & 0.8600	\\
-DoS / DDoS:DoS GoldenEye	&	0.8900 & 1.0000	&	0.5000 & 1.0000	&	0.4200 & 1.0000	&	0.9200 & 1.0000	\\
-DoS / DDoS:DoS Hulk	&	0.6700 & 1.0000	&	0.5400 & 1.0000	&	0.4700 & 1.0000	&	1.0000 & 1.0000	\\
-DoS / DDoS:DoS Slowhttptest	&	0.8700 & 1.0000	&	0.4100 & 1.0000	&	0.1900 & 1.0000	&	0.7400 & 1.0000	\\
-DoS / DDoS:DoS slowloris	&	0.4400 & 0.9700	&	0.3100 & 0.9700	&	0.1300 & 0.9700	&	0.9200 & 0.9700	\\
-DoS / DDoS:Heartbleed	&	0.0000 & 0.0000	&	0.0000 & 0.0000	&	0.0000 & 0.0000	&	0.0000 & 0.0000	\\
-Infiltration:Dropbox download 	&	0.6100 & 0.9600	&	0.4300 & 0.9600	&	0.3800 & 0.9600	&	0.7400 & 0.9600	\\
-PortScan:PortScan - Firewall off	&	0.6800 & 0.9100	&	0.3500 & 0.9100	&	0.3500 & 0.9100	&	0.7200 & 0.9100	\\
-PortScan:PortScan - Firewall on	&	0.7400 & 0.8700	&	0.6100 & 0.8700	&	0.5300 & 0.8700	&	0.7300 & 0.8700	\\
-Web Attack:Sql Injection	&	0.0000 & 0.6923	&	0.0000 & 0.6923	&	0.0000 & 0.6923	&	0.0000 & 0.6923	\\
-Web Attack:XSS	&	0.6700 & 0.9500	&	0.7200 & 0.9500	&	0.6200 & 0.9500	&	0.6700 & 0.9500	\\
-\bottomrule
+% \begin{table*}
+% \caption{Attack success ratios per attack family. Depicted are flow detection accuracies for adversarial flows (left numbers) and flow detection accuracies for unmodified flows (right numbers).}
+% \label{tab:adv_per_family}
+% \centering
+% \begin{tabular}{llllllllll}
+% \toprule
+% Attack type & \multicolumn{2}{l}{FGSM} & \multicolumn{2}{l}{Carlini-Wagner, $\ell_1$} & \multicolumn{2}{l}{Carlini-Wagner, $\ell_2$} & \multicolumn{2}{l}{$\ell_\infty$-bounded \gls{pgd}} \\
+% \cmidrule(r){2-3} \cmidrule(lr){4-5} \cmidrule(l){6-7} \cmidrule(l){8-9}
+% & adv. & orig. & adv. & orig. & adv. & orig. & adv. & orig. \\
+% \midrule
+% Botnet:ARES	&	0.0000 & 0.0000	&	0.0000 & 0.0000	&	0.0000 & 0.0000	&	0.0000 & 0.0000	\\
+% Brute Force:FTP-Patator	&	0.2800 & 0.5800	&	0.1600 & 0.5800	&	0.1600 & 0.5800	&	0.2800 & 0.5800	\\
+% Brute Force:SSH-Patator	&	0.8300 & 1.0000	&	0.0000 & 1.0000	&	0.0000 & 1.0000	&	0.8900 & 1.0000	\\
+% DDoS:LOIT	&	0.8300 & 0.8600	&	0.6400 & 0.8600	&	0.5600 & 0.8600	&	0.8200 & 0.8600	\\
+% DoS / DDoS:DoS GoldenEye	&	0.8900 & 1.0000	&	0.5000 & 1.0000	&	0.4200 & 1.0000	&	0.9200 & 1.0000	\\
+% DoS / DDoS:DoS Hulk	&	0.6700 & 1.0000	&	0.5400 & 1.0000	&	0.4700 & 1.0000	&	1.0000 & 1.0000	\\
+% DoS / DDoS:DoS Slowhttptest	&	0.8700 & 1.0000	&	0.4100 & 1.0000	&	0.1900 & 1.0000	&	0.7400 & 1.0000	\\
+% DoS / DDoS:DoS slowloris	&	0.4400 & 0.9700	&	0.3100 & 0.9700	&	0.1300 & 0.9700	&	0.9200 & 0.9700	\\
+% DoS / DDoS:Heartbleed	&	0.0000 & 0.0000	&	0.0000 & 0.0000	&	0.0000 & 0.0000	&	0.0000 & 0.0000	\\
+% Infiltration:Dropbox download 	&	0.6100 & 0.9600	&	0.4300 & 0.9600	&	0.3800 & 0.9600	&	0.7400 & 0.9600	\\
+% PortScan:PortScan - Firewall off	&	0.6800 & 0.9100	&	0.3500 & 0.9100	&	0.3500 & 0.9100	&	0.7200 & 0.9100	\\
+% PortScan:PortScan - Firewall on	&	0.7400 & 0.8700	&	0.6100 & 0.8700	&	0.5300 & 0.8700	&	0.7300 & 0.8700	\\
+% Web Attack:Sql Injection	&	0.0000 & 0.6923	&	0.0000 & 0.6923	&	0.0000 & 0.6923	&	0.0000 & 0.6923	\\
+% Web Attack:XSS	&	0.6700 & 0.9500	&	0.7200 & 0.9500	&	0.6200 & 0.9500	&	0.6700 & 0.9500	\\
+% \bottomrule
 
-\end{tabular}
+% \end{tabular}
 
-\todo{In the adv table, round the numbers!!! \\
-For Max: Introduce two-level table\\
-What's the tradeoff for CW?}
-\end{table*}
+% \todo{In the adv table, round the numbers!!! \\
+% For Max: Introduce two-level table\\
+% What's the tradeoff for CW?}
+% \end{table*}
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{../plots/adv_comparison/adv_comparison_17.pdf}
+\label{fig:adv_per_family}
+\caption{Attack success ratios per attack family. Depicted are flow detection accuracies for adversarial flows and for unmodified flows.}
+\todo{update numbers, shorten attack names, What's the tradeoff for CW?}
+\end{figure}
 
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{../plots/adv_training/avdistance.pdf}
-\caption{Distance of adversarial samples from original attack samples throughout adversarial training.}
+\includegraphics[width=\columnwidth]{../plots/ars/ars_15.pdf}
+\caption{\gls{ars} during adversarial training for UNSW-NB15.}
 \label{fig:avdistance}

### 2019-12-08 ex3/auxplot.py

@@ -1,0 +1,90 @@
+#!/usr/bin/env python3
+
+# An ad-hoc script for plotting tables copy-pasted from LibreOffice Calc
+
+import fileinput
+import numpy as np
+import matplotlib.pyplot as plt
+import sys
+from matplotlib.transforms import Affine2D
+
+group_names = []
+values = []
+
+plt.rcParams['font.family'] = 'serif'
+plt.rcParams['savefig.format'] = 'pdf'
+colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
+
+for line in fileinput.input('-'):
+	field = line.rstrip().split('\t')
+	group_names.append(field[0])
+	values.append(field[1:])
+
+feature_names = values[0]
+group_names = group_names[1:]
+values = values[1:]
+
+#  print (group_names)
+#  print (values)
+values = np.array(values, dtype=float)
+
+def feat_imp():
+	plt.figure(figsize=(5,4))
+	values[values < 0] = 0
+	values /= np.sum(values,axis=0)
+
+	width = 0.95 / values.shape[1]
+
+	order = np.argsort(-np.mean(values, axis=1))
+	x = np.arange(len(group_names))
+	for i in range(values.shape[1]):
+		plt.bar(x +width*(i-values.shape[1]/2), values[order,i], width, label=feature_names[i])
+	plt.legend()
+	plt.xticks(x, [group_names[i] for i in order], rotation=45, horizontalalignment='right')
+	for tick in plt.gca().xaxis.get_major_ticks():
+		label = tick.label1
+		label.set_transform(label.get_transform() + Affine2D().translate(8,0))
+	plt.ylabel('Normalized metric')
+	plt.tight_layout()
+
+	plt.savefig(sys.stdout.buffer, bbox_inches = 'tight', pad_inches = 0)
+	plt.show()
+	
+def adv_results():
+	global values
+	plt.figure(figsize=(5,4))
+	values = np.reshape(values, (-1,values.shape[1]//2,2))
+
+	order = np.argsort(np.mean(values[:,:,1], axis=1))
+	width = 0.95 / values.shape[1]
+
+	x = np.arange(len(group_names))
+	for i in range(values.shape[1]):
+		plt.bar(x +width*(i-values.shape[1]/2), values[order,i,1], width, color='gray', alpha=.5, **({'label': 'Original'} if i==0 else {}))
[…]

### 2019-12-08 ex3/barplot.py

@@ -1,49 +1,0 @@
-#!/usr/bin/env python3
-
-# An ad-hoc script for plotting tables copy-pasted from LibreOffice Calc
-
-import fileinput
-import numpy as np
-import matplotlib.pyplot as plt
-import sys
-from matplotlib.transforms import Affine2D
-
-group_names = []
-values = []
-
-plt.rcParams['font.family'] = 'serif'
-plt.rcParams['savefig.format'] = 'pdf'
-plt.figure(figsize=(5,4))
-
-for line in fileinput.input():
-	field = line.rstrip().split('\t')
-	group_names.append(field[0])
-	values.append(field[1:])
-
-feature_names = values[0]
-group_names = group_names[1:]
-values = values[1:]
-
-print (group_names)
-print (values)
-values = np.array(values, dtype=float)
-values[values < 0] = 0
-values /= np.sum(values,axis=0)
-
-width = 0.95 / values.shape[1]
-
-order = np.argsort(-np.mean(values, axis=1))
-x = np.arange(len(group_names))
-for i in range(values.shape[1]):
-	plt.bar(x +width*(i-values.shape[1]/2), values[order,i], width, label=feature_names[i])
-plt.legend()
-plt.xticks(x, [group_names[i] for i in order], rotation=45, horizontalalignment='right')
-for tick in plt.gca().xaxis.get_major_ticks():
-	label = tick.label1
-	label.set_transform(label.get_transform() + Affine2D().translate(8,0))
-plt.ylabel('Normalized metric')
-plt.tight_layout()
-
-plt.savefig(sys.stdout.buffer, bbox_inches = 'tight', pad_inches = 0)
-plt.show()
-

### 2019-12-09 ex3/auxplot.py
Modified importance plots

@@ -28,12 +28,13 @@ values = values[1:]
 #  print (values)
 values = np.array(values, dtype=float)
 
-def feat_imp():
+def importance():
+	global values
 	plt.figure(figsize=(5,4))
 	values[values < 0] = 0
 	values /= np.sum(values,axis=0)
 
-	width = 0.95 / values.shape[1]
+	width = 1/(1+values.shape[1])
 
 	order = np.argsort(-np.mean(values, axis=1))
 	x = np.arange(len(group_names))
@@ -43,12 +44,11 @@ def feat_imp():
 	plt.xticks(x, [group_names[i] for i in order], rotation=45, horizontalalignment='right')
 	for tick in plt.gca().xaxis.get_major_ticks():
 		label = tick.label1
-		label.set_transform(label.get_transform() + Affine2D().translate(8,0))
+		# don't know how to specify translation in axis coordinates, so pdf output
+		# looks slightly different from plot
+		label.set_transform(label.get_transform() + Affine2D().translate(5,0))
 	plt.ylabel('Normalized metric')
 	plt.tight_layout()
-
-	plt.savefig(sys.stdout.buffer, bbox_inches = 'tight', pad_inches = 0)
-	plt.show()
 	
 def adv_results():
 	global values
@@ -71,8 +71,6 @@ def adv_results():
 	plt.ylabel('Accuracy')
 	plt.tight_layout()
 
-	plt.savefig(sys.stdout.buffer, bbox_inches = 'tight', pad_inches = 0)
-	plt.show()
 	
 def ars():
 	plt.figure(figsize=(5,2))
@@ -80,11 +78,11 @@ def ars():
 	plt.legend(feature_names, ncol=2)
 	plt.xlabel('Training duration in epochs')
 	plt.ylabel('ARS')
-	plt.tight_layout()
 	ylim1,ylim2 = plt.ylim()
 	plt.ylim((ylim1,ylim2+20)) # move plots away from legend
-	plt.savefig(sys.stdout.buffer, bbox_inches = 'tight', pad_inches = 0)
-	plt.show()
+	plt.tight_layout()
 	
 globals()[sys.argv[1]]()
-
+if len(sys.argv) > 2:
+	plt.savefig(sys.argv[2], bbox_inches = 'tight', pad_inches = 0)
+plt.show()

### 2019-12-09 ex3/auxplot.py
Modified importance plots

@@ -35,11 +35,16 @@ def importance():
 	values /= np.sum(values,axis=0)
 
 	width = 1/(1+values.shape[1])
+	width = 0.85 / values.shape[1]
 
 	order = np.argsort(-np.mean(values, axis=1))
 	x = np.arange(len(group_names))
+
 	for i in range(values.shape[1]):
-		plt.bar(x +width*(i-values.shape[1]/2), values[order,i], width, label=feature_names[i])
+		plt.bar(x +width*(i-values.shape[1]/2), values[order,i], width, alpha=0.9, label=feature_names[i])
+	for i in range(values.shape[1]):
+		plt.bar(x +width*(i-values.shape[1]/2), np.mean(values[order,:], axis=1), width, color='gray', alpha=0.5, **({'label': 'Mean'} if i==0 else {}), zorder=0)
+
 	plt.legend()
 	plt.xticks(x, [group_names[i] for i in order], rotation=45, horizontalalignment='right')
 	for tick in plt.gca().xaxis.get_major_ticks():

### 2019-12-09 ex3/auxplot.py
Fix xticks location for the importance plot

@@ -34,19 +34,20 @@ def importance():
 	values[values < 0] = 0
 	values /= np.sum(values,axis=0)
 
-	width = 1/(1+values.shape[1])
-	width = 0.85 / values.shape[1]
+	FACTOR = 0.85
+	# width = 1/(1+values.shape[1])
+	width = FACTOR / values.shape[1]
 
 	order = np.argsort(-np.mean(values, axis=1))
 	x = np.arange(len(group_names))
 
 	for i in range(values.shape[1]):
-		plt.bar(x +width*(i-values.shape[1]/2), values[order,i], width, alpha=0.9, label=feature_names[i])
+		plt.bar(x + width*i, values[order,i], width, alpha=0.9, label=feature_names[i])
 	for i in range(values.shape[1]):
-		plt.bar(x +width*(i-values.shape[1]/2), np.mean(values[order,:], axis=1), width, color='gray', alpha=0.5, **({'label': 'Mean'} if i==0 else {}), zorder=0)
+		plt.bar(x + width*i, np.mean(values[order,:], axis=1), width, color='gray', alpha=0.5, **({'label': 'Mean'} if i==0 else {}), zorder=0)
 
 	plt.legend()
-	plt.xticks(x, [group_names[i] for i in order], rotation=45, horizontalalignment='right')
+	plt.xticks(x + width*values.shape[1]/2-width*0.5, [group_names[i] for i in order], rotation=45, horizontalalignment='right')
 	for tick in plt.gca().xaxis.get_major_ticks():
 		label = tick.label1
 		# don't know how to specify translation in axis coordinates, so pdf output
@@ -54,7 +55,7 @@ def importance():
 		label.set_transform(label.get_transform() + Affine2D().translate(5,0))
 	plt.ylabel('Normalized metric')
 	plt.tight_layout()
-	
+
 def adv_results():
 	global values
 	plt.figure(figsize=(5,4))
@@ -76,7 +77,7 @@ def adv_results():
 	plt.ylabel('Accuracy')
 	plt.tight_layout()
 
-	
+
 def ars():
 	plt.figure(figsize=(5,2))
 	plt.plot(group_names, values)
@@ -86,7 +87,7 @@ def ars():
 	ylim1,ylim2 = plt.ylim()
 	plt.ylim((ylim1,ylim2+20)) # move plots away from legend
 	plt.tight_layout()
-	
+
 globals()[sys.argv[1]]()
[…]

### 2019-12-09 ex3/paper/paper.tex
Added a reference

--- 2019-12-08 ex3/paper/paper.tex
@@
 
-We think that the ``ideal'' feature importance method for our scenario are Shapley values \cite{shapley_value_1953}: To analyze feature importance with them, the feature under investigation is completely omitted from the dataset and a model is trained without it.
+We think that the ``ideal'' feature importance method for our scenario are Shapley values \cite{shapley_value_1952}: To analyze feature importance with them, the feature under investigation is completely omitted from the dataset and a model is trained without it.
 The resulting accuracy is then compared to the baseline accuracy with all features.
@@
 While this seems tractable, when also analyzing the feature importance of combinations of features, the number of models required becomes even larger: Out of $n$ features, $2^n$ unique subsets can be chosen, meaning that the computational demand rises exponentially.
+
+\todo{Add SHAP}
 
@@
 We thus recommend training two models: One regular one and one with feature dropout to use for the feature importance.
+Another method that uses dropout for feature importance is \cite{chang_dropout_2017}, which, however, applies a technique called \textit{Variational Dropout} to learn the optimal dropout rate for each feature.
+It tries to leave out as many features as possible and at the same time keep accuracy high.
+Thus important features are going to be left out less often and one can then extract the dropout probabilities for each feature and assess their significance based on them.
 

### 2019-12-10 ex3/learn.py
Made adv output less confusing

@@ -1134,8 +1134,9 @@ def adv_internal(in_training = False, tradeoff=None, lr=None, iterations=None, a
 		per_flow_accuracy = (np.mean(np.round(numpy_sigmoid(np.array([item[-1] for item in per_attack_results])))))
 		# TODO: l1 or l2 norm?
 		dist = np.array([np.linalg.norm((per_attack_orig_item-per_attack_modified_item).flatten(), ord=1).mean() for per_attack_orig_item, per_attack_modified_item in zip(per_attack_orig, per_attack_modified)]).mean()
+		linf_dist = np.array([np.linalg.norm((per_attack_orig_item-per_attack_modified_item).flatten(), ord='inf').mean() for per_attack_orig_item, per_attack_modified_item in zip(per_attack_orig, per_attack_modified)]).mean()
 
-		print("Attack type: {}; number of samples: {}, average dist: {}, packet confidence: {}/{}, flow confidence: {}/{}".format(reverse_mapping[attack_number], len(per_attack_results), dist, per_packet_accuracy, per_packet_orig_accuracy, per_flow_accuracy, per_flow_orig_accuracy))
+		print("Attack type: {}; number of samples: {}, average dist: {}, average inf. dist: {}, packet accuracy: {}/{}, flow accuracy: {}/{}".format(reverse_mapping[attack_number], len(per_attack_results), dist, linf_dist, per_packet_accuracy, per_packet_orig_accuracy, per_flow_accuracy, per_flow_orig_accuracy))
 
 	file_name = opt.dataroot[:-7]+"_adv_{}{}_outcomes_{}_{}.pickle".format(tradeoff, "_notBidirectional" if not opt.canManipulateBothDirections else "", opt.fold, opt.nFold)
 	results_dict = {"results_by_attack_number": results_by_attack_number, "orig_results_by_attack_number": orig_results_by_attack_number, "modified_flows_by_attack_number": modified_flows_by_attack_number, "orig_flows_by_attack_number": orig_flows_by_attack_number}

### 2019-12-10 ex3/paper/paper.tex
Updated adv results

--- 2019-12-08 ex3/paper/paper.tex
@@
 
+\begin{table}
+\centering
+\caption{Total flow accuracy for adversarial flows.}
+\label{tab:adv_accuracies}
+\begin{tabular}{lrrr} \toprule
+& CW & FGSM & PGD \\ \midrule
+CIC-IDS-2017 &	0.346	 & 0.445	& 0.539 \\ 
+UNSW-NB15 &	0.326&	0.629	& 0.774 \\
+\bottomrule
+\end{tabular}
+\end{table}
 It is interesting to note that for very large values of $\epsilon$ the success ratio starts to decrease again.This phenomenon might be due to numerical effects and might be preventable by a smaller learning rate.
@@
 \label{tab:feat_importance}
+\centering
 \subfloat[Random replacement method\label{fig:featRandom}
@@
 \toprule
-Feature & Acc.~drop \\ \midrule
+Feature & \hspace*{-4mm}Accuracy~drop \\ \midrule
 Protocol & 0.207 \\
@@
 Destination port & 0.071 \\
-Source port & 0.06 \\
+Source port & 0.060 \\
 RST Flag & 0.057 \\
@@
 \toprule
-Feature & Acc.~drop \\ \midrule
+Feature & \hspace*{-4mm}Accuracy~drop \\ \midrule
 Destination port & 0.025 \\

### 2019-12-10 ex3/auxplot.py

@@ -58,7 +58,7 @@ def importance():
 
 def adv_results():
 	global values
-	plt.figure(figsize=(5,4))
+	plt.figure(figsize=(5,3))
 	values = np.reshape(values, (-1,values.shape[1]//2,2))
 
 	order = np.argsort(np.mean(values[:,:,1], axis=1))
@@ -69,7 +69,7 @@ def adv_results():
 		plt.bar(x +width*(i-values.shape[1]/2), values[order,i,1], width, color='gray', alpha=.5, **({'label': 'Original'} if i==0 else {}))
 	for i in range(values.shape[1]):
 		plt.bar(x +width*(i-values.shape[1]/2), values[order,i,0], width, color=colors[i], label=feature_names[i*2].rstrip(' adv'))
-	plt.legend()
+	plt.legend(loc='lower center', bbox_to_anchor=(0.5,1), ncol=4)
 	plt.xticks(x, [group_names[i] for i in order], rotation=45, horizontalalignment='right')
 	for tick in plt.gca().xaxis.get_major_ticks():
 		label = tick.label1

### 2019-12-10 ex3/auxplot.py
Fixed xticks for adv_comparison

@@ -66,11 +66,11 @@ def adv_results():
 
 	x = np.arange(len(group_names))
 	for i in range(values.shape[1]):
-		plt.bar(x +width*(i-values.shape[1]/2), values[order,i,1], width, color='gray', alpha=.5, **({'label': 'Original'} if i==0 else {}))
+		plt.bar(x +width*i, values[order,i,1], width, color='gray', alpha=.5, **({'label': 'Original'} if i==0 else {}))
 	for i in range(values.shape[1]):
-		plt.bar(x +width*(i-values.shape[1]/2), values[order,i,0], width, color=colors[i], label=feature_names[i*2].rstrip(' adv'))
+		plt.bar(x +width*i, values[order,i,0], width, color=colors[i], label=feature_names[i*2].rstrip(' adv'))
 	plt.legend(loc='lower center', bbox_to_anchor=(0.5,1), ncol=4)
-	plt.xticks(x, [group_names[i] for i in order], rotation=45, horizontalalignment='right')
+	plt.xticks(x + width*values.shape[1]/2-width*0.5, [group_names[i] for i in order], rotation=45, horizontalalignment='right')
 	for tick in plt.gca().xaxis.get_major_ticks():
 		label = tick.label1
 		label.set_transform(label.get_transform() + Affine2D().translate(8,0))

### 2019-12-10 ex3/paper/paper.tex
Forgot to commit: Remove more figures

--- 2019-12-10 ex3/paper/paper.tex
@@
 
-\begin{table}
-\centering
-\caption{Total flow accuracy for adversarial flows.}
-\label{tab:adv_accuracies}
-\begin{tabular}{lrrr} \toprule
-& CW & FGSM & PGD \\ \midrule
-CIC-IDS-2017 &	0.346	 & 0.445	& 0.539 \\ 
-UNSW-NB15 &	0.326&	0.629	& 0.774 \\
-\bottomrule
-\end{tabular}
-\end{table}
 It is interesting to note that for very large values of $\epsilon$ the success ratio starts to decrease again.This phenomenon might be due to numerical effects and might be preventable by a smaller learning rate.
@@
 \label{tab:feat_importance}
-\centering
 \subfloat[Random replacement method\label{fig:featRandom}
@@
 \toprule
-Feature & \hspace*{-4mm}Accuracy~drop \\ \midrule
+Feature & Acc.~drop \\ \midrule
 Protocol & 0.207 \\
@@
 Destination port & 0.071 \\
-Source port & 0.060 \\
+Source port & 0.06 \\
 RST Flag & 0.057 \\
@@
 \toprule
-Feature & \hspace*{-4mm}Accuracy~drop \\ \midrule
+Feature & Acc.~drop \\ \midrule
 Destination port & 0.025 \\
@@
 
-\begin{figure*}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot2_adv/flows_pred_plots2_outcomes_flows_adv_1.0_notBidirectional_outcomes_0_3.pickle_0_3.pickle_0_Botnet-ARES"}.pdf}
+\begin{figure}
+%\includegraphics[width=0.48\textwidth]{{"../plots/plot2_adv/flows_pred_plots2_outcomes_flows_adv_1.0_notBidirectional_outcomes_0_3.pickle_0_3.pickle_0_Botnet-ARES"}.pdf}
 \includegraphics[width=0.48\textwidth]{{"../plots/plot2_adv/flows_pred_plots2_outcomes_flows_adv_1.0_notBidirectional_outcomes_0_3.pickle_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
@@
 
-\caption{The solid line shows the mean value of the feature for all flows of a specific attack type. The shaded region shows the change in confidence that occurs when the feature is varied.}
+\caption{The solid line shows the mean value of the feature for all flows of the attack type. The shaded region shows the change in confidence that occurs when the feature is varied.}
 \label{fig:pred_plots2}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{figure*}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_2_Brute Force-SSH-Patator"}.pdf}
+\begin{figure}
+%\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_2_Brute Force-SSH-Patator"}.pdf}
 %\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_6_DoS - DDoS-DoS Slowhttptest"}.pdf}
@@
 \label{fig:plot_features}
-\end{figure*}
+\end{figure}
 

### 2019-12-10 ex3/paper/paper.tex
Fix labels in figures and hopefully clarify the feature importance section

@@
 \begin{figure*}
-\includegraphics[width=0.45\textwidth]{adv_plots/cwl1.pdf}
-\includegraphics[width=0.45\textwidth]{adv_plots/cwl2.pdf}
+\includegraphics[width=0.49\textwidth]{adv_plots/cwl1.pdf}
+\includegraphics[width=0.49\textwidth]{adv_plots/cwl2.pdf}
 \caption{Performance of the \gls{cw} with different $\epsilon$ values for $\ell_1$ and $\ell_2$ distances.}
@@
 \includegraphics[width=\columnwidth]{../plots/adv_comparison/adv_comparison_17.pdf}
+\caption{Attack success ratios per attack family. Depicted are flow detection accuracies for adversarial flows and for unmodified flows.}
 \label{fig:adv_per_family}
-\caption{Attack success ratios per attack family. Depicted are flow detection accuracies for adversarial flows and for unmodified flows.}
 \todo{update numbers, shorten attack names, What's the tradeoff for CW?}
@@
 
-\autoref{fig:fgsm} depicts the performance for FGSM and $\ell_\infty$-bounded \gls{pgd}. As expected, while FGSM is the fastest of all algorithms, it delivers the worst performance.
+\autoref{tab:adv_accuracies} depicts the performance for FGSM and $\ell_\infty$-bounded \gls{pgd}. 
+As expected, while FGSM is the fastest of all algorithms, it delivers the worst performance. Of the investigated algorithms, thus the \gls{cw} with $\ell_2$ distance delivered the best performance.
 
-Of the investigated algorithms, thus the \gls{cw} with $\ell_2$ distance delivered the best performance.
-
-\autoref{tab:adv_per_family} depicts how successful the attack is for individual attack families.
+\autoref{fig:adv_per_family} depicts how successful the attack is for individual attack families.
 For this table we used a $\ell_\infty$ bound of 0.3 for FGSM and $\ell_\infty$-bounded \gls{pgd} and a tradeoff of 0.5 for Carlini-Wagner.
@@
 
-\begin{figure*}
-\includegraphics[width=0.45\textwidth]{adv_plots/fgsm.pdf}
-\includegraphics[width=0.45\textwidth]{adv_plots/l_inf_pgd.pdf}
-\caption{Performance of FSGM and $\ell_\infty$-bounded \gls{pgd} for different bounds.
-\todo{Max: Omit}}
-\label{fig:fgsm}
-\end{figure*}
+%\begin{figure*}
+%\includegraphics[width=0.49\textwidth]{adv_plots/fgsm.pdf}
+%\includegraphics[width=0.49\textwidth]{adv_plots/l_inf_pgd.pdf}
+%\caption{Performance of FSGM and $\ell_\infty$-bounded \gls{pgd} for different bounds.
+%\todo{Max: Omit}}
+%\label{fig:fgsm}
+%\end{figure*}
 
@@
 
-As \autoref{tab:adv_per_family} shows, for most attack types, most samples can be modified to look benign by an adversary by using \gls{cw}.
+As \autoref{fig:adv_per_family} shows, for most attack types, most samples can be modified to look benign by an adversary by using \gls{cw}.
 High accuracy for a particular attack category does not imply that adversarial attacks are more likely to fail for these attacks.
@@
 
- The most commonly used techniques for feature importance which are used by practitioners for \glspl{rnn} \cite{stackexchange_cross_validated_neural_2019} and \gls{dl} \cite{stackexchange_cross_validated_feature_2016} in general are based on adding noise to a particular feature and then evaluating the drop in accuracy that occurs.
+ The most commonly used techniques for feature importance which are used by practitioners for \glspl{rnn} \cite{stackexchange_cross_validated_neural_2019} and \gls{dl} \cite{molnar_interpretable_2019,stackexchange_cross_validated_feature_2016} in general are based on adding noise to a particular feature and then evaluating the drop in accuracy that occurs.
 We argue that it is hard to find out the ``correct'' intensity of noise to add, we slightly vary the proposed approach: We do not add noise drawn from a probability distribution but instead sample the value for a feature from the distribution of all values of this feature.
@@
 
-We think that the ``ideal'' feature importance method for our scenario are Shapley values \cite{shapley_value_1952}: To analyze feature importance with them, the feature under investigation is completely omitted from the dataset and a model is trained without it.
+%We think that the ``ideal'' feature importance method for our scenario are Shapley values \cite{shapley_value_1952}: To analyze feature importance with them, the feature under investigation is completely omitted from the dataset and a model is trained without it.
 The resulting accuracy is then compared to the baseline accuracy with all features.
@@
 
-\todo{Add SHAP}
-
-Considering the difficulties that arise from computing Shapley values, we propose a light-weight version of them, called \textit{feature dropout}: When training a model, for each sample, we leave out each feature with a certain probability by setting it to zero.
+We thus develop a more sound method compared to input perturbation called \textit{feature dropout}: When training a model, for each sample, we leave out each feature with a certain probability by setting it to zero.
 On average, one feature gets zeroed out but it is also possible that none or more than one are left out.
@@
 Thus important features are going to be left out less often and one can then extract the dropout probabilities for each feature and assess their significance based on them.
+Feature dropout can also be used as a building block for computing the accuracy of datapoints with missing features that is necessary to compute Shapley values \cite{shapley_value_1952} or KernelSHAP \cite{lundberg_unified_2017}.
 
@@
 %\includegraphics[width=0.48\textwidth]{{"../plots/plot2_adv/flows_pred_plots2_outcomes_flows_adv_1.0_notBidirectional_outcomes_0_3.pickle_0_3.pickle_0_Botnet-ARES"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot2_adv/flows_pred_plots2_outcomes_flows_adv_1.0_notBidirectional_outcomes_0_3.pickle_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
+\includegraphics[width=\columnwidth]{{"../plots/plot2_adv/flows_pred_plots2_outcomes_flows_adv_1.0_notBidirectional_outcomes_0_3.pickle_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
 %\includegraphics[width=0.48\textwidth]{{"../plots/plot2_adv/flows_pred_plots2_outcomes_flows_adv_1.0_notBidirectional_outcomes_0_3.pickle_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
@@
 %\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_6_DoS - DDoS-DoS Slowhttptest"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
+\includegraphics[width=\columnwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
 %\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}

### 2019-12-10 ex3/paper/paper.tex
Shorten references, change todos and change dropout feature importance

@@
 \caption{Performance of the \gls{cw} with different $\epsilon$ values for $\ell_1$ and $\ell_2$ distances.}
+\todo{Max: Need to be updated with current results that do not decrease at some point}
 \label{fig:adv_cw}
+\todo{compress figures somehow}
 \end{figure*}
-\todo{compress figures somehow}
 \autoref{fig:adv_cw} depicts the ratio of attack flows which were successfully misclassified as benign with respect to the final packet along with the resulting distance from the original sample.
@@
 
-%We think that the ``ideal'' feature importance method for our scenario are Shapley values \cite{shapley_value_1952}: To analyze feature importance with them, the feature under investigation is completely omitted from the dataset and a model is trained without it.
+%We think that the ``ideal'' feature importance method for our scenario are Shapley values \cite{shapley_value_1953}: To analyze feature importance with them, the feature under investigation is completely omitted from the dataset and a model is trained without it.
 The resulting accuracy is then compared to the baseline accuracy with all features.
@@
 This procedure is equivalent to using dropout \cite{srivastava_dropout:_2014} with probability $\frac{1}{n}$ before the first layer.
-The result is the classifier learning to deal with missing features.
+An important implementation detail here is that for each feature we add another input which is 1 if the feature is suppressed and 0 otherwise.
+This is necessary so that the neural network can distinguish between a feature missing and a feature genuinely being zero.
+The overall outcome is a classifier being able to deal with missing features.
 As it can also happen that several features are left out for a sample, our method can also be used to analyze the effect of multiple features missing and can thus show possibly correlated features.
@@
 Thus important features are going to be left out less often and one can then extract the dropout probabilities for each feature and assess their significance based on them.
-Feature dropout can also be used as a building block for computing the accuracy of datapoints with missing features that is necessary to compute Shapley values \cite{shapley_value_1952} or KernelSHAP \cite{lundberg_unified_2017}.
+Feature dropout can also be used as a building block for computing the accuracy of datapoints with missing features that is necessary to compute Shapley values \cite{shapley_value_1953} or KernelSHAP \cite{lundberg_unified_2017}.
 
@@
 %\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_pred_plots2_outcomes_0_3.pickle_5_DoS - DDoS-DoS Hulk"}.pdf}
-\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_pred_plots2_outcomes_0_3.pickle_6_DoS - DDoS-DoS Slowhttptest"}.pdf}
+\includegraphics[width=\columnwidth]{{"../plots/plot/flows_pred_plots2_outcomes_0_3.pickle_6_DoS - DDoS-DoS Slowhttptest"}.pdf}
 % \includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_14_Web Attack-XSS"}.pdf}

### 2019-12-10 ex3/learn.py
Hopefully fix ARS calculation method. Fix feature plot

@@ -888,7 +888,7 @@ def adv_internal(in_training = False, tradeoff=None, lr=None, iterations=None, a
 	orig_indices, attack_indices = zip(*[(orig, i) for orig, i in zip(indices, range(len(subset_with_all_traffic))) if subset_with_all_traffic[i][1][0,0] == 1 and subset_with_all_traffic[i][2][0,0].item() not in attack_types_which_are_not_investigated_anymore])
 
 	subset = torch.utils.data.Subset(dataset, orig_indices)
-	print("len(subset)", len(subset))
+	# print("len(subset)", len(subset))
 
 	loader = torch.utils.data.DataLoader(subset, batch_size=opt.batchSize, shuffle=False, collate_fn=custom_collate)
 
@@ -1134,7 +1134,7 @@ def adv_internal(in_training = False, tradeoff=None, lr=None, iterations=None, a
 		per_flow_accuracy = (np.mean(np.round(numpy_sigmoid(np.array([item[-1] for item in per_attack_results])))))
 		# TODO: l1 or l2 norm?
 		dist = np.array([np.linalg.norm((per_attack_orig_item-per_attack_modified_item).flatten(), ord=1).mean() for per_attack_orig_item, per_attack_modified_item in zip(per_attack_orig, per_attack_modified)]).mean()
-		linf_dist = np.array([np.linalg.norm((per_attack_orig_item-per_attack_modified_item).flatten(), ord='inf').mean() for per_attack_orig_item, per_attack_modified_item in zip(per_attack_orig, per_attack_modified)]).mean()
+		linf_dist = np.array([np.linalg.norm((per_attack_orig_item-per_attack_modified_item).flatten(), ord=float("inf")).mean() for per_attack_orig_item, per_attack_modified_item in zip(per_attack_orig, per_attack_modified)]).mean()
 
 		print("Attack type: {}; number of samples: {}, average dist: {}, average inf. dist: {}, packet accuracy: {}/{}, flow accuracy: {}/{}".format(reverse_mapping[attack_number], len(per_attack_results), dist, linf_dist, per_packet_accuracy, per_packet_orig_accuracy, per_flow_accuracy, per_flow_orig_accuracy))
 
@@ -1168,7 +1168,7 @@ def adv_until_less_than_half():
 		results_dict = list(adv_internal(False, tradeoff, lr, iterations, next_filter))[0]
[…]

### 2019-12-10 ex3/plot_features.py

@@ -13,6 +13,9 @@ import json
 import pickle
 from learn import numpy_sigmoid
 
+plt.rcParams['font.family'] = 'serif'
+plt.rcParams['savefig.format'] = 'pdf'
+
 DIR_NAME = "plots/plot_features"
 
 dataroot_basename = sys.argv[1].split('_')[0]
@@ -51,7 +54,7 @@ colors_rgb_ranges = [matplotlib.colors.ListedColormap([brighten(color, item) for
 # print("colors", colors)
 # print("colors_rgb_ranges[0]", colors_rgb_ranges[0])
 # quit()
-FEATURE_NAMES = ["packet length", "iat"]
+FEATURE_NAMES = ["Packet length", "IAT"]
 
 for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, result_ranges_by_attack_number_item, sample_indices_by_attack_number_item) in enumerate(zip(results_by_attack_number, flows_by_attack_number, result_ranges_by_attack_number, sample_indices_by_attack_number)):
 
@@ -91,10 +94,10 @@ for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, re
 	# print("flow shapes", actual_flow_medians.shape, actual_flow_first_quartiles.shape, actual_flow_third_quartiles.shape)
 
 	# for i in range(medians.shape[1]):
-	plt.figure(attack_type)
+	# plt.figure(figsize=(5,4))
 	plt.title(reverse_mapping[attack_type])
 
-	fig, ax1 = plt.subplots()
+	fig, ax1 = plt.subplots(figsize=(5,4))
 	ax2 = ax1.twinx()
 
 	all_legends = []
@@ -114,15 +117,14 @@ for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, re
 		all_legends += [ret2]
 		# all_legends += [legend, legend+" 1st and 3rd quartile"]
 	all_labels = [item.get_label() for item in all_legends]
-	ax1.legend(all_legends, all_labels)
+	ax2.legend(all_legends, all_labels, loc="upper left")
 
 	ax1.set_xlabel('Sequence index')
 	plt.title(reverse_mapping[attack_type])
-	# plt.figure(attack_type)
 	plt.tight_layout()
 
 	os.makedirs(DIR_NAME, exist_ok=True)
[…]

### 2019-12-10 ex3/learn.py
Change ars generation function a bit

@@ -1222,8 +1222,8 @@ def adv_until_less_than_half():
 			"ratio", final_ratios[attack_index],
 			"flow_accuracy", np.mean(np.round(numpy_sigmoid(np.array([item[-1] for item in orig_results[attack_index]])))),
 			"packet_accuracy", np.mean(np.round(numpy_sigmoid(np.array([sublist for l in orig_results[attack_index] for sublist in l])))),
-			"flow_distance", distances_flows[attack_index],
-			"packet_distance", distances_packets[attack_index]
+			"flow_distance", distances_flows[attack_index] if ratio <= THRESHOLD else 0,
+			"packet_distance", distances_packets[attack_index] if ratio <= THRESHOLD else 0
 		)
 
 def eval_nn(data):

### 2019-12-10 ex3/paper/paper.tex
Incorporate Alex' text changes again

@@
 \label{fig:adv_per_family}
-\todo{update numbers, shorten attack names, What's the tradeoff for CW?}
+\todo{explain tradeoff for cw and bound for others}
 \end{figure}
@@
 
-As \autoref{fig:adv_per_family} shows, for most attack types, most samples can be modified to look benign by an adversary by using \gls{cw}.
-High accuracy for a particular attack category does not imply that adversarial attacks are more likely to fail for these attacks.
-Thus we argue that beside the classical metrics such as accuracy, false positives etc., a new metric is needed, which quantifies how easily an adversary can fool the classifier.
+As \autoref{fig:adv_per_family} and \autoref{tab:adv_accuracies} show, for most attack types, most samples can be modified by an adversary to be classified as benign by using \gls{cw}.
+High accuracy for a particular attack category does not imply that adversarial attacks are  likely to fail for these attacks.
+Thus we argue that beside the classical metrics such as accuracy, false positives etc., a new metric is needed, which quantifies how easily a classifier can be fooled.
 
@@
 
-Setting the threshold for our metric to 50\% is arbitrary but we think that it is reasonable since if an attacker can make the majority of samples adversarial, we think that its attack is ``successful''.
+Setting the threshold for the \gls{ars} to 50\% is arbitrary, but
+ reasonable as this way the \gls{ars} approximates the median of the distances of all samples and, additionally, marks the distance where an adversarial attack is more likely to be successful than unsuccessful.
+%we think that it is reasonable since if an attacker can make the majority of samples adversarial, we think that its attack is ``successful''.
 
@@
 \end{itemize}
+
+Many explainability methods proposed recently are local and thus provide an explaination for a particular data sample \cite{shapley_value_1953,lundberg_unified_2017,dhurandhar_model_nodate,ribeiro_why_2016}.
+However, for the particular problem of network traffic due to the high number of flows and characteristics of data, analyzing individual samples is of low interest.
+Thus explainability methods presented in this paper aim to understand a model by analyzing which features are important, at which time step they are important and which feature values lead to classification as attack.
 
@@
 \end{table}
-As a first step to understanding the neural network's decisions, we tried to estimate how important individual features are for the model's predictions.
+As a first step to understanding the neural network's decisions, we  estimate how important individual features are for the model's predictions.
 When investigating an \gls{ml} classifier, it is natural to ask which inputs have a large influence on the classifier's prediction.
@@
 \subsubsection{Neural Network Weights}
-In previous works, a simple method for determining feature importance in neural networks, has been summing up neural network weights leading from a certain input to the prediction outcome.
+In previous works~\cite{olden_accurate_2004}, a simple method for determining feature importance in neural networks, has been summing up neural network weights leading from a certain input to the prediction outcome.
 The weights method can be considered for both feature importance and feature sensitivity, however, clearly, especially in the case of complex network architectures, this method is likely to provide wrong results.
@@
 \subsection{Explainability Plots}
+Knowing which features to investigate, it is important to analyze which feature values lead to classification as attack.
 During flow extraction we used a usual 5-tuple flow key, which distinguishes flows based on the protocol they use and their source and destination port and IP address.
@@
 \section{Defences}
-
+We now propose two approaches to the \gls{rnn} more robust against adversarial attacks.
 \subsection{Reducing features}
@@
 \subsection{Adversarial training}
-In addition to omitting manipulable features we tried to make our network more robust against adversarial samples by augmenting the training set by adversarial flows generated using the Carlini-Wagner method, labeled as additional attacks.
+In addition to omitting manipulable features we tried to make our network more robust against adversarial samples by augmenting the training set by adversarial flows generated using the \gls{cw} method, labeled as additional attacks.
 
 For this purpose, we added one adversarial sample per attack sample to the training set.
-Note that, since our used method for obtaining adversarial samples is deterministic, this is the maximum number of adversarial samples possible.
+Since \gls{cw} is deterministic, this is the maximum number of adversarial samples possible.
 With this training set we then alternated retraining of the neural network and optimization of the adversarial samples using gradient descent.
@@
 
-\autoref{fig:avdistance} shows the average distance of a block of adversarial samples from to original, unmodified attack samples throughout the adversarial training procedure.
+\autoref{fig:avdistance} shows the \gls{ars} throughout the adversarial training procedure.
 The figure clearly indicates that adversarial training is effective, as the distance gradually increases.

### 2019-12-10 ex3/paper/paper.tex
Finish some citation issue

@@
 
-Many explainability methods proposed recently are local and thus provide an explaination for a particular data sample \cite{shapley_value_1953,lundberg_unified_2017,dhurandhar_model_nodate,ribeiro_why_2016}.
+Many explainability methods proposed recently are local and thus provide an explaination for a particular data sample \cite{shapley_value_1953,lundberg_unified_2017,dhurandhar_model_2018,ribeiro_why_2016}.
 However, for the particular problem of network traffic due to the high number of flows and characteristics of data, analyzing individual samples is of low interest.

### 2019-12-11 ex3/plot2_adv.py
Compressed plot2_adv a little

@@ -18,6 +18,7 @@ DIR_NAME = "plots/plot2_adv"
 USE_LOG_SCALE = True
 OMIT_UPPER_AXIS = True
 ONLY_ONE_LEGEND = True
+SHOW_TITLE = False
 
 plt.rcParams["font.family"] = "serif"
 dataroot_basename = sys.argv[1].split('_')[0]
@@ -50,7 +51,7 @@ adv_orig_flows_by_attack_number = adv_loaded["orig_flows_by_attack_number"]
 
 colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
 ORDERING = ["original", "adversarial"]
-FEATURE_NAMES = ["Packet length / B", "IAT / s"]
+FEATURE_NAMES = ["Pkt. length / B", "IAT / s"]
 FEATURE_SCALES = [1, 0.001] 
 
 def brighten(rgb, how_much=0.0):
@@ -153,7 +154,7 @@ for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, re
 
 
 	all_legends = []
-	plt.figure(attack_type, figsize=(5,4))
+	plt.figure(attack_type, figsize=(5,3))
 	plt.title(reverse_mapping[attack_type])
 
 	for feature_index_from_zero, (feature_name, feature_index, feature_scale) in enumerate(zip(FEATURE_NAMES, (3, 4), FEATURE_SCALES)):
@@ -197,11 +198,15 @@ for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, re
 		else:
 			ticks = plt.xticks()
 			plt.xticks([ tick for tick in ticks[0][1:-1] if tick.is_integer() ])
-
+			
+		plt.gca().yaxis.set_major_locator(plt.LogLocator(numticks=5))
+		
 	if ONLY_ONE_LEGEND:
 		plt.legend([Line2D([0], [0], c='k'), Line2D([0], [0], c='k', linestyle='dashed')], ['Original flows', 'Adversarial flows'], loc='lower right',  ncol=2)
-	plt.figure(attack_type)
-	plt.suptitle(reverse_mapping[attack_type])
+	fig = plt.figure(attack_type)
+	fig.align_ylabels()
+	if SHOW_TITLE:
+		plt.suptitle(reverse_mapping[attack_type])
 	plt.tight_layout()
 	plt.subplots_adjust(top=0.935)
 	if OMIT_UPPER_AXIS:

### 2019-12-11 ex3/learn.py
Added average distance to adv

@@ -1106,6 +1106,8 @@ def adv_internal(in_training = False, tradeoff=None, lr=None, iterations=None, a
 	print("Average confidence on original flows: {}".format(np.mean(numpy_sigmoid(np.array([item[-1] for item in original_results])))))
 	print("Average confidence on flows: {}".format(np.mean(numpy_sigmoid(np.array([item[-1] for item in results])))))
 	print("Ratio of successful adversarial attacks on flows: {}".format(1-np.mean(np.round(numpy_sigmoid(np.array([item[-1] for item in results]))))))
+	print("Average distance: {}".format(np.array([np.linalg.norm((orig_item-modified_item).flatten(), ord=1).mean() for ((orig_item,_,_), (modified_item,_,_)) in zip(original_dataset, subset)]).mean()))
+	print("Average inf. distance: {}".format(np.array([np.linalg.norm((orig_item-modified_item).flatten(), ord=float("inf")).mean() for ((orig_item,_,_), (modified_item,_,_)) in zip(original_dataset, subset)]).mean()))
 
 	attack_numbers = mapping.values()
 

### 2019-12-11 ex3/paper/paper.tex
Compressed plot2_adv a little
Further improved feature plot

@@
 
-\caption{The solid line shows the mean value of the feature for all flows of the attack type. The shaded region shows the change in confidence that occurs when the feature is varied.}
+\caption{Exemplary PD plot and adversarial flows for the DoS Slowloris attack in CIC-IDS-2017.
+The solid line shows the mean value of the feature for all flows of the attack type.
+The shaded region shows the change in confidence that occurs when the feature is varied.}
 \label{fig:pred_plots2}
@@
 \begin{figure}
-%\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_2_Brute Force-SSH-Patator"}.pdf}
+%AH. I think the 'Firewall off' plot is not a good example as it shows only 4 packets
+\includegraphics[width=\columnwidth]{{"../plots/plot_features/flows_pred_plots2_outcomes_0_3.pickle_2_Brute Force-SSH-Patator"}.pdf}
 %\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_6_DoS - DDoS-DoS Slowhttptest"}.pdf}
-\includegraphics[width=\columnwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
+%\includegraphics[width=\columnwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
 %\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
 
-\caption{The solid line shows the median value of the feature for all flows of a specific attack type. The shaded region are the first and third quartile.}
+\caption{\gls{iat} and packet length for SSH brute-force attacks in CIC-IDS-2017.}% The solid line shows the median value of the feature for all flows of a specific attack type.
+The shaded region are the first and third quartile.} % think legend is enough for unterstanding
 \todo{Max: Maybe here two figures are enough?}

### 2019-12-11 ex3/plot_features.py

@@ -11,6 +11,8 @@ import sys
 import os
 import json
 import pickle
+from matplotlib.lines import Line2D
+from matplotlib.patches import Rectangle
 from learn import numpy_sigmoid
 
 plt.rcParams['font.family'] = 'serif'
@@ -26,7 +28,7 @@ categories_mapping, mapping = categories_mapping_content["categories_mapping"], 
 reverse_mapping = {v: k for k, v in mapping.items()}
 # print("reverse_mapping", reverse_mapping)
 
-with open(dataroot_basename+"_full_no_ttl_normalization_data.pickle", "rb") as f:
+with open(dataroot_basename+"_normalization_data.pickle", "rb") as f:
 	means, stds = pickle.load(f)
 
 file_name = sys.argv[1]
@@ -54,7 +56,7 @@ colors_rgb_ranges = [matplotlib.colors.ListedColormap([brighten(color, item) for
 # print("colors", colors)
 # print("colors_rgb_ranges[0]", colors_rgb_ranges[0])
 # quit()
-FEATURE_NAMES = ["Packet length", "IAT"]
+FEATURE_NAMES = ["Pkt. length / B", "IAT / msec"]
 
 for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, result_ranges_by_attack_number_item, sample_indices_by_attack_number_item) in enumerate(zip(results_by_attack_number, flows_by_attack_number, result_ranges_by_attack_number, sample_indices_by_attack_number)):
 
@@ -97,7 +99,7 @@ for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, re
 	# plt.figure(figsize=(5,4))
 	plt.title(reverse_mapping[attack_type])
 
-	fig, ax1 = plt.subplots(figsize=(5,4))
+	fig, ax1 = plt.subplots(figsize=(5,2.4))
 	ax2 = ax1.twinx()
 
 	all_legends = []
@@ -105,22 +107,29 @@ for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, re
 		# if feature_index_from_zero > 0:
 		# 	continue
 		# plt.subplot("{}{}{}".format(len(FEATURE_NAMES), 1, feature_index_from_zero+1))
-		ax.set_ylabel(feature_name, color=colors[feature_index_from_zero])
+		ax.set_ylabel(feature_name) #, color=colors[feature_index_from_zero])
[…]

### 2019-12-11 ex3/auxplot.py
Prepare tradeoff plot

@@ -7,6 +7,7 @@ import numpy as np
 import matplotlib.pyplot as plt
 import sys
 from matplotlib.transforms import Affine2D
+from matplotlib.lines import Line2D
 
 group_names = []
 values = []
@@ -86,6 +87,22 @@ def ars():
 	plt.ylabel('ARS')
 	ylim1,ylim2 = plt.ylim()
 	plt.ylim((ylim1,ylim2+20)) # move plots away from legend
+	plt.tight_layout()
+	
+def adv():
+	plt.figure(figsize=(5,2))
+	x_values = [ float(value) for value in group_names ]
+	lines = plt.plot(x_values, values[:,:2])
+	plt.xlabel('Tradeoff')
+	plt.ylabel('Success ratio')
+	plt.gca().set_ylabel_legend(Line2D([0],[0], color='gray'))
+	plt.twinx()
+	plt.plot(x_values, values[:,2:], linestyle='--')
+	plt.ylabel('$\ell_1$ distance')
+	plt.gca().set_ylabel_legend(Line2D([0],[0], color='gray', linestyle='--'))
+	plt.legend(lines, ['CIC-IDS-2017', 'UNSW-NB15'])
+	#  ylim1,ylim2 = plt.ylim()
+	#  plt.ylim((ylim1,ylim2+20)) # move plots away from legend
 	plt.tight_layout()
 
 globals()[sys.argv[1]]()

### 2019-12-11 ex3/paper/paper.tex
Change ylabel in plots

@@
 Applying them shows that already the first packets of a communication flow are of crucial importance and are likely going to be targeted by attackers.
-Feature importance methods show that even though the modified features are not very relevant for the classifier, they can still be abused to generate adversarial samples.
+Feature importance methods show that even relatively unimportant features can be effectively abused to generate adversarial samples.
 As a countermeasure, we develop an easy-to-use adversarial training procedure %and can confirm that it
 which significantly and successfully reduces the attack surface and improves \gls{ars}.
-
-\todo{Max: Abstract too long}
 
@@
 \maketitle
-
-\todo{Legend in some figures is misplaced}
 
@@
 The solid line shows the mean value of the feature for all flows of the attack type.
-The shaded region shows the change in confidence that occurs when the feature is varied.}
+The shaded region shows the change in confidence that occurs when the feature is varied.
+\todo{IAT plot is wrong: There shouldn't be step 0.}}
 \label{fig:pred_plots2}
@@
 We have implemented a recurrent classifier based on LSTM to detect network attacks.
-While attack detection performance is not better than that of non-recurrent approaches, we can already detect attacks before they are over.
+While attack detection performance is not necessarily better than that of non-recurrent approaches, we can already detect attacks before they are over.
 Furthermore, the recurrent approach allows us to inspect the influence of single packets on the detection performance and shows which packets are \textit{characteristic} for attacks.
 
-We show that adversarial samples are possible even though the data we use are very different from those that are usually used when finding adversarial samples.
-
-\todo{Max: Need more conclusion. Shouldn't be the same like abstract and intro.}
+We showed that even though our use case is very different from computer vision, adversarial samples can be found even if only unimportant features are modified.
+We introduced the \gls{ars} as a new way to quantify adversarial threat and show that with our adversarial training procedure we can significantly reduce that threat.
 

### 2019-12-11 ex3/plot2_adv.py

@@ -51,8 +51,8 @@ adv_orig_flows_by_attack_number = adv_loaded["orig_flows_by_attack_number"]
 
 colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
 ORDERING = ["original", "adversarial"]
-FEATURE_NAMES = ["Pkt. length / B", "IAT / s"]
-FEATURE_SCALES = [1, 0.001] 
+FEATURE_NAMES = ["Pkt. length [B]", "IAT [s]"]
+FEATURE_SCALES = [1, 0.001]
 
 def brighten(rgb, how_much=0.0):
 	hls = list(colorsys.rgb_to_hls(*rgb))
@@ -190,7 +190,7 @@ for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, re
 			x, y = x[1:], y[1:]
 		ret = plot_function(x, y, label=legend, linestyle="dashed", color=colors[feature_index_from_zero])
 		all_legends += ret
-		
+
 		if not ONLY_ONE_LEGEND:
 			plt.legend()
 		if OMIT_UPPER_AXIS and feature_name != FEATURE_NAMES[-1]:
@@ -198,9 +198,9 @@ for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, re
 		else:
 			ticks = plt.xticks()
 			plt.xticks([ tick for tick in ticks[0][1:-1] if tick.is_integer() ])
-			
+
 		plt.gca().yaxis.set_major_locator(plt.LogLocator(numticks=5))
-		
+
 	if ONLY_ONE_LEGEND:
 		plt.legend([Line2D([0], [0], c='k'), Line2D([0], [0], c='k', linestyle='dashed')], ['Original flows', 'Adversarial flows'], loc='lower right',  ncol=2)
 	fig = plt.figure(attack_type)

### 2019-12-11 ex3/plot_features.py

@@ -56,7 +56,7 @@ colors_rgb_ranges = [matplotlib.colors.ListedColormap([brighten(color, item) for
 # print("colors", colors)
 # print("colors_rgb_ranges[0]", colors_rgb_ranges[0])
 # quit()
-FEATURE_NAMES = ["Pkt. length / B", "IAT / msec"]
+FEATURE_NAMES = ["Pkt. length [B]", "IAT [ms]"]
 
 for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, result_ranges_by_attack_number_item, sample_indices_by_attack_number_item) in enumerate(zip(results_by_attack_number, flows_by_attack_number, result_ranges_by_attack_number, sample_indices_by_attack_number)):
 
@@ -113,10 +113,10 @@ for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, re
 		legend = "{}".format(feature_name)
 		ret = ax.plot(range(max_length), actual_flow_medians[:,feature_index]*stds[feature_index]+means[feature_index], label=legend, color=colors[feature_index_from_zero])
 		ret2 = ax.fill_between(range(max_length), actual_flow_first_quartiles[:,feature_index]*stds[feature_index]+means[feature_index], actual_flow_third_quartiles[:,feature_index]*stds[feature_index]+means[feature_index], alpha=0.5, edgecolor=colors[feature_index_from_zero], facecolor=colors[feature_index_from_zero], label=legend+" 1st and 3rd quartile")
-		
+
 		ylim1,ylim2 = ax.get_ylim()
 		ax.set_ylim((ylim1,ylim2+(ylim2-ylim1)*0.25)) # move plots away from legend
-		
+
 		# print("ret", ret)
 		# print("ret2", ret2)
 		all_legends += ret

### 2019-12-12 ex3/paper/paper.tex
Modified paper

@@
 \label{sec:adv}
-We now investigate whether known \gls{aml} methods can be used for generating adversarial flows, i.e. minimally different attack flows which are classified as benign.
+We now investigate whether known \gls{aml} methods can be used for generating adversarial flows, i.e. minimally different attack flows which are classified as benign, for out \gls{rnn}.
 In our scenario, most features such as, e.g., \gls{tcp} flags cannot be easily manipulated because their manipulation might violate the \gls{tcp} and thus cause communication to fail.
@@
 \autoref{fig:adv_cw} depicts the ratio of attack flows which were successfully misclassified as benign with respect to the final packet along with the resulting distance from the original sample.
-Evidently, the larger the $\epsilon$ parameter, the better the attack works, but also the distances become larger.
-
+Evidently, the larger the tradeoff $\epsilon$, the better the attack works, but also average distances from original flows become larger.
 Hence, with acceptable distance values we were able to generate adversarial samples for about 80\% of attack samples.
@@
 \label{fig:adv_per_family}
-\todo{explain tradeoff for cw and bound for others}
 \end{figure}
@@
 \end{table}
-It is interesting to note that for very large values of $\epsilon$ the success ratio starts to decrease again.This phenomenon might be due to numerical effects and might be preventable by a smaller learning rate.
+%It is interesting to note that for very large values of $\epsilon$ the success ratio starts to decrease again.This phenomenon might be due to numerical effects and might be preventable by a smaller learning rate.
 
 \autoref{tab:adv_accuracies} depicts the performance for FGSM and $\ell_\infty$-bounded \gls{pgd}. 
-As expected, while FGSM is the fastest of all algorithms, it delivers the worst performance. Of the investigated algorithms, thus the \gls{cw} with $\ell_2$ distance delivered the best performance.
-
+As expected, while FGSM is the fastest of all algorithms, it delivers the worst performance. Of the investigated algorithms,  \gls{cw}  delivered the best performance.
 \autoref{fig:adv_per_family} depicts how successful the attack is for individual attack families.
-For this table we used a $\ell_\infty$ bound of 0.3 for FGSM and $\ell_\infty$-bounded \gls{pgd} and a tradeoff of 0.5 for Carlini-Wagner.
 The table shows significant differences for detecting different attack families in the first place.
 However, also for generating adversarial flows some attack families are more susceptible than others.
+
+For \autoref{tab:adv_accuracies} and \autoref{fig:adv_per_family} we used \gls{cw} with a tradeoff of 1.
+To provide a fair comparison, for \gls{fgsm} and \gls{pgd} we used an $\ell_\infty$ distance which we observed for \gls{cw}.
+
+It is interesting to note that also increasing the $\ell_\infty$ bound for \gls{pgd} did not result in significantly improved performance.
+We can thus confirm the recommendation by \cite{carlini_towards_2017} to perform gradient descent on the logit output instead of the loss.
 
@@
 
-Many explainability methods proposed recently are local and thus provide an explaination for a particular data sample \cite{shapley_value_1953,lundberg_unified_2017,dhurandhar_model_2018,ribeiro_why_2016}.
-However, for the particular problem of network traffic due to the high number of flows and characteristics of data, analyzing individual samples is of low interest.
-Thus explainability methods presented in this paper aim to understand a model by analyzing which features are important, at which time step they are important and which feature values lead to classification as attack.
+Many explainability methods proposed recently are local and thus provide explanations for a particular data sample \cite{shapley_value_1953,lundberg_unified_2017,dhurandhar_model_2018,ribeiro_why_2016}.
+However, for the particular problem of network traffic, due to the high number of flows and the characteristics of data, analyzing individual samples is of low interest.
+Explainability methods presented in this paper therefore aim to understand a model by analyzing which features are important, at which time step they are important and which feature values lead to classification as attack.
 
@@
 The shaded region are the first and third quartile.} % think legend is enough for unterstanding
-\todo{Max: Maybe here two figures are enough?}
 \label{fig:plot_features}
@@
 \section{Defences}
-We now propose two approaches to the \gls{rnn} more robust against adversarial attacks.
+We now propose two approaches to increase robustness of the \gls{rnn} against adversarial attacks.
 \subsection{Reducing features}
@@
 \subsection{Adversarial training}
-In addition to omitting manipulable features we tried to make our network more robust against adversarial samples by augmenting the training set by adversarial flows generated using the \gls{cw} method, labeled as additional attacks.
+As alternative to omitting manipulable features, we attempted to make the classifier more robust against adversarial samples by augmenting the training set by adversarial flows generated using  \gls{cw}, labeled as additional attacks.
 
@@
 
-One question which occurs in this process, is how many CPU cycles to spend on network training and adversarial sample optimization.
+A question which occurs in this process, is how many CPU cycles to spend on network training and adversarial sample optimization.
 As a natural approach, we decided to use as many backpropagation steps for neural network training as we use for adversarial sample optimization.
@@
 
-\autoref{fig:avdistance} shows the \gls{ars} throughout the adversarial training procedure.
-The figure clearly indicates that adversarial training is effective, as the distance gradually increases.
+\autoref{fig:avdistance} shows the \gls{ars} throughout the adversarial training procedure for several attack categories and clearly indicates that adversarial training is effective, as the distance gradually increases.
 Hence, an attacker would have to modify attack samples more and more, eventually rendering the attack unpractical.
 
-However, the figure also shows that an increase of distance can only be achieved after a considerable amount of training.
+%However, the figure also shows that an increase of distance can only be achieved after a considerable amount of training.
 It is, hence, questionable if the gained increase in robustness is enough for reliable attack detection.
-
-\todo{Max: Insert results that show that accuracy on normal data is still ok after adversarial training.}
 
@@
 
+For both datasets, after a small number of epochs, it was no longer possible to create a significant number of adversarial samples for most attack categories.
+For the datasets we considered, adversarial training left classification performance unaffected.
+\todo{Max: Insert results that show that accuracy on normal data is still ok after adversarial training.}
+
+% AH. guess this is no longer true
 %However, against our expectations, this approach did not lead to satisfactory results in the sense that the network would learn to classify adversarial samples as attacks.
@@
 
-A possible explanation for this observation might be that the adversarial samples assumed too much the shape of normal flows, so that the network could not reliable tell them apart.
+%A possible explanation for this observation might be that the adversarial samples assumed too much the shape of normal flows, so that the network could not reliable tell them apart.
 
@@
 
-We have implemented a recurrent classifier based on LSTM to detect network attacks.
-While attack detection performance is not necessarily better than that of non-recurrent approaches, we can already detect attacks before they are over.
+We have implemented a recurrent classifier based on LSTMs to detect network attacks, %While attack detection performance is not necessarily better than that of non-recurrent approaches,
+which can already detect attacks before they are over.
 Furthermore, the recurrent approach allows us to inspect the influence of single packets on the detection performance and shows which packets are \textit{characteristic} for attacks.
+Even though the interpretation of \glspl{rnn} poses several difficulties, we have demonstrated methods for gaining insights into the model's functioning.
 
-We showed that even though our use case is very different from computer vision, adversarial samples can be found even if only unimportant features are modified.
-We introduced the \gls{ars} as a new way to quantify adversarial threat and show that with our adversarial training procedure we can significantly reduce that threat.
+We showed that even though our use case is very different from computer vision, adversarial samples can be found efficiently, even if only ostensibly unimportant features can be modified.
+We introduced the \gls{ars} as a new way to quantify the adversarial threat.
+Deploying an adversarial training procedure, we could significantly reduce the adversarial threat.
 

### 2019-12-12 ex3/paper/paper.tex
Round brackets for units

@@
 \subsection{Evaluation}
-\begin{figure*}
-\includegraphics[width=0.49\textwidth]{adv_plots/cwl1.pdf}
-\includegraphics[width=0.49\textwidth]{adv_plots/cwl2.pdf}
-\caption{Performance of the \gls{cw} with different $\epsilon$ values for $\ell_1$ and $\ell_2$ distances.}
-\todo{Max: Need to be updated with current results that do not decrease at some point}
+\begin{figure}
+\includegraphics[width=\columnwidth]{../plots/adv_comparison/tradeoff.pdf}
+% \includegraphics[width=0.49\textwidth]{adv_plots/cwl1.pdf}
+% \includegraphics[width=0.49\textwidth]{adv_plots/cwl2.pdf}
+\caption{Performance of \gls{cw} with different $\epsilon$ values.}
+\todo{I guess the irregularity at eps=1 hints at too few iterations. Currently rerunning with 1000 base iterations}
 \label{fig:adv_cw}
-\todo{compress figures somehow}
-\end{figure*}
+\end{figure}
 \autoref{fig:adv_cw} depicts the ratio of attack flows which were successfully misclassified as benign with respect to the final packet along with the resulting distance from the original sample.

### 2019-12-12 ex3/auxplot.py

@@ -93,15 +93,16 @@ def adv():
 	plt.figure(figsize=(5,2))
 	x_values = [ float(value) for value in group_names ]
 	lines = plt.plot(x_values, values[:,:2])
-	plt.xlabel('Tradeoff')
+	plt.xlabel('Tradeoff $\epsilon$')
 	plt.ylabel('Success ratio')
-	plt.gca().set_ylabel_legend(Line2D([0],[0], color='gray'))
+	plt.gca().set_ylabel_legend(Line2D([0],[0], color='gray'), handlelength=1.4)
 	plt.twinx()
 	plt.plot(x_values, values[:,2:], linestyle='--')
 	plt.ylabel('$\ell_1$ distance')
-	plt.gca().set_ylabel_legend(Line2D([0],[0], color='gray', linestyle='--'))
+	plt.gca().set_ylabel_legend(Line2D([0],[0], color='gray', linestyle='--'), handlelength=1.4)
 	plt.legend(lines, ['CIC-IDS-2017', 'UNSW-NB15'])
-	#  ylim1,ylim2 = plt.ylim()
+	ylim1,ylim2 = plt.ylim()
+	plt.ylim((ylim1,12))
 	#  plt.ylim((ylim1,ylim2+20)) # move plots away from legend
 	plt.tight_layout()
 

### 2019-12-12 ex3/paper/paper.tex
Integrated total results into adv comparison

@@
 \includegraphics[width=\columnwidth]{../plots/adv_comparison/adv_comparison_17.pdf}
-\caption{Attack success ratios per attack family. Depicted are flow detection accuracies for adversarial flows and for unmodified flows.}
+\caption{Attack success ratios for both dataset and per attack family for CIC-IDS-2017. Depicted are flow detection accuracies for adversarial flows and for unmodified flows.}
 \label{fig:adv_per_family}
@@
 
-\begin{table}
-\centering
-\caption{Total flow accuracy for adversarial flows.}
-\label{tab:adv_accuracies}
-\begin{tabular}{lrrr} \toprule
-& CW & FGSM & PGD \\ \midrule
-CIC-IDS-2017 &	0.346	 & 0.445	& 0.539 \\ 
-UNSW-NB15 &	0.326&	0.629	& 0.774 \\
-\bottomrule
-\end{tabular}
-\end{table}
+% \begin{table}
+% \centering
+% \caption{Total flow accuracy for adversarial flows.}
+% \label{tab:adv_accuracies}
+% \begin{tabular}{lrrr} \toprule
+% & CW & FGSM & PGD \\ \midrule
+% CIC-IDS-2017 &	0.346	 & 0.445	& 0.539 \\ 
+% UNSW-NB15 &	0.326&	0.629	& 0.774 \\
+% \bottomrule
+% \end{tabular}
+% \end{table}
 %It is interesting to note that for very large values of $\epsilon$ the success ratio starts to decrease again.This phenomenon might be due to numerical effects and might be preventable by a smaller learning rate.
@@
 
+\begin{table}
+% Evaluated using
+% ./python learn.py --dataroot flows.pickle --net runs/Oct26_00-03-50_gpu/lstm_module_1284.pth --function test --batchSize 512
+% ./python learn.py --dataroot flows15.pickle --net runs/Oct28_15-41-46_gpu/lstm_module_997.pth --function test --batchSize 512
+% on Dec. 4th
+\caption{Detection performance after adversarial training.} \label{tab:adv_training_performance}
+\centering
+\begin{tabular}{l r r r r} \toprule
+& \multicolumn{2}{c}{CIC-IDS-17} & \multicolumn{2}{c}{UNSW-NB-15} \\
+	&	Packet	&	Flow	&	Packet	&	Flow	\\	\midrule
+Accuracy	&	0.991	&	0.997	&	0.994	&	0.981	\\
+Precision	&	0.971	&	0.996	&	0.761	&	0.704	\\
+Recall	&	0.979	&	0.992	&	0.970	&	0.824	\\
+F1	&	0.975	&	0.994	&	0.853	&	0.759	\\
+Youden	&	0.973	&	0.991	&	0.965	&	0.811	\\
+
+\bottomrule
+\end{tabular}
+\end{table}
+
 % AH. guess this is no longer true

### 2019-12-12 ex3/auxplot.py

@@ -62,16 +62,21 @@ def adv_results():
 	plt.figure(figsize=(5,3))
 	values = np.reshape(values, (-1,values.shape[1]//2,2))
 
-	order = np.argsort(np.mean(values[:,:,1], axis=1))
+	order = (2+np.argsort(np.mean(values[2:,:,1], axis=1))).tolist()
+	order = [0, 1] + order
 	width = 0.95 / values.shape[1]
 
-	x = np.arange(len(group_names))
+	x = np.arange(len(group_names), dtype=float)
+	x[:2] -= .5
 	for i in range(values.shape[1]):
 		plt.bar(x +width*i, values[order,i,1], width, color='gray', alpha=.5, **({'label': 'Original'} if i==0 else {}))
 	for i in range(values.shape[1]):
 		plt.bar(x +width*i, values[order,i,0], width, color=colors[i], label=feature_names[i*2].rstrip(' adv'))
 	plt.legend(loc='lower center', bbox_to_anchor=(0.5,1), ncol=4)
 	plt.xticks(x + width*values.shape[1]/2-width*0.5, [group_names[i] for i in order], rotation=45, horizontalalignment='right')
+	for i,label in  enumerate(plt.gca().get_xticklabels()):
+		if i < 2:
+			label.set_fontweight('bold')
 	for tick in plt.gca().xaxis.get_major_ticks():
 		label = tick.label1
 		label.set_transform(label.get_transform() + Affine2D().translate(8,0))

### 2019-12-12 ex3/auxplot.py
Changed 'accuracy' to 'recall'

@@ -80,7 +80,7 @@ def adv_results():
 	for tick in plt.gca().xaxis.get_major_ticks():
 		label = tick.label1
 		label.set_transform(label.get_transform() + Affine2D().translate(8,0))
-	plt.ylabel('Accuracy')
+	plt.ylabel('Recall')
 	plt.tight_layout()
 
 

### 2019-12-12 ex3/paper/paper.tex
Modified paper

@@
 
+% \begin{table}[b]
+% \caption{occurrence frequency of attack types.} \label{tab:occurrence}
+% \centering
+% \begin{tabular}{l r} \toprule
+% Attack type & Proportion \\
+% \midrule
+% Normal                                                         & $0.747468$ \\
+% DoS / DDoS:DoS Hulk                                            & $0.101014$ \\
+% PortScan:PortScan - Firewall off                               & $0.069020$ \\
+% DDoS:LOIT                                                      & $0.040784$ \\
+% Infiltration:Dropbox download                                  & $0.032982$ \\
+% DoS / DDoS:DoS GoldenEye                                       & $0.003224$ \\
+% DoS / DDoS:DoS Slowhttptest                                    & $0.001819$ \\
+% DoS / DDoS:DoS slowloris                                       & $0.001680$ \\
+% Brute Force:SSH-Patator                                        & $0.001107$ \\
+% Botnet:ARES                                                    & $0.000327$ \\
+% Web Attack:XSS                                                 & $0.000293$ \\
+% PortScan:PortScan - Firewall on                                & $0.000165$ \\
+% Brute Force:FTP-Patator                                        & $0.000110$ \\
+% Web Attack:Sql Injection                                       & $0.000006$ \\
+% DoS / DDoS:Heartbleed                                          & $0.000001$ \\
+% \bottomrule
+% \end{tabular}
+% \label{tab:occurrence}
+% \end{table}
+
 \begin{table}[b]
-\caption{Occurence probability of attack types.} \label{tab:occurrence}
+\caption{Occurrence frequency of attack types.}
+\label{tab:occurrence}
 \centering
-\begin{tabular}{l r} \toprule
-Attack type & Proportion \\
-\midrule
-Normal                                                         & $0.747468$ \\
-DoS / DDoS:DoS Hulk                                            & $0.101014$ \\
-PortScan:PortScan - Firewall off                               & $0.069020$ \\
-DDoS:LOIT                                                      & $0.040784$ \\
-Infiltration:Dropbox download                                  & $0.032982$ \\
-DoS / DDoS:DoS GoldenEye                                       & $0.003224$ \\
-DoS / DDoS:DoS Slowhttptest                                    & $0.001819$ \\
-DoS / DDoS:DoS slowloris                                       & $0.001680$ \\
-Brute Force:SSH-Patator                                        & $0.001107$ \\
-Botnet:ARES                                                    & $0.000327$ \\
-Web Attack:XSS                                                 & $0.000293$ \\
-PortScan:PortScan - Firewall on                                & $0.000165$ \\
-Brute Force:FTP-Patator                                        & $0.000110$ \\
-Web Attack:Sql Injection                                       & $0.000006$ \\
-DoS / DDoS:Heartbleed                                          & $0.000001$ \\
+\subfloat[CIC-IDS-2017\label{fig:cicids17proportions}
+]{
+\begin{tabular}{l r}
+\toprule
+Attack type & \hspace*{-4mm}Proportion \\ \midrule
+DoS Hulk	&	10.10\%	\\
+PortScan, Firewall	&	6.90\%	\\
+DDoS LOIT	&	4.08\%	\\
+Infiltration	&	3.30\%	\\
+DoS GoldenEye	&	0.32\%	\\
+DoS SlowHTTPTest	&	0.18\%	\\
+DoS Slowloris	&	0.17\%	\\
+Brute-force SSH	&	0.11\%	\\
+Botnet ARES	&	0.03\%	\\
+XSS attack	&	0.03\%	\\
+PortScan, no Fw.	&	0.02\%	\\
+Brute-force FTP	&	0.01\%	\\
+SQL injection	&	$<$0.01\%	\\
+Heartbleed	&	$<$0.01\%	\\
 \bottomrule
 \end{tabular}
-\label{tab:occurence}
+}{}
+\subfloat[UNSW-NB15\label{fig:unswnb15proportions}
+]{
+\begin{tabular}{l r}
+\toprule
+Attack type & \hspace*{-4mm}Proportion \\ \midrule
+Exploits	&	1.42\%	\\
+Fuzzers	&	1.01\%	\\
+Reconnaissance	&	0.58\%	\\
+Generic	&	0.21\%	\\
+DoS	&	0.19\%	\\
+Shellcode	&	0.08\%	\\
+Analysis	&	0.03\%	\\
+Backdoors	&	0.02\%	\\
+Worms	&	0.01\%	\\
+\bottomrule
+\end{tabular}
+}{}
 \end{table}
@@
 For evaluation, we use the \textit{CIC-IDS-2017} \cite{sharafaldin_toward_2018} and \textit{UNSW-NB15} \cite{moustafa_unsw-nb15:_2015} datasets which each include more than 2 million flows of network data, containing both benign traffic and a large number of different attacks.
-Attacks contained in CIC-IDS-2017 are shown in \autoref{tab:occurence}.
-\autoref{tab:performance_results} shows that our classifier achieves an accuracy that is similar to previous work based on these datasets~\cite{meghdouri_analysis_2018,bachl_walling_2019}.
-However, unlike these classifiers, our recurrent classifier has the advantage that it can detect attacks already before they are over.
+Attacks contained in the datasets are shown in \autoref{tab:occurrence}.
+\autoref{tab:performance_results} shows that our classifiers achieve an accuracy that is similar to previous work based on these datasets~\cite{meghdouri_analysis_2018,bachl_walling_2019}.
+However, unlike these classifiers, our recurrent classifier has the advantage of being able to detect attacks already before they are over.
 
@@
 \label{sec:adv}
-We now investigate whether known \gls{aml} methods can be used for generating adversarial flows, i.e. minimally different attack flows which are classified as benign, for out \gls{rnn}.
-In our scenario, most features such as, e.g., \gls{tcp} flags cannot be easily manipulated because their manipulation might violate the \gls{tcp} and thus cause communication to fail.
-Hence, we experiment whether by manipulating the \glspl{iat} and packet lengths of the first packets, the classifier might be fooled.
-Our scenario is quite different from how generating adversarial samples usually works: We have significantly less features than in, for example, an image.
+We now investigate whether known \gls{aml} methods can be used for generating adversarial flows, i.e. minimally different attack flows which are classified as benign, for our \gls{rnn}.
+% In our scenario, most features such as, e.g., \gls{tcp} flags cannot be easily manipulated because their manipulation might violate the \gls{tcp} and thus cause communication to fail.
+% Hence, we experiment whether by manipulating the \glspl{iat} and packet lengths of the first packets, the classifier might be fooled.
+% Our scenario is quite different from how generating adversarial samples usually works: We have significantly less features than in, for example, an image.
 Furthermore, only two out of those few features can be manipulated.
 And even these features cannot be manipulated freely: Besides the constraints that packet size and \gls{iat} cannot become negative there are also more problem-specific constraints:
+
+A network packet contains significantly less features than, e.g., an image and, furthermore, most features such as \gls{tcp} flags cannot be easily manipulated, as their manipulation might violate the protocol specifications and thus cause communication to fail.
+We identify the packet length and the \gls{iat} as features which are most likely to be exploited, but even these features cannot be manipulated freely.
+Beside the non-negativity constraint,  there are also more problem-specific constraints:
 \begin{itemize}[topsep=0pt,wide,labelwidth=!,labelindent=0pt]
-\item Only packets can be manipulated which are transmitted by the attacker. This means the attacker can only control one direction (except for botnets where both sides are attackers).
+\item Only packets can be manipulated which are transmitted by the attacker.
+Except for botnet and backdoor traffic, which is entirely controlled by an attacker, thus only packets travelling in one direction can be manipulated.
+ %This means the attacker can only control one direction (except for botnets where both sides are attackers).
 \item Packets must not be smaller than initially, as otherwise less information could be transmitted or packets might be rendered syntactically invalid.
 \item \glspl{iat} must not decrease, as otherwise the physical speed of data transmission can be violated in some cases.
-Since an in-depth analysis of when the reduction of \glspl{iat} is possible, would come with substantial effort, we generally disallowed a reduction of this feature.
+An in-depth analysis of when the reduction of \glspl{iat} is possible, would come with substantial effort, so we generally disallowed it.
+%a reduction of this feature.
 \end{itemize}
@@
 Several \gls{aml} methods have been proposed in the recent literature, achieving different speed-quality tradeoffs: 
-\cite{szegedy_intriguing_2014} show that it is possible to create adversarial samples for an image, which look similar to the original image but are classified wrongly.
-\cite{goodfellow_explaining_2015} develop the \gls{fgsm}, which makes easy generation of adversarial samples possible.
-\cite{papernot_crafting_2016} explore the use of \gls{aml} for \glspl{rnn}
+\cite{szegedy_intriguing_2014} shows that it is possible to create adversarial samples for an image, which look similar to the original image but are classified wrongly.
+\cite{goodfellow_explaining_2015} develops the \gls{fgsm}, which makes easy generation of adversarial samples possible.
+\cite{papernot_crafting_2016} explores the use of \gls{aml} for \glspl{rnn}, but lacks import \gls{aml} methods, which we consider in this paper.
 
-As finding adversarial flows is particularly difficult in our case for the reasons we discussed above, for the present research we preferred methods  producing high-quality adversarial samples while not being as fast as other methods.
-We implemented the following \gls{aml} methods.
+As finding adversarial flows is particularly difficult in our case, % for the reasons we discussed above,
+we preferred methods  producing high-quality adversarial samples while not being as fast as other methods. We implemented the following \gls{aml} methods.
 
@@
 For the datasets we considered, adversarial training left classification performance unaffected.
-\todo{Max: Insert results that show that accuracy on normal data is still ok after adversarial training.}
+%\todo{Max: Insert results that show that accuracy on normal data is still ok after adversarial training.}
+
+\autoref{tab:adv_training_performance} depicts classification on non-adversarial data after 50 epochs of adversarial training.
+Accuracy is essentially identical to results presented in \autoref{tab:performance_results}.
+Due to the higher proportion of attack samples in training data, we note, however, that recall increased at the expense of a slightly lower precision.
 
@@
 \caption{Detection performance after adversarial training.} \label{tab:adv_training_performance}
+\todo{Probably no space for table}
 \centering

### 2019-12-12 ex3/paper/paper.tex
Modified adv section

@@
 
+\begin{figure}[t]
+\includegraphics[width=\columnwidth]{../plots/adv_comparison/adv_comparison_17.pdf}
+\caption{Attack success ratios for both dataset and per attack family for CIC-IDS-2017. Depicted are flow detection accuracies for adversarial flows and for unmodified flows.}
+\label{fig:adv_per_family}
+\end{figure}
+
 \subsection{Evaluation}
-\begin{figure}
+\autoref{fig:adv_per_family} depicts the performance for the investigated algorithms for both datasets and for each attack family in CIC-IDS-2017. 
+We used \gls{cw} with a tradeoff of 1 and, to provide a fair comparison, for \gls{fgsm} and \gls{pgd} we used an average $\ell_\infty$ distance as observed for \gls{cw}.
+
+As expected, \gls{cw}  delivers the best performance and FGSM, while being the fastest of all investigated algorithms, delivers the worst performance.
+%\autoref{fig:adv_per_family} depicts how successful the attack is for individual attack families.
+The figure shows significant differences for detecting different attack families in the first place.
+However, also for generating adversarial flows some attack families are more susceptible than others.
+However, the results match to a high degree with our expectations.
+For example, SQL injection attacks apparently are closer to normal traffic than \gls{dos} attacks and, hence, finding adversarial samples should be easier.
+
+It is interesting to note that increasing the $\ell_\infty$ bound for \gls{pgd} did not yield significantly improved performance.
+We can thus confirm the recommendation by \cite{carlini_towards_2017} to perform gradient descent on the logit output instead of the loss for good results.
+
+\autoref{fig:adv_cw} depicts the success ratio and average distance for \gls{cw} for different tradeoffs $\epsilon$.
+%\autoref{fig:adv_cw} depicts the ratio of attack flows which were successfully misclassified as benign with respect to the final packet along with the resulting distance from the original sample.
+Evidently, the larger the tradeoff $\epsilon$, the better the attack works, but also distances from original flows become higher.
+Hence, with acceptable distance values we were able to generate adversarial samples for about 80\% of attack samples.
+The figure suggests that a tradeoff value between 0.5 and 1 would be ideal to find adversarial samples with a low enough distance from real samples.
+This conclusion might be misleading, however, as due to the objective function \eqref{eq:carliniWagner} the distance cannot become larger anymore for one sample as soon as an adversarial sample has been found.
+%, when performing optimization as depicted by the objective function .
+Hence, the depicted distance increase presumably is due to samples for which no adversarial sample could be found for lower $\epsilon$ values.
+
+\begin{figure}[b]
 \includegraphics[width=\columnwidth]{../plots/adv_comparison/tradeoff.pdf}
@@
 \end{figure}
-\autoref{fig:adv_cw} depicts the ratio of attack flows which were successfully misclassified as benign with respect to the final packet along with the resulting distance from the original sample.
-Evidently, the larger the tradeoff $\epsilon$, the better the attack works, but also average distances from original flows become larger.
-Hence, with acceptable distance values we were able to generate adversarial samples for about 80\% of attack samples.
-
-The figure suggests that a tradeoff value between 0.5 and 1 would be ideal to find adversarial samples with a low enough distance from real samples.
-This conclusion might be misleading, however, as the distance cannot become larger anymore as soon as an adversarial sample has been found, when performing optimization as depicted by the objective function \eqref{eq:carliniWagner}.
-Hence, the depicted increase of average distances presumably is due to samples for which no adversarial sample was found for lower $\epsilon$ values.
 
@@
 
-\begin{figure}
-\includegraphics[width=\columnwidth]{../plots/adv_comparison/adv_comparison_17.pdf}
-\caption{Attack success ratios for both dataset and per attack family for CIC-IDS-2017. Depicted are flow detection accuracies for adversarial flows and for unmodified flows.}
-\label{fig:adv_per_family}
-\end{figure}
-
 % \begin{table}
@@
 %It is interesting to note that for very large values of $\epsilon$ the success ratio starts to decrease again.This phenomenon might be due to numerical effects and might be preventable by a smaller learning rate.
-
-\autoref{tab:adv_accuracies} depicts the performance for FGSM and $\ell_\infty$-bounded \gls{pgd}. 
-As expected, while FGSM is the fastest of all algorithms, it delivers the worst performance. Of the investigated algorithms,  \gls{cw}  delivered the best performance.
-\autoref{fig:adv_per_family} depicts how successful the attack is for individual attack families.
-The table shows significant differences for detecting different attack families in the first place.
-However, also for generating adversarial flows some attack families are more susceptible than others.
-
-For \autoref{tab:adv_accuracies} and \autoref{fig:adv_per_family} we used \gls{cw} with a tradeoff of 1.
-To provide a fair comparison, for \gls{fgsm} and \gls{pgd} we used an $\ell_\infty$ distance which we observed for \gls{cw}.
-
-It is interesting to note that also increasing the $\ell_\infty$ bound for \gls{pgd} did not result in significantly improved performance.
-We can thus confirm the recommendation by \cite{carlini_towards_2017} to perform gradient descent on the logit output instead of the loss.
 

### 2019-12-12 ex3/paper/paper.tex
Change bib style to IEEE

@@
 \usepackage{graphicx}
-\usepackage{natbib}
+\usepackage[numbers]{natbib}
 \usepackage{subfig}
@@
 \begin{tabular}{l r r r r} \toprule
-& \multicolumn{2}{c}{CIC-IDS-17} & \multicolumn{2}{c}{UNSW-NB-15} \\
+& \multicolumn{2}{c}{CIC-IDS-2017} & \multicolumn{2}{c}{UNSW-NB15} \\
 	&	Packet	&	Flow	&	Packet	&	Flow	\\	\midrule
@@
 
-In particular,  in literature the use of Partial Dependency Plots (\gls{pdp}) is proposed.
-If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space and $f(X) \in [0,1]$ denotes the neural network's prediction, the \gls{pdp} for feature $i$ can be defined as
-\begin{equation} \label{eq:pdp}
-\text{\gls{pdp}}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right)
-\end{equation}
+In particular,  in literature the use of Partial Dependency Plots (\gls{pdp}) is proposed~\cite{friedman_greedy_2001}.
+To inspect attack types in detail, in this research we experiment with a modified definition of the \gls{pdp}.
+If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space, $f(X) \in [0,1]$ denotes the neural network's prediction and $c$ the attack class, the \gls{pdp} for feature $i$ can be defined as
+% \begin{equation} \label{eq:pdp}
+% \text{\gls{pdp}}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right)
+% \end{equation}
 
-To inspect attack types in detail, in this research we experiment with a modified definition of the \gls{pdp}  (with C being the attack classes):
+% To inspect attack types in detail, in this research we experiment with a modified definition of the \gls{pdp}  (with C being the attack classes):
 \begin{equation} \label{eq:pdp_conditional}
@@
 \begin{itemize}[topsep=0pt,wide,labelwidth=!,labelindent=0pt]
-\item Leaving out all features that are manipulable: packet size and \gls{iat}
+\item Leaving out all manipulable features: packet size and \gls{iat}
 \item Leaving out manipulable features when the attacker can manipulate them: only in the direction from the attacker to the victim (This, however does not prevent adversarial samples for botnets, for which both sides are malicious)
@@
 
-\bibliographystyle{plain}
-\bibliography{bibliography}
-
 \section*{Acknowledgements}
 The Titan Xp used for this research was donated by the NVIDIA Corporation.
+
+\renewcommand*{\bibfont}{\small}
+\bibliographystyle{ieeetr}
+\bibliography{bibliography}
 

### 2019-12-12 ex3/paper/paper.tex
Several text changes

@@
 \subsection{Carlini-Wagner}
-We implemented the \gls{cw} \cite{carlini_towards_2017}, which performs gradient descent on the optimization objective
+We implemented the \gls{cw} \cite{carlini_towards_2017}, performing gradient descent on the optimization objective
 \begin{equation} \label{eq:carliniWagner}
-d(X,\tilde X) + \epsilon  \max(Z(\tilde X), 0).
+d(X,\tilde X) + \epsilon  \max(Z(\tilde X), \delta).
 \end{equation}
@@
 
+$\delta \in \mathbb R$ determines the logit output of successful adversarial.
+In the original publication $\delta=0$, which means that the network's decision for adversarial samples is just between attack and normal traffic.
+In the present context, we need to make sure that the classifier's prediction would actually be normal traffic, even though a certain level of noise will be added to \glspl{iat} due to the network between attacker and victim.
+Hence, we introduced $\delta=-0.2$, corresponding to a confidence of 55\% for the sample to be benign.
+
 We used $\ell_1$ norm as distance metric and \gls{pgd} for meeting the real-world constraints discussed above.
 
-The method minimizes the adversarial sample's distance so that the network's decision is just between attack and normal traffic.
+% The method minimizes the adversarial sample's distance so that the network's decision for adversarial samples is just between attack and normal traffic.
 In the present context, we need to make sure that the classifier's prediction would actually be normal traffic, even though a certain level of noise will be added to \glspl{iat} due to the network between attacker and victim.
 Instead of the above equation \ref{eq:carliniWagner} we therefore used the optimization objective
-\begin{equation}
-d(X,\tilde X) + \epsilon  \max(Z(\tilde X), -0.2),
-\end{equation}
-i.e. we require a slight level of confidence that our adversarial samples are not classified as attack traffic. A logit value of $-0.2$ corresponds to a confidence of 55\% for the sample to be benign.
+% \begin{equation}
+% d(X,\tilde X) + \epsilon  \max(Z(\tilde X), -0.2),
+% \end{equation}
+% i.e.
+we require a slight level of confidence that our adversarial samples are not classified as attack traffic.
+A logit value of $-0.2$ corresponds to a confidence of 55\% for the sample to be benign.
 
@@
 In particular,  in literature the use of Partial Dependency Plots (\gls{pdp}) is proposed~\cite{friedman_greedy_2001}.
-To inspect attack types in detail, in this research we experiment with a modified definition of the \gls{pdp}.
-If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space, $f(X) \in [0,1]$ denotes the neural network's prediction and $c$ the attack class, the \gls{pdp} for feature $i$ can be defined as
+To inspect attack types in detail, in this research we use a \textit{conditional} form of the \gls{pdp}.
+If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space, $f(\boldsymbol X) \in [0,1]$ the neural network's prediction and $c$ the attack class, we define the conditional \gls{pdp} for feature $i$  as
 % \begin{equation} \label{eq:pdp}
@@
 \begin{equation} \label{eq:pdp_conditional}
-\text{\gls{pdp}}_{c,i}(y) = \mathbb E_{\boldsymbol X | C = c} \left(f \left( X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n \right) \right)
+\text{\gls{pdp}}_{c,i}(w) = \mathbb E_{\boldsymbol X | C} \Big(f \left( X_1,\ldots,X_{i-1},w,X_{i+1},\ldots X_n \right) | c\Big),
 \end{equation}
+empirically approximating the conditional distribution by the observed subset that belongs to a certain class.
+
+By using a classical \gls{pdp} we would be likely to loose information due to the averaging over very different types of traffic.
+However, for network traffic in particular, investigating each sample individually is not possible.
+Hence, the conditional \gls{pdp} provides the ideal compromise for our application.
+
 %
@@
 %
-In this variant of the \gls{pdp} that we call conditional \gls{pdp}, we don't take the whole dataset for the computation for the \gls{pdp} but only a subset that belongs to a certain class.
+%In this variant of the \gls{pdp} that we call conditional \gls{pdp}, we don't take the whole dataset for the computation for the \gls{pdp} but only a subset that belongs to a certain class.
 Specifically, we take each attack class and compute the \gls{pdp} only over samples of this attack class.
-The results (omitted for brevity) show that for some attack classes the port numbers play an important role.
-When looking at the class of normal traffic (non-attack traffic) it becomes apparent that traffic destined to a high source port is generally indicative of an attack.
-We argue that this is because most services that regular users use have low port numbers.
-A known issue about \glspl{pdp} in general is that dependencies of the feature under investigation and the remaining variables are not considered, hence potentially querying the neural network for feature combinations which cannot possibly occur.
+
+The results (omitted for brevity) show that for some attack classes the port numbers play an important role. When looking at the class of benign traffic %(non-attack traffic)
+it becomes apparent that traffic destined to a high source port is generally indicative of an attack. We argue that this is because most services that regular users use have low port numbers.
+
+%AH. As ALE is removed, I case this also isnt needed
+%A known issue about \glspl{pdp} in general is that dependencies of the feature under investigation and the remaining variables are not considered, hence potentially querying the neural network for feature combinations which cannot possibly occur.
 For a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.

### 2019-12-12 ex3/paper/paper.tex
Some shortening

@@
 
-As \autoref{fig:adv_per_family} and \autoref{tab:adv_accuracies} show, for most attack types, most samples can be modified by an adversary to be classified as benign by using \gls{cw}.
+As \autoref{fig:adv_per_family} shows, for most attack types, most samples can be modified by an adversary to be classified as benign by using \gls{cw}.
 High accuracy for a particular attack category does not imply that adversarial attacks are  likely to fail for these attacks.
@@
 Often an \gls{rnn} produces an output at each time step.
-When trying to apply explainability methods, hence, the question arises which output to base the methods on.
+When  applying explainability methods, the question arises which output to consider for the method.
 The natural choice is to base the methods on the prediction output which occurs at the same time step as the feature under investigation.
-Not only does this approach require a substantially lower amount of computational resources compared to predictions occurring at a later time step, but we also expect the immediately occurring prediction outcome to exhibit a stronger correlation to the feature under investigation, compared to a prediction outcome occurring many time steps later.
-However, due to the complex decision processes that can occur for deep neural networks, it is absolutely  uncertain if the network's final prediction is correctly depicted when following this approach.
+Not only does this approach require a substantially lower amount of computational resources compared to predictions occurring at a later time step, but we also expect the immediately occurring prediction outcome to exhibit a stronger dependence of the investigated feature, compared to a prediction outcome occurring many time steps later.
+However, due to the complex decision processes of deep neural networks,
+it is absolutely  uncertain if the network's final prediction is correctly depicted when following this approach.
 \end{itemize}
@@
 
- The most commonly used techniques for feature importance which are used by practitioners for \glspl{rnn} \cite{stackexchange_cross_validated_neural_2019} and \gls{dl} \cite{molnar_interpretable_2019,stackexchange_cross_validated_feature_2016} in general are based on adding noise to a particular feature and then evaluating the drop in accuracy that occurs.
-We argue that it is hard to find out the ``correct'' intensity of noise to add, we slightly vary the proposed approach: We do not add noise drawn from a probability distribution but instead sample the value for a feature from the distribution of all values of this feature.
-This makes the method non-parametric since the distribution of the noise does not need to be chosen.
+ In general, the most commonly deployed feature importance techniques, used by practitioners for \glspl{rnn} \cite{stackexchange_cross_validated_neural_2019} and \gls{dl} \cite{molnar_interpretable_2019,stackexchange_cross_validated_feature_2016,olden_accurate_2004}, are based on adding noise to a particular feature and evaluating the drop in accuracy that occurs.
+We argue that it is hard to determine the ``correct'' intensity of noise to add.
+%, we slightly vary the proposed approach: We do not add noise drawn from a probability distribution but instead
+ Hence, we sample the value for a feature from the distribution of all values of this feature in the dataset.
+This makes the method non-parametric since the noise distribution  does not need to be chosen.
 We ensured that features which stay constant for a flow, i.e.
-source port, destination port and protocol, also stay constant throughout the whole flow when randomizing the feature.
+source port, destination port and protocol, also stay constant throughout the flow when randomizing the feature.
 
 \subsubsection{Feature Dropout}
-While the method of adding noise to features/randomizing them is convenient since it is easy to implement and understand, we argue that it has some shortcomings: 
-The \gls{rnn} was never trained for dealing with ``garbage'' values that the noise/randomization methods create.
+While the perturbation method %of adding noise to features/randomizing them
+is convenient since it is easy to implement and understand, we argue that it has some shortcomings: 
+The \gls{rnn} was never trained for dealing with ``garbage'' values that the randomization creates.
 For example, completely unrealistic feature combinations could occur that were never observed during training.
@@
 This procedure is equivalent to using dropout \cite{srivastava_dropout:_2014} with probability $\frac{1}{n}$ before the first layer.
-An important implementation detail here is that for each feature we add another input which is 1 if the feature is suppressed and 0 otherwise.
+
+An important implementation detail is that for each feature we add another input which is 1 if the feature is suppressed and 0 otherwise.
 This is necessary so that the neural network can distinguish between a feature missing and a feature genuinely being zero.
@@
 As it can also happen that several features are left out for a sample, our method can also be used to analyze the effect of multiple features missing and can thus show possibly correlated features.
-The results in Table~\ref{fig:featDropout} show that, unlike the randomization feature importance method, the feature dropout method does not vastly overestimate features' importance.
-With feature dropout, it becomes apparent that there are very few features that actually contain unique information and that affect accuracy when left out: the destination port and the source port.
-A model trained with feature dropout typically yields slightly lower accuracy than a regularly trained model, even if no features are left out (flow accuracy of 0.9943 vs.~0.9965.
+The results in Table~\ref{fig:featDropout} show that, unlike input perturbation, feature dropout  does not vastly overestimate features' importance.
+With feature dropout, it becomes apparent that only very few features  actually contain unique information, affecting accuracy when left out: the destination port and the source port.
+
+A model trained with feature dropout typically yields slightly lower accuracy than a regularly trained model, even if no features are left out (flow accuracy of 0.9943 vs.~0.9965).
 We thus recommend training two models: One regular one and one with feature dropout to use for the feature importance.
-Another method that uses dropout for feature importance is \cite{chang_dropout_2017}, which, however, applies a technique called \textit{Variational Dropout} to learn the optimal dropout rate for each feature.
+
+Another method that uses dropout for feature importance is \cite{chang_dropout_2017}, applying a technique called \textit{Variational Dropout} to learn the optimal dropout rate for each feature.
 It tries to leave out as many features as possible and at the same time keep accuracy high.
@@
 
-Note that metrics for UNSW-NB15 differ substantially from CIC-IDS2017.
-However, due to the large number of different network attacks and network configurations, it is easily possible that different features are relevant for both datasets.
-We consider the question of model transferability of substantial importance for the presented application, but out of scope for the present research.
+Metrics for UNSW-NB15 differ substantially from CIC-IDS-2017.
+However, due to the large number of different network attacks and network configurations, it is easily possible that relevant features are very different.
+We consider the question of model transferability of substantial importance for \gls{ids} applications, but out of scope for the present research.
 

### 2019-12-12 ex3/paper/paper.tex
Small paper changes

@@
 Another interesting observation is that after the first couple of packets the classifier is certain about its decision and changing features doesn't change its confidence any longer.
-This allows the conclusion that if an attacker is able to change the first few packets, she could possibly change the decision of the classifier.
-We will further investigate this question in \autoref{sec:adv}.
+This allows the conclusion that if an attacker is able to change the first few packets, she could change the decision of the classifier.
+This is consistent with our observations in \autoref{sec:adv}.
+%We will further investigate this question in \autoref{sec:adv}.
 
@@
 For both datasets, after a small number of epochs, it was no longer possible to create a significant number of adversarial samples for most attack categories.
-For the datasets we considered, adversarial training left classification performance unaffected.
+%For the datasets we considered, adversarial training left classification performance unaffected.
 %\todo{Max: Insert results that show that accuracy on normal data is still ok after adversarial training.}
-
 \autoref{tab:adv_training_performance} depicts classification on non-adversarial data after 50 epochs of adversarial training.

### 2019-12-13 ex3/paper/paper.tex
Updated tradeoff plot
Changed intro and abstract

@@
 
-\glspl{rnn} yield attractive properties for performing intrusion detection on network data.
+\glspl{rnn} yield attractive properties for %performing intrusion detection on network data.
+constructing \glspl{ids} for network data.
 However, with the rise of ubiquitous \gls{ml} systems, malicious actors have been catching up quickly to find new ways to exploit \gls{ml} for profit.
-Recently developed \gls{aml} techniques focus on computer vision and their applicability to network traffic is not straightforward: Network packets expose fewer features than an image, are sequential and impose several constraints on their features.
-For example, destination ports cannot be modified since that could impair communication.
+Recently developed adversarial \gls{ml} techniques focus on computer vision and their applicability to network traffic is not straightforward: Network packets expose fewer features than an image, are sequential and impose several constraints on their features.
+%For example, destination ports cannot be modified since that could impair communication.
 
 We show that despite these completely different characteristics, adversarial samples can be reliably generated for \glspl{rnn}.
-Since traditional evaluation metrics such as accuracy are not sufficient for quantifying the adversarial threat, we propose the \gls{ars}.
+To understand a classifier's potential for misclassification, we extend existing explainability techniques and propose new ones, suitable particularly % specifically crafted %AH most can also be used for non-sequential models, so lets not restrict applicability too much
+for sequential data.
+Applying them shows that already the first packets of a communication flow are of crucial importance and are likely to be targeted by attackers.
+Feature importance methods show that even relatively unimportant features can be effectively abused to generate adversarial samples.
+Since traditional evaluation metrics such as accuracy are not sufficient for quantifying the adversarial threat, we propose the \gls{ars} for comparing \glspl{ids}, capturing a common notion of adversarial robustness, and show that an adversarial training procedure can significantly and successfully reduce the attack surface.% and improve \gls{ars}.
 %as a countermeasure.
@@
 the score alone is not yet a countermeasure
-To understand a classifier's potential for misclassification, we extend existing explainability techniques and propose new ones, particularly suitable % specifically crafted %AH most can also be used for non-sequential models, so lets not restrict applicability too much
-for sequential data.
-Applying them shows that already the first packets of a communication flow are of crucial importance and are likely going to be targeted by attackers.
-Feature importance methods show that even relatively unimportant features can be effectively abused to generate adversarial samples.
-As a countermeasure, we develop an easy-to-use adversarial training procedure %and can confirm that it
-which significantly and successfully reduces the attack surface and improves \gls{ars}.
+%As a countermeasure, we develop an easy-to-use adversarial training procedure %and can confirm that it
+%which significantly and successfully reduces the attack surface and improves \gls{ars}.
 
@@
 More recently, with the rise of omnipresent encryption, available features have changed.
-Since payload is no longer avilable, the focus now lies on features that are always available to network monitoring equipment such as port numbers, protocol flags, packet size etc.
+Since payload is no longer available, the focus now lies on features that are always available to network monitoring equipment such as port numbers, protocol flags or packet sizes.
 
-Network communication is usually grouped into \textit{flows}, which are commonly defined as a sequence of packets sharing certain properties, between two applications that may be located at different physical locations.
+Network communication is usually grouped into \textit{flows}, which are commonly defined as a sequence of packets sharing certain properties, between two applications potentially being located at different physical locations.
 When analyzing flows, not only the aforementioned features are available but also features related to the timing of the individual packets.
@@
 %AH: don't see why pkt size is not available without flows
-For example, a specific attack might require the attacker to send three small packets and the victim to reply with two large packets followed by a small one to the attacker.
+%For example, a specific attack might require the attacker to send three small packets and the victim to reply with two large packets followed by a small one to the attacker.
 Various approaches have been proposed to extract features from flows and then perform anomaly detection with the extracted flows \cite{meghdouri_analysis_2018}.
@@
 
+However, for practical use, high accuracy is not enough.
 With the recent rise of interest in \gls{aml} techniques, also \gls{aml} for \glspl{ids} has been investigated.
 For example, \cite{bachl_walling_2019} investigates remedies for poisoning attacks on \glspl{ids}.
-In this research, we investigate whether adversarial samples can be found for our \gls{rnn}-based model.
+In this research, we investigate whether adversarial samples can be found for our \gls{rnn}-based model, i.e.
+minimally different attack flows which are classified as benign.
 This is not a trivially answerable question since adversarial samples have mostly been analyzed in the context of computer vision.
@@
 In practice, a further crucial requirement for \gls{ml}-based \glspl{ids} 
-is that the classifiers decisions can be interpreted, recently giving rise to increasing interest in explainability methods.
+is that the classifier's decisions are interpretable, recently giving rise to increasing interest in explainability methods.
 Also in this case, common methods are designed to work with images or tabular data but not with sequences.
 Hence, we extend existing explainability methods such as \glspl{pdp} \cite{friedman_greedy_2001} for sequential data.
-These show the varying characteristics of different attacks and they aid with visualizing potentially vulnerable features for adversaries.
-Astonishingly, feature importance methods reveal that the manipulable features, that are left for modification by adversaries are not even particularly significant for the classification outcome.
-%Especially the first packets of a flow are critical for determining the decision of a recurrent \gls{ids}.
+%These show the varying characteristics of different attacks and they aid with visualizing potentially vulnerable features for adversaries.
+Astonishingly, feature importance methods reveal that
+features manipulable for adversarial flows,
+%the that are manipulable features, that are left for modification by adversaries
+are not even particularly important for the \gls{rnn}'s classification outcome. %Especially the first packets of a flow are critical for determining the decision of a recurrent \gls{ids}.
 
@@
 %develop a simple method for defending against adversaries who can manipulate arbitrary packets by simply leaving out potentially manipulable information from the classification process.
-On the one hand, by simply leaving out manipulable features, we obtain a classifier which is slightly less accurate but is still useful to be deployed in a real scenario.
-Furthermore, we develop a custom adversarial training procedure which especially focuses on ease of use and low resource consumption and aims to keep the number of additionally introduced training hyperparameters low.
-With this novel strategy we can reduce the attack surface and harden the resulting \gls{ids} while retaining similar classification performance.
-% like for the original \gls{ml} model.
+On the one hand, by simply leaving out manipulable features, we obtain a classifier which is slightly less accurate but is still useful to be deployed in a real scenario. Furthermore, 
+by deploying an adversarial training procedure,
+%AH. compared to other research about adversarial training, I dont feel that out approach is so special
+%develop a custom adversarial training procedure which especially focuses on ease of use and low resource consumption and aims to keep the number of additionally introduced training hyperparameters low.
+With this novel strategy
+we can reduce the attack surface and harden the resulting \gls{ids} while retaining similar classification performance. % like for the original \gls{ml} model.
 
@@
 \label{sec:adv}
-We now investigate whether known \gls{aml} methods can be used for generating adversarial flows, i.e. minimally different attack flows which are classified as benign, for our \gls{rnn}.
+We now investigate whether known \gls{aml} methods can be used for generating adversarial flows for our \gls{rnn}..
 % In our scenario, most features such as, e.g., \gls{tcp} flags cannot be easily manipulated because their manipulation might violate the \gls{tcp} and thus cause communication to fail.
@@
 \caption{Performance of \gls{cw} with different $\epsilon$ values.}
-\todo{I guess the irregularity at eps=1 hints at too few iterations. Currently rerunning with 1000 base iterations}
 \label{fig:adv_cw}
@@
 The solid line shows the mean value of the feature for all flows of the attack type.
-The shaded region shows the change in confidence that occurs when the feature is varied.
-\todo{IAT plot is wrong: There shouldn't be step 0.}}
+The shaded region shows the change in confidence that occurs when the feature is varied.}
+%\todo{IAT plot is wrong: There shouldn't be step 0.} %AH. guess this is fine now
 \label{fig:pred_plots2}

### 2019-12-13 ex3/auxplot.py
Updated ars plot

@@ -86,7 +86,8 @@ def adv_results():
 
 def ars():
 	plt.figure(figsize=(5,2))
-	plt.plot(group_names, values)
+	x_values = [ float(value) for value in group_names ]
+	plt.plot(x_values, values)
 	plt.legend(feature_names, ncol=2)
 	plt.xlabel('Training duration in epochs')
 	plt.ylabel('ARS')
@@ -107,7 +108,7 @@ def adv():
 	plt.gca().set_ylabel_legend(Line2D([0],[0], color='gray', linestyle='--'), handlelength=1.4)
 	plt.legend(lines, ['CIC-IDS-2017', 'UNSW-NB15'])
 	ylim1,ylim2 = plt.ylim()
-	plt.ylim((ylim1,12))
+	plt.ylim((ylim1,11))
 	#  plt.ylim((ylim1,ylim2+20)) # move plots away from legend
 	plt.tight_layout()
 

### 2019-12-13 ex3/paper/paper.tex
Undid ars motivation change

@@
 Setting the threshold for the \gls{ars} to 50\% is arbitrary, but
- reasonable as this way the \gls{ars} approximates the median of the distances of all samples and, additionally, marks the distance where an adversarial attack is more likely to be successful than unsuccessful.
-%we think that it is reasonable since if an attacker can make the majority of samples adversarial, we think that its attack is ``successful''.
+ % reasonable as this way the \gls{ars} approximates the median of the distances of all samples and, additionally, marks the distance where an adversarial attack is more likely to be successful than unsuccessful.
+we think that it is reasonable since if an attacker can make the majority of samples adversarial, we think that its attack is ``successful''.
 

### 2019-12-13 ex3/learn.py
Slight extension of adv_until_less_than_half

@@ -1197,7 +1197,9 @@ def adv_until_less_than_half():
 
 	reverse_mapping = {v: k for k, v in mapping.items()}
 	distances_packets = [None]*len(orig_results)
+	max_distance_packets = [None]*len(orig_results)
 	distances_flows = [None]*len(orig_results)
+	max_distance_flows = [None]*len(orig_results)
 	final_ratios = [None]*len(orig_results)
 	for i in range(len(prev_results)):
 		for attack_index, ratio in enumerate(prev_ratios[i]):
@@ -1214,7 +1216,9 @@ def adv_until_less_than_half():
 
 				distances_per_packet = [dist/len(flow) for dist, flow in zip(distances[lower_part], prev_flows[i][attack_index])]
 				distances_flows[attack_index] = float(np.mean(distances[lower_part]))
-				distances_packets[attack_index] = float(np.mean(distances_per_packet))
+				max_distance_flows[attack_index] = float(distances[lower_part[-1]])
+				distances_packets[attack_index] = float(distances_per_packet)
+				max_distance_packets[attack_index] = float(distances_per_packet[-1])
 
 	for attack_index in range(len(distances_packets)):
 		if len(orig_results[attack_index]) <= 0:
@@ -1225,7 +1229,9 @@ def adv_until_less_than_half():
 			"flow_accuracy", np.mean(np.round(numpy_sigmoid(np.array([item[-1] for item in orig_results[attack_index]])))),
 			"packet_accuracy", np.mean(np.round(numpy_sigmoid(np.array([sublist for l in orig_results[attack_index] for sublist in l])))),
 			"flow_distance", distances_flows[attack_index] if ratio <= THRESHOLD else 0,
-			"packet_distance", distances_packets[attack_index] if ratio <= THRESHOLD else 0
+			"packet_distance", distances_packets[attack_index] if ratio <= THRESHOLD else 0,
+			"max_flow_distance", max_distance_flows[attack_index] if ratio <= THRESHOLD else 0,
+			"max_packet_distance", max_distance_packets[attack_index] if ratio <= THRESHOLD else 0
 		)
 
 def eval_nn(data):

### 2019-12-13 ex3/learn.py
...and bug fix for previous commit

@@ -1215,10 +1215,10 @@ def adv_until_less_than_half():
 				lower_part = correct_indices[:int(math.ceil(len(distances)*min(1-ratio, THRESHOLD)))]
 
 				distances_per_packet = [dist/len(flow) for dist, flow in zip(distances[lower_part], prev_flows[i][attack_index])]
-				distances_flows[attack_index] = float(np.mean(distances[lower_part]))
-				max_distance_flows[attack_index] = float(distances[lower_part[-1]])
-				distances_packets[attack_index] = float(distances_per_packet)
-				max_distance_packets[attack_index] = float(distances_per_packet[-1])
+				distances_flows[attack_index] = float(np.mean(distances[lower_part]) if len(lower_part) else np.nan)
+				max_distance_flows[attack_index] = float(distances[lower_part[-1]] if len(lower_part) else np.nan)
+				distances_packets[attack_index] = float(np.mean(distances_per_packet) if len(distances_per_packet) else np.nan)
+				max_distance_packets[attack_index] = float(distances_per_packet[-1] if len(distances_per_packet) else np.nan)
 
 	for attack_index in range(len(distances_packets)):
 		if len(orig_results[attack_index]) <= 0:

### 2019-12-13 ex3/plot.py
Changed confidence plot

@@ -1,6 +1,7 @@
 #!/usr/bin/env python3
 
 import matplotlib.pyplot as plt
+from matplotlib.patches import Rectangle
 import numpy as np
 import sys
 import os
@@ -9,6 +10,12 @@ import pickle
 
 DIR_NAME = "plots/plot"
 ORDERING = ['', 'Length min', 'IAT min', 'Length max', 'IAT max']
+
+MAX_X = 18
+SHOW_TITLE = False
+
+plt.rcParams["font.family"] = "serif"
+
 
 dataroot_basename = sys.argv[1].split('_')[0]
 
@@ -33,7 +40,22 @@ assert lens_results == lens_indices
 
 colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
 
+all_seqs = []
+all_attacks = []
 for attack_type, seqs in enumerate(results_by_attack_number):
+	if reverse_mapping[attack_type] == 'Normal':
+		all_seqs.extend([ 1-seq for seq in seqs ])
+	else:
+		all_seqs.extend(seqs)
+		all_attacks.extend(seqs)
+	
+reverse_mapping[len(results_by_attack_number)] = 'All samples'
+results_by_attack_number.append(all_seqs)
+reverse_mapping[len(results_by_attack_number)] = 'All attacks'
+results_by_attack_number.append(all_attacks)
+	
+for attack_type, seqs in enumerate(results_by_attack_number):
+	print (attack_type, reverse_mapping[attack_type], len(seqs))
 	if len(seqs) <= 0:
 		continue
 
@@ -50,6 +72,9 @@ for attack_type, seqs in enumerate(results_by_attack_number):
 		for seq in seqs:
 			if len(seq) < i+1:
 				break
+				
+			if len(seq.shape) == 1:
+				seq = seq[:,None]
 
 			values_by_length[i].append(seq[i:i+1,:])
 
@@ -64,6 +89,8 @@ for attack_type, seqs in enumerate(results_by_attack_number):
 	# stds = np.array([np.std(item, axis=0) for item in values_by_length])
 	first_quartiles = np.array([np.quantile(item, 0.25, axis=0) for item in values_by_length])
 	third_quartiles = np.array([np.quantile(item, 0.75, axis=0) for item in values_by_length])
+	tens_percentiles = np.array([np.quantile(item, 0.1, axis=0) for item in values_by_length])
[…]

### 2019-12-13 ex3/paper/paper.tex
Changed confidence plot
Changed sequence plots

@@
 
-As apparent from the above discussion, trying to explain how packet features influence the model's predictions, poses various difficulties.
+%As apparent from the above discussion, trying to explain how packet features influence the model's predictions, poses various difficulties.
 As a first step to explaining our model's predictions, we investigate how important features at different time steps are for the model's decisions.
@@
 %\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_pred_plots2_outcomes_0_3.pickle_5_DoS - DDoS-DoS Hulk"}.pdf}
-\includegraphics[width=\columnwidth]{{"../plots/plot/flows_pred_plots2_outcomes_0_3.pickle_6_DoS - DDoS-DoS Slowhttptest"}.pdf}
+\includegraphics[width=\columnwidth]{{"../plots/plot/flows_pred_plots2_outcomes_0_3.pickle_15_All samples"}.pdf}
+% \includegraphics[width=\columnwidth]{{"../plots/plot/flows_pred_plots2_outcomes_0_3.pickle_6_DoS - DDoS-DoS Slowhttptest"}.pdf}
 % \includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_14_Web Attack-XSS"}.pdf}
@@
 
-\caption{Median confidence per time step and first and third quartiles. For the majority of attacks types, confidence increases in the first few steps and then stays almost constant at 1.}
+\caption{Classifier confidence per time step for CIC-IDS-2017. % and first and third quartiles.
+For the majority of attack types, confidence increases in the first few steps and then stays almost constant at 1.}
 \label{fig:confidence}
@@
 
+For investigating, how features influence the prediction outcome, we extend \glspl{pdp} to the sequential setting. 
+Denoting as $X=\{\boldsymbol X^1, \ldots, \boldsymbol X^m\}$ the series of packet features $\boldsymbol X^t \in \mathbb R^n$ and $h_t(X)$ the network's hidden state at time step $t$, we define the sequential \gls{pdp} as
+\begin{align}
+\text{PD}&\text{P}_{c,i}(t,w)= \\ &\mathbb E_{X | C} \Big(f \left(h_{t-1}( X), X_1^t,\ldots,X_{i-1}^t,w,X_{i+1}^t,\ldots X_n^t \right) | c\Big). \nonumber 
+\end{align}
+
+\autoref{fig:pred_plots2} shows an example together with the mean values for both unmodified samples and adversarial samples.
+Also in this figure we notice that mainly the first few packets are able to influence the prediction outcome and modifying features at a later time step does not change its confidence any longer.
+In many cases, the adversarial sample generation indeed moves packet features to areas where the network is less likely to be classified as attacks, confirming the effectiveness of \glspl{pdp}.
+In other cases, however, we did not observe an agreement between \gls{pdp} and adversarial modifications, hinting at dependencies which cannot be presented by \glspl{pdp}.
+
 \begin{figure}
@@
 
-\caption{Exemplary PD plot and adversarial flows for the DoS Slowloris attack in CIC-IDS-2017.
+\caption{Exemplary sequential PD plot and adversarial flows for the DoS Slowloris attack in CIC-IDS-2017.
 The solid line shows the mean value of the feature for all flows of the attack type.
@@
 
-Next, we want to investigate if our classifier just memorized its inputs.
+%Next, we want to investigate if our classifier just memorized its inputs.
 The most interesting packet features that can be used to visualize this in our use case are the \gls{iat} between packets in a flows and the packet length.
@@
 For a given flow $X$ we then varied feature values at each time step between minimum and maximum and plot how the confidence changes when the feature varies (\autoref{fig:pred_plots2}) depict the averaged predictions for different attack families.
-It seems clear that the classifier didn't just memorize the inputs but actually learned them since there is always a smooth decrease of confidence if we change the feature values.
-Instead, when memorizing, we would expect a sharp decline of confidence to occur.
-Another interesting observation is that after the first couple of packets the classifier is certain about its decision and changing features doesn't change its confidence any longer.
+
+Furthermore, it seems clear that the classifier did not just memorize the inputs but actually learned appropriate space representations, as there is always a smooth decrease of confidence when modifying feature values.
+When memorizing, we would  instead expect a sharp decline of confidence.
+% to occur.
+%Another interesting observation is that after the first couple of packets the classifier is certain about its decision and changing features doesn't change its confidence any longer.
 This allows the conclusion that if an attacker is able to change the first few packets, she could change the decision of the classifier.

### 2019-12-13 ex3/plot.py
Add tight bbox to confidence plot

@@ -151,6 +151,6 @@ for attack_type, seqs in enumerate(results_by_attack_number):
 	#plt.savefig('%s.pdf' % os.path.splitext(fn)[0])
 	# plt.show()
 	os.makedirs(DIR_NAME, exist_ok=True)
-	plt.savefig(DIR_NAME+'/{}_{}_{}.pdf'.format(file_name.split("/")[-1], attack_type, reverse_mapping[attack_type].replace("/", "-").replace(":", "-")))
+	plt.savefig(DIR_NAME+'/{}_{}_{}.pdf'.format(file_name.split("/")[-1], attack_type, reverse_mapping[attack_type].replace("/", "-").replace(":", "-")), bbox_inches = 'tight', pad_inches = 0)
 	plt.clf()
 

### 2019-12-13 ex3/paper/paper.tex
Changed sequence plots section

@@
 As a first step to explaining our model's predictions, we investigate how important features at different time steps are for the model's decisions.
-Intuitively, features at the beginning of a flow should be the most important while the model's predictions should not vary significantly anymore, as soon as the model has come to a decision.
-
+Intuitively, features at the beginning of a flow should be the most important while the classifier's predictions should not vary significantly anymore, as soon as it has come to a decision.
 \begin{figure}
@@
 % \includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
-
 \caption{Classifier confidence per time step for CIC-IDS-2017. % and first and third quartiles.
@@
 \end{figure}
+To evaluate this hypothesis, we plot the classifier's prediction confidence for each time step. % for each attack class.
+%averaging over all flows. %AH. there's no averaging happening
+% and also show the first and third quartiles.
+\autoref{fig:confidence} shows that %this hypothesis is correct for most attack classes. 
+while at the first couple of packets the confidence is not very high, towards the end of the flow it reaches values close to 1 and stays there.
+Hence, not only is the classifier able to make a reasonable classification after just a few packets, the figure also suggests that indeed later packets have a neglectible influence on the prediction.
 
-To determine if this hypothesis is correct, we plot the attack confidence of the classifier for each time step for each attack class.
-At each time step we average over all flows and also show the first and third quartiles.
-\autoref{fig:confidence} shows that this hypothesis is correct for most attack classes.
-At the first couple of packets the confidence for maliciousness is not very high yet but towards the end of the flow it reaches values close to 1 and stays there.
-
-For investigating, how features influence the prediction outcome, we extend \glspl{pdp} to the sequential setting. 
+For investigating in more detail how features influence the prediction outcome, we extend \glspl{pdp} to the sequential setting. 
 Denoting as $X=\{\boldsymbol X^1, \ldots, \boldsymbol X^m\}$ the series of packet features $\boldsymbol X^t \in \mathbb R^n$ and $h_t(X)$ the network's hidden state at time step $t$, we define the sequential \gls{pdp} as
@@
 
-\begin{figure}
+\begin{figure}[b]
 %\includegraphics[width=0.48\textwidth]{{"../plots/plot2_adv/flows_pred_plots2_outcomes_flows_adv_1.0_notBidirectional_outcomes_0_3.pickle_0_3.pickle_0_Botnet-ARES"}.pdf}
@@
 
-\caption{Exemplary sequential PD plot and adversarial flows for the DoS Slowloris attack in CIC-IDS-2017.
-The solid line shows the mean value of the feature for all flows of the attack type.
+\caption{Exemplary sequential PD plot and adversarial flows for the DoS Slowloris attack in CIC-IDS-2017. The lines show the feature's mean values. % for all flows of the attack type.
 The shaded region shows the change in confidence that occurs when the feature is varied.}
@@
 
+% AH. I guess this can only be properly formulated with a formula
 %Next, we want to investigate if our classifier just memorized its inputs.
@@
 \end{figure}
-
-Looking at the plots of the features that vary over time (\autoref{fig:plot_features}) we see that certain attack types have a characteristic pattern in which they send packets by which they are easily recognizable.
-We want to determine how the pattern of packet sizes and \glspl{iat} is significant for the classification.
+Finally, we investigated whether our classification task involves recognizing complex patterns in the feature space.
+As example, \autoref{fig:plot_features} shows that
+%Looking at the plots of the features that vary over time (\autoref{fig:plot_features}) we see that certain 
+attack types indeed have a characteristic pattern in which they send packets by which they are easily recognizable.
+Other attack families similarly show characteristic patterns.
+%We want to determine how the pattern of packet sizes and \glspl{iat} is significant for the classification.
 \autoref{fig:plot_features} clearly shows that some attack types have very characteristic patterns of packet sizes and \glspl{iat}.

### 2019-12-14 ex3/paper/paper.tex

@@
 High accuracy for a particular attack category does not imply that adversarial attacks are  likely to fail for these attacks.
-Thus we argue that beside the classical metrics such as accuracy, false positives etc., a new metric is needed, which quantifies how easily a classifier can be fooled.
+Thus we argue that beside the classical metrics such as accuracy, false positives etc., a new metric is needed, which quantifies how easily an \gls{ids} classifier can be fooled.
 
-Such a method should be easy to compute and easy to interpret.
-We thus propose the \gls{ars}, which roughly works as follows: Use \gls{cw} with a small tradeoff to try to generate adversarial samples for an attack type.
+Such a metric should be easy to compute, easy to interpret and robust with respect to outliers.
+We thus propose the \gls{ars}, as the average of the lower half  of minimum distances of adversarial samples.
+Since \gls{cw} is able to find adversarial samples with minimum distance, the \gls{ars} can
+  be found efficiently as follows: Use \gls{cw} with a small tradeoff to try to generate adversarial samples for an attack type.
 If more than half of the samples are now wrongly classified, stop.
@@
 
+%AH. TODO: I think this is quite technical curretnly and can be shortened a bit
 The result is a number of adversarial samples (more than half in the ideal case) and some that could not be made adversarial.
@@
  % reasonable as this way the \gls{ars} approximates the median of the distances of all samples and, additionally, marks the distance where an adversarial attack is more likely to be successful than unsuccessful.
-we think that it is reasonable since if an attacker can make the majority of samples adversarial, we think that its attack is ``successful''.
+ reasonable as outliers with very large distances are ignored and if an attacker can make the majority of samples adversarial, we think that its attack is ``successful''.
+%we think that it is reasonable since if an attacker can make the majority of samples adversarial, we think that its attack is ``successful''.
 
@@
 \begin{table}
-% Evaluated using
-% ./python learn.py --dataroot flows.pickle --net runs/Oct26_00-03-50_gpu/lstm_module_1284.pth --function test --batchSize 512
-% ./python learn.py --dataroot flows15.pickle --net runs/Oct28_15-41-46_gpu/lstm_module_997.pth --function test --batchSize 512
-% on Dec. 4th
 \caption{Detection performance after adversarial training.} \label{tab:adv_training_performance}

### 2019-12-14 ex3/paper/paper.tex
Some reference shortening

@@
 %\todo{Max: Insert results that show that accuracy on normal data is still ok after adversarial training.}
-\autoref{tab:adv_training_performance} depicts classification on non-adversarial data after 50 epochs of adversarial training.
-Accuracy is essentially identical to results presented in \autoref{tab:performance_results}.
-Due to the higher proportion of attack samples in training data, we note, however, that recall increased at the expense of a slightly lower precision.
+%\autoref{tab:adv_training_performance} depicts classification on non-adversarial data after 50 epochs of adversarial training.
+After 50 epochs of training, accuracy is essentially identical to results presented in \autoref{tab:performance_results}.
+Recall and Youden increased and precision slightly decreased.
+However, this is due to the higher proportion of attack samples in training data.
+We conclude that adversarial training is effective for reducing the attack surface for adversarial attacks in our scenario.
 
-\begin{table}
-\caption{Detection performance after adversarial training.} \label{tab:adv_training_performance}
-\todo{Probably no space for table}
-\centering
-\begin{tabular}{l r r r r} \toprule
-& \multicolumn{2}{c}{CIC-IDS-17} & \multicolumn{2}{c}{UNSW-NB-15} \\
-	&	Packet	&	Flow	&	Packet	&	Flow	\\	\midrule
-Accuracy	&	0.991	&	0.997	&	0.994	&	0.981	\\
-Precision	&	0.971	&	0.996	&	0.761	&	0.704	\\
-Recall	&	0.979	&	0.992	&	0.970	&	0.824	\\
-F1	&	0.975	&	0.994	&	0.853	&	0.759	\\
-Youden	&	0.973	&	0.991	&	0.965	&	0.811	\\
+% \begin{table}
+% \caption{Detection performance after adversarial training.} \label{tab:adv_training_performance}
+% \todo{Probably no space for table}
+% \centering
+% \begin{tabular}{l r r r r} \toprule
+% & \multicolumn{2}{c}{CIC-IDS-17} & \multicolumn{2}{c}{UNSW-NB-15} \\
+	% &	Packet	&	Flow	&	Packet	&	Flow	\\	\midrule
+% Accuracy	&	0.991	&	0.997	&	0.994	&	0.981	\\
+% Precision	&	0.971	&	0.996	&	0.761	&	0.704	\\
+% Recall	&	0.979	&	0.992	&	0.970	&	0.824	\\
+% F1	&	0.975	&	0.994	&	0.853	&	0.759	\\
+% Youden	&	0.973	&	0.991	&	0.965	&	0.811	\\
 
-\bottomrule
-\end{tabular}
-\end{table}
+% \bottomrule
+% \end{tabular}
+% \end{table}
 

### 2019-12-14 ex3/paper/paper.tex
Remove unnecessary table

--- 2019-12-14 ex3/paper/paper.tex
@@
 
-%AH. TODO: I think this is quite technical curretnly and can be shortened a bit
+\todo{Alex: I think this is quite technical curretnly and can be shortened a bit}
 The result is a number of adversarial samples (more than half in the ideal case) and some that could not be made adversarial.
@@
 \subsection{Feature Importance Metrics}
-\begin{table}
-\caption{Accuracy drop for:}
-\label{tab:feat_importance}
-\centering
-\subfloat[Random replacement method\label{fig:featRandom}
-]{
-\begin{tabular}{l r}
-\toprule
-Feature & \hspace*{-4mm}Accuracy~drop \\ \midrule
-Protocol & 0.207 \\
-Packet Length & 0.165 \\
-SYN Flag & 0.099 \\
-ACK Flag & 0.084 \\
-Direction & 0.073 \\
-Destination port & 0.071 \\
-Source port & 0.060 \\
-RST Flag & 0.057 \\
-PSH Flag & 0.056 \\
-Interarrival time & 0.024 \\
-FIN Flag & 0.012 \\
-URG Flag & 0 \\
-ECE Flag & 0 \\
-CWR Flag & 0 \\
-NS Flag & 0 \\
-\bottomrule
-\end{tabular}
-}{}
-\subfloat[Feature dropout\label{fig:featDropout}
-]{
-\begin{tabular}{l r}
-\toprule
-Feature & \hspace*{-4mm}Accuracy~drop \\ \midrule
-Destination port & 0.025 \\
-Source port & 0.003 \\
-RST Flag & 0.001 \\
-ACK Flag & 0.001 \\
-Protocol & 0.001 \\
-Packet Length & 0.001 \\
-Direction & 0.001 \\
-SYN Flag & 0 \\
-Interarrival time & 0 \\
-FIN Flag & 0 \\
-ECE Flag & 0 \\
-URG Flag & 0 \\
-CWR Flag & 0 \\
-NS Flag & 0 \\
-PSH Flag & 0 \\
-\bottomrule
-\end{tabular}
-}{}
-\end{table}
+%\begin{table}
+%\caption{Accuracy drop for:}
+%\label{tab:feat_importance}
+%\centering
+%\subfloat[Random replacement method\label{fig:featRandom}
+%]{
+%\begin{tabular}{l r}
+%\toprule
+%Feature & \hspace*{-4mm}Accuracy~drop \\ \midrule
+%Protocol & 0.207 \\
+%Packet Length & 0.165 \\
+%SYN Flag & 0.099 \\
+%ACK Flag & 0.084 \\
+%Direction & 0.073 \\
+%Destination port & 0.071 \\
+%Source port & 0.060 \\
+%RST Flag & 0.057 \\
+%PSH Flag & 0.056 \\
+%Interarrival time & 0.024 \\
+%FIN Flag & 0.012 \\
+%URG Flag & 0 \\
+%ECE Flag & 0 \\
+%CWR Flag & 0 \\
+%NS Flag & 0 \\
+%\bottomrule
+%\end{tabular}
+%}{}
+%\subfloat[Feature dropout\label{fig:featDropout}
+%]{
+%\begin{tabular}{l r}
+%\toprule
+%Feature & \hspace*{-4mm}Accuracy~drop \\ \midrule
+%Destination port & 0.025 \\
+%Source port & 0.003 \\
+%RST Flag & 0.001 \\
+%ACK Flag & 0.001 \\
+%Protocol & 0.001 \\
+%Packet Length & 0.001 \\
+%Direction & 0.001 \\
+%SYN Flag & 0 \\
+%Interarrival time & 0 \\
+%FIN Flag & 0 \\
+%ECE Flag & 0 \\
+%URG Flag & 0 \\
+%CWR Flag & 0 \\
+%NS Flag & 0 \\
+%PSH Flag & 0 \\
+%\bottomrule
+%\end{tabular}
+%}{}
+%\end{table}
 As a first step to understanding the neural network's decisions, we  estimate how important individual features are for the model's predictions.
@@
 As it can also happen that several features are left out for a sample, our method can also be used to analyze the effect of multiple features missing and can thus show possibly correlated features.
-The results in Table~\ref{fig:featDropout} show that, unlike input perturbation, feature dropout  does not vastly overestimate features' importance.
+The results in \autoref{fig:feat_imp} show that, unlike input perturbation, feature dropout  does not vastly overestimate features' importance.
 With feature dropout, it becomes apparent that only very few features  actually contain unique information, affecting accuracy when left out: the destination port and the source port.
@@
 \end{figure*}
-Table~\ref{fig:featRandom} and \autoref{fig:feat_imp} show the results, which match to a large extent with domain-specific expectations for classifying network flows.
+\autoref{fig:feat_imp} shows the results, which match to a large extent with domain-specific expectations for classifying network flows.
 In particular, rarely used \gls{tcp} flags like NS or URG are unimportant for the classifier.

### 2019-12-14 ex3/learn.py

@@ -1202,8 +1202,10 @@ def adv_until_less_than_half():
 	max_distance_flows = [None]*len(orig_results)
 	final_ratios = [None]*len(orig_results)
 	for i in range(len(prev_results)):
+		print(f"Looking at {i}th run")
 		for attack_index, ratio in enumerate(prev_ratios[i]):
 			if ratio > -float("inf"):
+				print(f"Looking at attack {attack_index} with a ratio of {ratio}")
 				final_ratios[attack_index] = ratio
 
 				successfully_changed_flows_mask = (np.round(numpy_sigmoid(np.array([item[-1] for item in prev_results[i][attack_index]]))) == 0).flatten()
@@ -1215,10 +1217,10 @@ def adv_until_less_than_half():
 				lower_part = correct_indices[:int(math.ceil(len(distances)*min(1-ratio, THRESHOLD)))]
 
 				distances_per_packet = [dist/len(flow) for dist, flow in zip(distances[lower_part], prev_flows[i][attack_index])]
-				distances_flows[attack_index] = float(np.mean(distances[lower_part]) if len(lower_part) else np.nan)
-				max_distance_flows[attack_index] = float(distances[lower_part[-1]] if len(lower_part) else np.nan)
-				distances_packets[attack_index] = float(np.mean(distances_per_packet) if len(distances_per_packet) else np.nan)
-				max_distance_packets[attack_index] = float(distances_per_packet[-1] if len(distances_per_packet) else np.nan)
+				distances_flows[attack_index] = float(np.mean(distances[lower_part]) if len(lower_part)>0 else np.nan)
+				max_distance_flows[attack_index] = float(distances[lower_part[-1]] if len(lower_part)>0 else np.nan)
+				distances_packets[attack_index] = float(np.mean(distances_per_packet) if len(distances_per_packet)>0 else np.nan)
+				max_distance_packets[attack_index] = float(distances_per_packet[-1] if len(distances_per_packet)>0 else np.nan)
 
 	for attack_index in range(len(distances_packets)):
 		if len(orig_results[attack_index]) <= 0:

### 2019-12-15 ex3/paper/paper.tex
Add ars results finally

--- 2019-12-14 ex3/paper/paper.tex
@@
 
-\todo{Insert results}
+\begin{figure}[h]
+\includegraphics[width=\columnwidth]{../plots/ars_original.pdf}
+\caption{Recall for each attack type and average modification per flow required by an adversary to make a flow appear benign for CIC-IDS-2017.}
+\label{fig:recall_ars}
+\end{figure}
+
+\autoref{fig:recall_ars} shows that certain attack types, like ``DoS Slowloris'' are easy to classify for an IDS but still are also surprisingly easy for an attacker to perform adversarial modifications to make them look benign.
 

### 2019-12-15 ex3/auxplot.py

@@ -25,8 +25,9 @@ feature_names = values[0]
 group_names = group_names[1:]
 values = values[1:]
 
-#  print (group_names)
-#  print (values)
+print("feature_names", feature_names)
+print("group_names", group_names)
+print("values", values)
 values = np.array(values, dtype=float)
 
 def importance():
@@ -83,6 +84,48 @@ def adv_results():
 	plt.ylabel('Recall')
 	plt.tight_layout()
 
+def ars_original():
+	global group_names
+	global values
+
+	y_labels = [None, "Recall", "Recall", "Distance", "Distance"]
+	value_indices_to_plot = (1,3)
+
+	# plt.figure(figsize=(5,3))
+	values = [[float(item) for item in sublist] for sublist in values]
+
+	index_by_which_to_sort = 1
+	old_len = len(group_names)
+	group_names, values = list(zip(*[(group_name, value) for group_name, value in zip(group_names, values) if value[0] <= 0.5]))
+	# print(f"Dropped {old_len-len(group_names)} elements.")
+	order = list(zip(*sorted(enumerate(values), key=lambda item: item[1][index_by_which_to_sort], reverse=True)))[0]
+	# print("order", order)
+	width = 0.75 / len(value_indices_to_plot)
+
+	values = np.array(values)
+
+	x = np.arange(len(group_names), dtype=float)
+
+	fig, ax1 = plt.subplots(figsize=(5,3))
+	plt.xticks(x + width/2, [group_names[i] for i in order], rotation=45, horizontalalignment='right')
+	for tick in plt.gca().xaxis.get_major_ticks():
+		label = tick.label1
+		label.set_transform(label.get_transform() + Affine2D().translate(8,0))
+	ax2 = ax1.twinx()
+	axes = [ax1, ax2]
+
+	all_labels = []
+	for index, i in enumerate(value_indices_to_plot):
+		label = axes[index].bar(x + width*index, values[order,i], width, color=colors[index], label=feature_names[i])
+		all_labels.append(label)
+		axes[index].set_ylabel(y_labels[i])
+
+	all_legends = [item.get_label() for item in all_labels]
+
+	# plt.legend(all_legends, all_labels, loc="upper right")
+	plt.legend(all_labels, all_legends, loc='lower center', bbox_to_anchor=(0.5,1), ncol=4)
+
+	plt.tight_layout()
 
 def ars():
[…]

### 2019-12-15 ex3/learn.py

@@ -1156,6 +1156,7 @@ def adv_until_less_than_half():
 	MAX = 10
 	SCALING = 0.5
 	THRESHOLD = 0.5
+	# FIXME: Max: I guess it should be < THRESHOLD not <= THRESHOLD as it is currently in the code
 	i = 0
 	prev_results = []
 	prev_flows = []

### 2019-12-15 ex3/learn.py
Fixed adv_util_less_than_half bug

@@ -1230,10 +1230,10 @@ def adv_until_less_than_half():
 			"ratio", final_ratios[attack_index],
 			"flow_accuracy", np.mean(np.round(numpy_sigmoid(np.array([item[-1] for item in orig_results[attack_index]])))),
 			"packet_accuracy", np.mean(np.round(numpy_sigmoid(np.array([sublist for l in orig_results[attack_index] for sublist in l])))),
-			"flow_distance", distances_flows[attack_index] if ratio <= THRESHOLD else 0,
-			"packet_distance", distances_packets[attack_index] if ratio <= THRESHOLD else 0,
-			"max_flow_distance", max_distance_flows[attack_index] if ratio <= THRESHOLD else 0,
-			"max_packet_distance", max_distance_packets[attack_index] if ratio <= THRESHOLD else 0
+			"flow_distance", distances_flows[attack_index] if final_ratios[attack_index] <= THRESHOLD else 0,
+			"packet_distance", distances_packets[attack_index] if final_ratios[attack_index] <= THRESHOLD else 0,
+			"max_flow_distance", max_distance_flows[attack_index] if final_ratios[attack_index] <= THRESHOLD else 0,
+			"max_packet_distance", max_distance_packets[attack_index] if final_ratios[attack_index] <= THRESHOLD else 0
 		)
 
 def eval_nn(data):

### 2019-12-15 ex3/paper/paper.tex
Capital L for distances (\ell only used for norms?)

@@
 
-We used $\ell_1$ norm as distance metric and \gls{pgd} for meeting the real-world constraints discussed above.
+We used $L_1$ norm as distance metric and \gls{pgd} for meeting the real-world constraints discussed above.
 
@@
 
-\subsection{$\ell_\infty$-bounded Projected Gradient Descent}
-\cite{madry_towards_2018} uses a method for generating adversarial samples which deploys \gls{pgd}  to minimize the network's negative loss function while constraining the achieved $\ell_\infty$ distance from the original samples.
+\subsection{$L_\infty$-bounded Projected Gradient Descent}
+\cite{madry_towards_2018} uses a method for generating adversarial samples which deploys \gls{pgd}  to minimize the network's negative loss function while constraining the achieved $L_\infty$ distance from the original samples.
 
@@
 Finally, we also tested the Fast Gradient Sign Method (FGSM), which was initially proposed in \cite{goodfellow_explaining_2015} and perhaps is the most frequently used method for generating adversarial samples.
-FGSM can be considered a single pass of \gls{pgd} on the loss function with an equality constraint on the $\ell_\infty$ distance, i.e.
+FGSM can be considered a single pass of \gls{pgd} on the loss function with an equality constraint on the $L_\infty$ distance, i.e.
 the adversarial sample is found as
@@
 \end{equation}
-where $L(X)$ denotes the network's loss function and $\epsilon$ denotes the achieved $\ell_\infty$ distance.
+where $L(X)$ denotes the network's loss function and $\epsilon$ denotes the achieved $L_\infty$ distance.
 
@@
 \autoref{fig:adv_per_family} depicts the performance for the investigated algorithms for both datasets and for each attack family in CIC-IDS-2017. 
-We used \gls{cw} with a tradeoff of 1 and, to provide a fair comparison, for \gls{fgsm} and \gls{pgd} we used an average $\ell_\infty$ distance as observed for \gls{cw}.
+We used \gls{cw} with a tradeoff of 1 and, to provide a fair comparison, for \gls{fgsm} and \gls{pgd} we used an average $L_\infty$ distance as observed for \gls{cw}.
 
@@
 
-It is interesting to note that increasing the $\ell_\infty$ bound for \gls{pgd} did not yield significantly improved performance.
+It is interesting to note that increasing the $L_\infty$ bound for \gls{pgd} did not yield significantly improved performance.
 We can thus confirm the recommendation by \cite{carlini_towards_2017} to perform gradient descent on the logit output instead of the loss for good results.
@@
 % \toprule
-% Attack type & \multicolumn{2}{l}{FGSM} & \multicolumn{2}{l}{Carlini-Wagner, $\ell_1$} & \multicolumn{2}{l}{Carlini-Wagner, $\ell_2$} & \multicolumn{2}{l}{$\ell_\infty$-bounded \gls{pgd}} \\
+% Attack type & \multicolumn{2}{l}{FGSM} & \multicolumn{2}{l}{Carlini-Wagner, $L_1$} & \multicolumn{2}{l}{Carlini-Wagner, $L_2$} & \multicolumn{2}{l}{$L_\infty$-bounded \gls{pgd}} \\
 % \cmidrule(r){2-3} \cmidrule(lr){4-5} \cmidrule(l){6-7} \cmidrule(l){8-9}
@@
 %\includegraphics[width=0.49\textwidth]{adv_plots/l_inf_pgd.pdf}
-%\caption{Performance of FSGM and $\ell_\infty$-bounded \gls{pgd} for different bounds.
+%\caption{Performance of FSGM and $L_\infty$-bounded \gls{pgd} for different bounds.
 %\todo{Max: Omit}}
@@
 %
-%To this end, we generated adversarial flows with the \gls{cw} with $\ell_2$ distance and a tradeoff of 0.5.
+%To this end, we generated adversarial flows with the \gls{cw} with $L_2$ distance and a tradeoff of 0.5.
 \autoref{adv_plots} depicts how the method manipulates flows for some exemplary attack categories.

### 2019-12-15 ex3/auxplot.py

@@ -147,7 +147,7 @@ def adv():
 	plt.gca().set_ylabel_legend(Line2D([0],[0], color='gray'), handlelength=1.4)
 	plt.twinx()
 	plt.plot(x_values, values[:,2:], linestyle='--')
-	plt.ylabel('$\ell_1$ distance')
+	plt.ylabel('$L_1$ distance')
 	plt.gca().set_ylabel_legend(Line2D([0],[0], color='gray', linestyle='--'), handlelength=1.4)
 	plt.legend(lines, ['CIC-IDS-2017', 'UNSW-NB15'])
 	ylim1,ylim2 = plt.ylim()

### 2019-12-15 ex3/paper/paper.tex
Text changes

@@
 
-\todo{Alex: I think this is quite technical curretnly and can be shortened a bit}
+\todo{Alex: I think this is quite technical curretnly and can be shortened a bit, if necessary}
 The result is a number of adversarial samples (more than half in the ideal case) and some that could not be made adversarial.
@@
 To distinguish, we propose the term \emph{feature sensitivity} for such metrics.
-
 To analyze features for our network, we use the following approaches.
@@
 
-We thus develop a more sound method compared to input perturbation called \textit{feature dropout}: When training a model, for each sample, we leave out each feature with a certain probability by setting it to zero.
+To evaluate true feature importance, we thus develop a more sound method %compared to input perturbation
+called \textit{feature dropout}: When training a model, for each sample, we leave out each feature  with independent probability $\frac{1}{n}$, $n \in \mathbb N$ being the number of features, by setting it to zero.
 On average, one feature gets zeroed out but it is also possible that none or more than one are left out.
@@
 An important implementation detail is that for each feature we add another input which is 1 if the feature is suppressed and 0 otherwise.
-This is necessary so that the neural network can distinguish between a feature missing and a feature genuinely being zero.
+This is necessary for the neural network to be able to distinguish between a feature missing and a feature genuinely being zero.
 The overall outcome is a classifier being able to deal with missing features.
-As it can also happen that several features are left out for a sample, our method can also be used to analyze the effect of multiple features missing and can thus show possibly correlated features.
 The results in \autoref{fig:feat_imp} show that, unlike input perturbation, feature dropout  does not vastly overestimate features' importance.
@@
 
-Feature dropout also allows us to find correlated features or -- more general -- features that contain common information:
-
+As it can also happen that several features are left out for a sample, our method can also be used to analyze the effect of multiple features missing and can thus show possibly correlated features or -- more general -- features that contain common information, relevant for the classification task:
+%Feature dropout also allows us to find correlated features or -- more general -- features that contain common information: 
 \begin{equation}
@@
 With $\text{acc}_\text{base}$ we denote the average accuracy of the classifier with all features included, with $\text{acc}_\text{-feature\_i}$ the average accuracy if feature $i$ is omitted and with $\text{acc}_\text{-both\_features}$ the average accuracy if both features are omitted.
-The higher the resulting number, the larger the information that both features share.
-If the score is one, this means that both features contain no joint information.
+Assuming a non-negative accuracy drop when omitting a feature, the resulting score is $\ge 0.5$.
+The higher it is, the larger the information that both features share.
+
+%AH. As it can become <1, I would say it this way
+%If the score is one,  both features contain no joint information.
 For example, the score between RST and the protocol identifier is 8.5; the highest of all pairs of features.
@@
 \end{figure}
-Input Perturbation has some shortcomings when considering feature sensitivity.
-Assuming that the test dataset is representative for production use, for feature importance it is reasonable to equate the distribution of the perturbed feature with the feature distribution.
-On the other hand, when evaluating feature sensitivity, e.g.
+Input perturbation and feature dropout mainly address feature importance.
+For example, assuming that the test dataset is representative for production use, for feature importance it is reasonable to equate the distribution for perturbing a feature with the feature distribution itself.
+However, when evaluating feature sensitivity, e.g.
 for analyzing potential for adversarial samples, the attacker is not limited by this distribution and frequently is able to choose arbitrary values in the feature space.
@@
 For investigating in more detail how features influence the prediction outcome, we extend \glspl{pdp} to the sequential setting. 
-Denoting as $X=\{\boldsymbol X^1, \ldots, \boldsymbol X^m\}$ the series of packet features $\boldsymbol X^t \in \mathbb R^n$ and $h_t(X)$ the network's hidden state at time step $t$, we define the sequential \gls{pdp} as
+Denoting as $X=\{\boldsymbol X^1, \ldots, \boldsymbol X^m\}$ the series of packet features $\boldsymbol X^t \in \mathbb R^n$ and $h_t(X)$ the network's hidden state after time step $t$, we define the sequential \gls{pdp} as
 \begin{align}
-\text{PD}&\text{P}_{c,i}(t,w)= \\ &\mathbb E_{X | C} \Big(f \left(h_{t-1}( X), X_1^t,\ldots,X_{i-1}^t,w,X_{i+1}^t,\ldots X_n^t \right) | c\Big). \nonumber 
+\text{se}&\text{qPDP}_{c,i}(t,w)= \\ &\mathbb E_{X | C} \Big(f \left(h_{t-1}( X), X_1^t,\ldots,X_{i-1}^t,w,X_{i+1}^t,\ldots X_n^t \right) | c\Big). \nonumber 
 \end{align}
@@
 We showed that even though our use case is very different from computer vision, adversarial samples can be found efficiently, even if only ostensibly unimportant features can be modified.
-We introduced the \gls{ars} as a new way to quantify the adversarial threat.
+We proposed the \gls{ars} for quantifying and comparing the adversarial threat for \glspl{ids}.
 Deploying an adversarial training procedure, we could significantly reduce the adversarial threat.

### 2019-12-16 ex3/paper/paper.tex
Add reference to ARS section

@@
 constructing \glspl{ids} for network data.
-However, with the rise of ubiquitous \gls{ml} systems, malicious actors have been catching up quickly to find new ways to exploit \gls{ml} for profit.
+With the rise of ubiquitous \gls{ml} systems, malicious actors have been catching up quickly to find new ways to exploit \gls{ml} for profit.
 Recently developed adversarial \gls{ml} techniques focus on computer vision and their applicability to network traffic is not straightforward: Network packets expose fewer features than an image, are sequential and impose several constraints on their features.
@@
 
-In practice, a further crucial requirement for \gls{ml}-based \glspl{ids} 
-is that the classifier's decisions are interpretable, recently giving rise to increasing interest in explainability methods.
+Due to the threat of adversarial attacks, a further crucial requirement for \gls{ml}-based \glspl{ids} is that the classifier's decisions are interpretable by humans.
+This recently gave increased interest in explainability methods.
 Also in this case, common methods are designed to work with images or tabular data but not with sequences.
@@
 % on Dec. 4th
-\caption{Detection performance on a packet and flow level.} \label{tab:performance_results}
+\caption{Performance metrics per packet and per flow.} \label{tab:performance_results}
 \centering
@@
 	&	Packet	&	Flow	&	Packet	&	Flow	\\	\midrule
-Accuracy	&	0.991	&	0.997	&	0.995	&	0.983	\\	
-Precision	&	0.970	&	0.997	&	0.834	&	0.786	\\	
-Recall	&	0.978	&	0.991	&	0.876	&	0.726	\\	
-F1	&	0.974	&	0.994	&	0.855	&	0.755	\\	
-Youden	&	0.972	&	0.990	&	0.873	&	0.719	\\
-
+Accuracy	&	99.1\%	&	99.7\%	&	99.5\%	&	98.3\%	\\	
+Precision	&	97.0\%	&	99.7\%	&	83.4\%	&	78.6\%	\\	
+Recall	&	97.8\%	&	99.1\%	&	87.6\%	&	72.6\%	\\	
+F1	&	97.4\%	&	99.4\%	&	85.5\%	&	75.5\%	\\	
+Youden	&	97.2\%	&	99.0\%	&	87.3\%	&	71.9\%	\\	
 \bottomrule
@@
 \label{sec:adv}
-We now investigate whether known \gls{aml} methods can be used for generating adversarial flows for our \gls{rnn}..
-% In our scenario, most features such as, e.g., \gls{tcp} flags cannot be easily manipulated because their manipulation might violate the \gls{tcp} and thus cause communication to fail.
-% Hence, we experiment whether by manipulating the \glspl{iat} and packet lengths of the first packets, the classifier might be fooled.
-% Our scenario is quite different from how generating adversarial samples usually works: We have significantly less features than in, for example, an image.
-Furthermore, only two out of those few features can be manipulated.
-And even these features cannot be manipulated freely: Besides the constraints that packet size and \gls{iat} cannot become negative there are also more problem-specific constraints:
+We now investigate whether known \gls{aml} methods can be used for generating adversarial flows for our \gls{rnn}.
+\cite{hashemi_towards_2019} previously studied the behavior of several Network \glspl{ids} under adversarial attacks but unlike us did not investigate \glspl{rnn}.
 
@@
 An in-depth analysis of when the reduction of \glspl{iat} is possible, would come with substantial effort, so we generally disallowed it.
-%a reduction of this feature.
 \end{itemize}
@@
 We used $L_1$ norm as distance metric and \gls{pgd} for meeting the real-world constraints discussed above.
-
-% The method minimizes the adversarial sample's distance so that the network's decision for adversarial samples is just between attack and normal traffic.
-In the present context, we need to make sure that the classifier's prediction would actually be normal traffic, even though a certain level of noise will be added to \glspl{iat} due to the network between attacker and victim.
-Instead of the above equation \ref{eq:carliniWagner} we therefore used the optimization objective
-% \begin{equation}
-% d(X,\tilde X) + \epsilon  \max(Z(\tilde X), -0.2),
-% \end{equation}
-% i.e.
-we require a slight level of confidence that our adversarial samples are not classified as attack traffic.
-A logit value of $-0.2$ corresponds to a confidence of 55\% for the sample to be benign.
 
@@
 
-% \begin{table*}
-% \caption{Attack success ratios per attack family. Depicted are flow detection accuracies for adversarial flows (left numbers) and flow detection accuracies for unmodified flows (right numbers).}
-% \label{tab:adv_per_family}
-% \centering
-% \begin{tabular}{llllllllll}
-% \toprule
-% Attack type & \multicolumn{2}{l}{FGSM} & \multicolumn{2}{l}{Carlini-Wagner, $L_1$} & \multicolumn{2}{l}{Carlini-Wagner, $L_2$} & \multicolumn{2}{l}{$L_\infty$-bounded \gls{pgd}} \\
-% \cmidrule(r){2-3} \cmidrule(lr){4-5} \cmidrule(l){6-7} \cmidrule(l){8-9}
-% & adv. & orig. & adv. & orig. & adv. & orig. & adv. & orig. \\
-% \midrule
-% Botnet:ARES	&	0.0000 & 0.0000	&	0.0000 & 0.0000	&	0.0000 & 0.0000	&	0.0000 & 0.0000	\\
-% Brute Force:FTP-Patator	&	0.2800 & 0.5800	&	0.1600 & 0.5800	&	0.1600 & 0.5800	&	0.2800 & 0.5800	\\
-% Brute Force:SSH-Patator	&	0.8300 & 1.0000	&	0.0000 & 1.0000	&	0.0000 & 1.0000	&	0.8900 & 1.0000	\\
-% DDoS:LOIT	&	0.8300 & 0.8600	&	0.6400 & 0.8600	&	0.5600 & 0.8600	&	0.8200 & 0.8600	\\
-% DoS / DDoS:DoS GoldenEye	&	0.8900 & 1.0000	&	0.5000 & 1.0000	&	0.4200 & 1.0000	&	0.9200 & 1.0000	\\
-% DoS / DDoS:DoS Hulk	&	0.6700 & 1.0000	&	0.5400 & 1.0000	&	0.4700 & 1.0000	&	1.0000 & 1.0000	\\
-% DoS / DDoS:DoS Slowhttptest	&	0.8700 & 1.0000	&	0.4100 & 1.0000	&	0.1900 & 1.0000	&	0.7400 & 1.0000	\\
-% DoS / DDoS:DoS slowloris	&	0.4400 & 0.9700	&	0.3100 & 0.9700	&	0.1300 & 0.9700	&	0.9200 & 0.9700	\\
-% DoS / DDoS:Heartbleed	&	0.0000 & 0.0000	&	0.0000 & 0.0000	&	0.0000 & 0.0000	&	0.0000 & 0.0000	\\
-% Infiltration:Dropbox download 	&	0.6100 & 0.9600	&	0.4300 & 0.9600	&	0.3800 & 0.9600	&	0.7400 & 0.9600	\\
-% PortScan:PortScan - Firewall off	&	0.6800 & 0.9100	&	0.3500 & 0.9100	&	0.3500 & 0.9100	&	0.7200 & 0.9100	\\
-% PortScan:PortScan - Firewall on	&	0.7400 & 0.8700	&	0.6100 & 0.8700	&	0.5300 & 0.8700	&	0.7300 & 0.8700	\\
-% Web Attack:Sql Injection	&	0.0000 & 0.6923	&	0.0000 & 0.6923	&	0.0000 & 0.6923	&	0.0000 & 0.6923	\\
-% Web Attack:XSS	&	0.6700 & 0.9500	&	0.7200 & 0.9500	&	0.6200 & 0.9500	&	0.6700 & 0.9500	\\
-% \bottomrule
-
-% \end{tabular}
-
-% \todo{In the adv table, round the numbers!!! \\
-% For Max: Introduce two-level table\\
-% What's the tradeoff for CW?}
-% \end{table*}
-
-% \begin{table}
-% \centering
-% \caption{Total flow accuracy for adversarial flows.}
-% \label{tab:adv_accuracies}
-% \begin{tabular}{lrrr} \toprule
-% & CW & FGSM & PGD \\ \midrule
-% CIC-IDS-2017 &	0.346	 & 0.445	& 0.539 \\ 
-% UNSW-NB15 &	0.326&	0.629	& 0.774 \\
-% \bottomrule
-% \end{tabular}
-% \end{table}
-%It is interesting to note that for very large values of $\epsilon$ the success ratio starts to decrease again.This phenomenon might be due to numerical effects and might be preventable by a smaller learning rate.
-
-%\begin{figure*}
-%\includegraphics[width=0.49\textwidth]{adv_plots/fgsm.pdf}
-%\includegraphics[width=0.49\textwidth]{adv_plots/l_inf_pgd.pdf}
-%\caption{Performance of FSGM and $L_\infty$-bounded \gls{pgd} for different bounds.
-%\todo{Max: Omit}}
-%\label{fig:fgsm}
-%\end{figure*}
-
-%Finally, we investigated what adversarial flows look like and which features are manipulated to circumvent detection by the neural network.
-%
-%To this end, we generated adversarial flows with the \gls{cw} with $L_2$ distance and a tradeoff of 0.5.
-\autoref{adv_plots} depicts how the method manipulates flows for some exemplary attack categories.
-Depicted is the mean of both unmodified flows and adversarial flows  of samples of individual attack families considering both \gls{iat} and packet length.
-%
-%The plots show that primarily features at the beginning of flows are modified.
-It is remarkable that it seems most promising to modify single packets while leaving most of the flow unchanged.
-In other cases, however, like depicted for the infiltration attack at the bottom, adversarial flows assume a complex pattern to circumvent detection.
-
-%\begin{figure*}[p]
-%\includegraphics[width=0.48\textwidth]{../plots/plot_adv/1.pdf}
-%\includegraphics[width=0.48\textwidth]{../plots/plot_adv/2.pdf}
-%\includegraphics[width=0.48\textwidth]{../plots/plot_adv/3.pdf}
-%\includegraphics[width=0.48\textwidth]{../plots/plot_adv/4.pdf}
-%\includegraphics[width=0.48\textwidth]{../plots/plot_adv/5.pdf}
-%\includegraphics[width=0.48\textwidth]{../plots/plot_adv/6.pdf}
-%\caption{Successful manipulations performed by the \gls{cw}. \todo{Max: Just keep one plot I'd say. Maximum two.}}
-%\label{adv_plots}
-%\end{figure*}
-
 \section{Adversarial Robustness Metric}
@@
 We thus propose the \gls{ars}, as the average of the lower half  of minimum distances of adversarial samples.
-Since \gls{cw} is able to find adversarial samples with minimum distance, the \gls{ars} can
-  be found efficiently as follows: Use \gls{cw} with a small tradeoff to try to generate adversarial samples for an attack type.
+Since \gls{cw} is able to find adversarial samples with minimum distance, the \gls{ars} can be found efficiently as follows: Use \gls{cw} with a small tradeoff to try to generate adversarial samples for an attack type.
 If more than half of the samples are now wrongly classified, stop.
@@
 
-\todo{Alex: I think this is quite technical curretnly and can be shortened a bit, if necessary}
-The result is a number of adversarial samples (more than half in the ideal case) and some that could not be made adversarial.
-For example, if in total, there are 100 samples, there might be 54 which could be made adversarial.
-Out of these 54, the 50 with the smallest distance to the original samples are taken.
-Of these samples, the distance of each adversarial sample to the original sample is taken and averaged over all 50 samples.
+The result is a number of adversarial samples and some that could not be made adversarial.
+If at least half of the samples are misclassified, we take 50\% of the samples, with those that required smaller modifications first.
+Of these samples, the distance of each adversarial sample to the original sample is taken and averaged.
 The resulting average distance is the \gls{ars}.
 The larger it is the better since then an adversary needs to modify the samples more to make them adversarial.
-If not more then half of the samples could be made adversarial, our metric is undefined since then apparently it is not possible for an adversary conduct adversarial attacks on the majority of samples.
-
+If not more then half of the samples could be made adversarial, our metric is undefined since then apparently it is not possible for an adversary to reliably conduct adversarial attacks on the majority of samples.
 Setting the threshold for the \gls{ars} to 50\% is arbitrary, but

### 2019-12-16 ex3/paper/paper.tex
Other small modification to ARS section
Normal->benign and minor text changes

@@
 $\delta \in \mathbb R$ determines the logit output of successful adversarial.
-In the original publication $\delta=0$, which means that the network's decision for adversarial samples is just between attack and normal traffic.
-In the present context, we need to make sure that the classifier's prediction would actually be normal traffic, even though a certain level of noise will be added to \glspl{iat} due to the network between attacker and victim.
+In the original publication $\delta=0$, which means that the network's decision for adversarial samples is just between attack and benign traffic.
+In the present context, we need to make sure that the classifier's prediction would actually be benign, even though a certain level of noise will be added to \glspl{iat} due to the network between attacker and victim.
 Hence, we introduced $\delta=-0.2$, corresponding to a confidence of 55\% for the sample to be benign.
@@
 However, the results match to a high degree with our expectations.
-For example, SQL injection attacks apparently are closer to normal traffic than \gls{dos} attacks and, hence, finding adversarial samples should be easier.
+For example, SQL injection attacks apparently are closer to benign traffic than \gls{dos} attacks and, hence, finding adversarial samples should be easier.
 
@@
 If not more then half of the samples could be made adversarial, our metric is undefined since then apparently it is not possible for an adversary to reliably conduct adversarial attacks on the majority of samples.
+Then the ratio of samples that could be made adversarial (<50\%) can be useful metric to determine the exact extent of the adversarial threat.
 Setting the threshold for the \gls{ars} to 50\% is arbitrary, but
@@
 As alternative to omitting manipulable features, we attempted to make the classifier more robust against adversarial samples by augmenting the training set by adversarial flows generated using  \gls{cw}, labeled as additional attacks.
-
 For this purpose, we added one adversarial sample per attack sample to the training set.
 Since \gls{cw} is deterministic, this is the maximum number of adversarial samples possible.
-With this training set we then alternated retraining of the neural network and optimization of the adversarial samples using gradient descent.
+With this augmented training set we then alternated retraining of the neural network and optimization of the adversarial samples.
+% using gradient descent.
 
@@
 
-\autoref{fig:avdistance} shows the \gls{ars} throughout the adversarial training procedure for several attack categories and clearly indicates that adversarial training is effective, as the distance gradually increases.
+\autoref{fig:avdistance} shows the \gls{ars} throughout  adversarial training  for several attack categories and clearly indicates that adversarial training is effective, as the distance gradually increases.
 Hence, an attacker would have to modify attack samples more and more, eventually rendering the attack unpractical.
@@
 %\autoref{tab:adv_training_performance} depicts classification on non-adversarial data after 50 epochs of adversarial training.
-After 50 epochs of training, accuracy is essentially identical to results presented in \autoref{tab:performance_results}.
+After 50 epochs of training, accuracy is essentially identical to the results presented in \autoref{tab:performance_results}.
 Recall and Youden increased and precision slightly decreased.
-However, this is due to the higher proportion of attack samples in training data.
+However, this is due to the higher proportion of attack samples in training data and, hence, expected.
 We conclude that adversarial training is effective for reducing the attack surface for adversarial attacks in our scenario.

### 2019-12-16 ex3/paper/paper.tex
Text changes

@@
 Among the used features, source port, destination port and protocol identifier are constant over the whole flow while the others vary.
+During flow extraction we used a usual 5-tuple flow key, which distinguishes flows based on the protocol they use and their source and destination port and IP address.
 We use Z-score normalization and a train/test split of 2:1.
@@
 \subsection{$L_\infty$-bounded Projected Gradient Descent}
-\cite{madry_towards_2018} uses a method for generating adversarial samples which deploys \gls{pgd}  to minimize the network's negative loss function while constraining the achieved $L_\infty$ distance from the original samples.
+\cite{madry_towards_2018} uses a method for generating adversarial samples which deploys \gls{pgd}  to maximizie the network's loss function while constraining the achieved $L_\infty$ distance from the original samples.
+We consider this method an interesting combination of \gls{cw} and \gls{fgsm}.
+Since its functioning is somewhat similar to \gls{cw}, we expected the generated adversarial samples to be of similar quality.
 
 \subsection{Fast Gradient Sign Method}
-Finally, we also tested the Fast Gradient Sign Method (FGSM), which was initially proposed in \cite{goodfellow_explaining_2015} and perhaps is the most frequently used method for generating adversarial samples.
+Finally, we also tested the \gls{fgsm} % which was initially proposed in
+ \cite{goodfellow_explaining_2015}, which perhaps is the most frequently used method for generating adversarial samples.
 FGSM can be considered a single pass of \gls{pgd} on the loss function with an equality constraint on the $L_\infty$ distance, i.e.
@@
 yielding the same accuracy, but clearly the right-side distribution has a larger influence on the prediction, as in the right-side case the prediction deterministically depends on the feature value.
-To capture such dependencies, we propose to use mutual information instead of accuracy drop to determine feature sensitivity.
+To capture such dependencies, we propose to use mutual information %instead of accuracy drop 
+to determine feature sensitivity.
 Mutual Information between two random variables $X,Y$ is defined as 
@@
 with $f_X(x), f_Y(y)$ and $f_{X,Y}(x,y)$ denoting the distribution of $X,Y$ and their joint distribution, respectively.
-To obtain feature sensitivity, we compute mutual information between an input variable and the prediction output for one flow at one particular time step, and average over the result for the test dataset.
+To obtain feature sensitivity, we compute mutual information between an input variable and the prediction output for one flow at one particular time step, averaging over the result for the test set.
 
-\subsubsection{Results}
+\subsubsection{Comparison}
 \begin{figure*}
@@
 \includegraphics[width=.49\textwidth]{../plots/importance/feat_imp_flow_2015.pdf}
-\caption{Feature importance metrics considering the final flow prediction for CIC-IDS-2017 (left side) and UNSW-NB15 (right side).}
+\caption{Feature importance metrics for the flow prediction for CIC-IDS-2017 (left side) and UNSW-NB15 (right side).}
 \label{fig:feat_imp}
@@
 Knowing which features to investigate, it is important to analyze which feature values lead to classification as attack.
-During flow extraction we used a usual 5-tuple flow key, which distinguishes flows based on the protocol they use and their source and destination port and IP address.
-Hence, these features are constant for all packets in one flow, which allows us to use explainability methods which have been proposed for non-recurrent classification methods.
-
-In particular,  in literature the use of Partial Dependency Plots (\gls{pdp}) is proposed~\cite{friedman_greedy_2001}.
+In literature, the use of Partial Dependency Plots (\gls{pdp}) has been proposed~\cite{friedman_greedy_2001}.
 To inspect attack types in detail, in this research we use a \textit{conditional} form of the \gls{pdp}.
@@
 
+Due to the use of a 5-tuple flow key, port numbers and the protocol identifier are constant for all packets in one flow, which allows us to use \glspl{pdp}, which have been proposed for non-recurrent classification methods.
 The results (omitted for brevity) show that for some attack classes the port numbers play an important role. When looking at the class of benign traffic %(non-attack traffic)
-it becomes apparent that traffic destined to a high source port is generally indicative of an attack. We argue that this is because most services that regular users use have low port numbers.
+it becomes apparent that traffic destined to a high destination port is generally indicative of an attack. We argue that this is because most services that regular users use have low port numbers.
 

### 2019-12-16 ex3/paper/paper.tex
Small text changes

--- 2019-12-16 ex3/paper/paper.tex
@@
 
-As finding adversarial flows is particularly difficult in our case, % for the reasons we discussed above,
-we preferred methods  producing high-quality adversarial samples while not being as fast as other methods. We implemented the following \gls{aml} methods.
+% Max: But FGSM is high-speed... So I removed this paragraph altogether
+%As finding adversarial flows is particularly difficult in our case, % for the reasons we discussed above,
+%we preferred methods  producing high-quality adversarial samples while not being as fast as other methods. We implemented the following \gls{aml} methods.
 
@@
 
-$\delta \in \mathbb R$ determines the logit output of successful adversarial.
+$\delta \in \mathbb R$ is a parameter that determines how far an adversary wants to go over the decision boundary.
 In the original publication $\delta=0$, which means that the network's decision for adversarial samples is just between attack and normal traffic.
@@
 If not more then half of the samples could be made adversarial, our metric is undefined since then apparently it is not possible for an adversary to reliably conduct adversarial attacks on the majority of samples.
+Then the ratio of samples that could be made adversarial ($<$50\%) can be useful metric to determine the exact extent of the adversarial threat.
 Setting the threshold for the \gls{ars} to 50\% is arbitrary, but
@@
 \caption{Recall for each attack type and average modification per flow required by an adversary to make a flow appear benign for CIC-IDS-2017.}
+\todo{Show for which attack class not enough adversarial samples could be found.}
 \label{fig:recall_ars}

### 2019-12-16 ex3/auxplot.py

@@ -96,8 +96,12 @@ def ars_original():
 
 	index_by_which_to_sort = 1
 	old_len = len(group_names)
+	orig_group_names, orig_values = group_names, values
 	group_names, values = list(zip(*[(group_name, value) for group_name, value in zip(group_names, values) if value[0] <= 0.5]))
 	# print(f"Dropped {old_len-len(group_names)} elements.")
+	robust_groups = set(orig_group_names)-set(group_names)
+	robust_indices = [orig_group_names.index(item) for item in robust_groups]
+	print("robust attack types:", "; ".join([f"{g}: ratio={v[0]}" for g, v in zip([orig_group_names[index] for index in robust_indices], [orig_values[index] for index in robust_indices])]))
 	order = list(zip(*sorted(enumerate(values), key=lambda item: item[1][index_by_which_to_sort], reverse=True)))[0]
 	# print("order", order)
 	width = 0.75 / len(value_indices_to_plot)

### 2019-12-16 ex3/paper/paper.tex

--- 2019-12-16 ex3/paper/paper.tex
@@
 
-$\delta \in \mathbb R$ determines the logit output of successful adversarial.
-In the original publication $\delta=0$, which means that the network's decision for adversarial samples is just between attack and benign traffic.
-In the present context, we need to make sure that the classifier's prediction would actually be benign, even though a certain level of noise will be added to \glspl{iat} due to the network between attacker and victim.
+$\delta \in \mathbb R$ is a parameter that determines how far an adversary wants to go over the decision boundary.
+In the original publication $\delta=0$, which means that the network's decision for adversarial samples is just between attack and normal traffic.
+In the present context, we need to make sure that the classifier's prediction would actually be normal traffic, even though a certain level of noise will be added to \glspl{iat} due to the network between attacker and victim.
 Hence, we introduced $\delta=-0.2$, corresponding to a confidence of 55\% for the sample to be benign.
@@
 If not more then half of the samples could be made adversarial, our metric is undefined since then apparently it is not possible for an adversary to reliably conduct adversarial attacks on the majority of samples.
-Then the ratio of samples that could be made adversarial (<50\%) can be useful metric to determine the exact extent of the adversarial threat.
+Then the ratio of samples that could be made adversarial ($<$50\%) can be useful metric to determine the exact extent of the adversarial threat.
 Setting the threshold for the \gls{ars} to 50\% is arbitrary, but

### 2019-12-16 ex3/paper/paper.tex
Added lr/iteration scaling to paper

@@
 Thus we argue that beside the classical metrics such as accuracy, false positives etc., a new metric is needed, which quantifies how easily an \gls{ids} classifier can be fooled.
-
 Such a metric should be easy to compute, easy to interpret and robust with respect to outliers.
 We thus propose the \gls{ars}, as the average of the lower half  of minimum distances of adversarial samples.
-Since \gls{cw} is able to find adversarial samples with minimum distance, the \gls{ars} can be found efficiently as follows: Use \gls{cw} with a small tradeoff to try to generate adversarial samples for an attack type.
+Since \gls{cw} is able to find adversarial samples with minimum distance, it can be used for finding the \gls{ars}  efficiently.
+
+When using a large $\epsilon$ in \autoref{eq:carliniWagner}, it is necessary to reduce the gradient descent learning rate  to avoid numerical issues.
+However, since the distance term is not affected by $\epsilon$, it is necessary to scale the iteration count inversely with the learning rate.
+\gls{cw} with a large $\epsilon$ therefore needs more time than with a low tradeoff.
+For neural networks, the \gls{ars} can be efficiently found as follows: Use \gls{cw} with a small tradeoff, trying to generate adversarial samples for an attack type.
 If more than half of the samples are now wrongly classified, stop.

### 2019-12-16 ex3/paper/paper.tex
Different symbols for fgsm bound and cw tradeoff

@@
 \begin{equation} \label{eq:carliniWagner}
-d(X,\tilde X) + \epsilon  \max(Z(\tilde X), \delta).
+d(X,\tilde X) + c  \max(Z(\tilde X), \delta).
 \end{equation}
-Here, $d(\cdot)$ is a distance metric and $\epsilon \in \mathbb R^+$ is a parameter governing the tradeoff achieved between attack success and distance from the original flow.
+Here, $d(\cdot)$ is a distance metric and $c \in \mathbb R^+$ is a parameter governing the tradeoff achieved between attack success and distance from the original flow.
 Furthermore, $Z(\cdot)$ denotes the neural network's logit output, $X$ denotes the original flow and $\tilde X$ the adversarial flow under optimization.
@@
 
-\autoref{fig:adv_cw} depicts the success ratio and average distance for \gls{cw} for different tradeoffs $\epsilon$.
+\autoref{fig:adv_cw} depicts the success ratio and average distance for \gls{cw} for different tradeoffs $c$.
 %\autoref{fig:adv_cw} depicts the ratio of attack flows which were successfully misclassified as benign with respect to the final packet along with the resulting distance from the original sample.
-Evidently, the larger the tradeoff $\epsilon$, the better the attack works, but also distances from original flows become higher.
+Evidently, the larger the tradeoff $c$, the better the attack works, but also distances from original flows become higher.
 Hence, with acceptable distance values we were able to generate adversarial samples for about 80\% of attack samples.
@@
 %, when performing optimization as depicted by the objective function .
-Hence, the depicted distance increase presumably is due to samples for which no adversarial sample could be found for lower $\epsilon$ values.
+Hence, the depicted distance increase presumably is due to samples for which no adversarial sample could be found for lower $c$ values.
 
@@
 % \includegraphics[width=0.49\textwidth]{adv_plots/cwl2.pdf}
-\caption{Performance of \gls{cw} with different $\epsilon$ values.}
+\caption{Performance of \gls{cw} with different $c$ values.}
 \label{fig:adv_cw}
@@
 
-When using a large $\epsilon$ in \autoref{eq:carliniWagner}, it is necessary to reduce the gradient descent learning rate  to avoid numerical issues.
-However, since the distance term is not affected by $\epsilon$, it is necessary to scale the iteration count inversely with the learning rate.
-\gls{cw} with a large $\epsilon$ therefore needs more time than with a low tradeoff.
+When using a large $c$ in \autoref{eq:carliniWagner}, it is necessary to reduce the gradient descent learning rate  to avoid numerical issues.
+However, since the distance term is not affected by $c$, it is necessary to scale the iteration count inversely with the learning rate.
+\gls{cw} with a large $c$ therefore needs more time than with a low tradeoff.
 For neural networks, the \gls{ars} can be efficiently found as follows: Use \gls{cw} with a small tradeoff, trying to generate adversarial samples for an attack type.

### 2019-12-16 ex3/auxplot.py

@@ -146,7 +146,7 @@ def adv():
 	plt.figure(figsize=(5,2))
 	x_values = [ float(value) for value in group_names ]
 	lines = plt.plot(x_values, values[:,:2])
-	plt.xlabel('Tradeoff $\epsilon$')
+	plt.xlabel('Tradeoff $c$')
 	plt.ylabel('Success ratio')
 	plt.gca().set_ylabel_legend(Line2D([0],[0], color='gray'), handlelength=1.4)
 	plt.twinx()

### 2019-12-16 ex3/paper/paper.tex
Changed feature dropout section

@@
 
-We used $L_1$ norm as distance metric and \gls{pgd} for meeting the real-world constraints discussed above.
+We used $L_1$ distance as distance metric, as we consider it to represent practically relevant differences of network flows best.
+We used \gls{pgd} for meeting the real-world constraints discussed above.
 
@@
 Thus important features are going to be left out less often and one can then extract the dropout probabilities for each feature and assess their significance based on them.
-Feature dropout can also be used as a building block for computing the accuracy of datapoints with missing features that is necessary to compute Shapley values \cite{shapley_value_1953} or KernelSHAP \cite{lundberg_unified_2017}.
+%Feature dropout can also be used as a building block for computing the accuracy of datapoints with missing features that is necessary to compute Shapley values \cite{shapley_value_1953} or KernelSHAP \cite{lundberg_unified_2017}.
 
@@
 \end{equation}
-With $\text{acc}_\text{base}$ we denote the average accuracy of the classifier with all features included, with $\text{acc}_\text{-feature\_i}$ the average accuracy if feature $i$ is omitted and with $\text{acc}_\text{-both\_features}$ the average accuracy if both features are omitted.
+With $\text{acc}_\text{base}$ we denote the accuracy of the classifier with all features included, with $\text{acc}_\text{-feature\_i}$ the accuracy if feature $i$ is omitted and with $\text{acc}_\text{-both\_features}$ the accuracy if both features are omitted.
 Assuming a non-negative accuracy drop when omitting a feature, the resulting score is $\ge 0.5$.
@@
 For example, the score between RST and the protocol identifier is 8.5; the highest of all pairs of features.
-While this may seem unintuitive at the first glance, it likely stems from the fact that if the protocol identifier (\gls{tcp} or UDP) is missing, then if RST is 1 at some point of the flow, the flow has to be \gls{tcp}.
+While this may be unintuitive at the first glance, it likely stems from the fact that if the protocol identifier (\gls{tcp} or UDP) is missing, then
+%if RST is 1 at some point of a flow, it has to be \gls{tcp}.
+RST being 1 at some point indicates that the flow is \gls{tcp}.
 
+Feature dropout might constitute a building block for methods based on Shapley values \cite{shapley_value_1953} like KernelSHAP \cite{lundberg_unified_2017}.
+Due to space limitations, we leave this for future work.
 \subsubsection{Mutual Information}

### 2019-12-16 ex3/paper/paper.tex
Update ars_original figure

--- 2019-12-16 ex3/paper/paper.tex
@@
 \includegraphics[width=\columnwidth]{../plots/ars_original.pdf}
-\caption{Recall for each attack type and average modification per flow required by an adversary to make a flow appear benign for CIC-IDS-2017.}
-\todo{Show for which attack class not enough adversarial samples could be found.}
+\caption{Recall for each attack type and average modification per flow required by an adversary to make a flow appear benign for CIC-IDS-2017. For ``PortScan, Firewall
+'' recall never falls under 66\%, meaning that an adversary does not succeed and that \gls{ars} is $\inf$.}
 \label{fig:recall_ars}

### 2019-12-16 ex3/auxplot.py

@@ -5,6 +5,7 @@
 import fileinput
 import numpy as np
 import matplotlib.pyplot as plt
+from matplotlib.patches import Rectangle
 import sys
 from matplotlib.transforms import Affine2D
 from matplotlib.lines import Line2D
@@ -88,10 +89,9 @@ def ars_original():
 	global group_names
 	global values
 
-	y_labels = [None, "Recall", "Recall", "Distance", "Distance"]
+	y_labels = [None, "Recall", "Recall", "ARS", "ARS"]
 	value_indices_to_plot = (1,3)
 
-	# plt.figure(figsize=(5,3))
 	values = [[float(item) for item in sublist] for sublist in values]
 
 	index_by_which_to_sort = 1
@@ -101,7 +101,7 @@ def ars_original():
 	# print(f"Dropped {old_len-len(group_names)} elements.")
 	robust_groups = set(orig_group_names)-set(group_names)
 	robust_indices = [orig_group_names.index(item) for item in robust_groups]
-	print("robust attack types:", "; ".join([f"{g}: ratio={v[0]}" for g, v in zip([orig_group_names[index] for index in robust_indices], [orig_values[index] for index in robust_indices])]))
+	print("robust attack types:", "; ".join([f"{g}: orig_flow_recall={v[1]}, adv_flow_recall={v[0]}" for g, v in zip([orig_group_names[index] for index in robust_indices], [orig_values[index] for index in robust_indices])]))
 	order = list(zip(*sorted(enumerate(values), key=lambda item: item[1][index_by_which_to_sort], reverse=True)))[0]
 	# print("order", order)
 	width = 0.75 / len(value_indices_to_plot)
@@ -126,8 +126,13 @@ def ars_original():
 
 	all_legends = [item.get_label() for item in all_labels]
 
-	# plt.legend(all_legends, all_labels, loc="upper right")
-	plt.legend(all_labels, all_legends, loc='lower center', bbox_to_anchor=(0.5,1), ncol=4)
+	# ax1.legend(all_legends, all_labels, loc='upper right', bbox_to_anchor=(1,0.95))
+	#  plt.ylim((min(first_quartiles), max(third_quartiles)))
+
+	ax2.set_ylabel_legend(Rectangle((0,0), 1, 1, fc=colors[1]), handlelength=1)
[…]

### 2019-12-16 ex3/paper/paper.tex
Change ars_original legend

--- 2019-12-16 ex3/paper/paper.tex
@@
 \includegraphics[width=\columnwidth]{../plots/ars_original.pdf}
-\caption{Recall for each attack type and average modification per flow required by an adversary to make a flow appear benign for CIC-IDS-2017.}
-\todo{Show for which attack class not enough adversarial samples could be found.}
+\caption{Recall for each attack type and \gls{ars} for CIC-IDS-2017. For ``PortScan, Firewall'' recall never falls under 66\%: The adversary does not succeed and that \gls{ars} is $\infty$.}
 \label{fig:recall_ars}

### 2019-12-16 ex3/paper/paper.tex
Moved figure

@@
 \includegraphics[width=\columnwidth]{../plots/ars_original.pdf}
-\caption{Recall for each attack type and \gls{ars} for CIC-IDS-2017. For ``PortScan, Firewall'' recall never falls under 66\%: The adversary does not succeed and that \gls{ars} is $\infty$.}
+\caption{Recall for each attack type and average modification per flow required by an adversary to make a flow appear benign for CIC-IDS-2017.}
+\todo{Show for which attack class not enough adversarial samples could be found.}
 \label{fig:recall_ars}
@@
 
-\begin{figure}
-\includegraphics[width=\columnwidth]{../plots/ars/ars_15.pdf}
-\caption{\gls{ars} during adversarial training for UNSW-NB15.}
-\label{fig:avdistance}
-\end{figure}
-
 For both datasets, after a small number of epochs, it was no longer possible to create a significant number of adversarial samples for most attack categories.
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.98\columnwidth]{../plots/ars/ars_15.pdf}
+\caption{\gls{ars} during adversarial training for UNSW-NB15.}
+\label{fig:avdistance}
+\end{figure}
+
 \renewcommand*{\bibfont}{\small}

### 2019-12-16 ex3/paper/paper.tex
Fit to 8 pages

--- 2019-12-16 ex3/paper/paper.tex
@@
 \subsection{Reducing features}
-The most obvious defence, which is only possible because of the nature of the dataset, is to simply leave out features which the attacker can manipulate. For this, we try two different approaches:
+The most obvious defence, which is only possible because of the nature of the dataset, is to simply omitting features an attacker can manipulate. We try two different approaches:
 \begin{itemize}[topsep=0pt,wide,labelwidth=!,labelindent=0pt]
@@
 It is, hence, questionable if the gained increase in robustness is enough for reliable attack detection.
-
-\begin{figure}
-\includegraphics[width=\columnwidth]{../plots/ars/ars_15.pdf}
-\caption{\gls{ars} during adversarial training for UNSW-NB15.}
-\label{fig:avdistance}
-\end{figure}
 

### 2019-12-16 ex3/plot.py

@@ -99,7 +99,7 @@ for attack_type, seqs in enumerate(results_by_attack_number):
 	# print("values_by_length", [item.shape for item in values_by_length])
 	lens = [item.shape[0] for item in values_by_length]
 
-	fig, ax1 = plt.subplots(figsize=(5,2.5))
+	fig, ax1 = plt.subplots(figsize=(5,2.4))
 	# print("lens", lens)
 	x_values = list(range(min(len(lens), MAX_X)))
 	ret = ax1.bar(x_values, lens[:MAX_X], width=1, color="gray", alpha=0.2, label="number of samples")
@@ -109,7 +109,7 @@ for attack_type, seqs in enumerate(results_by_attack_number):
 	ax2 = ax1.twinx()
 
 	ax2.set_ylabel('Confidence')
-	ax1.set_ylabel("Occurrence frequency")
+	ax1.set_ylabel("Number of samples")
 
 	ax1.yaxis.tick_right()
 	ax1.yaxis.set_label_position("right")

### 2019-12-16 ex3/paper/paper.tex
Typos

@@
 \subsection{$L_\infty$-bounded Projected Gradient Descent}
-\cite{madry_towards_2018} uses a method for generating adversarial samples which deploys \gls{pgd}  to maximizie the network's loss function while constraining the achieved $L_\infty$ distance from the original samples.
+\cite{madry_towards_2018} uses a method for generating adversarial samples which deploys \gls{pgd}  to maximize the network's loss function while constraining the achieved $L_\infty$ distance from the original samples.
 We consider this method an interesting combination of \gls{cw} and \gls{fgsm}.
@@
 \includegraphics[width=\columnwidth]{../plots/adv_comparison/adv_comparison_17.pdf}
-\caption{Attack success ratios for both dataset and per attack family for CIC-IDS-2017. Depicted are flow detection accuracies for adversarial flows and for unmodified flows.}
+\caption{Attack success ratios for both datasets and per attack family for CIC-IDS-2017. Depicted are flow detection accuracies for adversarial flows and for unmodified attack flows.}
 \label{fig:adv_per_family}
@@
 \begin{equation}
-\text{score} = \frac{\text{acc}_\text{base}-\text{acc}_\text{-both\_features}}{\left(\text{acc}_\text{base}-\text{acc}_\text{-feature\_1}\right) + \left(\text{acc}_\text{base}-\text{acc}_\text{-feature\_2}\right)}
+\text{score} = \frac{\text{acc}_{\text{base}}-\text{acc}_{\text{-both\_features}}}{\left(\text{acc}_{\text{base}}-\text{acc}_{\text{-feature\_1}}\right) + \left(\text{acc}_{\text{base}}-\text{acc}_{\text{-feature\_2}}\right)}
 \end{equation}
-With $\text{acc}_\text{base}$ we denote the accuracy of the classifier with all features included, with $\text{acc}_\text{-feature\_i}$ the accuracy if feature $i$ is omitted and with $\text{acc}_\text{-both\_features}$ the accuracy if both features are omitted.
+With $\text{acc}_{\text{base}}$ we denote the accuracy of the classifier with all features included, with $\text{acc}_{\text{-feature\_i}}$ the accuracy if feature $i$ is omitted and with $\text{acc}_{\text{-both\_features}}$ the accuracy if both features are omitted.
 Assuming a non-negative accuracy drop when omitting a feature, the resulting score is $\ge 0.5$.
@@
 while at the first couple of packets the confidence is not very high, towards the end of the flow it reaches values close to 1 and stays there.
-Hence, not only is the classifier able to make a reasonable classification after just a few packets, the figure also suggests that indeed later packets have a neglectible influence on the prediction.
+Hence, not only is the classifier able to make a reasonable classification after just a few packets, the figure also suggests that indeed later packets have a neglectable influence on the prediction.
 

### 2019-12-17 ex3/paper/paper.tex
Text changes

@@
 Due to the threat of adversarial attacks, a further crucial requirement for \gls{ml}-based \glspl{ids} is that the classifier's decisions are interpretable by humans.
-This recently gave increased interest in explainability methods.
+This recently stirred up increased interest in explainability methods.
 Also in this case, common methods are designed to work with images or tabular data but not with sequences.
@@
 The figure shows significant differences for detecting different attack families in the first place.
-However, also for generating adversarial flows some attack families are more susceptible than others.
+Also for generating adversarial flows some attack families are more susceptible than others.
 However, the results match to a high degree with our expectations.
@@
 Such a metric should be easy to compute, easy to interpret and robust with respect to outliers.
-We thus propose the \gls{ars}, as the average of the lower half  of minimum distances of adversarial samples.
+We thus propose the \gls{ars} as the average distance of the adversarial samples to the original samples for the 50\% of the samples which an adversary could successfully modified and which have the smallest distances.
 Since \gls{cw} is able to find adversarial samples with minimum distance, it can be used for finding the \gls{ars} efficiently.
 
-When using a large $c$ in \autoref{eq:carliniWagner}, it is necessary to reduce the gradient descent learning rate  to avoid numerical issues.
+When using a large $c$ in \autoref{eq:carliniWagner}, it is necessary to reduce the gradient descent learning rate to avoid overly large steps that impede convergence.
 However, since the distance term is not affected by $c$, it is necessary to scale the iteration count inversely with the learning rate.
@@
 If after a predefined number of iterations -- e.g.
-100 -- still not more than half of the samples are adversarial, also break.
+100 -- still not more than half of the samples are adversarial, stop.
 
 The result is a number of adversarial samples and some that could not be made adversarial.
-If at least half of the samples are misclassified, we take 50\% of the samples, with those that required smaller modifications first.
+If at least half of the samples are misclassified, we sort the adversarial samples by ascending distance to the original samples.
+Then we take samples by ascending distance until we took 50\% of all samples.
 Of these samples, the distance of each adversarial sample to the original sample is taken and averaged.
@@
 The larger it is the better since then an adversary needs to modify the samples more to make them adversarial.
-If not more then half of the samples could be made adversarial, our metric is undefined since then apparently it is not possible for an adversary to reliably conduct adversarial attacks on the majority of samples.
-Then the ratio of samples that could be made adversarial ($<$50\%) can be useful metric to determine the exact extent of the adversarial threat.
+If not more than half of the samples could be made adversarial, our metric is $\infty$ since then apparently it is not possible for an adversary to reliably conduct adversarial attacks on the majority of samples.
+Then the ratio of samples that could be made adversarial ($<$50\%) can be a useful metric to determine the exact extent of the adversarial threat.
 Setting the threshold for the \gls{ars} to 50\% is arbitrary, but
@@
 
-\autoref{fig:recall_ars} shows that certain attack types, like ``DoS Slowloris'' are easy to classify for an IDS but still are also surprisingly easy for an attacker to perform adversarial modifications to make them look benign.
+\autoref{fig:recall_ars} shows that certain attack types, like ``DoS SlowHTTPTest'', are easy to classify for an IDS but still are also surprisingly easy for an attacker to perform adversarial modifications to make them look benign.
 
@@
 However, due to the complex decision processes of deep neural networks,
-it is absolutely  uncertain if the network's final prediction is correctly depicted when following this approach.
+this does not hold true and a feature might influence a decision many time steps later.
 \end{itemize}
@@
 
-A model trained with feature dropout typically yields slightly lower accuracy than a regularly trained model, even if no features are left out (flow accuracy of 0.9943 vs.~0.9965).
+A model trained with feature dropout typically yields slightly lower accuracy than a regularly trained model, even if no features are left out (flow accuracy of 99.43\% vs.99.65\%).
 We thus recommend training two models: One regular one and one with feature dropout to use for the feature importance.
@@
 Feature dropout might constitute a building block for methods based on Shapley values \cite{shapley_value_1953} like KernelSHAP \cite{lundberg_unified_2017}. 
-Due to space limitations, we leave this for future work.
+%Due to space limitations, we leave this for future work.
 \subsubsection{Mutual Information}
@@
 
-By using a classical \gls{pdp} we would be likely to loose information due to the averaging over very different types of traffic.
+By using a classical \gls{pdp} we would be likely to lose information due to the averaging over very different types of traffic.
 However, for network traffic in particular, investigating each sample individually is not possible.
 Hence, the conditional \gls{pdp} provides the ideal compromise for our application.
-
-%
-%\begin{figure*}
-%\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_8_DoS - DDoS-Heartbleed"}.pdf}
-%\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_10_Normal"}.pdf}
-%
-%\caption{Examples of conditional PD plots for source and destination port.
-We only analyze these features with the \gls{pdp} since it is not straightforward to define the \gls{pdp} for features that can vary for each packet in a flow.
-The y-axis shows the probability for malicious traffic.}
-%\todo{Max: Remove these? Not really interesting or related to the rest of the paper. Refer to other paper where we talk about this already.}
-%\label{fig:pdp}
-%\end{figure*}
-%
-%In this variant of the \gls{pdp} that we call conditional \gls{pdp}, we don't take the whole dataset for the computation for the \gls{pdp} but only a subset that belongs to a certain class.
-Specifically, we take each attack class and compute the \gls{pdp} only over samples of this attack class.
 
@@
 it becomes apparent that traffic destined to a high destination port is generally indicative of an attack. We argue that this is because most services that regular users use have low port numbers.
-
-%AH. As ALE is removed, I case this also isnt needed
-%A known issue about \glspl{pdp} in general is that dependencies of the feature under investigation and the remaining variables are not considered, hence potentially querying the neural network for feature combinations which cannot possibly occur.
-For a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.
-
-%For this reason, Accumulated Local Effects (ALE) plots have been proposed in literature.
-The basic idea for ALE is that instead of using all samples, only the closest samples are used for assessing the effects of one feature.
-ALE plots therefore inherently require a distance measure for finding the closest samples, which, as revealed in the above discussion, in our case is not possible in a straight-forward manner.
 

### 2019-12-17 ex3/paper/paper.tex
Text changes

--- 2019-12-16 ex3/paper/paper.tex
@@
 
-$\delta \in \mathbb R$ is a parameter that determines how far an adversary wants to go over the decision boundary.
-In the original publication $\delta=0$, which means that the network's decision for adversarial samples is just between attack and normal traffic.
-In the present context, we need to make sure that the classifier's prediction would actually be normal traffic, even though a certain level of noise will be added to \glspl{iat} due to the network between attacker and victim.
-Hence, we introduced $\delta=-0.2$, corresponding to a confidence of 55\% for the sample to be benign.
+$\delta \in \mathbb R$ is a parameter that determines how far an adversary wants to exceed the decision boundary.
+In the original publication $\delta=0$, meaning that the network's decision for adversarial samples is just between attack and benign traffic.
+In the present context, we need to make sure that the classifier's prediction would actually be benign, even though a certain level of noise will be added to \glspl{iat} due to the network between attacker and victim.
+Hence, we introduced $\delta=-0.2$, corresponding to a prediction confidence of 55\% for the sample to be benign.
 
@@
 \cite{madry_towards_2018} uses a method for generating adversarial samples which deploys \gls{pgd}  to maximize the network's loss function while constraining the achieved $L_\infty$ distance from the original samples.
+Hence, an important difference to \gls{cw} is that for this method the network's loss instead of its logit output is considered.
 We consider this method an interesting combination of \gls{cw} and \gls{fgsm}.
@@
 
-It is interesting to note that increasing the $L_\infty$ bound for \gls{pgd} did not yield significantly improved performance.
-We can thus confirm the recommendation by \cite{carlini_towards_2017} to perform gradient descent on the logit output instead of the loss for good results.
+Interestingly, any increase of the $L_\infty$ bound for \gls{pgd} did not yield significantly improved performance. 
+\gls{cw} thus generally achieved superior results to $L_\infty$-bounded \gls{pgd}
+and we can confirm the recommendation by \cite{carlini_towards_2017} to perform gradient descent on the logit output instead of the loss for good results.
 
@@
 %\autoref{fig:adv_cw} depicts the ratio of attack flows which were successfully misclassified as benign with respect to the final packet along with the resulting distance from the original sample.
-Evidently, the larger the tradeoff $c$, the better the attack works, but also distances from original flows become higher.
-Hence, with acceptable distance values we were able to generate adversarial samples for about 80\% of attack samples.
-The figure suggests that a tradeoff value between 0.5 and 1 would be ideal to find adversarial samples with a low enough distance from real samples.
-This conclusion might be misleading, however, as due to the objective function \eqref{eq:carliniWagner} the distance cannot become larger anymore for one sample as soon as an adversarial sample has been found.
+Evidently, the larger $c$, the better the attack works, but also distances from original flows become higher.
+With acceptable distances, we were able to generate adversarial samples for about 80\% of attack samples.
+
+Note, however, that for successful adversarial samples, the second term in \autoref{eq:carliniWagner} becomes constant and, hence, is no longer relevant for optimization.
+Hence, $c$ is irrelevant for the achieved distance, but governs if for one sample an adversarial sample is found.
+However, to avoid numerical issues, when using a large $c$ in \autoref{eq:carliniWagner}, it is necessary to reduce the gradient descent learning rate.
+Since the distance term is not affected by $c$, it is furthermore necessary to scale the iteration count inversely with the learning rate.
+\gls{cw} with a large $c$ therefore needs more time than with a small $c$.
+
+% The figure suggests that a tradeoff value between 0.5 and 1 would be ideal to find adversarial samples with a low enough distance from real samples.
+% This conclusion might be misleading, however, as due to the objective function \eqref{eq:carliniWagner} the distance cannot become larger anymore for one sample as soon as an adversarial sample has been found.
 %, when performing optimization as depicted by the objective function .
-Hence, the depicted distance increase presumably is due to samples for which no adversarial sample could be found for lower $c$ values.
+% Hence, the depicted distance increase presumably is due to samples for which no adversarial sample could be found for lower $c$ values.
 
@@
 
-When using a large $c$ in \autoref{eq:carliniWagner}, it is necessary to reduce the gradient descent learning rate  to avoid numerical issues.
-However, since the distance term is not affected by $c$, it is necessary to scale the iteration count inversely with the learning rate.
-\gls{cw} with a large $c$ therefore needs more time than with a low tradeoff.
-For neural networks, the \gls{ars} can be efficiently found as follows: Use \gls{cw} with a small tradeoff, trying to generate adversarial samples for an attack type.
-If more than half of the samples are now wrongly classified, stop.
-Otherwise increase the tradeoff of \gls{cw} and repeat.
-If after a predefined number of iterations -- e.g.
-100 -- still not more than half of the samples are adversarial, also break.
+Since \gls{cw} becomes slow for large $c$, the \gls{ars} can be efficiently found as follows: Use \gls{cw} with a small tradeoff, trying to generate adversarial samples for an attack type.
+If, of a total of $N$ samples,
+ more than $N/2$  are wrongly classified, and
+ the smallest distance of non-adversarial samples is not smaller than  the $\lceil\frac{N}{2}\rceil$th smallest distance of adversarial samples,
+stop. Otherwise increase $c$ and repeat. If after a predefined number of iterations -- e.g. 100 -- still not more than half of the samples are adversarial, also break.
 
@@
 The larger it is the better since then an adversary needs to modify the samples more to make them adversarial.
-If not more then half of the samples could be made adversarial, our metric is undefined since then apparently it is not possible for an adversary to reliably conduct adversarial attacks on the majority of samples.
+If not more then half of the samples could be made adversarial, our metric is infinite since then apparently it is not possible for an adversary to reliably conduct adversarial attacks on the majority of samples.
 Then the ratio of samples that could be made adversarial ($<$50\%) can be useful metric to determine the exact extent of the adversarial threat.

### 2019-12-20 ex3/auxplot.py
Remove exercise 3

@@ -1,170 +1,0 @@
-#!/usr/bin/env python3
-
-# An ad-hoc script for plotting tables copy-pasted from LibreOffice Calc
-
-import fileinput
-import numpy as np
-import matplotlib.pyplot as plt
-from matplotlib.patches import Rectangle
-import sys
-from matplotlib.transforms import Affine2D
-from matplotlib.lines import Line2D
-
-group_names = []
-values = []
-
-plt.rcParams['font.family'] = 'serif'
-plt.rcParams['savefig.format'] = 'pdf'
-colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
-
-for line in fileinput.input('-'):
-	field = line.rstrip().split('\t')
-	group_names.append(field[0])
-	values.append(field[1:])
-
-feature_names = values[0]
-group_names = group_names[1:]
-values = values[1:]
-
-print("feature_names", feature_names)
-print("group_names", group_names)
-print("values", values)
-values = np.array(values, dtype=float)
-
-def importance():
-	global values
-	plt.figure(figsize=(5,4))
-	values[values < 0] = 0
-	values /= np.sum(values,axis=0)
-
-	FACTOR = 0.85
-	# width = 1/(1+values.shape[1])
-	width = FACTOR / values.shape[1]
-
-	order = np.argsort(-np.mean(values, axis=1))
-	x = np.arange(len(group_names))
-
-	for i in range(values.shape[1]):
-		plt.bar(x + width*i, values[order,i], width, alpha=0.9, label=feature_names[i])
-	for i in range(values.shape[1]):
-		plt.bar(x + width*i, np.mean(values[order,:], axis=1), width, color='gray', alpha=0.5, **({'label': 'Mean'} if i==0 else {}), zorder=0)
-
-	plt.legend()
-	plt.xticks(x + width*values.shape[1]/2-width*0.5, [group_names[i] for i in order], rotation=45, horizontalalignment='right')
-	for tick in plt.gca().xaxis.get_major_ticks():
-		label = tick.label1
-		# don't know how to specify translation in axis coordinates, so pdf output
-		# looks slightly different from plot
-		label.set_transform(label.get_transform() + Affine2D().translate(5,0))
-	plt.ylabel('Normalized metric')
-	plt.tight_layout()
-
-def adv_results():
[…]

### 2019-12-20 ex3/learn.py

@@ -1,1636 +1,0 @@
-#!/usr/bin/env python3
-
-import sys
-import os
-import pickle
-import json
-import math
-import argparse
-import random
-import time
-from functools import reduce
-from tqdm import tqdm
-
-import collections
-import itertools
-import numpy as np
-import torch
-import torch.nn as nn
-import torch.optim as optim
-from torch.utils.data import DataLoader, Dataset
-from tensorboardX import SummaryWriter
-import matplotlib.pyplot as plt
-from sklearn.metrics import accuracy_score,recall_score, precision_score, f1_score, balanced_accuracy_score
-
-ADVERSARIAL_THRESH = 50
-
-def output_scores(y_true, y_pred):
-	accuracy = accuracy_score(y_true, y_pred)
-	precision = precision_score(y_true, y_pred)
-	recall = recall_score(y_true, y_pred)
-	f1 = f1_score(y_true, y_pred)
-	youden = balanced_accuracy_score(y_true, y_pred, adjusted=True)
-	metrics = ['Accuracy', 'Precision', 'Recall', 'F1', 'Youden']
-	print (('{:>11}'*len(metrics)).format(*metrics))
-	print ((' {:.8f}'*len(metrics)).format(accuracy, precision, recall, f1, youden))
-
-def numpy_sigmoid(x):
-	return 1/(1+np.exp(-x))
-
-def pretty_print(*args, **kwargs):
-	# Print tables in a format which can be parsed easily
-	outputs = []
-	for arg in args:
-		try:
-			outputs.append(('%.6f' % arg).rstrip('0').rstrip('.'))
-		except:
-			outputs.append(arg)
-	if not "sep" in kwargs:
-		kwargs["sep"] = "\t"
-	print(*outputs, **kwargs)
-
-class OurDataset(Dataset):
-	def __init__(self, data, labels, categories):
-		self.data = data
-		self.labels = labels
-		self.categories = categories
-		assert(len(self.data) == len(self.labels) == len(self.categories))
-
-	def __getitem__(self, index):
-		data, labels, categories = torch.FloatTensor(self.data[index]), torch.FloatTensor(self.labels[index]), torch.FloatTensor(self.categories[index])
-		return data, labels, categories
-
-	def __len__(self):
-		return len(self.data)
-
-class AdvDataset(Dataset):
-	def __init__(self, base_dataset):
-		self.base_dataset = base_dataset
[…]

### 2019-12-20 ex3/mutinfo_example.py

@@ -1,34 +1,0 @@
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-import numpy as np
-
-plt.rcParams["font.family"] = "serif"
-
-resolution = 10000
-x = np.arange(0,1,1/resolution)
-
-y = (np.mod(np.arange(resolution), np.ceil(resolution/3.5)) < np.ceil(resolution/3.5)/2).astype(int)
-y2 = y.astype(float)
-y2[np.argmin(y):] = .5
-
-plt.figure(figsize=(5,1.3))
-plt.subplot(1,2,1)
-plt.plot(x,y2, c='k')
-plt.xlim((0,1))
-plt.yticks([0,.5,1])
-plt.ylim((-0.1,1.1))
-plt.xlabel('$x$')
-plt.ylabel('$f(A,x)$')
-
-plt.subplot(1,2,2)
-plt.plot(x,y, c='k')
-plt.xlim((0,1))
-plt.yticks([0,.5,1])
-plt.ylim((-0.1,1.1))
-plt.xlabel('$x$')
-plt.ylabel('$f(A,x)$')
-
-plt.tight_layout()
-plt.savefig('plots/mutinfo_example.pdf', bbox_inches = 'tight', pad_inches = 0)
-plt.show()

### 2019-12-20 ex3/parse.py

@@ -1,101 +1,0 @@
-#!/usr/bin/env python3
-
-import pandas as pd
-import numpy as np
-import itertools
-
-import pickle
-import json
-
-df = pd.read_csv('packet.csv')
-mapping = dict([(b, a) for a, b in enumerate(sorted(set(list(df["Attack"]))))])
-
-categories = ['Botnet', 'Brute Force', 'DoS', 'Infiltration', 'Normal', 'PortScan', 'Web Attack']
-
-categories_mapping = {}
-for key in mapping.keys():
-	for category in categories:
-		if category in key:
-			if not category in categories_mapping:
-				categories_mapping[category] = []
-			categories_mapping[category].append(key)
-			break
-	else:
-		if not key in categories_mapping:
-			categories_mapping[key] = []
-		categories_mapping[key].append(key)
-
-with open("categories_mapping.json", "w") as f:
-	json.dump({"categories_mapping": categories_mapping, "mapping": mapping}, f)
-
-minimal = False
-quic = False
-only_unchangeable = False
-
-all_flags = "SFRPAUECN"
-all_flags_name = ["SYN", "FIN", "RST", "PSH", "ACK", "URG", "ECE", "CWR", "NS"]
-assert len(all_flags) == len(all_flags_name)
-assert [item_long[0]==item for item_long, item in zip(all_flags_name, all_flags)]
-
-def read_list(l):
-	if isinstance(l, float) and np.isnan(l): return []
-	assert((l[0],l[-1]) == ('[', ']'))
-	return l[1:-1].split(' ')
-
-def read_numlist(l):
-	return ( [float(item)] for item in read_list(l) )
-
-def read_directionlist(l):
-	return ( [int(item == 'true')] for item in read_list(l) )
-
-def read_flaglist(l):
-	return ( [ int(f in flags) for f in all_flags] for flags in read_list(l) )
-
-const_feature_names = ["sourceTransportPort","destinationTransportPort","protocolIdentifier"]
-ip_total = 'accumulate(ipTotalLength)'
-inter_time = 'accumulate(_interPacketTimeNanoseconds)'
-direction = 'accumulate(flowDirection)'
-flags = "accumulate(_tcpFlags)"
-wrote_json = False
-
-def read_flow(row):
-	global wrote_json
-	features = []
-	const_features = [ row[item] for item in const_feature_names ]
-
-	generators = []
-	if not minimal or quic:
-		generators.append( (const_features for _ in itertools.count() ) )
[…]

### 2019-12-20 ex3/plot.py

@@ -1,156 +1,0 @@
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-from matplotlib.patches import Rectangle
-import numpy as np
-import sys
-import os
-import json
-import pickle
-
-DIR_NAME = "plots/plot"
-ORDERING = ['', 'Length min', 'IAT min', 'Length max', 'IAT max']
-
-MAX_X = 18
-SHOW_TITLE = False
-
-plt.rcParams["font.family"] = "serif"
-
-
-dataroot_basename = sys.argv[1].split('_')[0]
-
-with open(dataroot_basename + "_categories_mapping.json", "r") as f:
-	categories_mapping_content = json.load(f)
-categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
-reverse_mapping = {v: k for k, v in mapping.items()}
-print("reverse_mapping", reverse_mapping)
-
-file_name = sys.argv[1]
-with open(file_name, "rb") as f:
-	loaded = pickle.load(f)
-results_by_attack_number, sample_indices_by_attack_number = loaded["results_by_attack_number"], loaded["sample_indices_by_attack_number"]
-
-# print("results", results_by_attack_number)
-# print("sample_indices", sample_indices_by_attack_number)
-lens_results = [len(item) for item in results_by_attack_number]
-lens_indices = [len(item) for item in sample_indices_by_attack_number]
-
-print("lens_indices", "\n".join(["{}: {}".format(reverse_mapping[attack], length) for attack, length in zip(range(len(lens_indices)), lens_indices)]))
-assert lens_results == lens_indices
-
-colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
-
-all_seqs = []
-all_attacks = []
-for attack_type, seqs in enumerate(results_by_attack_number):
-	if reverse_mapping[attack_type] == 'Normal':
-		all_seqs.extend([ 1-seq for seq in seqs ])
-	else:
-		all_seqs.extend(seqs)
-		all_attacks.extend(seqs)
-	
-reverse_mapping[len(results_by_attack_number)] = 'All samples'
-results_by_attack_number.append(all_seqs)
-reverse_mapping[len(results_by_attack_number)] = 'All attacks'
-results_by_attack_number.append(all_attacks)
-	
[…]

### 2019-12-20 ex3/plot2.py

@@ -1,127 +1,0 @@
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-import matplotlib
-import matplotlib.colors as mc
-# from matplotlib import rcParams
-# rcParams.update({'figure.autolayout': True})
-import colorsys
-import numpy as np
-import sys
-import os
-import json
-import pickle
-from learn import numpy_sigmoid
-
-dataroot_basename = sys.argv[1].split('_')[0]
-
-with open(dataroot_basename + "_categories_mapping.json", "r") as f:
-	categories_mapping_content = json.load(f)
-categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
-reverse_mapping = {v: k for k, v in mapping.items()}
-# print("reverse_mapping", reverse_mapping)
-
-with open(dataroot_basename+"_full_no_ttl_normalization_data.pickle", "rb") as f:
-	means, stds = pickle.load(f)
-
-file_name = sys.argv[1]
-with open(file_name, "rb") as f:
-	loaded = pickle.load(f)
-results_by_attack_number = loaded["results_by_attack_number"]
-flows_by_attack_number = loaded["flows_by_attack_number"]
-result_ranges_by_attack_number = loaded["result_ranges_by_attack_number"]
-sample_indices_by_attack_number = loaded["sample_indices_by_attack_number"]
-features = loaded["features"]
-# print("features", features)
-
-def brighten(rgb, how_much=0.0):
-	hls = list(colorsys.rgb_to_hls(*rgb))
-	hls[1] = hls[1] + how_much*(1.0-hls[1])
-	return colorsys.hls_to_rgb(*hls)
-
-colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
-colors_rgb = [matplotlib.colors.to_rgb(item) for item in plt.rcParams['axes.prop_cycle'].by_key()['color']]
-
-COLOR_MAP_ELEMENTS = 100
-brightness_map = list(np.linspace(1.0, 0.5, num=COLOR_MAP_ELEMENTS))
-# print("brightness_map", brightness_map)
-colors_rgb_ranges = [matplotlib.colors.ListedColormap([brighten(color, item) for item in brightness_map]) for color in colors_rgb]
-# print("colors", colors)
-# print("colors_rgb_ranges[0]", colors_rgb_ranges[0])
-# quit()
-FEATURE_NAMES = ["packet length", "iat"]
-
[…]

### 2019-12-20 ex3/plot2_adv.py

@@ -1,219 +1,0 @@
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-import matplotlib
-import matplotlib.colors as mc
-from matplotlib.lines import Line2D
-import colorsys
-import numpy as np
-import sys
-import os
-import json
-import pickle
-from learn import numpy_sigmoid
-import math
-
-DIR_NAME = "plots/plot2_adv"
-
-USE_LOG_SCALE = True
-OMIT_UPPER_AXIS = True
-ONLY_ONE_LEGEND = True
-SHOW_TITLE = False
-
-plt.rcParams["font.family"] = "serif"
-dataroot_basename = sys.argv[1].split('_')[0]
-
-with open(dataroot_basename + "_categories_mapping.json", "r") as f:
-	categories_mapping_content = json.load(f)
-categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
-reverse_mapping = {v: k for k, v in mapping.items()}
-
-with open(dataroot_basename+"_normalization_data.pickle", "rb") as f:
-	means, stds = pickle.load(f)
-
-file_name = sys.argv[1]
-adv = "adv" in file_name
-with open(file_name, "rb") as f:
-	loaded = pickle.load(f)
-results_by_attack_number = loaded["results_by_attack_number"]
-flows_by_attack_number = loaded["flows_by_attack_number"]
-result_ranges_by_attack_number = loaded["result_ranges_by_attack_number"]
-sample_indices_by_attack_number = loaded["sample_indices_by_attack_number"]
-features = loaded["features"]
-
-adv_file_name = sys.argv[2]
-with open(adv_file_name, "rb") as f:
-	adv_loaded = pickle.load(f)
-adv_results_by_attack_number = adv_loaded["results_by_attack_number"]
-adv_orig_results_by_attack_number = adv_loaded["orig_results_by_attack_number"]
-adv_modified_flows_by_attack_number = adv_loaded["modified_flows_by_attack_number"]
-adv_orig_flows_by_attack_number = adv_loaded["orig_flows_by_attack_number"]
-
-colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
-ORDERING = ["original", "adversarial"]
-FEATURE_NAMES = ["Pkt. length (B)", "IAT (s)"]
-FEATURE_SCALES = [1, 0.001]
-
-def brighten(rgb, how_much=0.0):
[…]

### 2019-12-20 ex3/plot_adv.py

@@ -1,101 +1,0 @@
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-import numpy as np
-import sys
-import os
-import json
-import pickle
-from learn import numpy_sigmoid
-
-DIR_NAME = "plots/plot_adv"
-
-dataroot_basename = sys.argv[1].split('_')[0]
-
-with open(dataroot_basename + "_categories_mapping.json", "r") as f:
-	categories_mapping_content = json.load(f)
-categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
-reverse_mapping = {v: k for k, v in mapping.items()}
-# print("reverse_mapping", reverse_mapping)
-
-with open(dataroot_basename+"_full_no_ttl_normalization_data.pickle", "rb") as f:
-	means, stds = pickle.load(f)
-
-# TODO: Implement for more than one adv output so that different tradeoffs can be compared.
-file_name = sys.argv[1]
-with open(file_name, "rb") as f:
-	loaded = pickle.load(f)
-results_by_attack_number = loaded["results_by_attack_number"]
-orig_results_by_attack_number = loaded["orig_results_by_attack_number"]
-modified_flows_by_attack_number = loaded["modified_flows_by_attack_number"]
-orig_flows_by_attack_number = loaded["orig_flows_by_attack_number"]
-
-colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
-ORDERING = ["original", "adversarial"]
-FEATURE_NAMES = ["packet length", "iat"]
-
-for attack_type, (results_by_attack_number_item, orig_results_by_attack_number_item, modified_flows_by_attack_number_item, orig_flows_by_attack_number_item) in enumerate(zip(results_by_attack_number, orig_results_by_attack_number, modified_flows_by_attack_number, orig_flows_by_attack_number)):
-
-	assert len(results_by_attack_number_item) == len(orig_results_by_attack_number_item) == len(modified_flows_by_attack_number_item) == len(orig_flows_by_attack_number_item)
-	if len(results_by_attack_number_item) <= 0:
-		continue
-
[…]

### 2019-12-20 ex3/plot_features.py

@@ -1,139 +1,0 @@
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-import matplotlib
-import matplotlib.colors as mc
-# from matplotlib import rcParams
-# rcParams.update({'figure.autolayout': True})
-import colorsys
-import numpy as np
-import sys
-import os
-import json
-import pickle
-from matplotlib.lines import Line2D
-from matplotlib.patches import Rectangle
-from learn import numpy_sigmoid
-
-plt.rcParams['font.family'] = 'serif'
-plt.rcParams['savefig.format'] = 'pdf'
-
-DIR_NAME = "plots/plot_features"
-
-dataroot_basename = sys.argv[1].split('_')[0]
-
-with open(dataroot_basename + "_categories_mapping.json", "r") as f:
-	categories_mapping_content = json.load(f)
-categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
-reverse_mapping = {v: k for k, v in mapping.items()}
-# print("reverse_mapping", reverse_mapping)
-
-with open(dataroot_basename+"_normalization_data.pickle", "rb") as f:
-	means, stds = pickle.load(f)
-
-file_name = sys.argv[1]
-with open(file_name, "rb") as f:
-	loaded = pickle.load(f)
-results_by_attack_number = loaded["results_by_attack_number"]
-flows_by_attack_number = loaded["flows_by_attack_number"]
-result_ranges_by_attack_number = loaded["result_ranges_by_attack_number"]
-sample_indices_by_attack_number = loaded["sample_indices_by_attack_number"]
-features = loaded["features"]
-# print("features", features)
-
-def brighten(rgb, how_much=0.0):
-	hls = list(colorsys.rgb_to_hls(*rgb))
-	hls[1] = hls[1] + how_much*(1.0-hls[1])
-	return colorsys.hls_to_rgb(*hls)
-
-colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
-colors_rgb = [matplotlib.colors.to_rgb(item) for item in plt.rcParams['axes.prop_cycle'].by_key()['color']]
-
-COLOR_MAP_ELEMENTS = 100
-brightness_map = list(np.linspace(1.0, 0.5, num=COLOR_MAP_ELEMENTS))
-# print("brightness_map", brightness_map)
-colors_rgb_ranges = [matplotlib.colors.ListedColormap([brighten(color, item) for item in brightness_map]) for color in colors_rgb]
[…]

### 2019-12-20 ex3/plot_new.py

@@ -1,121 +1,0 @@
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-import matplotlib
-import matplotlib.colors as mc
-# from matplotlib import rcParams
-# rcParams.update({'figure.autolayout': True})
-import colorsys
-import numpy as np
-import sys
-import os
-import json
-import pickle
-from learn import numpy_sigmoid
-
-DIR_NAME = "plots/plot_new"
-
-dataroot_basename = sys.argv[1].split('_')[0]
-
-with open(dataroot_basename + "_categories_mapping.json", "r") as f:
-	categories_mapping_content = json.load(f)
-categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
-reverse_mapping = {v: k for k, v in mapping.items()}
-# print("reverse_mapping", reverse_mapping)
-
-with open(dataroot_basename+"_full_no_ttl_normalization_data.pickle", "rb") as f:
-	means, stds = pickle.load(f)
-
-file_name = sys.argv[1]
-with open(file_name, "rb") as f:
-	loaded = pickle.load(f)
-results_by_attack_number = loaded["results_by_attack_number"]
-flows_by_attack_number = loaded["flows_by_attack_number"]
-result_ranges_by_attack_number = loaded["result_ranges_by_attack_number"]
-sample_indices_by_attack_number = loaded["sample_indices_by_attack_number"]
-features = loaded["features"]
-
-lens_results = [len(item) for item in results_by_attack_number]
-lens_indices = [len(item) for item in sample_indices_by_attack_number]
-
-print("lens_indices", "\n".join(["{}: {}".format(reverse_mapping[attack], length) for attack, length in zip(range(len(lens_indices)), lens_indices)]))
-
-def brighten(rgb, how_much=0.0):
-	hls = list(colorsys.rgb_to_hls(*rgb))
-	hls[1] = hls[1] + how_much*(1.0-hls[1])
-	return colorsys.hls_to_rgb(*hls)
-
-colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
-colors_rgb = [matplotlib.colors.to_rgb(item) for item in plt.rcParams['axes.prop_cycle'].by_key()['color']]
-
-COLOR_MAP_ELEMENTS = 100
-brightness_map = list(np.linspace(1.0, 0.5, num=COLOR_MAP_ELEMENTS))
-# print("brightness_map", brightness_map)
[…]

### 2019-12-20 ex3/plot_pdp.py

@@ -1,81 +1,0 @@
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-import numpy as np
-import sys
-import os
-import json
-import pickle
-from collections import Counter
-
-DIR_NAME = "plots/plot_pdp"
-
-dataroot_basename = sys.argv[1].split('_')[0]
-
-with open(dataroot_basename + "_categories_mapping.json", "r") as f:
-	categories_mapping_content = json.load(f)
-categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
-reverse_mapping = {v: k for k, v in mapping.items()}
-print("reverse_mapping", reverse_mapping)
-
-file_name = sys.argv[1]
-with open(file_name, "rb") as f:
-	loaded = pickle.load(f)
-results_by_attack_number, feature_names, feature_values_by_attack_number = loaded["results_by_attack_number"], loaded["feature_names"], loaded["feature_values_by_attack_number"]
-
-# print("results", results_by_attack_number)
-
-colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
-
-for attack_type, (all_features, all_features_values) in enumerate(zip(results_by_attack_number, feature_values_by_attack_number)):
-
-	print("attack_type", attack_type)
-	fig, ax1 = plt.subplots()
-
-	ax2 = ax1.twinx()
-
-	ax2.set_ylabel('Prediction')
-	ax1.set_ylabel("Number of samples")
-
-	ax1.yaxis.tick_right()
-	ax1.yaxis.set_label_position("right")
-	ax2.yaxis.tick_left()
-	ax2.yaxis.set_label_position("left")
-
-	if all_features is None:
-		continue
-	# print("all_features.shape", all_features.shape)
-	all_legends = []
-	for feature_name, feature_index in zip(feature_names, range(all_features.shape[0])):
-
-		as_ints = list(all_features_values[feature_index].astype(np.int32))
-
-		# print("all_features_values[feature_index]", all_features_values[feature_index])
-		# ret1 = ax1.hist(all_features_values[feature_index], bins=range(int(round(all_features_values[feature_index].max())+1)), width=1, color=colors[feature_index], alpha=0.2, label="{} occurrence".format(feature_name))
-
-		counted = Counter(as_ints)
-		keys = counted.keys()
-		values = counted.values()
-
[…]

### 2019-12-20 ex3/plots/adv_training/plot.py

@@ -1,16 +1,0 @@
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-import pandas
-
-data_l1 = pandas.read_csv('run-Aug10_18-59-17_gpu-tag-adv_avdistance.csv')
-data_l2 = pandas.read_csv('run-Aug10_18-59-18_gpu-tag-adv_avdistance.csv')
-
-plt.plot(data_l1['Value'])
-plt.plot(data_l2['Value'])
-
-plt.legend(['Training with $\ell_1$ distance', 'Training with $\ell_2$ distance'])
-plt.ylabel('Average $\ell_1$ distance')
-plt.xlabel('Adversarial training step')
-plt.savefig('avdistance.pdf')
-#plt.show()

### 2019-12-20 ex3/weight_feat_imp.py

@@ -1,72 +1,0 @@
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-import torch
-import numpy as np
-import sys
-import json
-from learn import pretty_print
-
-hidden_size = 512
-n_layers = 3
-absolute_values = True
-
-assert len(sys.argv) > 1, 'No model specified'
-
-parameters = torch.load(sys.argv[1])
-
-with open('features_meaningful_names.json', 'r') as f:
-	feature_array = json.load(f)
-
-#	parameters = lstm_module.state_dict()
-weights_hh = [ parameters['lstm.weight_hh_l%d' % layer].cpu().numpy().reshape(4,hidden_size,hidden_size).sum(axis=0) for layer in range(n_layers) ]
-weights_ih = [ parameters['lstm.weight_ih_l%d' % layer].cpu().numpy().reshape(4,hidden_size,-1).sum(axis=0) for layer in range(n_layers) ]
-h2o = parameters['h2o.weight'].cpu().numpy()
-
-if absolute_values:
-	weights_hh = [ np.abs(weights) for weights in weights_hh ]
-	weights_ih = [ np.abs(weights) for weights in weights_ih ]
-	h2o = np.abs(h2o)
-	
-
-input_featimp = [ weights_ih[0] ]
-for i in range(1,n_layers):
-	input_featimp.append(np.matmul(weights_ih[i], input_featimp[-1]))
-input_featimp.append(np.matmul(h2o, input_featimp[-1]))
-
-output_featimp = [ h2o ]
-for i in range(n_layers-1,-1,-1):
-	output_featimp.append(np.matmul(output_featimp[-1], weights_ih[i]))
-output_featimp.reverse()
-#  assert (input_featimp[-1] == output_featimp[0]).all()
-
-fi_over_time = [ np.abs(input_featimp[-1][0,:]) ]
-hh_prod = weights_hh
-for _ in range(10):
-	hh_prod = [ np.matmul(hh_prod[i], weights_hh[i]) for i in range(n_layers) ]
-	fi_over_time.append(np.abs(np.sum([ np.matmul(output_featimp[i+1], np.matmul(hh_prod[i], input_featimp[i])) for i in range(n_layers)], axis=0)[0,:]))
-
-print ('Packet feature importance:')
-
-for feature_index, feature_importance in sorted(enumerate(fi_over_time[0]/np.sum(fi_over_time[0])), key=lambda item: item[1]):
[…]
